### 2019-12-04 draft.tex
First draft

\documentclass[prl,twocolumn,amsmath,amssymb,superscriptaddress]{revtex4-1}

\usepackage[pdftex]{graphics}
\usepackage{graphicx}
\graphicspath{{figures/}}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{physics}
\usepackage{subfig}
\usepackage{bm}
\usepackage{caption}

\newcommand{\carlos}[1]{{\color{red} #1}}
\newcommand{\maria}[1]{{\color{blue} #1}}

\begin{document}

\title{Something fancy...}
\author{Carlos Ortega Taberner}
\affiliation{Department of Physics, Stockholm University, AlbaNova University Center, SE-106 91 Stockholm, Sweden}
\affiliation{Nordita, KTH Royal Institute of Technology and Stockholm University, SE-106 91 Stockholm, Sweden}

\author{Maria Hermanns}
\affiliation{Department of Physics, Stockholm University, AlbaNova University Center, SE-106 91 Stockholm, Sweden}
\affiliation{Nordita, KTH Royal Institute of Technology and Stockholm University, SE-106 91 Stockholm, Sweden}
\date{\today}

\maketitle

\section{Introduction.}

\section{Model}

We will consider a modification of the model in the BDI class used in reference \cite{Song2014}, which shows a more interesting phase diagram and entanglement spectrum, given by the Hamiltonian
\begin{align}
H =& \sum_{i\alpha,j\beta} c_{i\alpha}^\dagger H_{ij,\alpha \beta} c_{j\beta} \\
H_{ij} =& (m \sigma_x + \kappa \sigma_z)\delta_{ij}  + \frac{1}{2i}\kappa'\sigma_z (\delta_{i-j,1}-\delta_{i-j,-1})\\
&+ \frac{1}{2} t \left[(\sigma_x + i \sigma_y)\delta_{i,j+1} + (\sigma_x - i \sigma_y) \delta_{i,j-1} \right] \\
&+  \frac{1}{2} t' \left[(\sigma_x + i \sigma_y)\delta_{i,j+2} + (\sigma_x - i \sigma_y) \delta_{i,j-2} \right].
\label{bdi_model}
\end{align}
The corresponding Bloch Hamiltonian is given by
\begin{align*}
H(k)=&\mqty( \kappa + \kappa' \sin(k) & t' e^{i2k} + t e^{ik}+m \\t' e^{-i2k} + t e^{-ik}+m & -\kappa-\kappa' \sin(k)  ) \\
=& (t' \cos(2k)+t\cos(k)+m)\sigma_x \\
&+ (-t' \sin(2k)-t\sin(k))\sigma_y + (\kappa+\kappa'\sin(k)) \sigma_z.
\end{align*}
For $\kappa = \kappa' = 0$ this Hamiltonian has the following symmetries
\begin{alignat*}{2}
&T = \mathcal{K} ; \quad &&T H(-k) T^{-1} = H(k) \\
&C = \sigma_z\mathcal{K} ; \quad &&C H(-k) C^{-1} = -H(k) \\
&S = \sigma_z ; \quad &&S H(k)S^{-1} = -H(k) 
\end{alignat*}
and therefore it is in the BDI class, where the system has the phase diagram shown in Fig.\ref{bdi_phase_diagram} for $t=0,1$. In Fig.\ref{bdi20} we show a cut of the transition $\nu = 2 \rightarrow 0$ for different values of $\kappa,\kappa'$. For $\kappa \neq 0$ only $T$ is preserved and the system is in the trivial AI class, where the midgap states split and the system acquires a finite geometric phase. For $\kappa' \neq 0$ only $C$ is preserved and the system belongs now to the $D$ class. The $Z$ invariant transforms into a $Z_2$ invariant and the original $\nu = 2,0$ phases become equivalent. We observe split midgap states but the Berry phase is still quantized and vanishes for this cut.

\begin{figure}[h!]
\centering
\subfloat[$t = 1$]{
  \includegraphics[width=35mm]{bdit1}
}
\subfloat[$t=0$]{\includegraphics[width=35mm]{bdit0}
}
\caption{Phase diagram for the system in the $BDI$ class showing the winding number of each region.}
\label{bdi_phase_diagram}
\end{figure}

\section{Homogeneous system}

For symmetry-protected topological systems in one dimension there is a well-known relation between the entanglement spectrum  (ES) and the geometric phase. The geometric phase, which is quantized, is zero whenever there is an even number per edge of eigenvalues at $\xi = 1/2$, and $\pi$ when the number is odd. Beyond symmetry protected systems there have been observations of similarities between the midgap eigenvalue of the entanglement spectrum and the geometric phase \cite{Ryu2006,Huang2012,Huang2012-2}, which in the case of the fully dimerized model becomes an exact identity \cite{Ryu2006}.

The entanglement spectrum eigenstates can be classified as bulk (with eigenvalues $\xi_{b} = 0,1 $) or edge states (with eigenvalues $\xi_{l,r} \neq 0,1$), which are localized in one of the edges. Note that any eigenstate with eigenvalue different from zero or one we refer to as edge state, and not only the ones at $\xi=1/2$. Consider the combination of all eigenvalues localized in one of the edges modulo one,
\begin{equation}
\chi = \sum_{i \in l} \xi_l \quad {\rm mod} \quad 1. 
\end{equation}
As we will show below, for systems where the ES has more than one edge state per edge one can find the geometric phase in the thermodynamic limit as
\begin{equation}
\lim_{L \rightarrow \infty}\gamma/2\pi =\lim_{L \rightarrow \infty} \chi
\label{mainl}
\end{equation}
In the examples studied in references \cite{Ryu2006,Huang2012,Huang2012-2} most of the edge states sit close to the bulk bands and therefore give a small contribution, which explains why the relation between the midgap eigenvalue and the geometric phase was almost an identity and why it breaks as we go deep into the trivial state. Alternatively, using that the entanglement spectrum is particle-hole symmetric for homogeneous systems we can express $\chi$ as
\begin{equation*}
\chi = \frac{1}{2} \left( \sum_{i\in l}\xi_i-\sum_{i\in r}\xi_i \pm \sum_{i\in b}\xi_i  + N \right) \quad {\rm mod} \quad 1,
\end{equation*}
where $N$ is the total number of electrons in the ground state and the sign on the bulk term is irrelevant since $-1/2 \, {\rm mod} \, 1 = 1/2 \, {\rm mod} \, 1$. This form is preferable when doing numerics because it is difficult to differentiate between edge modes and bulk modes since the latter can acquire a finite eigenvalue due to finite size effects or numerical errors. In Fig.\ref{bdi20} we show for several cases how Eq. \ref{mainl} holds up to finite size errors. The ES does not have therefore enough information to build up the geometric phase, but one needs to look at the localization structure of the edge states. One drastic example of this can be seen in figures \ref{bdi20}.b and \ref{bdi20}.c , where two ES that are almost identical result in a very different geometric phase.

\begin{figure}[h!]
\centering
\makebox[0pt]{
\subfloat[$t = 0, t' = 2, \kappa = 10^{-5}$]{
  \includegraphics[width=35mm]{bdi20.pdf}
}
\subfloat[$t = 0, t' = 2, \kappa = 0.3$]{
  \includegraphics[width=35mm]{ai20.pdf}
}
}\hspace{0mm}

\makebox[0pt]{
\subfloat[$t = 0, t' = 2, \kappa = 0.3$]{
  \includegraphics[width=35mm]{d20.pdf}
}
}
\caption{Entanglement spectrum (black), Berry phase (red) and $\chi$ (blue) for different cuts of the $H_{BDI}$. a) with $\kappa = 0,\kappa'=0$ in the $AIII$ class, b) with $\kappa'=0$ in the $AI$ class and c) with $\kappa = 0$ in the $D$ class. }
\label{bdi20}
\end{figure}

This relation between the geometric phase and the entanglement spectrum is a consequence of an identity between the geometric phase and the many-body entanglement spectrum found for infinite chains \cite{Zaletel2014}. In Appendix A we modified this derivation to account for the periodic boundary conditions we use and show how their result reduces to Eq. \ref{mainl} when expressing it in terms of the ES.

\begin{figure}[h]
\centering
\makebox[0pt]{
\subfloat[$t = 0, t' = 2, \kappa = 0.3$]{
  \includegraphics[width=65mm]{scaling_cut_disavg.png}
}
}
\caption{}
\label{disorder}
\end{figure}

\section{Inhomogeneous systems}

In reference \cite{Zaletel2014} translational invariance was used to derive Eq. \ref{zaletel} in Appendix A, and as a result Eq.\ref{mainl} is no longer correct when it is broken. This can be argued in the following way. Weak disorder will modify the entanglement spectrum but not the charge assignment. Without translational invariance the entanglement spectrum will depend on where is the cut placed on the ring, even though the ground state, and therefore the geometric phase, will not. Therefore there cannot be an equivalence between them. Instead, we found that one can still obtain the geometric phase by considering the average over all spatial cuts of expression \ref{mainl}.

In order to test this we take the model we used in the previous section and we promote every parameter to a position-dependent quantity, by adding a random part. In Fig.\ref{disorder} we show $\gamma-\expval{\chi}_{\rm cuts}$ for increasing system size. Note that we do the calculation for a particular disorder configuration for each size. We found that $\expval{\chi}_{\rm cuts}$ converges much faster than the geometric phase and that the difference scales as $O(L^{-2})$, which is the same scaling we find in the homogeneous system. We compared this result with a disorder average $\gamma - \expval{\chi}_{\rm{disorder}}$ and found that, even though we are comparing different disorder configurations, the disorder average and the cut average give the same error, which implies that it must come exclusively from finite size effects in computing the geometric phase, since is the only thing in common. One could think that this cut average is some sort of disorder average, but that is however not correct. In Fig.\ref{cuts} we show the case with a continuous position-dependent potential, plotted in Fig.\ref{cuts}.a, on top of the homogeneous model. In Fig.\ref{cuts}.b we show $\chi_i$ for two cuts where the left edge is either in the positive or negative potential region. As we can see in Fig.\ref{cuts}.c for any intermidiate cut, $\chi_i$ goes continuously from one value to the other. Here we encounter a new problem, since $\chi_i$ only makes physical sense as an angle, there are two paths between the two points. Below the phase transition we find that $\chi_i$ oscillates around $1/2$ while above the phase transition it does so around $0$. By redifining the range of values so that they do not cross $0$ and then averaging we obtain the same result as for the disordered system, that $\gamma-\expval{\chi}_{\rm cuts} \sim O(L^{-2})$. This is only well defined in the continuous limit and can potentially pose problems for finite systems and discontinuous terms, like with disorder. However, as long as the deviations are small and we have $|\expval{\gamma-\chi}|<1/2$, which has been the case so far, one only has to consider two averages, the one with $\chi_i \in [0,1)$ and the one with $\chi_i \in [-1/2,1/2)$, although at the moment there is no indication of which of the two numbers will result on the geometric phase.

\begin{figure}[h!]
\centering
\makebox[0pt]{
\subfloat[]{
  \includegraphics[width=40mm]{profile.pdf}
}
}\hspace{0mm}

\makebox[0pt]{
\subfloat[]{
  \includegraphics[width=40mm]{extreme_es.pdf}
}
\subfloat[]{
  \includegraphics[width=40mm]{cut_es.pdf}
}
}

\caption{a)Profile used for $\kappa_i$ . b)$\chi_i$ for two cases where we place the left cut deep in either the region with positive or negative $\kappa_i$, and geometric phase in red, showing a phase transition. c)$\chi_i$ for all different possible cuts using the parameters $m=1$ (blue) and $m=3$ (green)}
\label{cuts}
\end{figure}

\section{Superconducting systems}

The proof in Ref.\cite{Zaletel2014} also breaks down for superconducting systems. However, as we did for systems without translational invariance, one can extend expression \ref{mainl} for systems with pairing terms as well. Since one computes the geometric phase from the single-particle Bogoliubov-de Gennes Hamiltonian, we can obtain the correspondent entanglement spectrum and use expression \ref{mainl} to obtain the geometric phase. Take for example the kitaev chain, with the following Hamiltonian
\begin{align*}
H =& -\mu \sum_{i} c_i^\dagger c_i - t \sum_{i}\left( c_{i+1}^\dagger c_i + {\rm h.c.}\right) \\
&+  \sum_{i}\left( \Delta c_{i} c_{i+1} + {\rm h.c.}\right).
\end{align*}
Using the spinor $\phi_i^\dagger = (c_i^\dagger, c_i)$, we can rewrite the Hamiltonian as
\begin{align*}
&H = \frac{1}{2}\sum_i \phi^\dagger_i H_{BdG,ij} \phi_j,\\
&H_{BdG,ij} = -\mu \tau_z \delta_{ij} - (t \tau_z + i\Delta \tau_y )\delta_{i,j+1}- (t \tau_z - i\Delta \tau_y)\delta_{i,j-1}.
\end{align*}
Note that by taking $t' = \kappa = \kappa = 0$ in the BDI model, Eq. \ref{bdi_model}, we obtain
\begin{align*}
H_{ij} =& m \sigma_x\delta_{ij} + \frac{1}{2} t \left[(\sigma_x + i \sigma_y)\delta_{i,j+1} + (\sigma_x - i \sigma_y) \delta_{i,j-1} \right],
\end{align*}
which is just the Bogoliubov-de Gennes Hamiltonian of the Kitaev chain by rotating $x$ into $z$ and identifying $\mu = -m, t_{\rm Kit} = -t_{BDI}/2, \Delta = -t_{BDI}/2 $. From Fig.\ref{bdi_phase_diagram} we see that the $BDI$ model transitions from $\nu = 1$ to the trivial phase at $t=0, m=\pm t_{BDI}$, or alternatively, when $\mu = \pm 2 t_{Kit}$, which is the known result for the Kitaev chain. Therefore the results already shown for the BDI model studied here are also applicable to the Kitaev chain.

\bibliography{references}{}
\bibliographystyle{plain}

\appendix

\section{Appendix A}

Consider a chain with periodic boundary conditions. A Schmidt decomposition for two bipartite cuts gives the ground state
\begin{equation}
\ket{\psi} = \sum_{p,q} \ket{p,q}_A s_p \ket{p,q}_B s_q,
\end{equation}
where we assume that the size of each subsystem is big enough so that the cuts can be considered independent of each other. We can now take subsystem B and glue its ends together to obtain the ground state of a ring with a single cut. If we include a flux threading inside the ring we obtain
\begin{equation}
\ket{\psi^\phi} = \sum_p s_p \ket{p,p}_B e^{i\phi Q_p},
\end{equation}
The many-body states $\ket{p,p}_B$ can be built by occupying the different ES eigenstates and are therefore characterized by a set of occupying numbers $\{n\}_p$. The many-body entanglement spectrum can also be expressed in terms of the single-particle one \cite{Alexandrinata2011} as
\begin{align}
s_p^2 =& \prod_{i \in {\rm occ}} \xi_i \prod_{i \in {\rm empty}}(1-\xi_i) \\
=& \prod_i (1-\xi_i)\left(\frac{\xi_i}{1-\xi_i} \right)^{n_i}
\end{align}
And the charge can be assigned so that $Q_p = \sum_{\alpha = b,l,r} q_\alpha\sum_{i \in \alpha} n_i$ with the constrain that $q_r,\pm q_b = q_l-1$ \textcolor{red}{Why?}. As the main result of \cite{Zaletel2014} the geometric phase is then found to be
\begin{align}
e^{i\gamma} = e^{2\pi i \sum_p s_p^2 Q_p}  \bra{\psi^0}\ket{\psi^{2\pi}},
\label{zaletel}
\end{align}
where the monodromy is given by
\begin{align}
\bra{\psi^0}\ket{\psi^{2\pi}} =& \sum_p s_p^2 e^{2\pi i Q_p}.
\end{align}
Consider the exponential term
\begin{align*}
e^{2\pi i (q_L N^p_l + (1-q_l)(N^p_r+N^p_b))} =& e^{2\pi i (q_l N^p_l + (1-q_l)(N-N^p_l))} \\
=& e^{4\pi i q_l N^p_l}e^{2 \pi i (N-N^p_l)}e^{-2\pi i q_l N },
\end{align*}
where $N^p_\alpha = \sum_{i \in \alpha} n_i$ and the total number of electrons, $N = \sum_{\alpha = b,l,r} N^p_\alpha$, is independent of $p$. We will consider two natural choices, $q_l = 1,1/2$, for which we have
\begin{align}
\bra{\psi^0}\ket{\psi^{2\pi}} =& 
\begin{cases}
1 & \text{for} \quad q_l = 1 \\
e^{N\pi i} & \text{for} \quad q_l = 1/2,
\end{cases}
\end{align}
where we used that $\sum_p s_p^2 = 1$. The change in the monodromy between both choices is accompanied by a change in the exponential term in Eq.\ref{zaletel}, so that both charge choices are equivalent.

Expressing the contribution of the many-body entanglement spectrum to the geometric phase in terms of the ES, we have
\begin{align*}
\gamma' = 2\pi \sum_{n_i = 0,1} \left(\sum_{i } n_i q_i\right) \left(\prod_m (1- \xi_m)\left(\frac{\xi_m}{1-\xi_m} \right)^{n_m} \right) ,
\end{align*} 
Consider now its derivative with respect to a particular entanglement eigenvalue
\begin{align}
\frac{1}{2\pi}\frac{\partial \gamma'}{\partial \xi_p} =& \sum_{\{n_k\}=0,1}\left(\sum_{i } n_i q_i\right)\frac{n_p - \xi_p}{\xi_p^{1-n_p}(1-\xi_p)^{n_p}}\\
&\times \left(\prod_{m\neq p} (1- \xi_m)\left(\frac{\xi_m}{1-\xi_m} \right)^{n_m} \right)
\end{align}
where
\begin{equation}
 \frac{n_p - \xi_p}{\xi_p^{1-n_p}(1-\xi_p)^{n_p}} = 
  \begin{cases}
    -1, & \text{for } n_p=0 \\
    1, & \text{for } n_p = 1 \\
  \end{cases}	
\end{equation}
Therefore we have
\begin{align*}
\frac{1}{2\pi}\frac{\partial \gamma'}{\partial \xi_p} =& -\sum_{\substack{\{n_{k\neq p}\}=0,1 \\ n_p = 0}}\left(\sum_{i \neq p} n_i q_i\right)\prod_{m\neq p} (1-\xi_m)\left( \frac{\xi_m}{1-\xi_m} \right)^{n_m} \\
&+ \sum_{\substack{\{n_{k\neq p}\}=0,1 \\ n_p = 1}}\left(\sum_{i \neq p} n_i q_i +q_p\right) \prod_{m\neq p} (1-\xi_m)\left( \frac{\xi_m}{1-\xi_m} \right)^{n_m} \\
=&q_p\sum_{\{n_{k\neq p}\}=0,1} \prod_{k\neq p} (1-\xi_k)\left( \frac{\xi_k}{1-\xi_k} \right)^{n_k} \\
\end{align*}
Expand now the sum  for another occupation number $n_{p'} = 0,1$ 
\begin{align*}
\frac{1}{2\pi}\frac{\partial \gamma'}{\partial \xi_p} =& q_p\sum_{\substack{n_{k \neq p \neq p'}=0,1\\ n_{p'}= 0}} (1-\xi_{p'})\prod_{k\neq p,p'} (1-\xi_k)\left( \frac{\xi_k}{1-\xi_k} \right)^{n_k} \\
&+ q_p\sum_{\substack{n_{k \neq p \neq p'}=0,1\\ n_{p'}= 1}} \xi_{p'}\prod_{k\neq p,p'} (1-\xi_k)\left( \frac{\xi_k}{1-\xi_k} \right)^{n_k} \\
 &=q_p\sum_{\substack{n_{k \neq p \neq p'}=0,1}} \prod_{k\neq p,p'} (1-\xi_k)\left( \frac{\xi_k}{1-\xi_k} \right)^{n_k} 
\end{align*}
doing this for the rest of the occupation numbers we obtain 
\begin{equation}
\frac{1}{2\pi}\frac{\partial \gamma'}{\partial \xi_p} = q_p,
\end{equation}
so that 
\begin{align*}
\gamma =& \gamma' - i\log(\bra{\psi^0}\ket{\psi^{2\pi}})\\
=& \sum_{i}q_i \xi_i - i\log(\bra{\psi^0}\ket{\psi^{2\pi}})\\
\end{align*}
and we have recovered equation \ref{mainl}.

\end{document}

### 2020-01-15 draft.tex
changed bibtex style
..
prueba

@@
-\documentclass[prl,twocolumn,amsmath,amssymb,superscriptaddress]{revtex4-1}
+                       \documentclass[prb,twocolumn,amsmath,longbibliography,amssymb,superscriptaddress]{revtex4-1}
 
@@
 
-We will consider a modification of the model in the BDI class used in reference \cite{Song2014}, which shows a more interesting phase diagram and entanglement spectrum, given by the Hamiltonian
+As a starting point we consider the model in the BDI class used in reference \cite{Song2014} because it is the simplest model which supports topological phases with winding numbers $\nu = 0,1,2$.
+Additionaly, we include two symmetry breaking terms, $\kappa$ and $\kappa'$, to obtain an even richer phase diagram and show that our results are not limited to any topological class.
+The Hamiltonian is then given by
 \begin{align}
@@
 &+ \frac{1}{2} t \left[(\sigma_x + i \sigma_y)\delta_{i,j+1} + (\sigma_x - i \sigma_y) \delta_{i,j-1} \right] \\
-&+  \frac{1}{2} t' \left[(\sigma_x + i \sigma_y)\delta_{i,j+2} + (\sigma_x - i \sigma_y) \delta_{i,j-2} \right].
+&+  \frac{1}{2} t' \left[(\sigma_x + i \sigma_y)\delta_{i,j+2} + (\sigma_x - i \sigma_y) \delta_{i,j-2} \right],
 \label{bdi_model}
 \end{align}
-The corresponding Bloch Hamiltonian is given by
+with the corresponding Bloch Hamiltonian being
 \begin{align*}
@@
 and therefore it is in the BDI class, where the system has the phase diagram shown in Fig.\ref{bdi_phase_diagram} for $t=0,1$.
-In Fig.\ref{bdi20} we show a cut of the transition $\nu = 2 \rightarrow 0$ for different values of $\kappa,\kappa'$.
-For $\kappa \neq 0$ only $T$ is preserved and the system is in the trivial AI class, where the midgap states split and the system acquires a finite geometric phase.
+We also show cuts with the transitions $\nu = 2 \rightarrow 0$ and $\nu = 1 \rightarrow 2$ for different values of $\kappa,\kappa'$.
 For $\kappa' \neq 0$ only $C$ is preserved and the system belongs now to the $D$ class.
 The $Z$ invariant transforms into a $Z_2$ invariant and the original $\nu = 2,0$ phases become equivalent.
-We observe split midgap states but the Berry phase is still quantized and vanishes for this cut.
-
+For $\kappa \neq 0$ only $T$ is preserved and the system is in the trivial AI class, where all zero-energy modes split.
 \begin{figure}[h!]
 \centering
-\subfloat[$t = 1$]{
+\makebox[0pt]{
+\subfloat[$t = 1,\kappa =\kappa'=0$]{
   \includegraphics[width=35mm]{bdit1}
 }
-\subfloat[$t=0$]{\includegraphics[width=35mm]{bdit0}
+\subfloat[$t=0,\kappa = \kappa'=0$]{\includegraphics[width=35mm]{bdit0}
 }
-\caption{Phase diagram for the system in the $BDI$ class showing the winding number of each region.}
+}\hspace{0mm}
+
+\makebox[0pt]{
+\subfloat[$t = 1,t'=1,\kappa =\kappa'=0$]{
+  \includegraphics[width=35mm]{1_c.pdf}
+}
+\subfloat[$t=0,t'=1,\kappa = \kappa'=0$]{\includegraphics[width=35mm]{1_d.pdf}
+}
+}\hspace{0mm}
+
+\makebox[0pt]{
+\subfloat[$t = 1,t'=1$, $\kappa = 0,\kappa'=0.3$]{
+  \includegraphics[width=35mm]{1_f.pdf}
+}
+\subfloat[$t=0,t'=1$, $\kappa = 0.3,\kappa'=0$]{\includegraphics[width=35mm]{1_g.pdf}
+}
+}
+\caption{Phase diagram for the system in the $BDI$ class showing the winding number of each region and surface energy spectrum for two different cuts in the phase diagram for different values of $\kappa,\kappa'$.}
 \label{bdi_phase_diagram}
 \end{figure}
+
+\section{Geometric phase and Entanglement spectrum}
+Consider a quadratic Hamiltonian in one dimension
+\begin{equation}
+\mathcal{H} = \sum_{ij,\alpha\beta} c_{i\alpha}^\dagger H_{ij,\alpha \beta}c_{j\beta}.
+\end{equation}
+We can obtain the single-particle eigenstates as
+\begin{align*}
+\sum_{j\beta}H_{ij,\alpha\beta} u^{p\mu}_{j\beta} = E_{p\mu} u_{i\alpha}^{p\mu},
+\end{align*}
+where $[U]_{j\beta,p\mu} = u^{p\mu}_{j\beta}$ is the unitary matrix that diagonalizes $H$.
+The polarization can be obtained as the many-body average of position operator, regularized for periodic boudnary conditions \carlos{resta}
+\begin{align*}
+\expval{X} = -\frac{L}{2\pi} \rm{ Im\, Ln }\bra{\Psi_0}e^{i \frac{2\pi}{L}\hat{X}}\ket{\Psi_0},
+\end{align*}
+where $\ket{\Psi_0}$ is the many-body ground state and $\hat{X}$ is the sum of all the single-particle position operators.
+Using that the ground state is a Slater determinant of the single-particle eigenstates it can be rewritten as
+\begin{align*}
+\expval{\hat{X}} = -\frac{L}{2\pi} \rm{Im \, Ln}\, {\rm det} S,
+\end{align*}
+where 
+\begin{equation}
+S_{p\mu,q\nu} = \sum_{j\alpha}u^{p\mu\, \ast}_{j \alpha} e^{-i\frac{2\pi}{L}j}u^{q\nu}_{j \alpha}
+\end{equation}
+and the determinant is taken restricting $S$ to only the occupied eigenstates. The geometric phase can now be obtained as
+\begin{equation}
+\gamma = \expval{\hat{X}}\frac{2\pi}{L}
+\end{equation}
+
+The correlation function in position space is given by
+\begin{align*}
+C_{ij}^{\alpha \beta} = \expval{c_{i\alpha}^\dagger c_{j\beta}}.
+\end{align*}
+In terms of the fermions that diagonalize the Hamiltonian, 
+\begin{align*}
+& \gamma_{i\alpha} = \sum_{j\beta}u_{j\beta}^{i\alpha \, \ast} c_{j\beta} \\
+& c_{i\alpha} = \sum_{j\beta} u^{j\beta}_{i\alpha} \gamma_{j\beta}
+\end{align*}
+we have
+\begin{align*}
+C_{ij}^{\alpha \beta} =& \sum_{pq,\mu\nu} u^{q\nu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} \expval{\gamma^\dagger_{p\mu} \gamma_{q\nu} } \\
+=&  \sum_{p\mu} u^{p\mu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} \expval{\gamma^\dagger_{p\mu} \gamma_{p\mu} } \\
+=&  \sum_{p\mu} u^{p\mu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} [1-{\rm sign}(E_{p\mu})]/2.
+\end{align*}
+The first term is simply a kronecker delta between both sets of indeces. The second term can be rewritten as
+\begin{align*}
+UD(\abs{D})^{-1}U^\dagger =& UD(D^2)^{-1/2}U^\dagger \\
+=& U D U^\dagger (U D^2 U^\dagger)^{-1/2} \\
+=& H(H^2)^{-1/2},
+\end{align*}
+where $D$ is the diagonal Hamiltonian, $H=UDU^\dagger$. Finally the correlation function can be obtained in position space as
+\begin{align*}
+C_{ij}^{\alpha \beta} =& \frac{1}{2}\left[I - H/ (H^2)^{-1/2} \right]_{ij, \alpha \beta}
+\end{align*}
+
+\section{Entanglement spectrum}
+\carlos{motivation}
+
+We will obtain the entanglement spectrum by following Peschel's method. Consider a quadratic Hamiltonian in one dimension
+\begin{equation}
+H = \sum_{ij,\alpha\beta} c_{i\alpha}^\dagger H_{ij,\alpha \beta}c_{j\beta}.
+\end{equation}
+The correlation function in position space is given by
+\begin{align*}
+C_{ij}^{\alpha \beta} = \expval{c_{i\alpha}^\dagger c_{j\beta}}
+\end{align*}
+and the full correlation function is nothing else than the projector onto the occupied bands
+\begin{align*}
+C_{ij}^{\alpha \beta} =& \frac{1}{2}\left[I - H/ (H^2)^{-1/2} \right]_{ij, \alpha \beta}
+\end{align*}
+The entanglement spectrum is then obtained as the spectrum of the subsystem correlation function obtained by bisecting the system into two equal parts.
+
+\carlos{Explain a bit more ES and plots.}
 
@@
 
+The results in Fig.\ref{disorder} are obtained for weak disorder, where the ES is not modified much.
+For stronger disorder or other types or inhomogeneous perturbations that strongly modify the ES we run into the issue that $\chi$ is defined modulo $1$ and therefore the cut average is not well-defined.
+The cut average depends on the range considered for the values of $\chi$.
+We observed that $\chi$ oscillates around the value of the geometric phase.
+For weak disorder and continuous inhomogeneous perturbations, where one can follow the path of $\xi$ as we run through the different cuts one can use this information to set the numerical range and obtain the correct result for the geometric phase.
+However for stronger disorder or discontinious perturbations this does not work.
+The reason $\chi$ is defined modulo $1$ is to get rid of the contribution from the bulk states, as otherwise it is very difficult to numerically classify between bulk and edge state.
+If we found another way of getting rid of the bulk contribution then $\chi$ and its cut average would be uniquely defined.
+One can explot the fact that, even in the inhomogeneous case, the ES has a type of particle-hole symmetry.
+If $\xi$ is an eigenvalue for a cut in $s$ then $1-\xi$ is an eigenvalue of the cut in $s+L/2$.
+This means that all the information is already contained in half of the spectrum.
+Using the lower half of the spectrum ensures that we have all the information about the edge eigenstates but the bulk contribution vanishes.
+This is however only valid when there is a gap between the two halfs of the spectrum at any cut $s$.
+
+Alternative one can set a threshold for the eigenvalues we consider in order to avoid the bulk contribution, which will be a good approximation as long as the ES does not have eigenvalues close to the bulk bands, typically close to the trivial phase, and phase transitions.
+
 \section{Superconducting systems}
@@
 
-\bibliography{references}{}
-\bibliographystyle{plain}
+\bibliography{references}
 

### 2020-01-17 draft.tex
Update draft.tex
prubea
beginning intro

--- 2019-12-04 draft.tex
@@
-\documentclass[prl,twocolumn,amsmath,amssymb,superscriptaddress]{revtex4-1}
+\documentclass[twocolumn,amsmath,longbibliography,footinbib,bibnotes,amssymb,superscriptaddress]{revtex4-1}
 
@@
 \section{Introduction.}
+
+topological phases --- interacting and non-interacting
+
+How to identify interesting phases?--- entanglement entropy, entanglement spectrum
+
+Kitaev honeycomb model as example of entanglement spectrum
+
+What information is encoded in the ES? How to extract it?
+
+naively resp. easily accessible: information about edge spectrum
+
+But, other information can be accessed as well : e.g. compactification radius
+
+in general, still unclear what info is encoded and whether it can uniquely specify the topological order at hand
+
+situation not much better for non-interacting systems--- even here, unclear what information is encoded and how one can read it off from the spectrum
+
+here, we show that the information about the Berry phase can be completely recovered from the entanglement spectrum and provide a simple formula to do so
+
+This can be related to earlier attempts, REFS
 

### 2020-01-17 draft.tex

--- 2020-01-15 draft.tex
@@
-                       \documentclass[prb,twocolumn,amsmath,longbibliography,amssymb,superscriptaddress]{revtex4-1}
+\documentclass[twocolumn,amsmath,longbibliography,amssymb,superscriptaddress]{revtex4-1}
 
@@
 
-\section{Introduction.}
+\section{Introduction}
 
-\section{Model}
+topological phases --- interacting and non-interacting
+
+How to identify interesting phases?--- entanglement entropy, entanglement spectrum
+
+Kitaev honeycomb model as example of entanglement spectrum
+
+What information is encoded in the ES? How to extract it?
+
+naively resp. easily accessible: information about edge spectrum
+
+But, other information can be accessed as well : e.g. compactification radius
+
+in general, still unclear what info is encoded and whether it can uniquely specify the topological order at hand
+
+situation not much better for non-interacting systems--- even here, unclear what information is encoded and how one can read it off from the spectrum
+
+here, we show that the information about the Berry phase can be completely recovered from the entanglement spectrum and provide a simple formula to do so
+
+This can be related to earlier attempts, REFS
+
+\section{Model.}
 

### 2020-01-20 draft.tex
incomplete

@@
 \section{Introduction}
+Topological phases of matter have attracted more and more attention during the last decades, not the least because many more relevant systems have been realized experimentally. 
+The early focus was mainly on \emph{topologically ordered} systems \cite{Wen}, where interaction effects are crucial for stabilizing the phase. 
+The most notable examples are the fractional quantum Hall effect~\cite{} and quantum spin liquids~\cite{}. 
+However, since 2005 \maria{?}~\cite{KaneMele, etc} the focus has shifted to non-interacting topological phases, which can be characterized in terms of topological invariants~\cite{}. 
+Symmetries are often necessary to protect these topological phases, and determine which distinct topological phases can be realized for a given dimensionality.
 
-topological phases --- interacting and non-interacting
-
+XXX
 How to identify interesting phases?--- entanglement entropy, entanglement spectrum

### 2020-01-20 draft.tex
new figures

--- 2020-01-17 draft.tex
@@
-\documentclass[twocolumn,amsmath,longbibliography,amssymb,superscriptaddress]{revtex4-1}
-
+\documentclass[prb,twocolumn,amsmath,longbibliography,amssymb,superscriptaddress]{revtex4-1}
 \usepackage[pdftex]{graphics}
@@
 
-\section{Introduction}
+\section{Introduction.}
 
-topological phases --- interacting and non-interacting
-
-How to identify interesting phases?--- entanglement entropy, entanglement spectrum
-
-Kitaev honeycomb model as example of entanglement spectrum
-
-What information is encoded in the ES? How to extract it?
-
-naively resp. easily accessible: information about edge spectrum
-
-But, other information can be accessed as well : e.g. compactification radius
-
-in general, still unclear what info is encoded and whether it can uniquely specify the topological order at hand
-
-situation not much better for non-interacting systems--- even here, unclear what information is encoded and how one can read it off from the spectrum
-
-here, we show that the information about the Berry phase can be completely recovered from the entanglement spectrum and provide a simple formula to do so
-
-This can be related to earlier attempts, REFS
-
-\section{Model.}
+\section{Model}
 
@@
   \includegraphics[width=35mm]{bdit1}
-}
-\subfloat[$t=0,\kappa = \kappa'=0$]{\includegraphics[width=35mm]{bdit0}
-}
-}\hspace{0mm}
+}}\hspace{0mm}
 
 \makebox[0pt]{
-\subfloat[$t = 1,t'=1,\kappa =\kappa'=0$]{
-  \includegraphics[width=35mm]{1_c.pdf}
+\subfloat[$t = 1,t'=1$, \, $\kappa =\kappa'=0$]{
+  \includegraphics[width=35mm]{1_b.pdf}
 }
-\subfloat[$t=0,t'=1,\kappa = \kappa'=0$]{\includegraphics[width=35mm]{1_d.pdf}
+\subfloat[$t=1,t'=1$, $\kappa = 0.3, \kappa'=0$]{\includegraphics[width=35mm]{1_c.pdf}
 }
@@
 \subfloat[$t = 1,t'=1$, $\kappa = 0,\kappa'=0.3$]{
-  \includegraphics[width=35mm]{1_f.pdf}
+  \includegraphics[width=35mm]{1_d.pdf}
 }
-\subfloat[$t=0,t'=1$, $\kappa = 0.3,\kappa'=0$]{\includegraphics[width=35mm]{1_g.pdf}
+\subfloat[$t=1,t'=1$, $\kappa = 0.3,\kappa'=0.3$]{\includegraphics[width=35mm]{1_e.pdf}
 }
@@
 where $[U]_{j\beta,p\mu} = u^{p\mu}_{j\beta}$ is the unitary matrix that diagonalizes $H$.
-The polarization can be obtained as the many-body average of position operator, regularized for periodic boudnary conditions \carlos{resta}
+The polarization can be obtained as the many-body average of position operator, regularized for periodic boudnary conditions \cite{Resta1997}
 \begin{align*}
@@
 where $\ket{\Psi_0}$ is the many-body ground state and $\hat{X}$ is the sum of all the single-particle position operators.
-Using that the ground state is a Slater determinant of the single-particle eigenstates it can be rewritten as
+Using that the ground state is a Slater determinant of the single-particle eigenstates, $\ket{u^{p\mu}}$, it can be rewritten as
 \begin{align*}
-\expval{\hat{X}} = -\frac{L}{2\pi} \rm{Im \, Ln}\, {\rm det} S,
+\expval{\hat{X}} = -\frac{L}{2\pi} \rm{Im \, Ln}\, {\rm det} \, \tilde{S},
 \end{align*}
-where 
+where the matrix $\tilde{S}$ is obtained from another matrix S with elements
 \begin{equation}
@@
 \end{equation}
-and the determinant is taken restricting $S$ to only the occupied eigenstates. The geometric phase can now be obtained as
+by restricting it to the space of occupied single-particle states. The geometric phase can now be obtained as
 \begin{equation}
@@
 \end{align*}
+The entanglement spectrum (ES) is then obtained as the spectrum of the subsystem correlation function obtained after bisecting the system into two equal parts.
 
-\section{Entanglement spectrum}
-\carlos{motivation}
-
-We will obtain the entanglement spectrum by following Peschel's method. Consider a quadratic Hamiltonian in one dimension
-\begin{equation}
-H = \sum_{ij,\alpha\beta} c_{i\alpha}^\dagger H_{ij,\alpha \beta}c_{j\beta}.
-\end{equation}
-The correlation function in position space is given by
-\begin{align*}
-C_{ij}^{\alpha \beta} = \expval{c_{i\alpha}^\dagger c_{j\beta}}
-\end{align*}
-and the full correlation function is nothing else than the projector onto the occupied bands
-\begin{align*}
-C_{ij}^{\alpha \beta} =& \frac{1}{2}\left[I - H/ (H^2)^{-1/2} \right]_{ij, \alpha \beta}
-\end{align*}
-The entanglement spectrum is then obtained as the spectrum of the subsystem correlation function obtained by bisecting the system into two equal parts.
-
-\carlos{Explain a bit more ES and plots.}
+In the ES we can differentiate between two types of eigenvalues, which we denote by $\xi$.
+The ones at $\xi_b = 0,1$ correspond to bulk states while for the ones in between, $0<\xi_{l,r}<1$, we find their correspondent eigenstate localized in either the left, $\xi_l$, or right, $\xi_r$, virtual edge \carlos{Cite Peschel}.
 
@@
 
-For symmetry-protected topological systems in one dimension there is a well-known relation between the entanglement spectrum  (ES) and the geometric phase.
+For symmetry-protected topological systems in one dimension there is a well-known relation between ES and the geometric phase.
 The geometric phase, which is quantized, is zero whenever there is an even number per edge of eigenvalues at $\xi = 1/2$, and $\pi$ when the number is odd.
@@
 
-The entanglement spectrum eigenstates can be classified as bulk (with eigenvalues $\xi_{b} = 0,1 $) or edge states (with eigenvalues $\xi_{l,r} \neq 0,1$), which are localized in one of the edges.
-Note that any eigenstate with eigenvalue different from zero or one we refer to as edge state, and not only the ones at $\xi=1/2$.
 Consider the combination of all eigenvalues localized in one of the edges modulo one,
@@
 \end{equation}
-In the examples studied in references \cite{Ryu2006,Huang2012,Huang2012-2} most of the edge states sit close to the bulk bands and therefore give a small contribution, which explains why the relation between the midgap eigenvalue and the geometric phase was almost an identity and why it breaks as we go deep into the trivial state.
+In the examples studied in references \cite{Ryu2006,Huang2012,Huang2012-2} the main contribution comes from one midgap state, while the rest of the edge states sit close to the bulk bands and therefore give a small contribution, which explains why the relation between the midgap eigenvalue and the geometric phase was almost an identity and why it breaks as we go deep into the trivial state.
 Alternatively, using that the entanglement spectrum is particle-hole symmetric for homogeneous systems we can express $\chi$ as
@@
 where $N$ is the total number of electrons in the ground state and the sign on the bulk term is irrelevant since $-1/2 \, {\rm mod} \, 1 = 1/2 \, {\rm mod} \, 1$.
-This form is preferable when doing numerics because it is difficult to differentiate between edge modes and bulk modes since the latter can acquire a finite eigenvalue due to finite size effects or numerical errors.
+This form is preferable when doing numerics because it is difficult to differentiate between edge modes and bulk modes, since the latter can acquire a finite eigenvalue due to finite size effects or numerical errors.
 In Fig.\ref{bdi20} we show for several cases how Eq.
 \ref{mainl} holds up to finite size errors.
-The ES does not have therefore enough information to build up the geometric phase, but one needs to look at the localization structure of the edge states.
-One drastic example of this can be seen in figures \ref{bdi20}.b and \ref{bdi20}.c , where two ES that are almost identical result in a very different geometric phase.
+$\chi$ is equivalent to the trace of the ES of a semi-infinite chain, but one can extract the information from the ES of a ring in the thermodynamic limit by looking at the localization structure of the eigenstates, only looking at the spectrum is not enough.
+One drastic example of this can be seen in figures \ref{bdi20}.b and \ref{bdi20}.c , where two ES that look almost identical result in a very different geometric phase.
 
@@
 \subfloat[$t = 0, t' = 2, \kappa = 10^{-5}$]{
-  \includegraphics[width=35mm]{bdi20.pdf}
+  \includegraphics[width=35mm]{2_a.pdf}
 }
 \subfloat[$t = 0, t' = 2, \kappa = 0.3$]{
-  \includegraphics[width=35mm]{ai20.pdf}
+  \includegraphics[width=35mm]{2_b.pdf}
 }
@@
 \makebox[0pt]{
+\subfloat[$t = 0, t' = 2, \kappa = 10^{-5}$]{
+  \includegraphics[width=35mm]{2_c.pdf}
+}
 \subfloat[$t = 0, t' = 2, \kappa = 0.3$]{
-  \includegraphics[width=35mm]{d20.pdf}
+  \includegraphics[width=35mm]{2_d.pdf}
 }
@@
 \end{figure}
-
-This relation between the geometric phase and the entanglement spectrum is a consequence of an identity between the geometric phase and the many-body entanglement spectrum found for infinite chains \cite{Zaletel2014}.
+as numerically is
+This relation between the geometric phase and the entanglement spectrum is a consequence of an identity between the geometric phase and the many-body entanglement spectrum previously found for infinite chains \cite{Zaletel2014}.
 In Appendix A we modified this derivation to account for the periodic boundary conditions we use and show how their result reduces to Eq.
@@
 This can be argued in the following way.
-Weak disorder will modify the entanglement spectrum but not the charge assignment.
+Weak disorder will modify slightly the entanglement spectrum but not the charge assignment.
 Without translational invariance the entanglement spectrum will depend on where is the cut placed on the ring, even though the ground state, and therefore the geometric phase, will not.
 Therefore there cannot be an equivalence between them.
-Instead, we found that one can still obtain the geometric phase by considering the average over all spatial cuts of expression \ref{mainl}.
+One thing one could do to get rid of the cut-dependence is to perform and average over all the spatial cuts of expression \ref{mainl}, however, we run into problems.
+Since we defined $\chi$ as a circular quantity its average is not well-defined.
+We can take the modulo $1$ outside the average.
+Consider then
+\begin{equation*}
+\chi_s = \frac{1}{2} \left( \sum_{i\in l}\xi^s_i-\sum_{i\in r}\xi^s_i \pm \sum_{i\in b}\xi^s_i  + N \right) \quad,
+\end{equation*}
+where $\xi^s$ are the eigenvalues of the ES for a cut in positions $s$ and $s+L/2$. Then we have
+\begin{equation}
+\gamma = \expval{\chi_s }_{s}\quad {\rm mod} \quad 1,
+\label{main_inh}
+\end{equation}
+however one of the reasons we defined $\chi$ as a circular quantity is that it gets rid of the bulk contribution, since as mentioned above it is very difficult to differenciate between bulk and edge states and an inconsistent classification of the eigenstates for different cuts will lead to a wrong result.
+We will show different cases that point to expression \ref{main_inh} being correct but we do not have a practical way of correctly computing it in general.
 
-In order to test this we take the model we used in the previous section and we promote every parameter to a position-dependent quantity, by adding a random part.
+In order for the system to remain gapped we consider $\kappa_i >0$ .
 In Fig.\ref{disorder} we show $\gamma-\expval{\chi}_{\rm cuts}$ for increasing system size.

### 2020-01-23 draft.tex
added refs and draft of intro

--- 2020-01-17 draft.tex
@@
 \newcommand{\maria}[1]{{\color{blue} #1}}
+\newcommand{\mariac}[1]{{\it\color{cyan}#1}}
 
@@
 \section{Introduction}
+Topological phases of matter have attracted more and more attention during the last decades, not the least because many more relevant systems have been realized experimentally. 
+The early focus was mainly on \emph{topologically ordered} systems \cite{wenbook}, where interaction effects are crucial for stabilizing the phase. 
+The most notable examples are the fractional quantum Hall effect~\cite{Tsui1982} and quantum spin liquids~\cite{Balents2010spin}. 
+However, since 2005~\cite{kane2005quantum,roy2009topological} the focus has shifted to non-interacting topological phases, which can be characterized in terms of topological invariants~\cite{ryu2010topological}.
+Symmetries are often necessary to protect these topological phases, and determine which distinct topological phases can be realized for a given dimensionality.
 
-topological phases --- interacting and non-interacting
+There are a few tools to identify topological phases. 
+A very efficient one is the entanglement entropy~\cite{}, which allows one to identify the total quantum dimension of the underlying topological field theory. 
+As such, the entanglement entropy can only be used for interacting systems. 
+Its main drawback is the fact that it can only provide one of the quantum numbers, and is not able to uniquely determine the topologically ordered phase at hand. 
+In addition, it requires a scaling analysis, which can be challenging for strongly correlated systems. 
+Another, closely related tool, is the entanglement spectrum (ES), originally introduced for fractional quantum Hall systems~\cite{Li2008entanglement}. 
+It was conjectured to provide information about the edge spectrum, which was later shown to be a general feature of ground states with an effective topological field theory description~\cite{Qi2012general}.
+It also proved to be an effective tool for fractional Chern insulators~\cite{Regnault2011fractional} and certain quantum spin liquids~\cite{yao2010entanglement}. \mariac{Others?}
 
-How to identify interesting phases?--- entanglement entropy, entanglement spectrum
+Generically, any of the proposed tools provides some information about the phase naturally, while others seem inaccessible. 
+It is not known, whether there exists a method to obtain the full topological data from the ground state. 
+For the simplest quantum Hall states, the Laughlin states,  it was shown that the ES does, in fact, contain the full information about the topological phase~\cite{hermanns2011haldane}.
+However, it is far from clear if this holds for more complicated/nonabelian states as well.
 
-Kitaev honeycomb model as example of entanglement spectrum
+Here, we address a much simpler question: what information can be extracted from the ES of \emph{non-interacting} systems?
+The latter can be computed very efficiently, using methods developed by Peschel and others~\cite{Peschel2003}. 
+It was subsequently shown for topological insulators/superconductors that the ES for (gapped) periodic systems is equivalent to the flat-band energy spectrum of the corresponding system with open boundaries~\cite{Fidkowski2010entanglement}.
+The same correspondence was also found for closely related gapless systems~\cite{matern2018entanglement}
+However, even for non-interacting systems, it is unclear which information (beyond the  `edge' spectrum) is encoded in the spectrum.
 
-What information is encoded in the ES? How to extract it?
+In this manuscript, we show that the information about the Berry phase can be completely recovered from the entanglement spectrum and provide a simple formula to do so. 
+We relate our results to earlier ones XXX
 
-naively resp. easily accessible: information about edge spectrum
-
-But, other information can be accessed as well : e.g. compactification radius
-
-in general, still unclear what info is encoded and whether it can uniquely specify the topological order at hand
-
-situation not much better for non-interacting systems--- even here, unclear what information is encoded and how one can read it off from the spectrum
-
-here, we show that the information about the Berry phase can be completely recovered from the entanglement spectrum and provide a simple formula to do so
-
-This can be related to earlier attempts, REFS
-
+\emph{Outline of the paper}
 \section{Model.}

### 2020-01-23 draft.tex

--- 2020-01-20 draft.tex
@@
-\documentclass[prb,twocolumn,amsmath,longbibliography,amssymb,superscriptaddress]{revtex4-1}
+\documentclass[twocolumn,amsmath,longbibliography,amssymb,superscriptaddress]{revtex4-1}
 \usepackage[pdftex]{graphics}
@@
 
-\section{Introduction.}
+\section{Introduction}
+Topological phases of matter have attracted more and more attention during the last decades, not the least because many more relevant systems have been realized experimentally. 
+The early focus was mainly on \emph{topologically ordered} systems \cite{Wen}, where interaction effects are crucial for stabilizing the phase. 
+The most notable examples are the fractional quantum Hall effect~\cite{} and quantum spin liquids~\cite{}. 
+However, since 2005 \maria{?}~\cite{KaneMele, etc} the focus has shifted to non-interacting topological phases, which can be characterized in terms of topological invariants~\cite{}. 
+Symmetries are often necessary to protect these topological phases, and determine which distinct topological phases can be realized for a given dimensionality.
+
+XXX
+How to identify interesting phases?--- entanglement entropy, entanglement spectrum
+
+Kitaev honeycomb model as example of entanglement spectrum
+
+What information is encoded in the ES? How to extract it?
+
+naively resp. easily accessible: information about edge spectrum
+
+But, other information can be accessed as well : e.g. compactification radius
+
+in general, still unclear what info is encoded and whether it can uniquely specify the topological order at hand
+
+situation not much better for non-interacting systems--- even here, unclear what information is encoded and how one can read it off from the spectrum
+
+here, we show that the information about the Berry phase can be completely recovered from the entanglement spectrum and provide a simple formula to do so
+
+This can be related to earlier attempts, REFS
 
@@
 }
-\caption{Entanglement spectrum (black), Berry phase (red) and $\chi$ (blue) for different cuts of the $H_{BDI}$.
-a) with $\kappa = 0,\kappa'=0$ in the $AIII$ class, b) with $\kappa'=0$ in the $AI$ class and c) with $\kappa = 0$ in the $D$ class.
-}
+\caption{fill. }
 \label{bdi20}
@@
 In reference \cite{Zaletel2014} translational invariance was used to derive Eq.
-\ref{zaletel} in Appendix A, and as a result Eq.\ref{mainl} is no longer correct when it is broken.
-This can be argued in the following way.
-Weak disorder will modify slightly the entanglement spectrum but not the charge assignment.
+\ref{zaletel} in Appendix A.
+Say we break translational invariance by adding weak disorder to the system.
+This will modify slightly the entanglement spectrum but not the charge assignment.
 Without translational invariance the entanglement spectrum will depend on where is the cut placed on the ring, even though the ground state, and therefore the geometric phase, will not.
-Therefore there cannot be an equivalence between them.
-One thing one could do to get rid of the cut-dependence is to perform and average over all the spatial cuts of expression \ref{mainl}, however, we run into problems.
-Since we defined $\chi$ as a circular quantity its average is not well-defined.
-We can take the modulo $1$ outside the average.
-Consider then
+One thing one could do to get rid of the cut-dependence is to perform an average over all the spatial cuts of expression \ref{mainl}.
+Defining now the cut-dependent $\chi$ as
 \begin{equation*}
@@
 \end{equation*}
-where $\xi^s$ are the eigenvalues of the ES for a cut in positions $s$ and $s+L/2$. Then we have
+for a cut in positions $s$ and $s+L/2$. We can write
 \begin{equation}
-\gamma = \expval{\chi_s }_{s}\quad {\rm mod} \quad 1,
-\label{main_inh}
+\gamma = \expval{\chi_s \, {\rm mod}\, 1}_s \, {\rm mod}\, 1.
+\label{modmod}
 \end{equation}
-however one of the reasons we defined $\chi$ as a circular quantity is that it gets rid of the bulk contribution, since as mentioned above it is very difficult to differenciate between bulk and edge states and an inconsistent classification of the eigenstates for different cuts will lead to a wrong result.
-We will show different cases that point to expression \ref{main_inh} being correct but we do not have a practical way of correctly computing it in general.
+The average of a circular quantity is however not well-defined.
+For the case when the width of $\chi_s$ is lower than one, there is a range of values $[x,x+1]$ where $\chi_s$ does not cross $0,1$, and averaging over this range makes sure than once we take the second modulo $1$ all the different possible results for the cut-average are reduced to a single one.
+The shortcomings of expression \ref{modmod} will be discussed later, but to show that this cut-average is correct we will consider now a system in the topological phase to which we add weak disorder, so one can then take the cut-average around $[0,1]$.
+In Fig.\ref{disorder} we show $1 - \expval{\chi_s \, {\rm mod}\, 1}_s \, {\rm mod}\, 1 /\gamma$ in red for different system sizes.
+In blue we have shown the same difference using a disorder average.
+Both values match fairly well and they seem to vanish in the thermodynamic limit, scaling as $O(L^{-2})$.
+The fact that different disorder configurations match so well points to a systematic finite-size error in the way we calculate either the ES or the geometric phase that is system-independent.
+From this result one could think that the cut-average is some sort of disorder average.
+The implication however goes in the opposite direction.
+If the cut-average is assumed to be correct, that implies that in the thermodynamic limit we can group all the infinite different disorder configurations into an infinite set of $L$ configurations relating to different cuts of the same system, which average individually to the correct $\gamma_{\rm conf}$.
 
-In order for the system to remain gapped we consider $\kappa_i >0$ .
-In Fig.\ref{disorder} we show $\gamma-\expval{\chi}_{\rm cuts}$ for increasing system size.
-Note that we do the calculation for a particular disorder configuration for each size.
-We found that $\expval{\chi}_{\rm cuts}$ converges much faster than the geometric phase and that the difference scales as $O(L^{-2})$, which is the same scaling we find in the homogeneous system.
-We compared this result with a disorder average $\gamma - \expval{\chi}_{\rm{disorder}}$ and found that, even though we are comparing different disorder configurations, the disorder average and the cut average give the same error, which implies that it must come exclusively from finite size effects in computing the geometric phase, since is the only thing in common.
-One could think that this cut average is some sort of disorder average, but that is however not correct.
-In Fig.\ref{cuts} we show the case with a continuous position-dependent potential, plotted in Fig.\ref{cuts}.a, on top of the homogeneous model.
-In Fig.\ref{cuts}.b we show $\chi_i$ for two cuts where the left edge is either in the positive or negative potential region.
-As we can see in Fig.\ref{cuts}.c for any intermidiate cut, $\chi_i$ goes continuously from one value to the other.
-Here we encounter a new problem, since $\chi_i$ only makes physical sense as an angle, there are two paths between the two points.
-Below the phase transition we find that $\chi_i$ oscillates around $1/2$ while above the phase transition it does so around $0$.
-By redifining the range of values so that they do not cross $0$ and then averaging we obtain the same result as for the disordered system, that $\gamma-\expval{\chi}_{\rm cuts} \sim O(L^{-2})$.
-This is only well defined in the continuous limit and can potentially pose problems for finite systems and discontinuous terms, like with disorder.
-However, as long as the deviations are small and we have $|\expval{\gamma-\chi}|<1/2$, which has been the case so far, one only has to consider two averages, the one with $\chi_i \in [0,1)$ and the one with $\chi_i \in [-1/2,1/2)$, although at the moment there is no indication of which of the two numbers will result on the geometric phase.
+For stronger disorder one cannot be sure if $\chi_s$ crosses $0,1$ or not so this method cannot be applied but for continuous inhomogeneous perturbations one can follow $\chi_s$, so we can use expression \ref{modmod} even when they are strong.
+In Fig.\ref{4} we show the results using a position dependent $\kappa$ parameter, shown in figure \ref{4}.a for different values of $m$ and the amplitude of $\kappa$.
+In cases Fig.\ref{4}.b,c the width of $\chi_s$ is small so we can average taking the value ranges $[0,1]$ and $[-1/2,1/2]$ respectively and we see that the cut average matches the geometric phase perfectly.
+However for stronger perturbations, as shown in Fig.\ref{4}.d, one can have that the width of $\chi_s$ is bigger than 1 and there is no meaningful range of values one can average.
+Instead what one can do is to recover $\chi_s$ (\carlos{ only edge modes}) by constructing a new object $\tilde{\chi}_s$ by adding or substracting $1$ each time $\chi_s$ crosses $0,1$, taking the cut-average of $\tilde{\chi}_s$ and then take the modulo 1.
 
@@
 \subfloat[]{
-  \includegraphics[width=40mm]{profile.pdf}
+  \includegraphics[width=40mm]{4_a.pdf}
+}
+\subfloat[]{
+  \includegraphics[width=40mm]{4_b.pdf}
 }
@@
 \subfloat[]{
-  \includegraphics[width=40mm]{extreme_es.pdf}
+  \includegraphics[width=40mm]{4_c.pdf}
 }
 \subfloat[]{
-  \includegraphics[width=40mm]{cut_es.pdf}
+  \includegraphics[width=40mm]{4_d.pdf}
 }
@@
 
-\caption{a)Profile used for $\kappa_i$ .
-b)$\chi_i$ for two cases where we place the left cut deep in either the region with positive or negative $\kappa_i$, and geometric phase in red, showing a phase transition.
-c)$\chi_i$ for all different possible cuts using the parameters $m=1$ (blue) and $m=3$ (green)}
-\label{cuts}
+\caption{ fill. put .5 line, also break the symmetry. with an envelope.}
+\label{4}
 \end{figure}
-
-The results in Fig.\ref{disorder} are obtained for weak disorder, where the ES is not modified much.
-For stronger disorder or other types or inhomogeneous perturbations that strongly modify the ES we run into the issue that $\chi$ is defined modulo $1$ and therefore the cut average is not well-defined.
-The cut average depends on the range considered for the values of $\chi$.
-We observed that $\chi$ oscillates around the value of the geometric phase.
-For weak disorder and continuous inhomogeneous perturbations, where one can follow the path of $\xi$ as we run through the different cuts one can use this information to set the numerical range and obtain the correct result for the geometric phase.
-However for stronger disorder or discontinious perturbations this does not work.
-The reason $\chi$ is defined modulo $1$ is to get rid of the contribution from the bulk states, as otherwise it is very difficult to numerically classify between bulk and edge state.
-If we found another way of getting rid of the bulk contribution then $\chi$ and its cut average would be uniquely defined.
+(We use a profile which is invariant for $x\rightarrow x+L/2$ for practical reasons, so we can illustrate all cases, as it enhances the effect of the nonhomogeneity.
+But this symmetry is not necesarry for anything we discuss here)
+The last case with an easy solution is when the system is fully gapped, so that the ES doesn't have modes crossing $1/2$ for any cut.
+In this case one can make use of the particle-hole symmetry of the ES and use only the lower half of the spectrum to build $\chi_s$ with no issues due to the bulk contribution.
 One can explot the fact that, even in the inhomogeneous case, the ES has a type of particle-hole symmetry.
@@
 This is however only valid when there is a gap between the two halfs of the spectrum at any cut $s$.
-
-Alternative one can set a threshold for the eigenvalues we consider in order to avoid the bulk contribution, which will be a good approximation as long as the ES does not have eigenvalues close to the bulk bands, typically close to the trivial phase, and phase transitions.
+MERGE
+Another interesting point is that the two charge choices now leave to different $\chi_s$ since the particle-hole is no longer fulfilled for individual cuts.
+However both still lead to the same result because there is still a symmetry involving cuts related by a half-system translation.
+Interestingly for $\pm 1/2$ the quantity $\chi_s$ seems to be smaller, leading to fewer issues when averaging.
 

### 2020-01-27 draft.tex
Update inhomogeneous

@@
 Topological phases of matter have attracted more and more attention during the last decades, not the least because many more relevant systems have been realized experimentally. 
-The early focus was mainly on \emph{topologically ordered} systems \cite{Wen}, where interaction effects are crucial for stabilizing the phase. 
-The most notable examples are the fractional quantum Hall effect~\cite{} and quantum spin liquids~\cite{}. 
-However, since 2005 \maria{?}~\cite{KaneMele, etc} the focus has shifted to non-interacting topological phases, which can be characterized in terms of topological invariants~\cite{}. 
+The early focus was mainly on \emph{topologically ordered} systems \cite{wenbook}, where interaction effects are crucial for stabilizing the phase. 
+The most notable examples are the fractional quantum Hall effect~\cite{Tsui1982} and quantum spin liquids~\cite{Balents2010spin}. 
+However, since 2005~\cite{kane2005quantum,roy2009topological} the focus has shifted to non-interacting topological phases, which can be characterized in terms of topological invariants~\cite{ryu2010topological}.
 Symmetries are often necessary to protect these topological phases, and determine which distinct topological phases can be realized for a given dimensionality.
 
-XXX
-How to identify interesting phases?--- entanglement entropy, entanglement spectrum
+There are a few tools to identify topological phases. 
+A very efficient one is the entanglement entropy~\cite{}, which allows one to identify the total quantum dimension of the underlying topological field theory. 
+As such, the entanglement entropy can only be used for interacting systems. 
+Its main drawback is the fact that it can only provide one of the quantum numbers, and is not able to uniquely determine the topologically ordered phase at hand. 
+In addition, it requires a scaling analysis, which can be challenging for strongly correlated systems. 
+Another, closely related tool, is the entanglement spectrum (ES), originally introduced for fractional quantum Hall systems~\cite{Li2008entanglement}. 
+It was conjectured to provide information about the edge spectrum, which was later shown to be a general feature of ground states with an effective topological field theory description~\cite{Qi2012general}.
+It also proved to be an effective tool for fractional Chern insulators~\cite{Regnault2011fractional} and certain quantum spin liquids~\cite{yao2010entanglement}. \mariac{Others?}
 
-Kitaev honeycomb model as example of entanglement spectrum
+Generically, any of the proposed tools provides some information about the phase naturally, while others seem inaccessible. 
+It is not known, whether there exists a method to obtain the full topological data from the ground state. 
+For the simplest quantum Hall states, the Laughlin states,  it was shown that the ES does, in fact, contain the full information about the topological phase~\cite{hermanns2011haldane}.
+However, it is far from clear if this holds for more complicated/nonabelian states as well.
 
-What information is encoded in the ES? How to extract it?
+Here, we address a much simpler question: what information can be extracted from the ES of \emph{non-interacting} systems?
+The latter can be computed very efficiently, using methods developed by Peschel and others~\cite{Peschel2003}. 
+It was subsequently shown for topological insulators/superconductors that the ES for (gapped) periodic systems is equivalent to the flat-band energy spectrum of the corresponding system with open boundaries~\cite{Fidkowski2010entanglement}.
+The same correspondence was also found for closely related gapless systems~\cite{matern2018entanglement}
+However, even for non-interacting systems, it is unclear which information (beyond the  `edge' spectrum) is encoded in the spectrum.
 
-naively resp. easily accessible: information about edge spectrum
+In this manuscript, we show that the information about the Berry phase can be completely recovered from the entanglement spectrum and provide a simple formula to do so. 
+We relate our results to earlier ones XXX
 
-But, other information can be accessed as well : e.g. compactification radius
-
-in general, still unclear what info is encoded and whether it can uniquely specify the topological order at hand
-
-situation not much better for non-interacting systems--- even here, unclear what information is encoded and how one can read it off from the spectrum
-
-here, we show that the information about the Berry phase can be completely recovered from the entanglement spectrum and provide a simple formula to do so
-
-This can be related to earlier attempts, REFS
-
-\section{Model}
+\emph{Outline of the paper}
+\section{Model.}
 
@@
 \begin{align*}
-\expval{X} = -\frac{L}{2\pi} \rm{ Im\, Ln }\bra{\Psi_0}e^{i \frac{2\pi}{L}\hat{X}}\ket{\Psi_0},
+\expval{X} = -\frac{L}{2\pi} \rm{ Im\, Ln }\bra{\Psi_0}e^{-i \frac{2\pi}{L}\hat{X}}\ket{\Psi_0},
 \end{align*}
@@
 \begin{equation}
-\gamma = \expval{\hat{X}}\frac{2\pi}{L}
+\gamma = \expval{\hat{X}}\frac{2\pi}{L} + \frac{1}{2}
 \end{equation}
-
+\carlos{There is an extra 1/2 factor compared to Kivelson's formula which I don't know where it is coming from.}
 The correlation function in position space is given by
@@
 The geometric phase, which is quantized, is zero whenever there is an even number per edge of eigenvalues at $\xi = 1/2$, and $\pi$ when the number is odd.
-Beyond symmetry protected systems there have been observations of similarities between the midgap eigenvalue of the entanglement spectrum and the geometric phase \cite{Ryu2006,Huang2012,Huang2012-2}, which in the case of the fully dimerized model becomes an exact identity \cite{Ryu2006}.
+Beyond symmetry protected systems there have been observations, in simple systems with one edge-mode per edge, of similarities between the midgap eigenvalue of the entanglement spectrum and the geometric phase \cite{Ryu2006,Huang2012,Huang2012-2}, which in the case of the fully dimerized model becomes an exact identity \cite{Ryu2006}.
 
@@
 \begin{equation}
-\chi = \sum_{i \in l} \xi_l \quad {\rm mod} \quad 1. 
+\chi = \sum_{i \in r} \xi_i \quad {\rm mod} \quad 1. 
 \end{equation}
-As we will show below, for systems where the ES has more than one edge state per edge one can find the geometric phase in the thermodynamic limit as
+As we will show below, one can find the geometric phase in the thermodynamic limit as
 \begin{equation}
-\lim_{L \rightarrow \infty}\gamma/2\pi =\lim_{L \rightarrow \infty} \chi
+\lim_{L \rightarrow \infty}\gamma/2\pi =\lim_{L \rightarrow \infty} \chi.
 \label{mainl}
@@
 \begin{equation*}
-\chi = \frac{1}{2} \left( \sum_{i\in l}\xi_i-\sum_{i\in r}\xi_i \pm \sum_{i\in b}\xi_i  + N \right) \quad {\rm mod} \quad 1,
+\chi = \frac{1}{2} \left( \sum_{i\in r}\xi_i-\sum_{i\in l}\xi_i \pm \sum_{i\in b}\xi_i  + N_A \right) \quad {\rm mod} \quad 1,
 \end{equation*}
-where $N$ is the total number of electrons in the ground state and the sign on the bulk term is irrelevant since $-1/2 \, {\rm mod} \, 1 = 1/2 \, {\rm mod} \, 1$.
-This form is preferable when doing numerics because it is difficult to differentiate between edge modes and bulk modes, since the latter can acquire a finite eigenvalue due to finite size effects or numerical errors.
+where $N_A$ is the total number of electrons in subsystem A  in the ground state and the sign on the bulk term is irrelevant since $-1/2 \, {\rm mod} \, 1 = 1/2 \, {\rm mod} \, 1$.
+This form is preferable when doing numerics because due to finite size effects the bulk modes can acquire a finite eigenvalue and therefore in general it is difficult to classify between bulk and edge modes.
+\carlos{There is also the issue edge-modes and bulk-modes mixing in the inhomogenous case, introduce later?} In practice we diagonalize $C_A + \lambda C_A\hat{X}C_A$, where $\hat{X}$ is the position operator and $\lambda$ is a small parameter.
+The second term ensures that the originally degerate bulk states are also localized, so we can divide all eigenvalues in right and left sectors and assign the charges $1/2$ and $-1/2$, respectively.
 In Fig.\ref{bdi20} we show for several cases how Eq.
@@
 $\chi$ is equivalent to the trace of the ES of a semi-infinite chain, but one can extract the information from the ES of a ring in the thermodynamic limit by looking at the localization structure of the eigenstates, only looking at the spectrum is not enough.
-One drastic example of this can be seen in figures \ref{bdi20}.b and \ref{bdi20}.c , where two ES that look almost identical result in a very different geometric phase.
+
+One drastic example of this can be seen in figure \ref{old} , where two ES that look almost identical result in a very different geometric phase.
 
@@
 \end{figure}
-as numerically is
+\begin{figure}[h!]
+\centering
+\makebox[0pt]{
+\subfloat[$t = 0, t' = 2, \kappa = 10^{-5}$]{
+  \includegraphics[width=35mm]{a20.pdf}
+}
+\subfloat[$t = 0, t' = 2, \kappa = 0.3$]{
+  \includegraphics[width=35mm]{d20.pdf}
+}
+}
+\caption{Old plot, do I keep it? }
+\label{old}
+\end{figure}
+
 This relation between the geometric phase and the entanglement spectrum is a consequence of an identity between the geometric phase and the many-body entanglement spectrum previously found for infinite chains \cite{Zaletel2014}.
-In Appendix A we modified this derivation to account for the periodic boundary conditions we use and show how their result reduces to Eq.
+In Appendix A (\carlos{I have to take a look at it again}) we modified this derivation to account for the periodic boundary conditions we use and show how their result reduces to Eq.
 \ref{mainl} when expressing it in terms of the ES.
@@
 \begin{equation*}
-\chi_s = \frac{1}{2} \left( \sum_{i\in l}\xi^s_i-\sum_{i\in r}\xi^s_i \pm \sum_{i\in b}\xi^s_i  + N \right) \quad,
+\chi_s = \frac{1}{2} \left( \sum_{i\in r}\xi^s_i-\sum_{i\in l}\xi^s_i \pm \sum_{i\in b}\xi^s_i  + N_A \right) \quad,
 \end{equation*}
@@
 \begin{equation}
-\gamma = \expval{\chi_s \, {\rm mod}\, 1}_s \, {\rm mod}\, 1.
-\label{modmod}
+\gamma = \expval{\chi_s}_{\rm cuts} \, {\rm mod}\, 1.
+\label{gphase_ih_mod}
 \end{equation}
-The average of a circular quantity is however not well-defined.
+However this expression is not practical.
+In order for the bulk contribution to vanish with the modulo $1$ it has to be constant for different cuts.
+Last section we divided the bulk modes in right and left, but now we encounter the problem that their localization changes with the cut, so that there is the possibility that one bulk mode around $1/2$ changes its charge, leading to a wrong result.
+To get around this problem we can get rid of the bulk contribution before averaging,
+\begin{equation}
+\gamma = \expval{\chi_s \, {\rm mod}\, 1}_s \, {\rm mod}\, 1,
+\label{gphase_ih_modmod}
+\end{equation}
+however the average of a circular quantity is not well defined.
+We will explain how one can perform the cut-average for different cases so that one can recover expression \ref{gphase_ih_mod} from expression \ref{gphase_ih_modmod}.
+
 For the case when the width of $\chi_s$ is lower than one, there is a range of values $[x,x+1]$ where $\chi_s$ does not cross $0,1$, and averaging over this range makes sure than once we take the second modulo $1$ all the different possible results for the cut-average are reduced to a single one.
@@
 
-For stronger disorder one cannot be sure if $\chi_s$ crosses $0,1$ or not so this method cannot be applied but for continuous inhomogeneous perturbations one can follow $\chi_s$, so we can use expression \ref{modmod} even when they are strong.
-In Fig.\ref{4} we show the results using a position dependent $\kappa$ parameter, shown in figure \ref{4}.a for different values of $m$ and the amplitude of $\kappa$.
+For stronger disorder one cannot be sure if $\chi_s$ crosses $0,1$ or not so this method cannot be applied but for continuous inhomogeneous perturbations one can follow $\chi_s$, so we can use expression \ref{modmod} even when the inhomogeneity is strong.
+In Fig.\ref{4} we show the results using a position dependent $\kappa$ parameter, shown in figure \ref{4}.a.
 In cases Fig.\ref{4}.b,c the width of $\chi_s$ is small so we can average taking the value ranges $[0,1]$ and $[-1/2,1/2]$ respectively and we see that the cut average matches the geometric phase perfectly.
 However for stronger perturbations, as shown in Fig.\ref{4}.d, one can have that the width of $\chi_s$ is bigger than 1 and there is no meaningful range of values one can average.
-Instead what one can do is to recover $\chi_s$ (\carlos{ only edge modes}) by constructing a new object $\tilde{\chi}_s$ by adding or substracting $1$ each time $\chi_s$ crosses $0,1$, taking the cut-average of $\tilde{\chi}_s$ and then take the modulo 1.
+Instead what one can do is to "double the periodicity of $\chi_s$" by adding or substracting $1$ each time it crosses $0,1$, taking the cut-average of and then take the modulo 1.
 
@@
 \end{figure}
-(We use a profile which is invariant for $x\rightarrow x+L/2$ for practical reasons, so we can illustrate all cases, as it enhances the effect of the nonhomogeneity.
-But this symmetry is not necesarry for anything we discuss here)
-The last case with an easy solution is when the system is fully gapped, so that the ES doesn't have modes crossing $1/2$ for any cut.
-In this case one can make use of the particle-hole symmetry of the ES and use only the lower half of the spectrum to build $\chi_s$ with no issues due to the bulk contribution.
+
+In Fig.\ref{4} we use a profile which is invariant for $x\rightarrow x+L/2$ because it enhances the effect of the nonhomogeneity, but this symmetry is not necesarry for anything we discuss here.
+
+Another case where one can find a solution to the issue when cut-averaging is when the system is fully gapped, so that the ES doesn't have modes crossing $1/2$ for any cut.
+In this case one can make use of the particle-hole symmetry of the ES (If $\xi$ is an eigenvalue for a cut in $s$ then $1-\xi$ is an eigenvalue of the cut in $s+L/2$) and use only the lower half of the spectrum to build $\chi_s$ so that the bulk contribution vanishes enitrely.
 One can explot the fact that, even in the inhomogeneous case, the ES has a type of particle-hole symmetry.
-If $\xi$ is an eigenvalue for a cut in $s$ then $1-\xi$ is an eigenvalue of the cut in $s+L/2$.
-This means that all the information is already contained in half of the spectrum.
-Using the lower half of the spectrum ensures that we have all the information about the edge eigenstates but the bulk contribution vanishes.
-This is however only valid when there is a gap between the two halfs of the spectrum at any cut $s$.
-MERGE
-Another interesting point is that the two charge choices now leave to different $\chi_s$ since the particle-hole is no longer fulfilled for individual cuts.
-However both still lead to the same result because there is still a symmetry involving cuts related by a half-system translation.
-Interestingly for $\pm 1/2$ the quantity $\chi_s$ seems to be smaller, leading to fewer issues when averaging.
+
+Another interesting point to note is that the two charge choices now leave to different $\chi_s$ since the particle-hole symmetry is no longer fulfilled for individual cuts.
+However both still lead to the same result because of the particle-hole time $L/2$ translation symmetry.
+Interestingly for the half-integer charge choice the quantity $\chi_s$ seems to be smaller, leading to fewer issues when averaging.
 

### 2020-02-11 draft.tex

@@
 
-As a starting point we consider the model in the BDI class used in reference \cite{Song2014} because it is the simplest model which supports topological phases with winding numbers $\nu = 0,1,2$.
+As a starting point we consider the model in the BDI class used in reference \cite{Song2014} because it is a very simple model which supports topological phases with winding numbers up to $\nu =2$.
 Additionaly, we include two symmetry breaking terms, $\kappa$ and $\kappa'$, to obtain an even richer phase diagram and show that our results are not limited to any topological class.
@@
 \end{alignat*}
-and therefore it is in the BDI class, where the system has the phase diagram shown in Fig.\ref{bdi_phase_diagram} for $t=0,1$.
+and therefore it is in the BDI class, where the system has the phase diagram shown in Fig.\ref{bdi_phase_diagram} for $t=1$.
 We also show cuts with the transitions $\nu = 2 \rightarrow 0$ and $\nu = 1 \rightarrow 2$ for different values of $\kappa,\kappa'$.
-For $\kappa' \neq 0$ only $C$ is preserved and the system belongs now to the $D$ class.
+For $\kappa' \neq 0$ only $C$ is preserved and the system belongs to the $D$ class.
 The $Z$ invariant transforms into a $Z_2$ invariant and the original $\nu = 2,0$ phases become equivalent.
@@
 \subfloat[$t = 1,\kappa =\kappa'=0$]{
-  \includegraphics[width=35mm]{bdit1}
+  \includegraphics[width=45mm]{phase_diagram2.pdf}
 }}\hspace{0mm}
@@
 \subfloat[$t = 1,t'=1$, \, $\kappa =\kappa'=0$]{
-  \includegraphics[width=35mm]{1_b.pdf}
+  \includegraphics[width=35mm]{1_a.pdf}
 }
-\subfloat[$t=1,t'=1$, $\kappa = 0.3, \kappa'=0$]{\includegraphics[width=35mm]{1_c.pdf}
+\subfloat[$t=1,t'=1$, $\kappa = 0.3, \kappa'=0$]{\includegraphics[width=35mm]{1_b.pdf}
 }
@@
 \subfloat[$t = 1,t'=1$, $\kappa = 0,\kappa'=0.3$]{
-  \includegraphics[width=35mm]{1_d.pdf}
+  \includegraphics[width=35mm]{1_b.pdf}
 }
-\subfloat[$t=1,t'=1$, $\kappa = 0.3,\kappa'=0.3$]{\includegraphics[width=35mm]{1_e.pdf}
+\subfloat[$t=1,t'=1$, $\kappa = 0.3,\kappa'=0.3$]{\includegraphics[width=35mm]{1_b.pdf}
 }
 }
-\caption{Phase diagram for the system in the $BDI$ class showing the winding number of each region and surface energy spectrum for two different cuts in the phase diagram for different values of $\kappa,\kappa'$.}
+\caption{Phase diagram for the system in the $BDI$ class showing the winding number of each region and surface energy spectrum for two different cuts in the phase diagram for different values of $\kappa,\kappa'$.
+\carlos{I'm not sure what to include here.}}
 \label{bdi_phase_diagram}
@@
 
-\section{Geometric phase and Entanglement spectrum}
+\section{Zak phase and Entanglement spectrum}
 Consider a quadratic Hamiltonian in one dimension
@@
 \end{equation}
-by restricting it to the space of occupied single-particle states. The geometric phase can now be obtained as
+by restricting it to the space of occupied single-particle states. The Zak phase can now be obtained as
 \begin{equation}
@@
 In the ES we can differentiate between two types of eigenvalues, which we denote by $\xi$.
-The ones at $\xi_b = 0,1$ correspond to bulk states while for the ones in between, $0<\xi_{l,r}<1$, we find their correspondent eigenstate localized in either the left, $\xi_l$, or right, $\xi_r$, virtual edge \carlos{Cite Peschel}.
+The ones at $\xi_b = 0,1$ correspond to bulk states while for the ones in between, $0<\xi_{l,r}<1$, we find their correspondent eigenstate localized in either the left, $\xi_l$, or right, $\xi_r$, virtual edge \cite{Peschel2008}.
 
@@
 
-For symmetry-protected topological systems in one dimension there is a well-known relation between ES and the geometric phase.
-The geometric phase, which is quantized, is zero whenever there is an even number per edge of eigenvalues at $\xi = 1/2$, and $\pi$ when the number is odd.
-Beyond symmetry protected systems there have been observations, in simple systems with one edge-mode per edge, of similarities between the midgap eigenvalue of the entanglement spectrum and the geometric phase \cite{Ryu2006,Huang2012,Huang2012-2}, which in the case of the fully dimerized model becomes an exact identity \cite{Ryu2006}.
+\subsubsection{One edge mode case}
 
-Consider the combination of all eigenvalues localized in one of the edges modulo one,
+For symmetry-protected topological systems in one dimension there is a well-known relation between ES and the Zak phase.
+The Zak phase, which is quantized, is zero whenever there is an even number per edge of eigenvalues at $\xi = 1/2$, and $\pi$ when the number is odd.
+Beyond symmetry protected systems, there was observed that in the fully dimerized SSH chain, with only a single mode present in the ES, the Zak phase is equal to the midgap eigenvalue \cite{Ryu2006}.
+For slighly more complex ES it has been shown that the Zak phase follows the midgap eigenvalues qualitatively \cite{Huang2012,Huang2012-2}.
+It is known \carlos{Cite peschel} that the entanglement energies of 1D chains are equispaced sets of distinct values.
+The first result we show is that in simple systems like the SSH chain, where there is only one "distinct" mode, one can obtain the Zak phase by ordering the eigenvalues of the ES, $\xi_1 < \xi_2 < ...$, and computing
 \begin{equation}
-\chi = \sum_{i \in r} \xi_i \quad {\rm mod} \quad 1. 
+\chi = \frac{1}{2}\left(\sum_j (-1)^j \xi_j+N_A\right) \, {\rm mod} \, 1,
+\label{eqplusminus}
 \end{equation}
-As we will show below, one can find the geometric phase in the thermodynamic limit as
+where $N_A$ is the number of electrons in the ground state in region A, which in the thermodynamic limit fulfills
 \begin{equation}
-\lim_{L \rightarrow \infty}\gamma/2\pi =\lim_{L \rightarrow \infty} \chi.
-\label{mainl}
+\lim_{L \rightarrow \infty} \gamma = \lim_{L \rightarrow \infty} \chi.
 \end{equation}
-In the examples studied in references \cite{Ryu2006,Huang2012,Huang2012-2} the main contribution comes from one midgap state, while the rest of the edge states sit close to the bulk bands and therefore give a small contribution, which explains why the relation between the midgap eigenvalue and the geometric phase was almost an identity and why it breaks as we go deep into the trivial state.
-Alternatively, using that the entanglement spectrum is particle-hole symmetric for homogeneous systems we can express $\chi$ as
-\begin{equation*}
+In the case where there is only one finite eigenvalue in  the ES, like in the fully dimerized limit of the SSH chain, we recover the equality between the midgap eigenvalue and the Zak phase.
+In Fig.\ref{huang} we show the case were it is not fully dimerized.
+We see in Fig.\ref{huang}.a that even though in the symmetry protected case the ES has finite eigenvalues in the trivial phase these are double-degenerate so that they cancel each other.
+In the case where we break the symmetry we see that there is one main midgap eigenvalue, so that the Zak phase follows it very closely, but as the midgap eigenvalue approaches the bulk the contribution from the other modes becomes more relevant and the Zak phase, and therefore $\chi$, start to deviate from the midgap eigenvalue.
+\begin{figure}[h!]
+\centering
+\makebox[0pt]{
+\subfloat[$t = 1,t'=0$, \, \,$\kappa =\kappa'=0$]{
+  \includegraphics[width=35mm]{2_a.pdf}
+}
+\subfloat[$t=1,t'=0$, $\kappa = 0.3, \kappa'=0$]{\includegraphics[width=35mm]{2_b.pdf}
+}
+}
+\caption{Phase diagram for the system in the $BDI$ class showing the winding number of each region and surface energy spectrum for two different cuts in the phase diagram for different values of $\kappa,\kappa'$.}
+\label{huang}
+\end{figure}
+
+\subsubsection{Localization dependent charges}
+
+In order to generalize this to more complex systems one needs to know how to properly assign this $\pm 1/2$ charge to each entanglement mode.
+This is done according to their localization classification, so we have
+\begin{equation}
 \chi = \frac{1}{2} \left( \sum_{i\in r}\xi_i-\sum_{i\in l}\xi_i \pm \sum_{i\in b}\xi_i  + N_A \right) \quad {\rm mod} \quad 1,
-\end{equation*}
-where $N_A$ is the total number of electrons in subsystem A  in the ground state and the sign on the bulk term is irrelevant since $-1/2 \, {\rm mod} \, 1 = 1/2 \, {\rm mod} \, 1$.
-This form is preferable when doing numerics because due to finite size effects the bulk modes can acquire a finite eigenvalue and therefore in general it is difficult to classify between bulk and edge modes.
-\carlos{There is also the issue edge-modes and bulk-modes mixing in the inhomogenous case, introduce later?} In practice we diagonalize $C_A + \lambda C_A\hat{X}C_A$, where $\hat{X}$ is the position operator and $\lambda$ is a small parameter.
-The second term ensures that the originally degerate bulk states are also localized, so we can divide all eigenvalues in right and left sectors and assign the charges $1/2$ and $-1/2$, respectively.
-In Fig.\ref{bdi20} we show for several cases how Eq.
-\ref{mainl} holds up to finite size errors.
-$\chi$ is equivalent to the trace of the ES of a semi-infinite chain, but one can extract the information from the ES of a ring in the thermodynamic limit by looking at the localization structure of the eigenstates, only looking at the spectrum is not enough.
+\label{loceq}
+\end{equation}
+the charge of the bulk modes so far is irrelevant since their eigenvalue is 1 in the thermodynamic limit, but in the next section we will have to take a closer look at it.
+In the case the spectrum is double degenerate the left and right edge states might mix but one will always classify oneof the pair as left and the other one as right states, which is why the expression Eq.~\eqref{eqplusminus} simplifies so much.
+Even in the case where the degeneracies are broken the pairs of entanglement modes split in the same way (all left modes go up, right ones go down, or the other way around) so that the formula still applies as long as there are no crossings.
+This is however not the case when the degeneracies are higher than two-fold, or when we have more modes present that can cross.
+In order to always obtain localized modes we diagonalize instead $C_A + \lambda C_A\hat{X}C_A$, where $\hat{X}$ is the position operator and $\lambda$ is a small parameter that vanishes in the thermodynamic limit.
+This relation between the Zak phase and the entanglement spectrum is a consequence of an identity between the Zak phase and the many-body entanglement spectrum previously found for infinite chains \cite{Zaletel2014}.
+In Appendix A (\carlos{I have to take a look at it again}) we modified this derivation to account for the periodic boundary conditions we use and show how their result reduces to Eq.~\eqref{loceq} when expressing it in terms of the ES.
 
-One drastic example of this can be seen in figure \ref{old} , where two ES that look almost identical result in a very different geometric phase.
+In Fig.\ref{2},a,b.
+we show an example that highlights the importance of the localization structure of the ES.
+Two ES that seem at first very similar result in very different Zak phases.
+This happens because of how the two symmetry breaking terms split the four-fold degenerate midgap states.
+While $\kappa'$ splits them into two two-fold degenerate pairs whose contribution cancels, $\kappa$ split thems into two $left-left$ and $right-right$ pairs, with a finite contribution to the Zak phase.
+In Fig.\ref{2},c,d.
+we also show the $\nu = 2 \rightarrow 1$ cut in the phase-diagram, to illustrate a more complex ES.
+
+One drastic example of this can be seen in figure \ref{old} , where two ES that look almost identical result in a very different Zak phase.
+\carlos{This was an old plot I don't know if we want to keep or not}
 
@@
 \makebox[0pt]{
-\subfloat[$t = 0, t' = 2, \kappa = 10^{-5}$]{
-  \includegraphics[width=35mm]{2_a.pdf}
+\subfloat[$t = 1, t' = 2,$   $\kappa =0.3, \kappa' = 0$]{
+  \includegraphics[width=35mm]{3_a.pdf}
 }
-\subfloat[$t = 0, t' = 2, \kappa = 0.3$]{
-  \includegraphics[width=35mm]{2_b.pdf}
+\subfloat[$t = 1, t' = 2,$ $\kappa = 0, \kappa' = 0.3$]{
+  \includegraphics[width=35mm]{3_b.pdf}
 }
@@
 \makebox[0pt]{
-\subfloat[$t = 0, t' = 2, \kappa = 10^{-5}$]{
-  \includegraphics[width=35mm]{2_c.pdf}
+\subfloat[$t = 1, t' = -2,$ $\kappa = 0.3, \kappa' = 0$]{
+  \includegraphics[width=35mm]{3_c.pdf}
 }
-\subfloat[$t = 0, t' = 2, \kappa = 0.3$]{
-  \includegraphics[width=35mm]{2_d.pdf}
+\subfloat[$t = 1, t' = -2,$ $\kappa = 0, \kappa' = 0.3$]{
+  \includegraphics[width=35mm]{3_d.pdf}
 }
 }
-\caption{fill. }
-\label{bdi20}
+\caption{Entanglement spectrum (black), $\chi$ (green) and $\gamma$ (red)  }
+\label{2}
 \end{figure}
+
+\section{Inhomogeneous systems}
+
+\subsubsection{Simple case, continuous ES.}
+
+In reference \cite{Zaletel2014} translational invariance was used to derive Eq.
+\ref{zaletel} in Appendix A.
+Say we break translational invariance by adding weak disorder to the system.
+This will modify slightly the entanglement spectrum but not the charge assignment.
+Without translational invariance the entanglement spectrum will depend on where is the cut placed on the ring, even though the ground state, and therefore the Zak phase, will not.
+The first quantity one can construct that is cut-independent is the cut-average of $\chi$.
+Naively one would define
+\begin{equation}
+\bar{\chi} = \frac{1}{L}\sum_{s=1}^L \chi_s
+\label{batcutavg}
+\end{equation}
+where
+\begin{equation}
+\chi_s = \frac{1}{2} \left( \sum_{i\in r}\xi^s_i-\sum_{i\in l}\xi^s_i \pm \sum_{i\in b}\xi^s_i  + N_A \right) \quad {\rm mod} \quad 1,
+\end{equation}
+is defined for cuts at $s$ and $s+L/2$.
+However the average of a circular quantity is not well defined.
+The reason we took the modulo 1 was to get rid of the bulk contribution, which should vanish in the end result anyway.
+If we take the modulo 1 after averaging we still might have an issue due to the bulk classification being inconsistent.
+A bulk mode near the center of A, at $L/4$, might cross $L/4$ at different cuts, changing its assigned charge.
+The bulk contribution will not be constant across cuts and that will lead to errors when averaging.
+
+One can get away with expression Eq.~\eqref{batcutavg} in a few cases.
+For a system with a continuous position-dependent Hamiltonian in a trivial phase, $\chi_s$ should also be continuous.
+This tell us how we have to perform the cut-average.
+Even when discontinuities are present as long they are weak enough so that we still can tell when $\chi_s$ crosses $0,1$, we can use this information to perform the cut average.
+In the case that the width of $\chi_s$ is lower than 1 one can select the range of values so that $\chi_s$ does not cross the "$0,1$" line and then average.
+We show this in Fig.\ref{inh}.a for a system with disorder on $\kappa_s$.
+Note that even though we have random disorder the ES, and therefore $\chi_s$ shows some sign of continuity so that the cut-averaged can be performed this way even for relatively strong disorder.
+
 \begin{figure}[h!]
@@
 \makebox[0pt]{
-\subfloat[$t = 0, t' = 2, \kappa = 10^{-5}$]{
-  \includegraphics[width=35mm]{a20.pdf}
+\subfloat[$\kappa_s$]{
+  \includegraphics[width=40mm]{4_c.pdf}
 }
-\subfloat[$t = 0, t' = 2, \kappa = 0.3$]{
-  \includegraphics[width=35mm]{d20.pdf}
+\subfloat[$t=1,t'=2,m =1$ ]{
+  \includegraphics[width=40mm]{4_d.pdf}
+}
+}\hspace{0mm}
+
+\makebox[0pt]{
+\subfloat[$\kappa_s$]{
+  \includegraphics[width=40mm]{4_a.pdf}
+}
+\subfloat[$t=1,t'=2,m =1$]{
+  \includegraphics[width=40mm]{4_b.pdf}
 }
 }
-\caption{Old plot, do I keep it? }
-\label{old}
+\label{inh}
 \end{figure}
 
-This relation between the geometric phase and the entanglement spectrum is a consequence of an identity between the geometric phase and the many-body entanglement spectrum previously found for infinite chains \cite{Zaletel2014}.
-In Appendix A (\carlos{I have to take a look at it again}) we modified this derivation to account for the periodic boundary conditions we use and show how their result reduces to Eq.
-\ref{mainl} when expressing it in terms of the ES.
+In the case where $\chi_s$ is continuous but the width is larger than 1 can construct a new quantity $\chi_s'$ that is not defined modulo 1, by extending the range of $\chi_s \, {\rm mod} \, 1$ every time it crosses $0$ or $1$, such that
+\begin{equation}
+\chi_{s+1}' = {\rm Min}\{\chi'_{s+1}-\chi'_s,\chi'_{s+1}-\chi'_s+1,\chi'_{s+1}-\chi'_s-1\}
+\end{equation}
+where at each step $\chi_{s+1}'$ is chosen by minimizing the absolute value of the difference.
+In Fig.\ref{inh}.b we show an example of this.
+In Fig.\ref{disorder} we show $1-\expval{\chi_s}_{\rm cuts}/\gamma$ for different system lengths for system with weak disorder, as well as the disorder average of $1-\chi_0/\gamma$.
+We observe that both averages lead to the same difference, which scales as $O(L^{-2})$ (\carlos{I have to redo de plot, maybe scaling is different}), vanishing in the thermodynamic limit.
+The fact that both methods lead to the same error suggests that it comes strictly from a systematic finite size error in the computation of either the Zak phase or the ES which is system-independent.
 
@@
 }
-\caption{}
+\caption{\carlos{Redo}}
 \label{disorder}
@@
 
-\section{Inhomogeneous systems}
+\subsubsection{Wannier states, L/4 crossing fix}
 
-In reference \cite{Zaletel2014} translational invariance was used to derive Eq.
-\ref{zaletel} in Appendix A.
-Say we break translational invariance by adding weak disorder to the system.
-This will modify slightly the entanglement spectrum but not the charge assignment.
-Without translational invariance the entanglement spectrum will depend on where is the cut placed on the ring, even though the ground state, and therefore the geometric phase, will not.
-One thing one could do to get rid of the cut-dependence is to perform an average over all the spatial cuts of expression \ref{mainl}.
-Defining now the cut-dependent $\chi$ as
-\begin{equation*}
-\chi_s = \frac{1}{2} \left( \sum_{i\in r}\xi^s_i-\sum_{i\in l}\xi^s_i \pm \sum_{i\in b}\xi^s_i  + N_A \right) \quad,
-\end{equation*}
-for a cut in positions $s$ and $s+L/2$. We can write
+In a general case where the discontinuities are too strong and the discussed above does not apply one has to take a closer look into the bulk states to ensure a consistent classification.
+As we mentioned above the bulk states farthest away from the edges are the problematic ones.
+They are also the ones best approximated by the Wannier states obtained from the chain with periodic boundary conditions as the eigenstates of
 \begin{equation}
-\gamma = \expval{\chi_s}_{\rm cuts} \, {\rm mod}\, 1.
-\label{gphase_ih_mod}
+C\hat{X}_{\rm PBC}C
 \end{equation}
-However this expression is not practical.
-In order for the bulk contribution to vanish with the modulo $1$ it has to be constant for different cuts.
-Last section we divided the bulk modes in right and left, but now we encounter the problem that their localization changes with the cut, so that there is the possibility that one bulk mode around $1/2$ changes its charge, leading to a wrong result.
-To get around this problem we can get rid of the bulk contribution before averaging,
+where $\hat{X}_{\rm PBC} = \exp\{i 2\pi \hat{X}/L\}$.
+Similar to when we go through different cuts labeled by $s$, we now consider the Wannier states dependent on where we define our lattice site 1, shifted by $s$.
+The obtained Wannier states for $s$ and $s+1$ will be therefore exctly the same, but shifted one lattice site.
+By considering their average position $\expval{\hat{X}_{PBC}}_{i,s}$ we can define a set of $2L$ Wannier states that depend continuously on $s$ (provided that the Hamiltonian is continuous).
+These Wannier states are then a good approximation for the Bulk states of $C_A$ we find for the different cuts.
+Consider now the quantity $\sum_i\expval{\hat{X}}_{i,s}$.
+So far everything was $s$ independent as our system was periodic, but by considering the operator $\hat{X}$ we introduce an unphysical boundary so that it will depend on where we put the first and last sites.
+If the states do not cross this unphysical boundary then $\sum_i\expval{\hat{X}}_{i,s} = \sum_i\expval{\hat{X}_{\rm PBC}}_{i,s} $ and it will be a constant.
+However, when a Wannier state crosses the boundary we obtain the same constant plus $L$.
+If we put this boundary at $L/4$ the object
 \begin{equation}
-\gamma = \expval{\chi_s \, {\rm mod}\, 1}_s \, {\rm mod}\, 1,
-\label{gphase_ih_modmod}
+\frac{1}{L}\left(\sum_i\expval{\hat{X}}_{i,s}-\sum_i\expval{\hat{X}}_{i,0}\right)
 \end{equation}
-however the average of a circular quantity is not well defined.
-We will explain how one can perform the cut-average for different cases so that one can recover expression \ref{gphase_ih_mod} from expression \ref{gphase_ih_modmod}.
-
-For the case when the width of $\chi_s$ is lower than one, there is a range of values $[x,x+1]$ where $\chi_s$ does not cross $0,1$, and averaging over this range makes sure than once we take the second modulo $1$ all the different possible results for the cut-average are reduced to a single one.
-The shortcomings of expression \ref{modmod} will be discussed later, but to show that this cut-average is correct we will consider now a system in the topological phase to which we add weak disorder, so one can then take the cut-average around $[0,1]$.
-In Fig.\ref{disorder} we show $1 - \expval{\chi_s \, {\rm mod}\, 1}_s \, {\rm mod}\, 1 /\gamma$ in red for different system sizes.
-In blue we have shown the same difference using a disorder average.
-Both values match fairly well and they seem to vanish in the thermodynamic limit, scaling as $O(L^{-2})$.
-The fact that different disorder configurations match so well points to a systematic finite-size error in the way we calculate either the ES or the geometric phase that is system-independent.
-From this result one could think that the cut-average is some sort of disorder average.
-The implication however goes in the opposite direction.
-If the cut-average is assumed to be correct, that implies that in the thermodynamic limit we can group all the infinite different disorder configurations into an infinite set of $L$ configurations relating to different cuts of the same system, which average individually to the correct $\gamma_{\rm conf}$.
-
-For stronger disorder one cannot be sure if $\chi_s$ crosses $0,1$ or not so this method cannot be applied but for continuous inhomogeneous perturbations one can follow $\chi_s$, so we can use expression \ref{modmod} even when the inhomogeneity is strong.
-In Fig.\ref{4} we show the results using a position dependent $\kappa$ parameter, shown in figure \ref{4}.a.
-In cases Fig.\ref{4}.b,c the width of $\chi_s$ is small so we can average taking the value ranges $[0,1]$ and $[-1/2,1/2]$ respectively and we see that the cut average matches the geometric phase perfectly.
-However for stronger perturbations, as shown in Fig.\ref{4}.d, one can have that the width of $\chi_s$ is bigger than 1 and there is no meaningful range of values one can average.
-Instead what one can do is to "double the periodicity of $\chi_s$" by adding or substracting $1$ each time it crosses $0,1$, taking the cut-average of and then take the modulo 1.
+will be 1 (0) when a states has crossed the boundary or 0 (-1) when it has not, depending on the number of states in A at $s=0$, which we take as reference.
+Depending which state we take as reference will introduce a global integer shift for all cuts, so it is not relevant as it will go away after we average and take the modulo 1.
+By adding this term to our cut-average we will compensate for when our bulk state in the middle of A crosses $L/4$ and its charge changes.
+Leading to a continuous $\chi_s$ which can be averaged.
+In Fig.\ref{wan} we show how this works for one case.
+In Fig.\ref{wan}.a we show the expectation value of the position operator for both the entanglement eigenstates (in black) and the Wannier states (in red) for different cuts.
+In Fig.\ref{wan}.b and c we show $\chi_s$, which displays jumps due to a bulk state crossing $L/4$, and
+\begin{equation}
+\chi_s' = \chi_s + \frac{1}{L}\left(\sum_i\expval{\hat{X}}_{i,s}-\sum_i\expval{\hat{X}}_{i,0}\right),
+\end{equation}
+where the position is defined in $[-3L/4,L/4]$ to correctly compensate for these jumps.
+This method conceptually works perfectly whenever the bulk states of the ES match the Wannier states, i.e.
+in the thermodynamic limit, but it is not clear what the small parameter is.
+It depends on the $\lambda$ used (it should be small enough to not affect ES but large enough to localize the ES eigenstates) but also on the model.
+Models with longe-range hopping, for example, lead to a higher number of edge modes in the ES which makes the effect of the edge go deeper into the bulk.
+\carlos{We should think again if there is a better method to get this}
 
@@
 \makebox[0pt]{
-\subfloat[]{
-  \includegraphics[width=40mm]{4_a.pdf}
-}
-\subfloat[]{
-  \includegraphics[width=40mm]{4_b.pdf}
-}
-}\hspace{0mm}
+\subfloat[$\expval{\hat{X}}_{i,s}$ for different cuts, for the entanglement eigenstates (black) and the Wannier states (red)]{
+  \includegraphics[width=90mm]{loca.pdf}
+}}\hspace{0mm}
 
 \makebox[0pt]{
-\subfloat[]{
-  \includegraphics[width=40mm]{4_c.pdf}
+\subfloat[$\chi_s$]{
+  \includegraphics[width=45mm]{wan1.pdf}
 }
-\subfloat[]{
-  \includegraphics[width=40mm]{4_d.pdf}
+\subfloat[$\chi_s'$ ]{
+  \includegraphics[width=45mm]{wan2.pdf}
 }
@@
 
-\caption{ fill. put .5 line, also break the symmetry. with an envelope.}
-\label{4}
+\caption{a)Profile used for the inhomogeneous part $\kappa_s$. b,c,d) ES (black), $\chi_s$ mod $1$ (green), $\gamma$ (red), $\expval{\chi_s}_{\rm cuts}$ mod $1$ (dashed blue).}
+\label{wan}
 \end{figure}
-
-In Fig.\ref{4} we use a profile which is invariant for $x\rightarrow x+L/2$ because it enhances the effect of the nonhomogeneity, but this symmetry is not necesarry for anything we discuss here.
-
-Another case where one can find a solution to the issue when cut-averaging is when the system is fully gapped, so that the ES doesn't have modes crossing $1/2$ for any cut.
-In this case one can make use of the particle-hole symmetry of the ES (If $\xi$ is an eigenvalue for a cut in $s$ then $1-\xi$ is an eigenvalue of the cut in $s+L/2$) and use only the lower half of the spectrum to build $\chi_s$ so that the bulk contribution vanishes enitrely.
-One can explot the fact that, even in the inhomogeneous case, the ES has a type of particle-hole symmetry.
-
-Another interesting point to note is that the two charge choices now leave to different $\chi_s$ since the particle-hole symmetry is no longer fulfilled for individual cuts.
-However both still lead to the same result because of the particle-hole time $L/2$ translation symmetry.
-Interestingly for the half-integer charge choice the quantity $\chi_s$ seems to be smaller, leading to fewer issues when averaging.
 
@@
 However, as we did for systems without translational invariance, one can extend expression \ref{mainl} for systems with pairing terms as well.
-Since one computes the geometric phase from the single-particle Bogoliubov-de Gennes Hamiltonian, we can obtain the correspondent entanglement spectrum and use expression \ref{mainl} to obtain the geometric phase.
+Since one computes the Zak phase from the single-particle Bogoliubov-de Gennes Hamiltonian, we can obtain the correspondent entanglement spectrum and use expression \ref{mainl} to obtain the Zak phase.
 Take for example the kitaev chain, with the following Hamiltonian
@@
 And the charge can be assigned so that $Q_p = \sum_{\alpha = b,l,r} q_\alpha\sum_{i \in \alpha} n_i$ with the constrain that $q_r,\pm q_b = q_l-1$ \textcolor{red}{Why?}.
-As the main result of \cite{Zaletel2014} the geometric phase is then found to be
+As the main result of \cite{Zaletel2014} the Zak phase is then found to be
 \begin{align}
@@
 
-Expressing the contribution of the many-body entanglement spectrum to the geometric phase in terms of the ES, we have
+Expressing the contribution of the many-body entanglement spectrum to the Zak phase in terms of the ES, we have
 \begin{align*}

### 2020-02-11 draft.tex
update 11-2
..

@@
 The most notable examples are the fractional quantum Hall effect~\cite{Tsui1982} and quantum spin liquids~\cite{Balents2010spin}. 
-However, since 2005~\cite{kane2005quantum,roy2009topological} the focus has shifted to non-interacting topological phases, which can be characterized in terms of topological invariants~\cite{ryu2010topological}.
+However, since 2005~\cite{kane2005quantum, roy2009topological} the focus has shifted to non-interacting topological phases, which can be characterized in terms of topological invariants~\cite{ryu2010topological}.
 Symmetries are often necessary to protect these topological phases, and determine which distinct topological phases can be realized for a given dimensionality.
@@
 There are a few tools to identify topological phases. 
-A very efficient one is the entanglement entropy~\cite{}, which allows one to identify the total quantum dimension of the underlying topological field theory. 
+A very efficient one is the entanglement entropy~\cite{Kitaev2006topological, Levin2006detecting}, which allows one to identify the total quantum dimension of the underlying topological field theory. 
 As such, the entanglement entropy can only be used for interacting systems. 
@@
 }
-\caption{Phase diagram for the system in the $BDI$ class showing the winding number of each region and surface energy spectrum for two different cuts in the phase diagram for different values of $\kappa,\kappa'$.
+\caption{\maria{axis labels} Phase diagram for the system in the $BDI$ class showing the winding number of each region and surface energy spectrum for two different cuts in the phase diagram for different values of $\kappa,\kappa'$.
 \carlos{I'm not sure what to include here.}}
@@
 The ones at $\xi_b = 0,1$ correspond to bulk states while for the ones in between, $0<\xi_{l,r}<1$, we find their correspondent eigenstate localized in either the left, $\xi_l$, or right, $\xi_r$, virtual edge \cite{Peschel2008}.
-
+\mariac{introduce entanglement energy, relation between C and energies}
 \section{Homogeneous system}
 
-\subsubsection{One edge mode case}
+\subsubsection{Single edge mode}
+\mariac{find better wording?}
 
@@
 
-\subsubsection{Localization dependent charges}
+\subsubsection{Multiple edge modes}
+%\subsubsection{Localization dependent charges}
 
@@
 we also show the $\nu = 2 \rightarrow 1$ cut in the phase-diagram, to illustrate a more complex ES.
-
-One drastic example of this can be seen in figure \ref{old} , where two ES that look almost identical result in a very different Zak phase.
-\carlos{This was an old plot I don't know if we want to keep or not}
 

### 2020-02-13 draft.tex
Update homogeneous

@@
 
-\subsubsection{Single edge mode}
+\subsubsection{Nearest-neighbor hopping}
 \mariac{find better wording?}
 
-For symmetry-protected topological systems in one dimension there is a well-known relation between ES and the Zak phase.
+For symmetry-protected topological systems in one dimension there is a well-known relation between ES and the Zak phase \carlos{cite}.
 The Zak phase, which is quantized, is zero whenever there is an even number per edge of eigenvalues at $\xi = 1/2$, and $\pi$ when the number is odd.
@@
 For slighly more complex ES it has been shown that the Zak phase follows the midgap eigenvalues qualitatively \cite{Huang2012,Huang2012-2}.
-It is known \carlos{Cite peschel} that the entanglement energies of 1D chains are equispaced sets of distinct values.
-The first result we show is that in simple systems like the SSH chain, where there is only one "distinct" mode, one can obtain the Zak phase by ordering the eigenvalues of the ES, $\xi_1 < \xi_2 < ...$, and computing
+The first result we show is that in systems with nearest-neighbor hopping, like the SSH chain, one can obtain the Zak phase by ordering the eigenvalues of the ES, $\xi_1 < \xi_2 < ...$, and computing
 \begin{equation}
@@
 \end{equation}
+\carlos{Note: Peschel only finds the entanglement energies to be equispaced in systems with nearest-neighbor hopping.
+The general picture is seems much more complex.
+So far it worked for the two-band models I considered because they effectively only have nearest-neighbor hopping, adding a $\sigma_0$ hopping for example breaks this.
+See in Fig.\ref{eh} that for $t'\neq 0$ this is not true anymore, and even in the $\nu=1$ phase there can be crossings.
+}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=80mm]{eh1.pdf}
+\caption{Entanglement energy spectrum for a cut with $t'=-2$.}
+\label{eh}
+\end{figure}
+
 In the case where there is only one finite eigenvalue in  the ES, like in the fully dimerized limit of the SSH chain, we recover the equality between the midgap eigenvalue and the Zak phase.

### 2020-02-13 draft.tex
Update homogeneous

@@
 
-As a starting point we consider the model in the BDI class used in reference \cite{Song2014} because it is a very simple model which supports topological phases with winding numbers up to $\nu =2$.
+To showcase our results we consider as a starting point the model in the BDI class used in reference \cite{Song2014} because it is a very simple model which supports topological phases with winding numbers up to $\nu =2$.
 Additionaly, we include two symmetry breaking terms, $\kappa$ and $\kappa'$, to obtain an even richer phase diagram and show that our results are not limited to any topological class.
@@
 and therefore it is in the BDI class, where the system has the phase diagram shown in Fig.\ref{bdi_phase_diagram} for $t=1$.
-We also show cuts with the transitions $\nu = 2 \rightarrow 0$ and $\nu = 1 \rightarrow 2$ for different values of $\kappa,\kappa'$.
-For $\kappa' \neq 0$ only $C$ is preserved and the system belongs to the $D$ class.
-The $Z$ invariant transforms into a $Z_2$ invariant and the original $\nu = 2,0$ phases become equivalent.
+We also a cut through the three phases (dotted line in the phase diagram) for different values of $\kappa,\kappa'$.
+For $\kappa = \kappa'= 0$ when the system is in the BDI class we the system has a number of zero modes equal to the winding number, as it is well known.
+For $\kappa' \neq 0$ only $C$ is preserved and the system belongs to the $D$ class As the zero modes in the $\nu=2$ phase split from zero pairwise, the $\nu = 2$ and $\nu=0$ phases become equivalent.
 For $\kappa \neq 0$ only $T$ is preserved and the system is in the trivial AI class, where all zero-energy modes split.
+With both $\kappa \neq 0$ and $\kappa' \neq0$ the system has no symmetry, belonging to the $A$ class and the edge modes that were doubly degenerate in the $D$ class split as well.
 \begin{figure}[h!]
@@
 \makebox[0pt]{
-\subfloat[$t = 1,t'=1$, \, $\kappa =\kappa'=0$]{
+\subfloat[$t = 1,t'=-2$, \, $\kappa =\kappa'=0$]{
   \includegraphics[width=35mm]{1_a.pdf}
 }
-\subfloat[$t=1,t'=1$, $\kappa = 0.3, \kappa'=0$]{\includegraphics[width=35mm]{1_b.pdf}
+\subfloat[$t=1,t'=-2$, $\kappa = 0.3, \kappa'=0$]{\includegraphics[width=35mm]{1_b.pdf}
 }
@@
 \makebox[0pt]{
-\subfloat[$t = 1,t'=1$, $\kappa = 0,\kappa'=0.3$]{
-  \includegraphics[width=35mm]{1_b.pdf}
+\subfloat[$t = 1,t'=-2$, $\kappa = 0,\kappa'=0.3$]{
+  \includegraphics[width=35mm]{1_c.pdf}
 }
-\subfloat[$t=1,t'=1$, $\kappa = 0.3,\kappa'=0.3$]{\includegraphics[width=35mm]{1_b.pdf}
+\subfloat[$t=1,t'=-2$, $\kappa = 0.3,\kappa'=0.3$]{\includegraphics[width=35mm]{1_d.pdf}
 }
 }
-\caption{\maria{axis labels} Phase diagram for the system in the $BDI$ class showing the winding number of each region and surface energy spectrum for two different cuts in the phase diagram for different values of $\kappa,\kappa'$.
-\carlos{I'm not sure what to include here.}}
+\caption{Phase diagram for the system in the $BDI$ class showing the winding number of each region and surface energy spectrum for two different cuts in the phase diagram for different values of $\kappa,\kappa'$.}
 \label{bdi_phase_diagram}
@@
 where $[U]_{j\beta,p\mu} = u^{p\mu}_{j\beta}$ is the unitary matrix that diagonalizes $H$.
-The polarization can be obtained as the many-body average of position operator, regularized for periodic boudnary conditions \cite{Resta1997}
+The polarization can be obtained as the many-body average of the position operator, but it has to be regularized for periodic boundary conditions \cite{Resta1997}.
+We can define a position operator for PBC as
+\begin{equation}
+\hat{X}_{\rm PBC} = e^{-i\frac{2\pi}{L}\hat{X}},
+\end{equation}
+where $\hat{X}$ is the sum of all the single-particle position operators. We can consider then the many-body average of the position operator over the ground state, defined as
 \begin{align*}
-\expval{X} = -\frac{L}{2\pi} \rm{ Im\, Ln }\bra{\Psi_0}e^{-i \frac{2\pi}{L}\hat{X}}\ket{\Psi_0},
+\expval{\hat{X}_{\rm PBC}}_0 \equiv -\frac{L}{2\pi} \rm{ Im\, Ln }\bra{\Psi_0}\hat{X}_{\rm PBC}\ket{\Psi_0},
 \end{align*}
-where $\ket{\Psi_0}$ is the many-body ground state and $\hat{X}$ is the sum of all the single-particle position operators.
-Using that the ground state is a Slater determinant of the single-particle eigenstates, $\ket{u^{p\mu}}$, it can be rewritten as
+\carlos{This is the notation used in the Resta paper, but I'm not sure if it's confusing.}
+
+where $\ket{\Psi_0}$ is the many-body ground state. Using that the ground state is a Slater determinant of the single-particle eigenstates, $\ket{u^{p\mu}}$, it can be rewritten as
 \begin{align*}
-\expval{\hat{X}} = -\frac{L}{2\pi} \rm{Im \, Ln}\, {\rm det} \, \tilde{S},
+\expval{\hat{X}_{\rm PBC}}_0 = -\frac{L}{2\pi} \rm{Im \, Ln}\, {\rm det} \, \tilde{S},
 \end{align*}
@@
 \begin{equation}
-\gamma = \expval{\hat{X}}\frac{2\pi}{L} + \frac{1}{2}
+\gamma = \expval{\hat{X}_{\rm PBC}}_0\frac{2\pi}{L} + \frac{1}{2}
 \end{equation}
-\carlos{There is an extra 1/2 factor compared to Kivelson's formula which I don't know where it is coming from.}
+\carlos{There is an extra 1/2 factor compared to Restas's formula which I don't know where it is coming from.}
 The correlation function in position space is given by
@@
 \end{align*}
-The entanglement spectrum (ES) is then obtained as the spectrum of the subsystem correlation function obtained after bisecting the system into two equal parts.
+The single-particle entanglement spectrum (ES) is then obtained as the spectrum of the subsystem correlation function obtained after bisecting the system into two equal parts.
 
@@
 Beyond symmetry protected systems, there was observed that in the fully dimerized SSH chain, with only a single mode present in the ES, the Zak phase is equal to the midgap eigenvalue \cite{Ryu2006}.
-For slighly more complex ES it has been shown that the Zak phase follows the midgap eigenvalues qualitatively \cite{Huang2012,Huang2012-2}.
+For cases with more edge modes present in the ES it has been shown that the Zak phase follows the midgap eigenvalues qualitatively \cite{Huang2012,Huang2012-2}.
 The first result we show is that in systems with nearest-neighbor hopping, like the SSH chain, one can obtain the Zak phase by ordering the eigenvalues of the ES, $\xi_1 < \xi_2 < ...$, and computing
@@
 In the case where there is only one finite eigenvalue in  the ES, like in the fully dimerized limit of the SSH chain, we recover the equality between the midgap eigenvalue and the Zak phase.
-In Fig.\ref{huang} we show the case were it is not fully dimerized.
-We see in Fig.\ref{huang}.a that even though in the symmetry protected case the ES has finite eigenvalues in the trivial phase these are double-degenerate so that they cancel each other.
-In the case where we break the symmetry we see that there is one main midgap eigenvalue, so that the Zak phase follows it very closely, but as the midgap eigenvalue approaches the bulk the contribution from the other modes becomes more relevant and the Zak phase, and therefore $\chi$, start to deviate from the midgap eigenvalue.
+In Fig.\ref{huang} we show show the dashed cut in the phase diagram shown above, which is equivalent to an SSH chain.
+We see that even though the ES has finite eigenvalues in the trivial phase these are double-degenerate so that they cancel each other.
+In the case where we break the symmetry we see that there is one main midgap eigenvalue, so that the Zak phase follows it very closely, but as the midgap eigenvalue approaches the bulk, the contribution from the other modes becomes more relevant and the Zak phase start to deviate from the midgap eigenvalue.
 \begin{figure}[h!]
@@
 %\subsubsection{Localization dependent charges}
-
-In order to generalize this to more complex systems one needs to know how to properly assign this $\pm 1/2$ charge to each entanglement mode.
-This is done according to their localization classification, so we have
+\carlos{change title.}
+In expression 	\eqref{eqplusminus} we assigned a charge of $\pm1/2$ to each entanglement eigenstate depending on the eigenvalue ordering. 
+In order to generalize this to systems with longer-range hopping one needs to know how to properly assign this $\pm 1/2$ charge to each entanglement mode.
+We found that this is done according to their localization, so we have
 \begin{equation}
@@
 \end{equation}
-the charge of the bulk modes so far is irrelevant since their eigenvalue is 1 in the thermodynamic limit, but in the next section we will have to take a closer look at it.
-In the case the spectrum is double degenerate the left and right edge states might mix but one will always classify oneof the pair as left and the other one as right states, which is why the expression Eq.~\eqref{eqplusminus} simplifies so much.
-Even in the case where the degeneracies are broken the pairs of entanglement modes split in the same way (all left modes go up, right ones go down, or the other way around) so that the formula still applies as long as there are no crossings.
-This is however not the case when the degeneracies are higher than two-fold, or when we have more modes present that can cross.
-In order to always obtain localized modes we diagonalize instead $C_A + \lambda C_A\hat{X}C_A$, where $\hat{X}$ is the position operator and $\lambda$ is a small parameter that vanishes in the thermodynamic limit.
+by defining $ \chi$ modulo $1$ we make the charge of the bulk modes irrelevant but in the next section we will have to take a closer look at it.
+We find the charge of the edge modes by computing $\expval{\hat{X}}_i$ and seeing if it is above or below $L/4$.
+In the case the ES is double degenerate at most, left and right edge states might mix but one will always classify one state of the pair as left and the other one as right, this is however not the case when the degeneracies are higher than two-fold.
+In order to ensure that we obtain edge modes localized at one edge, instead of $C_A$ we diagonalize $C_A + \lambda C_A\hat{X}C_A$, where $\hat{X}$ is the position operator and $\lambda$ is a small parameter that vanishes in the thermodynamic limit.
+
+The reason expression~\eqref{loceq} simplifies so much for systems with nearest-neighbor hopping is that in this case it was found that all the eigenvalues of the ES are related to each other.
+The entanglement energies  (\carlos{Peschel always refers to them as entanglement energies or eigenvalues of the entanglement Hamiltonian.
+Entanglement spectrum he uses for the spectrum of the reduced density matrix, so it makes sense to use single-particle ES for the spectrum of CA}) come in pairs of left and right modes that are equispaced in the thermodynamic limit, so that the charges assigned are always $ \pm(+-+-+-...)$.
+In the case that the modes cross each other they do so for every pair, so that it just ammounts to a global minus sign in the charge assignment.
+
 This relation between the Zak phase and the entanglement spectrum is a consequence of an identity between the Zak phase and the many-body entanglement spectrum previously found for infinite chains \cite{Zaletel2014}.
@@
 
-In Fig.\ref{2},a,b.
-we show an example that highlights the importance of the localization structure of the ES.
-Two ES that seem at first very similar result in very different Zak phases.
+In Fig.\ref{2} we show an example that highlights the importance of the localization structure of the ES.
+For the solid line cut in the phase-diagram, with the $\nu = 2 \rightarrow 0 $ transition, two ES that seem at first very similar result in very different Zak phases.
 This happens because of how the two symmetry breaking terms split the four-fold degenerate midgap states.
 While $\kappa'$ splits them into two two-fold degenerate pairs whose contribution cancels, $\kappa$ split thems into two $left-left$ and $right-right$ pairs, with a finite contribution to the Zak phase.
-In Fig.\ref{2},c,d.
-we also show the $\nu = 2 \rightarrow 1$ cut in the phase-diagram, to illustrate a more complex ES.
+We also show the cut through the dotted line that shows all possible phases to illustrate how our results also apply to more complex ES.
 
@@
 Therefore the results already shown for the BDI model studied here are also applicable to the Kitaev chain.
+\carlos{check again becase I though that the kitaev chain had left-left right-right pairs, so in this case the entnsglement energies are not equispaced}
 

### 2020-02-28 draft.tex
.

@@
 
-For symmetry-protected topological systems in one dimension there is a well-known relation between ES and the Zak phase \carlos{cite}.
+For symmetry-protected topological systems in one dimension there is a well-known relation between ES and the Zak phase \cite{Peschel2008}.
 The Zak phase, which is quantized, is zero whenever there is an even number per edge of eigenvalues at $\xi = 1/2$, and $\pi$ when the number is odd.
@@
 \end{equation}
-\carlos{Note: Peschel only finds the entanglement energies to be equispaced in systems with nearest-neighbor hopping.
-The general picture is seems much more complex.
-So far it worked for the two-band models I considered because they effectively only have nearest-neighbor hopping, adding a $\sigma_0$ hopping for example breaks this.
-See in Fig.\ref{eh} that for $t'\neq 0$ this is not true anymore, and even in the $\nu=1$ phase there can be crossings.
-}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=80mm]{eh1.pdf}
-\caption{Entanglement energy spectrum for a cut with $t'=-2$.}
-\label{eh}
-\end{figure}
 
@@
 
-\subsubsection{Multiple edge modes}
-%\subsubsection{Localization dependent charges}
-\carlos{change title.}
+\subsubsection{General case}
+
 In expression 	\eqref{eqplusminus} we assigned a charge of $\pm1/2$ to each entanglement eigenstate depending on the eigenvalue ordering. 
@@
 
-The reason expression~\eqref{loceq} simplifies so much for systems with nearest-neighbor hopping is that in this case it was found that all the eigenvalues of the ES are related to each other.
-The entanglement energies  (\carlos{Peschel always refers to them as entanglement energies or eigenvalues of the entanglement Hamiltonian.
-Entanglement spectrum he uses for the spectrum of the reduced density matrix, so it makes sense to use single-particle ES for the spectrum of CA}) come in pairs of left and right modes that are equispaced in the thermodynamic limit, so that the charges assigned are always $ \pm(+-+-+-...)$.
-In the case that the modes cross each other they do so for every pair, so that it just ammounts to a global minus sign in the charge assignment.
+The reason expression~\eqref{loceq} simplifies so much for systems with only nearest-neighbor hopping is that it is known that in this case the eigenstates of $C_A$ always come with an alternating edge localization (left-right-left...), so that the signs assigned are always $ \pm(+-+-+-...)$.
+Their eigenvalues are also related, so that the modes can only cross each other if all of them do so two by two, so that it just ammounts to a global minus sign.
 
 This relation between the Zak phase and the entanglement spectrum is a consequence of an identity between the Zak phase and the many-body entanglement spectrum previously found for infinite chains \cite{Zaletel2014}.
-In Appendix A (\carlos{I have to take a look at it again}) we modified this derivation to account for the periodic boundary conditions we use and show how their result reduces to Eq.~\eqref{loceq} when expressing it in terms of the ES.
+In Appendix A we modified this derivation to account for the periodic boundary conditions we use and show how their result reduces to Eq.~\eqref{loceq} when expressing it in terms of the ES.
 
@@
 
-In reference \cite{Zaletel2014} translational invariance was used to derive Eq.
-\ref{zaletel} in Appendix A.
-Say we break translational invariance by adding weak disorder to the system.
-This will modify slightly the entanglement spectrum but not the charge assignment.
-Without translational invariance the entanglement spectrum will depend on where is the cut placed on the ring, even though the ground state, and therefore the Zak phase, will not.
-The first quantity one can construct that is cut-independent is the cut-average of $\chi$.
-Naively one would define
+In reference \cite{Zaletel2014} translational invariance was used to derive the relation between the many-body entanglement spectrum and the Zak phase.
+If this symmetry is broken by, for example, adding weak disorder to the system, this will modify slightly the eigenvalues of the reduced density matrix but not the charge assignment in Eq.~\eqref{zaletel} in the Appendix.
+The left side of Eq.~\eqref{zaletel} will be cut independent while the right side is not, and therefore cannot be correct.
+
+We are interested therefore in finding a cut-independent quantity from the ES that might still result in the Zak phase. The first thing one can do is a cut-average of $\chi$. Naively one would define
 \begin{equation}
-\bar{\chi} = \frac{1}{L}\sum_{s=1}^L \chi_s
+\bar{\chi} = \frac{1}{L}\sum_{s=1}^L (\chi_s \, {\rm mod} \, 1 )
 \label{batcutavg}
@@
 \begin{equation}
-\chi_s = \frac{1}{2} \left( \sum_{i\in r}\xi^s_i-\sum_{i\in l}\xi^s_i \pm \sum_{i\in b}\xi^s_i  + N_A \right) \quad {\rm mod} \quad 1,
+\chi_s = \frac{1}{2} \left( \sum_{i\in r}\xi^s_i-\sum_{i\in l}\xi^s_i \pm \sum_{i\in b}\xi^s_i  + N_A \right) 
 \end{equation}
-is defined for cuts at $s$ and $s+L/2$.
+is obtained for cuts at $s$ and $s+L/2$.
 However the average of a circular quantity is not well defined.
-The reason we took the modulo 1 was to get rid of the bulk contribution, which should vanish in the end result anyway.
-If we take the modulo 1 after averaging we still might have an issue due to the bulk classification being inconsistent.
+The reason we defined $\chi$ with modulo 1 in Eq.~\eqref{loceq} was to get rid of the choice of sign for the bulk, which should not matter for the end result anyway.
+Instead, we can calculate
+\begin{equation}
+\bar{\chi} = \frac{1}{L}\sum_{s=1}^L \chi_s \quad {\rm mod} \quad 1 
+\label{goodcutavg}
+\end{equation}
+but in order to obtain the correct result we need the contribution from the bulk to be constant across different cuts,$\xi_{i \in b}^s = \xi_{i \in b}$, so that in the end
+\begin{align}
+\bar{\chi} =& \frac{1}{2L}\sum_{s=1}^L \left( \sum_{i\in r}\xi^s_i-\sum_{i\in l}\xi^s_i \right) \quad {\rm mod} \quad 1 \\
+&+\frac{1}{2} \left(\sum_{i\in b}\xi^s_i + N_A \right) \quad {\rm mod} \quad 1 \nonumber
+\label{goodcutavg}
+\end{align}
+If instead we only take the modulo 1 after averaging we still might have an issue due to the charge assignment of the bulk modes being inconsistent.
 A bulk mode near the center of A, at $L/4$, might cross $L/4$ at different cuts, changing its assigned charge.
@@
 One can get away with expression Eq.~\eqref{batcutavg} in a few cases.
-For a system with a continuous position-dependent Hamiltonian in a trivial phase, $\chi_s$ should also be continuous.
+For a system with a continuous position-dependent Hamiltonian $\chi_s$ should also be continuous.
 This tell us how we have to perform the cut-average.
-Even when discontinuities are present as long they are weak enough so that we still can tell when $\chi_s$ crosses $0,1$, we can use this information to perform the cut average.
+Even when discontinuities are present as long they are weak enough so that we still can tell when $\chi_s$ crosses "$0,1$" (\carlos{write properly}), the cut-average can be performed.
 In the case that the width of $\chi_s$ is lower than 1 one can select the range of values so that $\chi_s$ does not cross the "$0,1$" line and then average.
-We show this in Fig.\ref{inh}.a for a system with disorder on $\kappa_s$.
-Note that even though we have random disorder the ES, and therefore $\chi_s$ shows some sign of continuity so that the cut-averaged can be performed this way even for relatively strong disorder.
+We show this in Fig.\ref{inh} for a system with disorder on $\kappa_s$.
+Note that even though we have random disorder, the ES and $\chi_s$ show some sign of continuity so that the cut-averaged can be performed this way even for relatively strong disorder.
 
@@
 }
-}
+}\hspace{0mm}
+
+\makebox[0pt]{
+\subfloat[$\chi_s'$]{
+  \includegraphics[width=40mm]{xicfix.pdf}
+}}
 \label{inh}
@@
 
-In the case where $\chi_s$ is continuous but the width is larger than 1 can construct a new quantity $\chi_s'$ that is not defined modulo 1, by extending the range of $\chi_s \, {\rm mod} \, 1$ every time it crosses $0$ or $1$, such that
-\begin{equation}
-\chi_{s+1}' = {\rm Min}\{\chi'_{s+1}-\chi'_s,\chi'_{s+1}-\chi'_s+1,\chi'_{s+1}-\chi'_s-1\}
-\end{equation}
-where at each step $\chi_{s+1}'$ is chosen by minimizing the absolute value of the difference.
-In Fig.\ref{inh}.b we show an example of this.
+In the case where $\chi_s$ is continuous but the width is larger than 1 can construct a new quantity $\chi_s'$ that is not defined modulo 1, by extending the range of $\chi_s \, {\rm mod} \, 1$ every time it crosses $0$ or $1$.
+In Fig.\ref{inh}.d we show an example of this.
 In Fig.\ref{disorder} we show $1-\expval{\chi_s}_{\rm cuts}/\gamma$ for different system lengths for system with weak disorder, as well as the disorder average of $1-\chi_0/\gamma$.
@@
 
-In a general case where the discontinuities are too strong and the discussed above does not apply one has to take a closer look into the bulk states to ensure a consistent classification.
-As we mentioned above the bulk states farthest away from the edges are the problematic ones.
-They are also the ones best approximated by the Wannier states obtained from the chain with periodic boundary conditions as the eigenstates of
-\begin{equation}
-C\hat{X}_{\rm PBC}C
-\end{equation}
-where $\hat{X}_{\rm PBC} = \exp\{i 2\pi \hat{X}/L\}$.
-Similar to when we go through different cuts labeled by $s$, we now consider the Wannier states dependent on where we define our lattice site 1, shifted by $s$.
-The obtained Wannier states for $s$ and $s+1$ will be therefore exctly the same, but shifted one lattice site.
-By considering their average position $\expval{\hat{X}_{PBC}}_{i,s}$ we can define a set of $2L$ Wannier states that depend continuously on $s$ (provided that the Hamiltonian is continuous).
-These Wannier states are then a good approximation for the Bulk states of $C_A$ we find for the different cuts.
-Consider now the quantity $\sum_i\expval{\hat{X}}_{i,s}$.
-So far everything was $s$ independent as our system was periodic, but by considering the operator $\hat{X}$ we introduce an unphysical boundary so that it will depend on where we put the first and last sites.
-If the states do not cross this unphysical boundary then $\sum_i\expval{\hat{X}}_{i,s} = \sum_i\expval{\hat{X}_{\rm PBC}}_{i,s} $ and it will be a constant.
-However, when a Wannier state crosses the boundary we obtain the same constant plus $L$.
-If we put this boundary at $L/4$ the object
-\begin{equation}
-\frac{1}{L}\left(\sum_i\expval{\hat{X}}_{i,s}-\sum_i\expval{\hat{X}}_{i,0}\right)
-\end{equation}
-will be 1 (0) when a states has crossed the boundary or 0 (-1) when it has not, depending on the number of states in A at $s=0$, which we take as reference.
-Depending which state we take as reference will introduce a global integer shift for all cuts, so it is not relevant as it will go away after we average and take the modulo 1.
-By adding this term to our cut-average we will compensate for when our bulk state in the middle of A crosses $L/4$ and its charge changes.
-Leading to a continuous $\chi_s$ which can be averaged.
-In Fig.\ref{wan} we show how this works for one case.
-In Fig.\ref{wan}.a we show the expectation value of the position operator for both the entanglement eigenstates (in black) and the Wannier states (in red) for different cuts.
-In Fig.\ref{wan}.b and c we show $\chi_s$, which displays jumps due to a bulk state crossing $L/4$, and
-\begin{equation}
-\chi_s' = \chi_s + \frac{1}{L}\left(\sum_i\expval{\hat{X}}_{i,s}-\sum_i\expval{\hat{X}}_{i,0}\right),
-\end{equation}
-where the position is defined in $[-3L/4,L/4]$ to correctly compensate for these jumps.
-This method conceptually works perfectly whenever the bulk states of the ES match the Wannier states, i.e.
-in the thermodynamic limit, but it is not clear what the small parameter is.
-It depends on the $\lambda$ used (it should be small enough to not affect ES but large enough to localize the ES eigenstates) but also on the model.
-Models with longe-range hopping, for example, lead to a higher number of edge modes in the ES which makes the effect of the edge go deeper into the bulk.
-\carlos{We should think again if there is a better method to get this}
-
-\begin{figure}[h!]
+\begin{figure}[t]
 \centering
@@
 
-\caption{a)Profile used for the inhomogeneous part $\kappa_s$. b,c,d) ES (black), $\chi_s$ mod $1$ (green), $\gamma$ (red), $\expval{\chi_s}_{\rm cuts}$ mod $1$ (dashed blue).}
+\caption{}
 \label{wan}
 \end{figure}
+
+In a general case where the discontinuities are too strong and the discussed above does not apply one has to take a closer look into the bulk states to ensure a consistent charge assignment.
+As we mentioned above the bulk states farthest away from the edges are the problematic ones.
+They are also the ones best approximated by the Wannier states obtained from the chain with periodic boundary conditions, as the eigenstates of
+\begin{equation}
+C\hat{X}_{\rm PBC}C
+\end{equation}
+where $\hat{X}_{\rm PBC} = \exp\{i 2\pi \hat{X}/L\}$.
+Similar to when we go through different cuts labeled by $s$, we now consider the Wannier states dependent on where we define our lattice site 1, shifted by $s$.
+The obtained Wannier states for $s$ and $s+1$ will be therefore exactly the same, but shifted one lattice site.
+By considering their average position $\expval{\hat{X}_{PBC}}_{i,s}$ we can define a set of $2L$ Wannier states that depend continuously on $s$, provided that the Hamiltonian is continuous.
+These Wannier states are then a good approximation for the Bulk states of $C_A$ we find for the different cuts.
+Consider now the quantity $\sum_i\expval{\hat{X}}_{i,s}$.
+So far everything was $s$ independent as our system was periodic, but by considering the operator $\hat{X}$ we introduce an unphysical boundary so that it will depend on where we put the first and last sites.
+If the states do not cross this unphysical boundary then $\sum_i\expval{\hat{X}}_{i,s} = \sum_i\expval{\hat{X}_{\rm PBC}}_{i,s} $ and it will be a constant with $s$.
+However, when a Wannier state crosses the boundary we obtain the same constant plus $L$.
+If we put this boundary at $L/4$ the quantity
+\begin{equation}
+\frac{1}{L}\left(\sum_i\expval{\hat{X}}_{i,s}-\sum_i\expval{\hat{X}}_{i,0}\right)
+\end{equation}
+will be 1 (0) when a states has crossed the boundary or 0 (-1) when it has not, depending on the number of states in A at $s=0$, which we take as reference.
+Depending which state we take as reference will introduce a global integer shift for all cuts, so it is not relevant as it will go away after we average and take the modulo 1.
+By adding this term to our cut-average we will compensate for when our bulk state in the middle of A crosses $L/4$ and its charge changes, leading to a continuous $\chi_s$ which can be averaged.
+In Fig.\ref{wan} we show how this works for one case.
+In Fig.\ref{wan}.a we show the expectation value of the position operator for both the entanglement eigenstates (in black) and the Wannier states (in red) for different cuts.
+In Fig.\ref{wan}.b and c we show $\chi_s$, which displays jumps due to a bulk state crossing $L/4$, and
+\begin{equation}
+\chi_s' = \chi_s + \frac{1}{L}\left(\sum_i\expval{\hat{X}}_{i,s}-\sum_i\expval{\hat{X}}_{i,0}\right),
+\end{equation}
+where the position is defined in $[-3L/4,L/4]$ to correctly compensate for these jumps.
+This method conceptually works perfectly whenever the bulk states of the ES match the Wannier states, i.e.
+in the thermodynamic limit, but it is not clear what the small parameter is.
+It depends on the $\lambda$ used (it should be small enough to not affect ES but large enough to localize the ES eigenstates) but also on the system.
+Systems with many ES edge modes close to the bulk tend to have a higher finite size effect.
+
+This method is however not general.
+We find that whenever there is an interphase in our system, and therefore a zero mode in the energy spectrum, the Wannier states are in general no longer a good approximation of the bulk of finite systems.
+However, in the cases where we have an interphase but $\chi_s$ is continuous and its width is lower than one, we observe that Eq.\eqref{batcutavg} still gives the correct Zak phase, suggesting that this relation between the ES and the Zak phase exists in general for inhomogeneous systems even though we do not know how to compute it in some cases.
 
@@
 The proof in Ref.\cite{Zaletel2014} also breaks down for superconducting systems.
-However, as we did for systems without translational invariance, one can extend expression \ref{mainl} for systems with pairing terms as well.
-Since one computes the Zak phase from the single-particle Bogoliubov-de Gennes Hamiltonian, we can obtain the correspondent entanglement spectrum and use expression \ref{mainl} to obtain the Zak phase.
+However, as we did for systems without translational invariance, one can extend this identity for systems with pairing terms as well.
+Since one computes the Zak phase from the single-particle Bogoliubov-de Gennes Hamiltonian, we can obtain the correspondent ES and everything discussed above should apply as well.
 Take for example the kitaev chain, with the following Hamiltonian
@@
 Therefore the results already shown for the BDI model studied here are also applicable to the Kitaev chain.
-\carlos{check again becase I though that the kitaev chain had left-left right-right pairs, so in this case the entnsglement energies are not equispaced}
+
+\section{Notes}
+
+So far we show examples of two-band models, where the Zak phase is obtained from the lower occupied band.
+If we consider systems with a higher number of bands each band carries a Zak phase.
+We find that if we substitute the projector onto the occupied states, $C$, by a projector to any band $\alpha$ (or set of states), $P_\alpha$, the relations discussed above still hold.
+
+There is one thing one has to be careful about.
+The Zak phase seems not well-defined for two-band models at gapless points, or at least for finite systems.
+This seems to generalize to wether or not there is a gap between the set of states chosen to calculate the Zak phase over and the rest.
+
+We encounter another issue.
+When an inhomogeneous system presents interphases between two different topological phases, and therefore zero-modes, the localization of the entanglement eigenstates does not work as discussed above anymore.
+The Wannier states are no longer always good approximations in the bulk and we do not have a way of defining a consistent clasification of the bulk states.
+However, for the cases where $\chi_s$'s width is lower than 1 this classification is not needed and the can still compute the Zak phase.
 

### 2020-03-12 draft.tex
Add files via upload

@@
 \usepackage{caption}
+\usepackage{amsmath}
 
@@
 
-\emph{Outline of the paper}
+\emph{Outline of the paper}: In section II we introduce the model used throughout the paper and we discuss its phase diagram.
+In section III we introduce the Zak phase and the entanglement spectrum and set up the notation that will be used.
+In section IV and V we show our results for homogeneous and inhomogeneous systems, respectively.
+In section VI we will discuss how our results also generalize to the case of models with pairing terms and finally in section VII we present the conclusions and possible outlook.
 \section{Model.}
@@
 }
-\caption{Phase diagram for the system in the $BDI$ class showing the winding number of each region and surface energy spectrum for two different cuts in the phase diagram for different values of $\kappa,\kappa'$.}
+\caption{Phase diagram for the system in the $BDI$ class showing the winding number of each region and surface energy spectrum for a cut going through the three differet regions, for different values of $\kappa,\kappa'$.}
 \label{bdi_phase_diagram}
@@
 \subsubsection{Nearest-neighbor hopping}
-\mariac{find better wording?}
 
@@
 }
-\caption{Phase diagram for the system in the $BDI$ class showing the winding number of each region and surface energy spectrum for two different cuts in the phase diagram for different values of $\kappa,\kappa'$.}
+\caption{Entanglement spectrum (black), $\chi$ (green) and $\gamma$ (red) for a cut in the phase diagram through the $\nu = 1 \rightarrow \nu = 0$ transition (dashed line in the phase diagram) for $t'=0$ and different values of $\kappa$.}
 \label{huang}
@@
 In the case the ES is double degenerate at most, left and right edge states might mix but one will always classify one state of the pair as left and the other one as right, this is however not the case when the degeneracies are higher than two-fold.
-In order to ensure that we obtain edge modes localized at one edge, instead of $C_A$ we diagonalize $C_A + \lambda C_A\hat{X}C_A$, where $\hat{X}$ is the position operator and $\lambda$ is a small parameter that vanishes in the thermodynamic limit.
+In order to ensure that we obtain edge modes localized at one edge, instead of $C_A$ we diagonalize $C_A + \lambda C_A\hat{X}C_A$ (\carlos{Explain why it does not work in general}), where $\hat{X}$ is the position operator and $\lambda$ is a small parameter that vanishes in the thermodynamic limit.
 
@@
 }
-\caption{Entanglement spectrum (black), $\chi$ (green) and $\gamma$ (red)  }
+\caption{Entanglement spectrum (black), $\chi$ (green) and $\gamma$ (red) for two different cuts,  continuous line for a,b, and dotted line for c,d plots, Showing a more diverse ES where more than one midgap state is present.
+}
 \label{2}
@@
 
+Using the fact that $N_A = \Tr C_A$ we can rewrite $\chi$ as
+\begin{equation}
+\chi = \left(\sum_{i\ in r} \xi_i+ \sum_{i\in b}\xi_i\right) \, {\rm mod} \, 1.
+\end{equation}
+A simpler way one can compute this quantity is by obtaining $C_A$ for an open chain, labeled as $C_A^{ o}$.
+For a long enough chain the right eigenvalues will remain the same, as we still place a virtual cut in the middle of the chain but the left eigenvalues will no longer be present, so that one can simply compute
+\begin{equation}
+\chi = \Tr C_A^{o} \, {\rm mod} \, 1.
+\end{equation}
+This method however will introduce a stronger finite size effect and as we will discuss below it only works for homogeneous systems.
+\carlos{I don't know if its worth it to introduce in this paper.
+Then I dont know what to do for the ES = winding.
+}
+
 \section{Inhomogeneous systems}
-
-\subsubsection{Simple case, continuous ES.}
 
@@
 
-We are interested therefore in finding a cut-independent quantity from the ES that might still result in the Zak phase. The first thing one can do is a cut-average of $\chi$. Naively one would define
+We are interested therefore in finding a cut-independent quantity from the ES that might still result in the Zak phase.
+The first thing one can do is to do an average over the different cuts.
+Naively one would cut-average the expression in Eq.~\eqref{loceq}, however the average of a circular quantity is not well defined.
+We defined $\chi$ modulo 1 because the Zak phase is also defined modulo 1, but it also served the purpose of getting rid of the sign choice for the bulk modes, which is problematic.
+As we do for the edge state, we can choose a sign for the bulk modes in the following way
 \begin{equation}
-\bar{\chi} = \frac{1}{L}\sum_{s=1}^L (\chi_s \, {\rm mod} \, 1 )
+g_i=
+\begin{cases}
+ +1 &,\rm{if} \expval{\hat{X}}_i > L/4  \\
+ -1 &,\rm{if} \expval{\hat{X}}_i < L/4.
+\end{cases}
+\end{equation}
+So that now we define
+\begin{equation}
+\chi_s = \frac{1}{2} \left( \sum_{i\in r}\xi^s_i-\sum_{i\in l}\xi^s_i + \sum_{i\in b}g_i^s \xi^s_i  + N_A \right),
+\end{equation}
+obtained for cuts at $s$ and $s+L/2$ and the cut-average is given by
+\begin{equation}
+\bar{\chi} = \frac{1}{L}\sum_{s=1}^L \chi_s.
 \label{batcutavg}
 \end{equation}
-where
+By adding the perturbation proportional to $C_A\hat{X} C_A$ the bulk modes also localize.
+As we move through the different cuts the average position of the bulk modes moves slightly.
+For most modes this will not affect the assigned sign, but for the modes near the middle of subsystem $A$ it might be that for strong enough inhomogeneity that the mode crosses $L/4$ and the sign changes.
+We find that in order for us to get the correct result the contribution from the bulk modes should be constant across cuts, so we have to find a way of correcting for these jumps of $1$ in $\chi_s$.
+
+\subsubsection{Continuous inhomogeneity}
+
+In the case we have a continuous inhomogeneity one can obtain a new quantity $\chi'_s$ with a constant bulk contribution, by assuming that the discontinuities of $\chi_s$ above some threshold, $h$, are due to these unphisical changes of sign for the bulk states and eliminating them by shifting it by an integer number so that
 \begin{equation}
-\chi_s = \frac{1}{2} \left( \sum_{i\in r}\xi^s_i-\sum_{i\in l}\xi^s_i \pm \sum_{i\in b}\xi^s_i  + N_A \right) 
+\chi_{s+1}'-\chi'_s < h
 \end{equation}
-is obtained for cuts at $s$ and $s+L/2$.
-However the average of a circular quantity is not well defined.
-The reason we defined $\chi$ with modulo 1 in Eq.~\eqref{loceq} was to get rid of the choice of sign for the bulk, which should not matter for the end result anyway.
-Instead, we can calculate
-\begin{equation}
-\bar{\chi} = \frac{1}{L}\sum_{s=1}^L \chi_s \quad {\rm mod} \quad 1 
-\label{goodcutavg}
-\end{equation}
-but in order to obtain the correct result we need the contribution from the bulk to be constant across different cuts,$\xi_{i \in b}^s = \xi_{i \in b}$, so that in the end
-\begin{align}
-\bar{\chi} =& \frac{1}{2L}\sum_{s=1}^L \left( \sum_{i\in r}\xi^s_i-\sum_{i\in l}\xi^s_i \right) \quad {\rm mod} \quad 1 \\
-&+\frac{1}{2} \left(\sum_{i\in b}\xi^s_i + N_A \right) \quad {\rm mod} \quad 1 \nonumber
-\label{goodcutavg}
-\end{align}
-If instead we only take the modulo 1 after averaging we still might have an issue due to the charge assignment of the bulk modes being inconsistent.
-A bulk mode near the center of A, at $L/4$, might cross $L/4$ at different cuts, changing its assigned charge.
-The bulk contribution will not be constant across cuts and that will lead to errors when averaging.
-
-One can get away with expression Eq.~\eqref{batcutavg} in a few cases.
-For a system with a continuous position-dependent Hamiltonian $\chi_s$ should also be continuous.
-This tell us how we have to perform the cut-average.
-Even when discontinuities are present as long they are weak enough so that we still can tell when $\chi_s$ crosses "$0,1$" (\carlos{write properly}), the cut-average can be performed.
-In the case that the width of $\chi_s$ is lower than 1 one can select the range of values so that $\chi_s$ does not cross the "$0,1$" line and then average.
-We show this in Fig.\ref{inh} for a system with disorder on $\kappa_s$.
-Note that even though we have random disorder, the ES and $\chi_s$ show some sign of continuity so that the cut-averaged can be performed this way even for relatively strong disorder.
+and it is continuous.
+$\chi'_s$ can now be cut-averaged to obtain the Zak phase.
+In figure \ref{inh} we show this for a continuous profile for $\kappa_s$.
+We see that $\chi_s$ in fact presents jumps of $1$ but once we correct them in $\chi'_s$, its cut average results in the Zak phase.
+We also see the same for a disordered $\kappa_s$.
+Even though the disorder is relatively strong we see that $\chi_s'$, after the artificial jumps are eliminated, is fairly continuous, which allows us to use this method even for strong inhomogeneity, as long as we can identify this jumps.
 
 \begin{figure}[h!]
+
 \centering
 \makebox[0pt]{
-\subfloat[$\kappa_s$]{
+\subfloat[]{
+  \includegraphics[width=40mm]{4_a.pdf}
+  \label{inha}
+}
+\subfloat[ ]{
+  \includegraphics[width=40mm]{4_b.pdf}
+}
+}\hspace{0mm}
+
+\makebox[0pt]{
+\subfloat[]{
   \includegraphics[width=40mm]{4_c.pdf}
 }
-\subfloat[$t=1,t'=2,m =1$ ]{
+\subfloat[]{
   \includegraphics[width=40mm]{4_d.pdf}
@@
 \makebox[0pt]{
-\subfloat[$\kappa_s$]{
-  \includegraphics[width=40mm]{4_a.pdf}
+\subfloat[]{
+  \includegraphics[width=40mm]{4_e.pdf}
 }
-\subfloat[$t=1,t'=2,m =1$]{
-  \includegraphics[width=40mm]{4_b.pdf}
+\subfloat[]{
+  \includegraphics[width=40mm]{4_f.pdf}
 }
-}\hspace{0mm}
+}
 
-\makebox[0pt]{
-\subfloat[$\chi_s'$]{
-  \includegraphics[width=40mm]{xicfix.pdf}
-}}
-\label{inh}
+\caption{Using the parameters \label{inh}$t=1,t'=2,m =1$, we show our results for two different profiles for $\kappa_s$, one continuous (a) and one with disorder (b).
+In (c) and (d) we show the ES (black) and $ \chi_s \, {\rm mod} \, 1$ (green).
+In (e) and (f) we show $\chi_s$ (green), $\chi'_s$ (blue dots), $\gamma$ (red) and $\expval{\chi'_s}_{\rm cuts}$ (blue dashed) }
+
 \end{figure}
 
-In the case where $\chi_s$ is continuous but the width is larger than 1 can construct a new quantity $\chi_s'$ that is not defined modulo 1, by extending the range of $\chi_s \, {\rm mod} \, 1$ every time it crosses $0$ or $1$.
-In Fig.\ref{inh}.d we show an example of this.
+In some cases, this calculation simplifies a bit more.
+Note that, in the case where one can tell a priori that the width of $\chi'_s$ is lower than 1, as in the example with disorder, to recover $\chi_s'$ one can simply take the modulo 1 and adding an offset, so for instance in the example with disorder we can obtain
+\begin{equation}
+\chi'_s = ([\chi_s-1/2] \, {\rm mod} \, 1)+1/2.
+\end{equation}
+One can then cut-average the resulting $\chi_s \, \rm{mod} \,1$ in the same way to obtain the Zak phase.
+This cannot be done in the first example because as we already see in Fig.\ref{inh}.c, $\chi_s \, \rm{mod} \,1$ wraps aroud the full range, so that there is no offset that will allow us to average it in a well-defined way.
+
 In Fig.\ref{disorder} we show $1-\expval{\chi_s}_{\rm cuts}/\gamma$ for different system lengths for system with weak disorder, as well as the disorder average of $1-\chi_0/\gamma$.
-We observe that both averages lead to the same difference, which scales as $O(L^{-2})$ (\carlos{I have to redo de plot, maybe scaling is different}), vanishing in the thermodynamic limit.
-The fact that both methods lead to the same error suggests that it comes strictly from a systematic finite size error in the computation of either the Zak phase or the ES which is system-independent.
+For both cases the error seems to vanish in the thermodynamic limit.
+For small systems both errors are very similar, which suggests that it comes strictly from a systematic finite size error in the computation of either the Zak phase or the ES which is system-independent.
+For larger systems the disoder average is more dispersed, probably due to the number of configurations considered not being high enough.
 
@@
 \makebox[0pt]{
-\subfloat[$t = 0, t' = 2, \kappa = 0.3$]{
-  \includegraphics[width=65mm]{scaling_cut_disavg.png}
+\subfloat[$t = 1, t' = 2, m=1, \kappa_s \in (-1,1)$]{
+  \includegraphics[width=65mm]{disvscut.pdf}
 }
 }
-\caption{\carlos{Redo}}
+\caption{We show $1-\expval{\chi_s'}_{\rm cuts}/\gamma$ (red) and $1-\expval{\chi_s'}_{\rm dis}/\expval{\gamma}_{\rm dis}$.
+The cut average seems to do as good or better than the disorder average.
+For small systems it seems that both errors are similar, pointing to a systematic finite size error in the computation of either the Zak phase or the ES.
+For larger systems the error in the disorder average is more spread, which suggests that one should average more configurations.}
 \label{disorder}
@@
 \makebox[0pt]{
-\subfloat[$\expval{\hat{X}}_{i,s}$ for different cuts, for the entanglement eigenstates (black) and the Wannier states (red)]{
+\subfloat[$\expval{\hat{X}}_{i,s}$ for different cuts, for the entanglement eigenstates such that $\xi_i < 1/2$ (black) and the Wannier states (red)]{
   \includegraphics[width=90mm]{loca.pdf}
@@
 
-\caption{}
+\caption{For the same parameters as used in Fig.5.a,c,e we show the expectation value of the position for the ES eigenstates (black) and the Wannier states (red), showing that the latter are a good approximation of the former in the bulk.
+We highlight by a continuous red line the bulk mode that crosses $L/4$ which causes the jumps in $\chi_s$.
+In (b) we show again $\chi_s$ (black) and the correction for the sign switch (red).
+In (c) $\chi_s'$ is reconstructed via this method, showing the same result as we had before.}
 \label{wan}
@@
 
-In a general case where the discontinuities are too strong and the discussed above does not apply one has to take a closer look into the bulk states to ensure a consistent charge assignment.
+In a general case where the discontinuities are too strong and we do not know how to correct for these jumps as explained before one has to take a closer look into the bulk states to ensure a consistent sign assignment.
 As we mentioned above the bulk states farthest away from the edges are the problematic ones.
@@
 Consider now the quantity $\sum_i\expval{\hat{X}}_{i,s}$.
-So far everything was $s$ independent as our system was periodic, but by considering the operator $\hat{X}$ we introduce an unphysical boundary so that it will depend on where we put the first and last sites.
+So far everything was $s$ independent as our system was periodic, but by considering the operator $\hat{X}$ we introduce an unphysical boundary so that it will depend on how the lattice sites are defined.
 If the states do not cross this unphysical boundary then $\sum_i\expval{\hat{X}}_{i,s} = \sum_i\expval{\hat{X}_{\rm PBC}}_{i,s} $ and it will be a constant with $s$.
@@
 \begin{equation}
-\frac{1}{L}\left(\sum_i\expval{\hat{X}}_{i,s}-\sum_i\expval{\hat{X}}_{i,0}\right)
+\delta_s = \frac{1}{L}\left(\sum_i\expval{\hat{X}}_{i,s}-\sum_i\expval{\hat{X}}_{i,0}\right)
 \end{equation}
@@
 Depending which state we take as reference will introduce a global integer shift for all cuts, so it is not relevant as it will go away after we average and take the modulo 1.
-By adding this term to our cut-average we will compensate for when our bulk state in the middle of A crosses $L/4$ and its charge changes, leading to a continuous $\chi_s$ which can be averaged.
-In Fig.\ref{wan} we show how this works for one case.
-In Fig.\ref{wan}.a we show the expectation value of the position operator for both the entanglement eigenstates (in black) and the Wannier states (in red) for different cuts.
-In Fig.\ref{wan}.b and c we show $\chi_s$, which displays jumps due to a bulk state crossing $L/4$, and
+By adding this term to our cut-average we will compensate for when our bulk state in the middle of A crosses $L/4$ and its charge changes, leading to a continuous $\chi_s'$ which can be averaged,
 \begin{equation}
-\chi_s' = \chi_s + \frac{1}{L}\left(\sum_i\expval{\hat{X}}_{i,s}-\sum_i\expval{\hat{X}}_{i,0}\right),
+\chi_s' = \chi_s + \delta_s
 \end{equation}
-where the position is defined in $[-3L/4,L/4]$ to correctly compensate for these jumps.
-This method conceptually works perfectly whenever the bulk states of the ES match the Wannier states, i.e.
+
+In Fig.\ref{wan} we show how this works for the same continuous example as in \ref{inh}.
+In Fig.\ref{wan}.a we show the expectation value of the position operator for both the entanglement eigenstates (in black) and the Wannier states (in red) for different values of $s$.
+In Fig.\ref{wan}.b and c we show $\chi_s$, which displays jumps due to a bulk state crossing $L/4$, and $\delta_s$, showing that it properly corrects the sign switch.
+
+This method conceptually works perfectly whenever the bulk states of the ES are equal to the Wannier states, i.e.
 in the thermodynamic limit, but it is not clear what the small parameter is.
@@
 
-This method is however not general.
-We find that whenever there is an interphase in our system, and therefore a zero mode in the energy spectrum, the Wannier states are in general no longer a good approximation of the bulk of finite systems.
-However, in the cases where we have an interphase but $\chi_s$ is continuous and its width is lower than one, we observe that Eq.\eqref{batcutavg} still gives the correct Zak phase, suggesting that this relation between the ES and the Zak phase exists in general for inhomogeneous systems even though we do not know how to compute it in some cases.
+We also find that whenever there is an interphase in our system, and therefore a zero mode in the energy spectrum, the Wannier states are not necesseraly a good approximation of the bulk of finite systems.
+However, in the cases where we have an interphase but $\chi_s$ is fairly continuous, so that we can correct these jumps, cut-averaging results again in the Zak phase.
+
+As the sign assigned to the bulk modes becomes relevant in the inhomogeneous case one might think that it is better to use the open chain, $\chi_s = \Tr C_A^{o,s}$, as one avoids the issue of having to assign any signs at all.
+Although this is true we find another problem.
+By adiabatically tuning down the coupling between the $L$ and $1$ sites we see that the eigenvalues of the $left$ modes flow to either $0$ or $1$.
+In the case where this eigenvalues oscillate around $1/2$ for different cuts this leads to discontinuities in the number of bulk modes at $\xi=1$ which again result in an incorrect cut-average and we found no practical way of compensating for these discontinuitiesyfgh.
 
@@
 
-\section{Notes}
+\section{Conclusion}
 
-So far we show examples of two-band models, where the Zak phase is obtained from the lower occupied band.
-If we consider systems with a higher number of bands each band carries a Zak phase.
-We find that if we substitute the projector onto the occupied states, $C$, by a projector to any band $\alpha$ (or set of states), $P_\alpha$, the relations discussed above still hold.
-
-There is one thing one has to be careful about.
-The Zak phase seems not well-defined for two-band models at gapless points, or at least for finite systems.
-This seems to generalize to wether or not there is a gap between the set of states chosen to calculate the Zak phase over and the rest.
-
-We encounter another issue.
-When an inhomogeneous system presents interphases between two different topological phases, and therefore zero-modes, the localization of the entanglement eigenstates does not work as discussed above anymore.
-The Wannier states are no longer always good approximations in the bulk and we do not have a way of defining a consistent clasification of the bulk states.
-However, for the cases where $\chi_s$'s width is lower than 1 this classification is not needed and the can still compute the Zak phase.
+It was previously known that the Zak phase was encoded in the many-body entanglement spectrum.
+Here we have shown how it is encoded in the single-body entanglement spectrum.
+This relation is simpler than the one involving the MBES and it provides more insight into the ES.
+We have furthermore shown how this relation generalizes to inhomogeneous and superconducting systems, where it was not previously shown to be present.
+Our result also provides a new way of computing the Zak phase for inhomogeneous systems where instead of diagonalizing a $2L\times 2L$ matrix one has to diagonalize $L$ matrices of size $L\times L$, which in some cases might be computationally favorable.
+(\carlos{can we get insight on this?}).
 

### 2020-04-16 draft.tex
changes in intro and section I

@@
 \section{Introduction}
-Topological phases of matter have attracted more and more attention during the last decades, not the least because many more relevant systems have been realized experimentally. 
+Topological phases of matter have attracted a lot of attention during the last decades, not the least because a large variety of relevant systems have been realized experimentally. 
 The early focus was mainly on \emph{topologically ordered} systems \cite{wenbook}, where interaction effects are crucial for stabilizing the phase. 
@@
 
-There are a few tools to identify topological phases. 
-A very efficient one is the entanglement entropy~\cite{Kitaev2006topological, Levin2006detecting}, which allows one to identify the total quantum dimension of the underlying topological field theory. 
+There are few tools to identify topological phases. 
+A very efficient one is the entanglement entropy~\cite{Kitaev2006topological, Levin2006detecting}, which allows one to identify the total quantum dimension of the underlying topological quantum field theory.
 As such, the entanglement entropy can only be used for interacting systems. 
-Its main drawback is the fact that it can only provide one of the quantum numbers, and is not able to uniquely determine the topologically ordered phase at hand. 
-In addition, it requires a scaling analysis, which can be challenging for strongly correlated systems. 
+Its main drawback is the fact that it can only provide one of the pertinent quantum numbers that characterizes the system, and is, thus,  not able to uniquely determine the topologically ordered phase at hand.
+In addition, it requires a careful scaling analysis, which can be challenging for strongly correlated systems. 
 Another, closely related tool, is the entanglement spectrum (ES), originally introduced for fractional quantum Hall systems~\cite{Li2008entanglement}. 
-It was conjectured to provide information about the edge spectrum, which was later shown to be a general feature of ground states with an effective topological field theory description~\cite{Qi2012general}.
+It was conjectured to provide information about the edge spectrum, which was later shown to be a general feature of ground states with an effective topological quantum field theory description~\cite{Qi2012general}.
 It also proved to be an effective tool for fractional Chern insulators~\cite{Regnault2011fractional} and certain quantum spin liquids~\cite{yao2010entanglement}. \mariac{Others?}
@@
 For the simplest quantum Hall states, the Laughlin states,  it was shown that the ES does, in fact, contain the full information about the topological phase~\cite{hermanns2011haldane}.
-However, it is far from clear if this holds for more complicated/nonabelian states as well.
+However, it is far from clear if this holds for more complicated, in particular nonabelian, states as well.
 
@@
 It was subsequently shown for topological insulators/superconductors that the ES for (gapped) periodic systems is equivalent to the flat-band energy spectrum of the corresponding system with open boundaries~\cite{Fidkowski2010entanglement}.
-The same correspondence was also found for closely related gapless systems~\cite{matern2018entanglement}
+The same correspondence was also found for closely related gapless systems~\cite{Matern2018entanglement}
 However, even for non-interacting systems, it is unclear which information (beyond the  `edge' spectrum) is encoded in the spectrum.
 
-In this manuscript, we show that the information about the Berry phase can be completely recovered from the entanglement spectrum and provide a simple formula to do so. 
+In this manuscript, we show that the information about the Zak phase can be completely recovered from the entanglement spectrum and provide a simple formula to do so. 
 We relate our results to earlier ones XXX
@@
 In section VI we will discuss how our results also generalize to the case of models with pairing terms and finally in section VII we present the conclusions and possible outlook.
+
+% - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % 
 \section{Model.}
-
-To showcase our results we consider as a starting point the model in the BDI class used in reference \cite{Song2014} because it is a very simple model which supports topological phases with winding numbers up to $\nu =2$.
-Additionaly, we include two symmetry breaking terms, $\kappa$ and $\kappa'$, to obtain an even richer phase diagram and show that our results are not limited to any topological class.
-The Hamiltonian is then given by
+\mariac{ref needed to 10-fold way}
+To showcase our results we consider as a starting point the model in the BDI class used in reference~\cite{Song2014}. 
+This is a very simple model which nevertheless supports a rich phase diagram containing topological phases with winding numbers up to $\nu =2$. 
+In order to show that our results are not limited to any symmetry class, we also include two symmetry breaking terms. 
+The Hamiltonian is  given by
 \begin{align}
@@
 \end{align*}
-For $\kappa = \kappa' = 0$ this Hamiltonian has the following symmetries
+For $\kappa = \kappa' = 0$ this Hamiltonian is in the BDI class, i.e. it has time-reversal ($T$), particle-hole ($C$), and chiral symmetry ($S$):
 \begin{alignat*}{2}
@@
 &C = \sigma_z\mathcal{K} ; \quad &&C H(-k) C^{-1} = -H(k) \\
-&S = \sigma_z ; \quad &&S H(k)S^{-1} = -H(k) 
+&S = \sigma_z ; \quad &&S H(k)S^{-1} = -H(k) .
 \end{alignat*}
-and therefore it is in the BDI class, where the system has the phase diagram shown in Fig.\ref{bdi_phase_diagram} for $t=1$.
-We also a cut through the three phases (dotted line in the phase diagram) for different values of $\kappa,\kappa'$.
-For $\kappa = \kappa'= 0$ when the system is in the BDI class we the system has a number of zero modes equal to the winding number, as it is well known.
-For $\kappa' \neq 0$ only $C$ is preserved and the system belongs to the $D$ class As the zero modes in the $\nu=2$ phase split from zero pairwise, the $\nu = 2$ and $\nu=0$ phases become equivalent.
-For $\kappa \neq 0$ only $T$ is preserved and the system is in the trivial AI class, where all zero-energy modes split.
-With both $\kappa \neq 0$ and $\kappa' \neq0$ the system has no symmetry, belonging to the $A$ class and the edge modes that were doubly degenerate in the $D$ class split as well.
-\begin{figure}[h!]
+The corresponding phase diagram is shown in Fig.\ref{fig:bdi_phase_diagram} for $t=1$. 
+\begin{figure}[t]
 	\centering
@@
 	}}\hspace{0mm}
+\caption{\mariac{need more visible lines .}Phase diagram for the system in the $BDI$ class showing the winding number of each region.}
+\label{fig:bdi_phase_diagram}
+\end{figure}
+When imposing open boundary conditions on the system harbors, the number of symmetry-protected zero energy edge modes is equal to the winding number. 
+%We also a cut through the three phases (dotted line in the phase diagram) for different values of $\kappa,\kappa'$.
+
+For $\kappa' \neq 0$ only particle-hole symmetry $C$ is preserved and the system belongs to symmetry class $D$. 
+In one spatial dimension, the latter is characterized \mariac{better expression?} by a $\mathbb{Z}_2$ instead of the $\mathbb{Z}$ invariant of the BDI class. 
+Consequently, adding such a term causes the zero modes in the $\nu=2$ phase to split from zero pairwise and the $\nu = 2$ and $\nu=0$ phases become equivalent. 
+For $\kappa \neq 0$ only time-reversal $T$ is preserved and the system is in the AI class. 
+In one dimensions, this class is trivial and, thus,  all zero-energy modes split. 
+With both $\kappa \neq 0$ and $\kappa' \neq0$ the system has no symmetry, thus belonging to the $A$ class.  Also this symmetry class is topologically trivial in one dimension.
+
+The behavior of the zero-energy edge modes in the various symmetry classes is  shown in Fig.~\ref{fig:zero_E_modes}, using the dotted path ($t=1,t'=-2$) marked in the phase diagram~\ref{fig:bdi_phase_diagram}.
+Subfigure (a) shows the edge spectrum in the BDI class, with 4 (resp. 2) symmetry-protected zero modes for winding number 2 (1). 
+In subfigure (b) and (d), where the system is in a topologically trivial phase, all edge modes are gapped. 
+Subfigure (c) shows the edge spectrum of the system being in class D, where the $\nu=2 $ phase becomes trivial. 
+Consequently, symmetry-protected zero modes only exist for $1<m<2.5$\mariac{?} (corresponding to $\nu=1$ for $\kappa'=0$) , whereas the modes are split for $m<1$ (corresponding to $\nu=2$ for $\kappa'=0$).
+
+\begin{figure}[h!]
+\centering
+%\makebox[0pt]{
+%\subfloat[$t = 1,\kappa =\kappa'=0$]{
+%  \includegraphics[width=45mm]{phase_diagram2.pdf}
+%}}\hspace{0mm}
 
 \makebox[0pt]{
-\subfloat[$t = 1,t'=-2$, \, $\kappa =\kappa'=0$]{
+\subfloat[$\kappa =\kappa'=0$]{
   \includegraphics[width=35mm]{1_a.pdf}
 }
-\subfloat[$t=1,t'=-2$, $\kappa = 0.3, \kappa'=0$]{\includegraphics[width=35mm]{1_b.pdf}
+\subfloat[$\kappa = 0.3, \kappa'=0$]{\includegraphics[width=35mm]{1_b.pdf}
 }
@@
 \makebox[0pt]{
-\subfloat[$t = 1,t'=-2$, $\kappa = 0,\kappa'=0.3$]{
+\subfloat[$\kappa = 0,\kappa'=0.3$]{
   \includegraphics[width=35mm]{1_c.pdf}
 }
-\subfloat[$t=1,t'=-2$, $\kappa = 0.3,\kappa'=0.3$]{\includegraphics[width=35mm]{1_d.pdf}
+\subfloat[$\kappa = 0.3,\kappa'=0.3$]{\includegraphics[width=35mm]{1_d.pdf}
 }
 }
-\caption{Phase diagram for the system in the $BDI$ class showing the winding number of each region and surface energy spectrum for a cut going through the three differet regions, for different values of $\kappa,\kappa'$.}
-\label{bdi_phase_diagram}
+\caption{Surface energy spectrum along the dotted path (i.e. $t=1,t'=-2$) in the phase diagram of Fig.~\ref{fig:bdi_phase_diagram} for different values of $\kappa,\kappa'$.}
+\label{fig:zero_E_modes}
 \end{figure}
 
+% - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % 
 \section{Zak phase and Entanglement spectrum}
@@
 \mariac{introduce entanglement energy, relation between C and energies}
+
+% - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % 
 \section{Homogeneous system}
@@
 }
+
+% - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - %
 
@@
 
+% - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - %
+
 \section{Superconducting systems}
@@
 \appendix
+
+% - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - %
 

### 2020-04-20 draft.tex
changed section III

@@
 H =& \sum_{i\alpha,j\beta} c_{i\alpha}^\dagger H_{ij,\alpha \beta} c_{j\beta} \\
-H_{ij} =& (m \sigma_x + \kappa \sigma_z)\delta_{ij}  + \frac{1}{2i}\kappa'\sigma_z (\delta_{i-j,1}-\delta_{i-j,-1})\\
-&+ \frac{1}{2} t \left[(\sigma_x + i \sigma_y)\delta_{i,j+1} + (\sigma_x - i \sigma_y) \delta_{i,j-1} \right] \\
+H_{ij} =& (m \sigma_x + \kappa \sigma_z)\delta_{ij}  + \frac{1}{2i}\kappa'\sigma_z (\delta_{i-j,1}-\delta_{i-j,-1})\nonumber\\
+&+ \frac{1}{2} t \left[(\sigma_x + i \sigma_y)\delta_{i,j+1} + (\sigma_x - i \sigma_y) \delta_{i,j-1} \right] \nonumber\\
 &+  \frac{1}{2} t' \left[(\sigma_x + i \sigma_y)\delta_{i,j+2} + (\sigma_x - i \sigma_y) \delta_{i,j-2} \right],
@@
 with the corresponding Bloch Hamiltonian being
-\begin{align*}
-H(k)=&\mqty( \kappa + \kappa' \sin(k) & t' e^{i2k} + t e^{ik}+m \\t' e^{-i2k} + t e^{-ik}+m & -\kappa-\kappa' \sin(k)  ) \\
+\begin{align}
+H(k)=&\mqty( \kappa + \kappa' \sin(k) & t' e^{i2k} + t e^{ik}+m \nonumber\\t' e^{-i2k} + t e^{-ik}+m & -\kappa-\kappa' \sin(k)  )\nonumber \\
 =& (t' \cos(2k)+t\cos(k)+m)\sigma_x \\
-&+ (-t' \sin(2k)-t\sin(k))\sigma_y + (\kappa+\kappa'\sin(k)) \sigma_z.
-\end{align*}
+&+ (-t' \sin(2k)-t\sin(k))\sigma_y + (\kappa+\kappa'\sin(k)) \sigma_z.\nonumber
+\end{align}
 For $\kappa = \kappa' = 0$ this Hamiltonian is in the BDI class, i.e. it has time-reversal ($T$), particle-hole ($C$), and chiral symmetry ($S$):
-\begin{alignat*}{2}
-&T = \mathcal{K} ; \quad &&T H(-k) T^{-1} = H(k) \\
-&C = \sigma_z\mathcal{K} ; \quad &&C H(-k) C^{-1} = -H(k) \\
+\begin{alignat}{2}
+&T = \mathcal{K} ; \quad &&T H(-k) T^{-1} = H(k) \nonumber\\
+&C = \sigma_z\mathcal{K} ; \quad &&C H(-k) C^{-1} = -H(k) \nonumber\\
 &S = \sigma_z ; \quad &&S H(k)S^{-1} = -H(k) .
-\end{alignat*}
+\end{alignat}
 The corresponding phase diagram is shown in Fig.\ref{fig:bdi_phase_diagram} for $t=1$. 
@@
 % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % 
-\section{Zak phase and Entanglement spectrum}
-Consider a quadratic Hamiltonian in one dimension
+\section{Zak phase and entanglement spectrum}
+In this section, we review some relevant information about firstly the Zak phase and secondly the entanglement spectrum in non-interacting systems. 
+We consider a generic, quadratic Hamiltonian in one dimension
 \begin{equation}
@@
 We can obtain the single-particle eigenstates as
-\begin{align*}
+\begin{align}
 \sum_{j\beta}H_{ij,\alpha\beta} u^{p\mu}_{j\beta} = E_{p\mu} u_{i\alpha}^{p\mu},
-\end{align*}
-where $[U]_{j\beta,p\mu} = u^{p\mu}_{j\beta}$ is the unitary matrix that diagonalizes $H$.
-The polarization can be obtained as the many-body average of the position operator, but it has to be regularized for periodic boundary conditions \cite{Resta1997}.
+\end{align}
+where $[U]_{j\beta,p\mu} = u^{p\mu}_{j\beta}$ is the unitary matrix that diagonalizes $H$, i.e. 
+\begin{align}\label{eq:D}
+H&=UDU^\dagger& \mbox{with } D=\mbox{diag}(E_{p\mu}).
+\end{align} 
+The polarization can be obtained as the many-body average of the position operator, but the latter has to be regularized for periodic boundary conditions \cite{Resta1997}. 
 We can define a position operator for PBC as
@@
 \end{equation}
-where $\hat{X}$ is the sum of all the single-particle position operators. We can consider then the many-body average of the position operator over the ground state, defined as
-\begin{align*}
-\expval{\hat{X}_{\rm PBC}}_0 \equiv -\frac{L}{2\pi} \rm{ Im\, Ln }\bra{\Psi_0}\hat{X}_{\rm PBC}\ket{\Psi_0},
-\end{align*}
-\carlos{This is the notation used in the Resta paper, but I'm not sure if it's confusing.}
+where $\hat{X}$ is the sum of all the single-particle position operators.
 
-where $\ket{\Psi_0}$ is the many-body ground state. Using that the ground state is a Slater determinant of the single-particle eigenstates, $\ket{u^{p\mu}}$, it can be rewritten as
-\begin{align*}
-\expval{\hat{X}_{\rm PBC}}_0 = -\frac{L}{2\pi} \rm{Im \, Ln}\, {\rm det} \, \tilde{S},
-\end{align*}
-where the matrix $\tilde{S}$ is obtained from another matrix S with elements
+We can now consider then the many-body average of the position operator over the ground state, defined as
+\begin{align}
+\expval{\hat{X}_{\rm PBC}}_0 \equiv -\frac{L}{2\pi} \rm{ Im }\ln\bra{\Psi_0}\hat{X}_{\rm PBC}\ket{\Psi_0},
+\end{align}
+%\carlos{This is the notation used in the Resta paper, but I'm not sure if it's confusing.}
+where $\ket{\Psi_0}$ is the many-body ground state. 
+Using that the ground state is a Slater determinant of the occupied single-particle eigenstates, $\ket{u^{p\mu}}$, it can be rewritten as
+\begin{align}
+\expval{\hat{X}_{\rm PBC}}_0 = -\frac{L}{2\pi} \rm{Im }\ln \, {\rm det} \, S,
+\end{align}
+where the matrix $S$ is given by
 \begin{equation}
-S_{p\mu,q\nu} = \sum_{j\alpha}u^{p\mu\, \ast}_{j \alpha} e^{-i\frac{2\pi}{L}j}u^{q\nu}_{j \alpha}
+S_{p\mu,q\nu} = \sum_{j\alpha}\!^{'} u^{p\mu\, \ast}_{j \alpha} e^{-i\frac{2\pi}{L}j}u^{q\nu}_{j \alpha}
 \end{equation}
-by restricting it to the space of occupied single-particle states. The Zak phase can now be obtained as
+and the $\sum\!{'}$ indicates that the sum is restricted to the space of \emph{occupied} single-particle states. 
+The Zak phase can then be obtained as
 \begin{equation}
-\gamma = \expval{\hat{X}_{\rm PBC}}_0\frac{2\pi}{L} + \frac{1}{2}
+\gamma = \expval{\hat{X}_{\rm PBC}}_0\frac{2\pi}{L} + \frac{1}{2}.
 \end{equation}
 \carlos{There is an extra 1/2 factor compared to Restas's formula which I don't know where it is coming from.}
-The correlation function in position space is given by
-\begin{align*}
+
+Let us now proceed to discuss the entanglement spectrum. 
+An important quantity is the correlation function, which in position space is given by
+\begin{align}
 C_{ij}^{\alpha \beta} = \expval{c_{i\alpha}^\dagger c_{j\beta}}.
-\end{align*}
+\end{align}
 In terms of the fermions that diagonalize the Hamiltonian, 
-\begin{align*}
-& \gamma_{i\alpha} = \sum_{j\beta}u_{j\beta}^{i\alpha \, \ast} c_{j\beta} \\
+\begin{align}
+& \gamma_{i\alpha} = \sum_{j\beta}u_{j\beta}^{i\alpha \, \ast} c_{j\beta} \nonumber\\
 & c_{i\alpha} = \sum_{j\beta} u^{j\beta}_{i\alpha} \gamma_{j\beta}
-\end{align*}
-we have
-\begin{align*}
-C_{ij}^{\alpha \beta} =& \sum_{pq,\mu\nu} u^{q\nu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} \expval{\gamma^\dagger_{p\mu} \gamma_{q\nu} } \\
-=&  \sum_{p\mu} u^{p\mu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} \expval{\gamma^\dagger_{p\mu} \gamma_{p\mu} } \\
+\end{align}
+we find that the correlation matrix can be written as 
+\begin{align}\label{eq:corr_mat}
+C_{ij}^{\alpha \beta} =& \sum_{pq,\mu\nu} u^{q\nu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} \expval{\gamma^\dagger_{p\mu} \gamma_{q\nu} }\nonumber \\
+=&  \sum_{p\mu} u^{p\mu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} \expval{\gamma^\dagger_{p\mu} \gamma_{p\mu} }\nonumber \\
 =&  \sum_{p\mu} u^{p\mu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} [1-{\rm sign}(E_{p\mu})]/2.
-\end{align*}
-The first term is simply a kronecker delta between both sets of indeces. The second term can be rewritten as
-\begin{align*}
-UD(\abs{D})^{-1}U^\dagger =& UD(D^2)^{-1/2}U^\dagger \\
-=& U D U^\dagger (U D^2 U^\dagger)^{-1/2} \\
-=& H(H^2)^{-1/2},
-\end{align*}
-where $D$ is the diagonal Hamiltonian, $H=UDU^\dagger$. Finally the correlation function can be obtained in position space as
-\begin{align*}
-C_{ij}^{\alpha \beta} =& \frac{1}{2}\left[I - H/ (H^2)^{-1/2} \right]_{ij, \alpha \beta}
-\end{align*}
-The single-particle entanglement spectrum (ES) is then obtained as the spectrum of the subsystem correlation function obtained after bisecting the system into two equal parts.
+\end{align}
+The first term of the last line is simply a kronecker delta between both sets of indices. 
+The second term can be rewritten, using Eq.~\ref{eq:D}, as
+\begin{align}\label{eq:C_rewritten}
+UD(\abs{D})^{-1}U^\dagger =& UD(D^2)^{-1/2}U^\dagger\nonumber \\
+=& U D U^\dagger (U D^2 U^\dagger)^{-1/2}\nonumber \\
+=& H(H^2)^{-1/2}.
+\end{align}
+\mariac{I am sorry, but I do not see how to move from the first to the second line}
+%where $D$ is the diagonal Hamiltonian, $H=UDU^\dagger$. 
+Consequently,  the correlation function in position space can be written  as
+\begin{align}\label{eq:corr_mat2}
+C_{ij}^{\alpha \beta} =& \frac{1}{2}\left[I - H/ (H^2)^{-1/2} \right]_{ij, \alpha \beta}.
+\end{align}
+This form is used for all subsequent calculations.
 
-In the ES we can differentiate between two types of eigenvalues, which we denote by $\xi$.
+In this manuscript, we are mainly interested in the spectrum of the reduced\mariac{better expression?} correlation matrix, i.e.
+the correlation matrix restricted to a spatial sub-region $A$ (its complement will be denoted by $B$ in the following).
+For simplicity, we generically choose $A$ n as half the system, but none of our results depends on this fact. 
+The resulting correlation matrix spectrum (CMS) is  denoted by the set of eigenvalues $\xi_j$. 
+We will differentiate between two types of eigenvalues: those at $\xi_j=0 (1)$ correspond to ``bulk'' states, by which we mean that the corresponding eigenstates lie (mostly) in  $A$ ($B$). 
+Values $0<\xi_{l,r}<1$ indicate states that are located at the virtual boundaries -- labeled `left' and `right' -- between $A$ and $B$~\cite{Peschel2008}. 
+Topological phases are characterized by (symmetry-protected) zero-energy modes in the edge spectrum, which is in one-to-one correspondence to $\xi=\frac 1 2$ modes in the CMS~\cite{Fidkowski2010entanglement}.
+
+The CMS is in one-to-one correspondence to the single-particle entanglement spectrum (ES). 
+Following Ref.~\cite{Peschel2008}, one can write the  reduced density matrix as 
+\begin{align}\label{eq:red_dens_mat}
+\rho&=\mathcal{K} \exp(-\mathcal H),
+\end{align}
+where $\mathcal{K}$ is a normalization constant and $\mathcal{H}$ is a quadratic Hamiltonian, referred to as entanglement Hamiltonian. 
+The eigenvalues $\epsilon_j$ of $\mathcal{H}$ are related to those of $C$ by 
+\begin{align}
+\xi_j &=\left(e^{\epsilon_j}+1\right)^{-1}. 
+\end{align}
+Thus, the full information of the \emph{many-particle} ES is fully contained in the single-particle spectrum of the correlation matrix. 
+The latter is much simpler to interpret. 
+Consequently, we will focus on the CMS in the remainder of the manuscript and only mention the ES when necessary. 
+\mariac{not happy with the flow yet}
+%The single-particle entanglement spectrum (ES) is then obtained as the spectrum of the subsystem correlation function obtained after bisecting the system into two equal parts.
+
+%In the ES we can differentiate between two types of eigenvalues, which we denote by $\xi$.
 The ones at $\xi_b = 0,1$ correspond to bulk states while for the ones in between, $0<\xi_{l,r}<1$, we find their correspondent eigenstate localized in either the left, $\xi_l$, or right, $\xi_r$, virtual edge \cite{Peschel2008}.
-\mariac{introduce entanglement energy, relation between C and energies}
+%\mariac{introduce entanglement energy, relation between C and energies}
 

### 2020-04-27 draft.tex
some changes

@@
 
-\title{Something fancy...}
+\title{On the relation of the entanglement spectrum to the Zak-phase}
 \author{Carlos Ortega Taberner}
@@
 \section{Model.}
-\mariac{ref needed to 10-fold way}
 To showcase our results we consider as a starting point the model in the BDI class used in reference~\cite{Song2014}. 
 This is a very simple model which nevertheless supports a rich phase diagram containing topological phases with winding numbers up to $\nu =2$. 
-In order to show that our results are not limited to any symmetry class, we also include two symmetry breaking terms. 
+In order to show that our results are not limited to any symmetry class~\cite{ryu2010topological}, we also include two symmetry breaking terms. 
 The Hamiltonian is  given by
@@
 \end{align}
+
 For $\kappa = \kappa' = 0$ this Hamiltonian is in the BDI class, i.e. it has time-reversal ($T$), particle-hole ($C$), and chiral symmetry ($S$):
@@
 \end{figure}
-When imposing open boundary conditions on the system harbors, the number of symmetry-protected zero energy edge modes is equal to the winding number. 
+The BDI class  is characterized by a $\mathbb{Z}$ invariant, the winding number,  in one spatial dimension.
+When imposing open boundary conditions on the system, the number of symmetry-protected zero energy edge modes is equal to the winding number. 
+The behavior of the  edge modes  is  shown in Fig.~\ref{fig:zero_E_modes}, using the dotted path ($t=1,t'=-2$) marked in the phase diagram. 
+Subfigure (a) shows the edge spectrum in the BDI class, with 4 (resp. 2) symmetry-protected zero modes for winding number 2 (1). 
 %We also a cut through the three phases (dotted line in the phase diagram) for different values of $\kappa,\kappa'$.
@@
 For $\kappa' \neq 0$ only particle-hole symmetry $C$ is preserved and the system belongs to symmetry class $D$. 
-In one spatial dimension, the latter is characterized \mariac{better expression?} by a $\mathbb{Z}_2$ instead of the $\mathbb{Z}$ invariant of the BDI class. 
-Consequently, adding such a term causes the zero modes in the $\nu=2$ phase to split from zero pairwise and the $\nu = 2$ and $\nu=0$ phases become equivalent. 
+In one spatial dimension, the latter is characterized by a $\mathbb{Z}_2$ invariant.  
+Consequently, adding such a term causes the zero modes in the $\nu=2$ phase to split  pairwise and the $\nu = 2$ and $\nu=0$ phases become equivalent. 
+This is shown in subfigure (c), where one clearly sees that the zero modes below $m=1$ are split.
+
 For $\kappa \neq 0$ only time-reversal $T$ is preserved and the system is in the AI class. 
 In one dimensions, this class is trivial and, thus,  all zero-energy modes split. 
-With both $\kappa \neq 0$ and $\kappa' \neq0$ the system has no symmetry, thus belonging to the $A$ class.  Also this symmetry class is topologically trivial in one dimension.
+With both $\kappa \neq 0$ and $\kappa' \neq0$ the system has no symmetry, thus belonging to the $A$ class.  
+Also this symmetry class is topologically trivial in one dimension. 
+Subfigures (b) and (d) show the edge spectrum for class AI and A, respectively. In both cases, the edge modes are split from zero.
 
-The behavior of the zero-energy edge modes in the various symmetry classes is  shown in Fig.~\ref{fig:zero_E_modes}, using the dotted path ($t=1,t'=-2$) marked in the phase diagram~\ref{fig:bdi_phase_diagram}.
-Subfigure (a) shows the edge spectrum in the BDI class, with 4 (resp. 2) symmetry-protected zero modes for winding number 2 (1). 
-In subfigure (b) and (d), where the system is in a topologically trivial phase, all edge modes are gapped. 
-Subfigure (c) shows the edge spectrum of the system being in class D, where the $\nu=2 $ phase becomes trivial. 
-Consequently, symmetry-protected zero modes only exist for $1<m<2.5$\mariac{?} (corresponding to $\nu=1$ for $\kappa'=0$) , whereas the modes are split for $m<1$ (corresponding to $\nu=2$ for $\kappa'=0$).
+%
+%In subfigure (b) and (d), where the system is in a topologically trivial phase, all edge modes are gapped. 
+%Subfigure (c) shows the edge spectrum of the system being in class D, where the $\nu=2 $ phase becomes trivial. 
+%Consequently, symmetry-protected zero modes only exist for $1<m<2.5$\mariac{?} (corresponding to $\nu=1$ for $\kappa'=0$) , whereas the modes are split for $m<1$ (corresponding to $\nu=2$ for $\kappa'=0$).
 
@@
 }
-\caption{Surface energy spectrum along the dotted path (i.e. $t=1,t'=-2$) in the phase diagram of Fig.~\ref{fig:bdi_phase_diagram} for different values of $\kappa,\kappa'$.}
+\caption{\mariac{increase system size and colorcode edge modes.}Edge spectrum along the dotted path (i.e.
+$t=1,t'=-2$) of Fig.~\ref{fig:bdi_phase_diagram} for different values of $\kappa$ and $\kappa'$.\mariac{note on numerical system size}}
 \label{fig:zero_E_modes}
@@
 Beyond symmetry protected systems, there was observed that in the fully dimerized SSH chain, with only a single mode present in the ES, the Zak phase is equal to the midgap eigenvalue \cite{Ryu2006}.
-For cases with more edge modes present in the ES it has been shown that the Zak phase follows the midgap eigenvalues qualitatively \cite{Huang2012,Huang2012-2}.
-The first result we show is that in systems with nearest-neighbor hopping, like the SSH chain, one can obtain the Zak phase by ordering the eigenvalues of the ES, $\xi_1 < \xi_2 < ...$, and computing
+\mariac{correct?}For cases with more edge modes present in the ES it has been shown that the Zak phase follows the midgap eigenvalues qualitatively \cite{Huang2012,Huang2012-2}.
+\mariac{too long}The first result we show is that in systems with nearest-neighbor hopping, like the SSH chain, one can obtain the Zak phase by ordering the eigenvalues of the ES, $\xi_1 < \xi_2 < ...$, and computing
 \begin{equation}
@@
 In the case where there is only one finite eigenvalue in  the ES, like in the fully dimerized limit of the SSH chain, we recover the equality between the midgap eigenvalue and the Zak phase.
-In Fig.\ref{huang} we show show the dashed cut in the phase diagram shown above, which is equivalent to an SSH chain.
+In Fig.\ref{huang} we show show the ES along the dashed cut in the phase diagram of Fig.~\ref{fig:bdi_phase_diagram}, which is equivalent to an SSH chain.
 We see that even though the ES has finite eigenvalues in the trivial phase these are double-degenerate so that they cancel each other.

### 2020-04-27 draft.tex
added appendix

@@
 \end{alignat}
-The corresponding phase diagram is shown in Fig.\ref{fig:bdi_phase_diagram} for $t=1$. 
+Using the results of Ref.~\cite{Song2014}, the corresponding phase diagram can be calculated and is shown in Fig.\ref{fig:bdi_phase_diagram} for $t=1$. 
 \begin{figure}[t]
@@
 	}}\hspace{0mm}
-\caption{\mariac{need more visible lines .}Phase diagram for the system in the $BDI$ class showing the winding number of each region.}
+\caption{\mariac{need more visible lines. Get rid of (a).}Phase diagram for the system in the $BDI$ class showing the winding number of each region.}
 \label{fig:bdi_phase_diagram}
@@
 Also this symmetry class is topologically trivial in one dimension. 
-Subfigures (b) and (d) show the edge spectrum for class AI and A, respectively. In both cases, the edge modes are split from zero.
+Subfigures (b) and (d) show the edge spectrum for class AI and A, respectively. In both cases, the edge modes are split from zero for all values of $m$.
 
@@
 We consider a generic, quadratic Hamiltonian in one dimension
-\begin{equation}
+\begin{equation}\label{eq:quadr_Ham}
 \mathcal{H} = \sum_{ij,\alpha\beta} c_{i\alpha}^\dagger H_{ij,\alpha \beta}c_{j\beta}.
@@
 \end{align}
-where $[U]_{j\beta,p\mu} = u^{p\mu}_{j\beta}$ is the unitary matrix that diagonalizes $H$, i.e. 
-\begin{align}\label{eq:D}
-H&=UDU^\dagger& \mbox{with } D=\mbox{diag}(E_{p\mu}).
-\end{align} 
+where $[U]_{j\beta,p\mu} = u^{p\mu}_{j\beta}$ is the unitary matrix that diagonalizes $H$.
+
+\mariac{Carlos wants to add a paragraph on Zak phase. }
+
 The polarization can be obtained as the many-body average of the position operator, but the latter has to be regularized for periodic boundary conditions \cite{Resta1997}. 
@@
 where $\ket{\Psi_0}$ is the many-body ground state. 
-Using that the ground state is a Slater determinant of the occupied single-particle eigenstates, $\ket{u^{p\mu}}$, it can be rewritten as
+Using that the ground state is a Slater determinant of the occupied single-particle eigenstates, $\ket{u^{p\mu}}$, it can be rewritten as~\cite{Resta1997}
 \begin{align}
-\expval{\hat{X}_{\rm PBC}}_0 = -\frac{L}{2\pi} \rm{Im }\ln \, {\rm det} \, S,
+\expval{\hat{X}_{\rm PBC}}_0 = -\frac{L}{2\pi} \rm{Im }\ln \, {\rm det }\!' \, S,
 \end{align}
@@
 \begin{equation}
-S_{p\mu,q\nu} = \sum_{j\alpha}\!^{'} u^{p\mu\, \ast}_{j \alpha} e^{-i\frac{2\pi}{L}j}u^{q\nu}_{j \alpha}
+S_{p\mu,q\nu} = \sum_{j\alpha} u^{p\mu\, \ast}_{j \alpha} e^{-i\frac{2\pi}{L}j}u^{q\nu}_{j \alpha}
 \end{equation}
-and the $\sum\!{'}$ indicates that the sum is restricted to the space of \emph{occupied} single-particle states. 
+and  ${\rm det }\!' $ indicates that the determinant is restricted to the space of \emph{occupied} single-particle states. 
 The Zak phase can then be obtained as
@@
 Let us now proceed to discuss the entanglement spectrum. 
-An important quantity is the correlation function, which in position space is given by
+An important quantity is the correlation matrix, which in position space is defined by
 \begin{align}
@@
 \end{align}
-In terms of the fermions that diagonalize the Hamiltonian, 
-\begin{align}
-& \gamma_{i\alpha} = \sum_{j\beta}u_{j\beta}^{i\alpha \, \ast} c_{j\beta} \nonumber\\
-& c_{i\alpha} = \sum_{j\beta} u^{j\beta}_{i\alpha} \gamma_{j\beta}
-\end{align}
-we find that the correlation matrix can be written as 
-\begin{align}\label{eq:corr_mat}
-C_{ij}^{\alpha \beta} =& \sum_{pq,\mu\nu} u^{q\nu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} \expval{\gamma^\dagger_{p\mu} \gamma_{q\nu} }\nonumber \\
-=&  \sum_{p\mu} u^{p\mu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} \expval{\gamma^\dagger_{p\mu} \gamma_{p\mu} }\nonumber \\
-=&  \sum_{p\mu} u^{p\mu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} [1-{\rm sign}(E_{p\mu})]/2.
-\end{align}
-The first term of the last line is simply a kronecker delta between both sets of indices. 
-The second term can be rewritten, using Eq.~\ref{eq:D}, as
-\begin{align}\label{eq:C_rewritten}
-UD(\abs{D})^{-1}U^\dagger =& UD(D^2)^{-1/2}U^\dagger\nonumber \\
-=& U D U^\dagger (U D^2 U^\dagger)^{-1/2}\nonumber \\
-=& H(H^2)^{-1/2}.
-\end{align}
-\mariac{I am sorry, but I do not see how to move from the first to the second line}
-%where $D$ is the diagonal Hamiltonian, $H=UDU^\dagger$. 
-Consequently,  the correlation function in position space can be written  as
+Using the results of appendix~\ref{app:CM}, it  can be written  as
 \begin{align}\label{eq:corr_mat2}
@@
 \end{align}
-This form is used for all subsequent calculations.
+This form is particularly useful for all our subsequent calculations, since it does not rely on translational symmetry.
 
-In this manuscript, we are mainly interested in the spectrum of the reduced\mariac{better expression?} correlation matrix, i.e.
-the correlation matrix restricted to a spatial sub-region $A$ (its complement will be denoted by $B$ in the following).
-For simplicity, we generically choose $A$ n as half the system, but none of our results depends on this fact. 
+In this manuscript, we are mainly interested in the spectrum of the correlation matrix, when restricted to a spatial sub-region $A$ (its complement will be denoted by $B$ in the following). 
+For simplicity, we generically choose $A$ as half the system. \mariac{expand this to a paragraph.}
+
 The resulting correlation matrix spectrum (CMS) is  denoted by the set of eigenvalues $\xi_j$. 
@@
 
+\section{Correlation matrix}\label{app:CM}
+We consider the generic, quadratic Hamiltonian in one dimension of Eq.~\ref{eq:quadr_Ham}, which is diagonalized by a unitary matrix $U$ with 
+	\begin{align}\label{eq:D}
+	H&=UDU^\dagger& \mbox{with } D=\mbox{diag}(E_{p\mu}).
+	\end{align} 
+In terms of the fermionic operators that diagonalize the Hamiltonian, 
+\begin{align}
+& \gamma_{i\alpha} = \sum_{j\beta}u_{j\beta}^{i\alpha \, \ast} c_{j\beta},
+%& c_{i\alpha} = \sum_{j\beta} u^{j\beta}_{i\alpha} \gamma_{j\beta}
+\end{align}
+we find that the correlation matrix can be written as 
+\begin{align}\label{eq:corr_mat}
+C_{ij}^{\alpha \beta} =& \sum_{pq,\mu\nu} u^{q\nu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} \expval{\gamma^\dagger_{p\mu} \gamma_{q\nu} }\nonumber \\
+=&  \sum_{p\mu} u^{p\mu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} \expval{\gamma^\dagger_{p\mu} \gamma_{p\mu} }\nonumber \\
+=&  \sum_{p\mu} u^{p\mu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} [1-{\rm sign}(E_{p\mu})]/2.
+\end{align}
+The first term of the last line is simply a kronecker delta between both sets of indices.
+
+The second term can be rewritten, using Eq.~\ref{eq:D}, as
+\begin{align}\label{eq:C_rewritten}
+UD(\abs{D})^{-1}U^\dagger =& UDU^\dagger U(D^2)^{-1/2}U^\dagger
+\end{align}
+Consider
+\begin{align*}
+[U (D^2)^{-1/2} U^\dagger]^2 =& U (D^2)^{-1/2} U^\dagger U  (D^2)^{-1/2} U^\dagger \\
+=&  U (D^2)^{-1/2}  (D^2)^{-1/2} U^\dagger \\
+=&  U (D^2)^{-1} U^\dagger \\
+=&  (U D^2 U^\dagger)^{-1} \\
+\end{align*}
+Therefore
+\begin{align*}
+U (D^2)^{-1/2} U^\dagger =&(U D^2 U^\dagger)^{-1/2} 
+\end{align*}
+and we have
+\begin{align}\label{eq:C_rewritten}
+UD(\abs{D})^{-1}U^\dagger =& UDU^\dagger(U D^2 U^\dagger)^{-1/2} \\ 
+=& H(H^2)^{-1/2}.
+\end{align}
+
 \end{document}

### 2020-04-27 draft.tex
added appendix

@@
 
-\mariac{Carlos wants to add a paragraph on Zak phase. }
+In translational invariant systems the Zak phase is defined as the Berry phase acquired by a state $\ket{u_k}$ as it winds around the Brillouin zone,
+\begin{equation}
+\gamma = \int_{-\pi}^\pi dk \, \bra{u_k}\partial_k \ket{u_k} \, {\rm mod}\, 2\pi.
+\end{equation}
+The Zak phase is equal to $\pi$ times the parity of the winding number, and it is also related to the polarization \cite{Resta1997}.
+Since the latter has a position space formulation we will use  it in all calculations below to compute the Zak phase.
 
@@
 \end{align}
-This form is particularly useful for all our subsequent calculations, since it does not rely on translational symmetry.
+This form is particularly useful for all our subsequent calculations, since it does not rely on translation symmetry.
 
 In this manuscript, we are mainly interested in the spectrum of the correlation matrix, when restricted to a spatial sub-region $A$ (its complement will be denoted by $B$ in the following). 
-For simplicity, we generically choose $A$ as half the system. \mariac{expand this to a paragraph.}
-
-The resulting correlation matrix spectrum (CMS) is  denoted by the set of eigenvalues $\xi_j$. 
-We will differentiate between two types of eigenvalues: those at $\xi_j=0 (1)$ correspond to ``bulk'' states, by which we mean that the corresponding eigenstates lie (mostly) in  $A$ ($B$). 
-Values $0<\xi_{l,r}<1$ indicate states that are located at the virtual boundaries -- labeled `left' and `right' -- between $A$ and $B$~\cite{Peschel2008}. 
+Following Ref.s~\cite{Huang2012,Huang2012-2}, we refer to this spectrum as the entanglement occupancy spectrum (EOS). 
+The resulting eigenvalues  are denoted by $\xi_j$. 
+We will differentiate between two types of eigenvalues: those (exponentially) close to $\xi_j=0 (1)$ correspond to ``bulk'' states, by which we mean that the corresponding eigenstates lie (up to exponentially small corrections) in  $A$ ($B$).
+Values $0<\xi_{l,r}<1$ indicate states that are located at the virtual boundaries -- labeled `left' and `right' in the following-- between $A$ and $B$~\cite{Peschel2008}. 
+There is, of course, an ambiguity here on dividing eigenstates of the EOS into ``bulk'' and ``edge'' states and `how close' the eigenvalue has to be to $0,1$ in order to qualify as a bulk state. 
+XXXX
 Topological phases are characterized by (symmetry-protected) zero-energy modes in the edge spectrum, which is in one-to-one correspondence to $\xi=\frac 1 2$ modes in the CMS~\cite{Fidkowski2010entanglement}.
@@
 
+A few comments are needed regarding  bipartition. 
+In the following sections, we show that the Zak phase can be recovered from the EOS.
+We also give a simple formula that is identical to the Zak phase in the thermodyamic limit.  
+However, for a generic, finite size system, there will be finite-size discrepancies between our formula and the Zak phase.
+In order to reduce these, we are going to choose $A$ as half the system in the reminder of the manuscript.
+The results (in the thermodynamic limit) do not depend on this choice. 
 % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % 

### 2020-04-27 draft.tex
added appendix

@@
 In this manuscript, we show that the information about the Zak phase can be completely recovered from the entanglement spectrum and provide a simple formula to do so. 
-We relate our results to earlier ones XXX
+We relate our results to earlier ones, \mariac{mention Pollman~\cite{Zaletel2014}, Ryu hatsugai~\cite{Ryu2006} and Huang Arovas~\cite{Huang2012,Huang2012-2}} XXX
 
@@
 The resulting eigenvalues  are denoted by $\xi_j$. 
-We will differentiate between two types of eigenvalues: those (exponentially) close to $\xi_j=0 (1)$ correspond to ``bulk'' states, by which we mean that the corresponding eigenstates lie (up to exponentially small corrections) in  $A$ ($B$).
-Values $0<\xi_{l,r}<1$ indicate states that are located at the virtual boundaries -- labeled `left' and `right' in the following-- between $A$ and $B$~\cite{Peschel2008}. 
-There is, of course, an ambiguity here on dividing eigenstates of the EOS into ``bulk'' and ``edge'' states and `how close' the eigenvalue has to be to $0,1$ in order to qualify as a bulk state. 
-XXXX
-Topological phases are characterized by (symmetry-protected) zero-energy modes in the edge spectrum, which is in one-to-one correspondence to $\xi=\frac 1 2$ modes in the CMS~\cite{Fidkowski2010entanglement}.
+%We will differentiate between two types of eigenvalues: those (exponentially) close to $\xi_j=0 (1)$ correspond to ``bulk'' states, by which we mean that the corresponding eigenstates lie (up to exponentially small corrections) in  $A$ ($B$).
+%Values $0<\xi_{l,r}<1$ indicate states that are located at the virtual boundaries -- labeled `left' and `right' in the following-- between $A$ and $B$~\cite{Peschel2008}. 
+%There is, of course, an ambiguity here on dividing eigenstates of the EOS into ``bulk'' and ``edge'' states and `how close' the eigenvalue has to be to $0,1$ in order to qualify as a bulk state. 
+%XXXX
+Topological phases are characterized by (symmetry-protected) zero-energy modes in the edge spectrum and $\xi=\frac 1 2$ modes in the OES~\cite{Fidkowski2010entanglement}.
 
@@
 \begin{align}\label{eq:red_dens_mat}
-\rho&=\mathcal{K} \exp(-\mathcal H),
+\rho_A&=\mathcal{K} \exp(-\mathcal H),
 \end{align}
 where $\mathcal{K}$ is a normalization constant and $\mathcal{H}$ is a quadratic Hamiltonian, referred to as entanglement Hamiltonian. 
-The eigenvalues $\epsilon_j$ of $\mathcal{H}$ are related to those of $C$ by 
+The eigenvalues $\epsilon_j$ of $\mathcal{H}$ are related to the EOS by 
 \begin{align}
@@
 \end{align}
-Thus, the full information of the \emph{many-particle} ES is fully contained in the single-particle spectrum of the correlation matrix. 
+Thus, the full information of the \emph{many-particle} ES is fully contained in the single-particle spectrum of the reduced correlation matrix. 
 The latter is much simpler to interpret. 
-Consequently, we will focus on the CMS in the remainder of the manuscript and only mention the ES when necessary. 
-\mariac{not happy with the flow yet}
+Consequently, we will focus on the EOS in the remainder of the manuscript and only mention the ES when necessary. 
 %The single-particle entanglement spectrum (ES) is then obtained as the spectrum of the subsystem correlation function obtained after bisecting the system into two equal parts.
@@
 A few comments are needed regarding  bipartition. 
-In the following sections, we show that the Zak phase can be recovered from the EOS.
+In the following sections, we show how the Zak phase can be recovered from the EOS.
 We also give a simple formula that is identical to the Zak phase in the thermodyamic limit.  
 However, for a generic, finite size system, there will be finite-size discrepancies between our formula and the Zak phase.
-In order to reduce these, we are going to choose $A$ as half the system in the reminder of the manuscript.
+In order to reduce these, we are going to choose $A$ as half the system in the remainder of the manuscript.
 The results (in the thermodynamic limit) do not depend on this choice. 
@@
 
+Let us first review previous results on the relation between the entanglement spectrum and the Zak phase. 
+For symmetry-protected topological systems in one dimension, the quantized Zak phase is zero whenever there is an even number  of eigenvalues per edge at $\xi = 1/2$, and $\pi$ when the number is odd~\cite{Peschel2008}.
+Ryu and Hatsugai considered the fully dimerized SSH chain with broken chiral symmetry~\cite{Ryu2006}.  
+This model is rather special in that there is only a single pair of eigenvalues in the EOS that is not identical to $0$ or $1$. 
+In fact, the authors could show that the value of this `midgap state' is identical to the Zak phase divided by $2\pi$. 
+Their result is a particular limit of Eq.~\eqref{eqplusminus}. 
+%
+%Beyond symmetry protected systems, there was observed that in the fully dimerized SSH chain, with only a single mode present in the ES, the Zak phase is equal to the midgap eigenvalue \cite{Ryu2006}.
+Huang and Arovas considered a two-dimensional chern insulator and noted that the Zak phase qualitatively follows the `virtual edge mode' that connects the EOS values at 0 to those at 1~\cite{Huang2012,Huang2012-2}.
+Interpreting this as a one-dimensional system with a parameter, we can explain this behavior by Eq.~\eqref{eqplusminus}, noting that there is a single pair of eigenvalues that dominates the sum. 
+Note that the observation of Ref.s~\cite{Huang2012,Huang2012-2} is particular to a system with Chern number 0 or $\pm 1$, and it generically fails for systems with higher Chern numbers.
+In the latter case, there are several terms in Eq.~\eqref{eqplusminus} with comparably large contributions. 
+%For cases with more edge modes present in the ES it has been shown that the Zak phase follows the midgap eigenvalues qualitatively \cite{Huang2012,Huang2012-2}.
+
 \subsubsection{Nearest-neighbor hopping}
-
-For symmetry-protected topological systems in one dimension there is a well-known relation between ES and the Zak phase \cite{Peschel2008}.
-The Zak phase, which is quantized, is zero whenever there is an even number per edge of eigenvalues at $\xi = 1/2$, and $\pi$ when the number is odd.
-Beyond symmetry protected systems, there was observed that in the fully dimerized SSH chain, with only a single mode present in the ES, the Zak phase is equal to the midgap eigenvalue \cite{Ryu2006}.
-\mariac{correct?}For cases with more edge modes present in the ES it has been shown that the Zak phase follows the midgap eigenvalues qualitatively \cite{Huang2012,Huang2012-2}.
 \mariac{too long}The first result we show is that in systems with nearest-neighbor hopping, like the SSH chain, one can obtain the Zak phase by ordering the eigenvalues of the ES, $\xi_1 < \xi_2 < ...$, and computing

### 2020-04-28 draft.tex
changes in intro

@@
 In this manuscript, we show that the information about the Zak phase can be completely recovered from the entanglement spectrum and provide a simple formula to do so. 
-We relate our results to earlier ones, \mariac{mention Pollman~\cite{Zaletel2014}, Ryu hatsugai~\cite{Ryu2006} and Huang Arovas~\cite{Huang2012,Huang2012-2}} XXX
+The relation between the entanglement spectrum and various topological invariants has already been studied in the literature. 
+In Ref.~\cite{Zaletel2014}, the authors show that the Zak phase can be computed from the Schmidt decomposition of a translation invariant, infinite chain. 
+This demonstrates that there should be a direct relation between the single-particle entanglement energies and the Zak phase, although deriving that relation turned out to be non-trivial. 
+An added benefit of our approach is that it can be generalized to systems that are not translationally invariant in a numerically efficient way. 
+A particular limit of our result for homogeneous systems~\ref{eqplusminus} was derived by Ryu and Hatsugai for a dimerized SSH chain~\cite{Ryu2006}. 
+The similarity between the behavior of the Zak phase and the 'virtual edge modes' in the entanglement spectrum of a Chern insulator, which was observed in Ref.~\cite{Huang2012,Huang2012-2}, can also be explained by our results.
 
@@
 \subsubsection{Nearest-neighbor hopping}
-\mariac{too long}The first result we show is that in systems with nearest-neighbor hopping, like the SSH chain, one can obtain the Zak phase by ordering the eigenvalues of the ES, $\xi_1 < \xi_2 < ...$, and computing
+Let us first focus on systems with nearest-neighbor hopping, such as the SSH chain.
+ one can obtain the Zak phase by ordering the eigenvalues of the ES, $\xi_1 < \xi_2 < ...$, and computing
 \begin{equation}

### 2020-04-28 draft.tex
changes in intro

@@
 An added benefit of our approach is that it can be generalized to systems that are not translationally invariant in a numerically efficient way. 
-A particular limit of our result for homogeneous systems~\ref{eqplusminus} was derived by Ryu and Hatsugai for a dimerized SSH chain~\cite{Ryu2006}. 
+A particular limit of our result for homogeneous systems~\ref{eqplusminus} was derived by Ryu and Hatsugai for a fully dimerized SSH chain~\cite{Ryu2006}. 
 The similarity between the behavior of the Zak phase and the 'virtual edge modes' in the entanglement spectrum of a Chern insulator, which was observed in Ref.~\cite{Huang2012,Huang2012-2}, can also be explained by our results.
@@
 	}}\hspace{0mm}
-\caption{\mariac{need more visible lines. Get rid of (a).}Phase diagram for the system in the $BDI$ class showing the winding number of each region.}
+\caption{\mariac{ Get rid of (a).}Phase diagram for the system in the $BDI$ class showing the winding number of each region.}
 \label{fig:bdi_phase_diagram}
@@
 
-The CMS is in one-to-one correspondence to the single-particle entanglement spectrum (ES). 
+The EOS is in one-to-one correspondence to the entanglement spectrum (ES). 
 Following Ref.~\cite{Peschel2008}, one can write the  reduced density matrix as 
@@
 where $\mathcal{K}$ is a normalization constant and $\mathcal{H}$ is a quadratic Hamiltonian, referred to as entanglement Hamiltonian. 
-The eigenvalues $\epsilon_j$ of $\mathcal{H}$ are related to the EOS by 
+The eigenvalues $\epsilon_j$ of $\mathcal{H}$, referred to as entanglement energies,  are related to the EOS by 
 \begin{align}
@@
 \end{align}
+XXXX
 Thus, the full information of the \emph{many-particle} ES is fully contained in the single-particle spectrum of the reduced correlation matrix. 
@@
 
-\subsubsection{Nearest-neighbor hopping}
-Let us first focus on systems with nearest-neighbor hopping, such as the SSH chain.
- one can obtain the Zak phase by ordering the eigenvalues of the ES, $\xi_1 < \xi_2 < ...$, and computing
+\subsubsection{Systems with equidistant entanglement energies}
+We first discuss systems, for which the entanglement energies of a \emph{single} virtual edge are equidistant. 
+This is a common feature for e.g. nearest neighbor hopping models\mariac{ref to Peschel}, such as the SSH chain, and applies to the Hamiltonian~\ref{bdi_model} when $t'=0$. 
+In this case, one can obtain the Zak phase in a very simple fashion: We reorder the eigenvalues of the ES by magnitude, $\xi_1 < \xi_2 < ...$, and compute
 \begin{equation}
@@
 \end{equation}
-where $N_A$ is the number of electrons in the ground state in region A, which in the thermodynamic limit fulfills
+where $N_A$ is the number of electrons in the ground state in region A. 
+In the thermodynamic limit, $\chi$ becomes identical to the Zak phase divided by $2\pi$:
 \begin{equation}
-\lim_{L \rightarrow \infty} \gamma = \lim_{L \rightarrow \infty} \chi.
+\lim_{L \rightarrow \infty} \frac{\gamma}{2\pi} = \lim_{L \rightarrow \infty} \chi.
 \end{equation}
 
-In the case where there is only one finite eigenvalue in  the ES, like in the fully dimerized limit of the SSH chain, we recover the equality between the midgap eigenvalue and the Zak phase.
-In Fig.\ref{huang} we show show the ES along the dashed cut in the phase diagram of Fig.~\ref{fig:bdi_phase_diagram}, which is equivalent to an SSH chain.
-We see that even though the ES has finite eigenvalues in the trivial phase these are double-degenerate so that they cancel each other.
-In the case where we break the symmetry we see that there is one main midgap eigenvalue, so that the Zak phase follows it very closely, but as the midgap eigenvalue approaches the bulk, the contribution from the other modes becomes more relevant and the Zak phase start to deviate from the midgap eigenvalue.
+In the case where there is only one eigenvalue per edge that is $\neq 0,1$ in  the EOS, like in the fully dimerized limit of the SSH chain, we recover the equality between the midgap eigenvalue and the Zak phase~\cite{Ryu2006}.
+
+In Fig.~\ref{huang} we show  the EOS along the dashed cut in the phase diagram of Fig.~\ref{fig:bdi_phase_diagram}. 
+Along this line, our model is equivalent to an SSH chain. 
+In the presence of chiral and translation symmetry, all eigenvalues of the EOS are doubly degenerate --- one eigenvalue from each edge. 
+The difference of $\pi$ between the trivial and topological regime originates from the rearrangement of a few dominant eigenvalues in the EOS. 
+In the topological regime, there are a pair of eigenvalues at $\xi=1/2$, while the rest are close to $0$ and $1$. 
+At the phase transition, the two $\xi=1/2$ modes mix with one mode coming from $1$ and $0$ respectively, leading to a pair of modes close to $0$ and $1$, see Fig.~\ref{huang}(a). 
+Consequently, there is one less mode at $\xi=1$, which yields a difference of $\pi$ in the Zak phase. 
+Note that the full sum~\ref{eqplusminus}  evaluates to zero in the trivial regime, up to finite-size corrections, despite the presence of modes with entanglement occupancy $0<\xi<1$.
+
+When breaking chiral symmetry, the EOS will no longer be doubly degenerate. 
+For small $m$, there are two dominant midgap eigenvalues, and the Zak phase follows the lower of them very closely, see Fig.~\ref{huang}(b).
+When this midgap eigenvalue approaches  0, the contribution from the other modes becomes more relevant and the Zak phase starts to deviate from the midgap eigenvalue, \mariac{see inset}.
 \begin{figure}[h!]
@@
 }
-\caption{Entanglement spectrum (black), $\chi$ (green) and $\gamma$ (red) for a cut in the phase diagram through the $\nu = 1 \rightarrow \nu = 0$ transition (dashed line in the phase diagram) for $t'=0$ and different values of $\kappa$.}
+\caption{\mariac{system size?}Entanglement spectrum (black), $\chi$ (green) and $\frac{\gamma}{2\pi}$ (red) for a cut in the phase diagram through the $\nu = 1 \rightarrow \nu = 0$ transition (dashed line in 	\ref{fig:bdi_phase_diagram}
+	) for $t'=0$ and different values of $\kappa$.}
 \label{huang}

### 2020-04-29 draft.tex
..

@@
 \begin{align}
-\xi_j &=\left(e^{\epsilon_j}+1\right)^{-1}. 
+\xi_j &=\left(e^{\epsilon_j}+1\right)^{-1}, 
 \end{align}
+the corresponding eigenstates are the same. 
 XXXX
@@
 
-In expression 	\eqref{eqplusminus} we assigned a charge of $\pm1/2$ to each entanglement eigenstate depending on the eigenvalue ordering. 
-In order to generalize this to systems with longer-range hopping one needs to know how to properly assign this $\pm 1/2$ charge to each entanglement mode.
-We found that this is done according to their localization, so we have
+In expression 	\eqref{eqplusminus} we assigned a `charge'  of $\pm1/2$ to each entanglement eigenstate depending on the eigenvalue ordering, see also appendix~\ref{app:pollmann} and the related discussion in Ref.~\cite{Zaletel2014}.
+In order to generalize this to systems for which the ES is not equidistant  one needs to know how to properly assign this $\pm 1/2$ charge to each entanglement mode.
+We found that this is done according to their localization.
+We separate the eigenstates of the entanglement Hamiltonian (or alternatively the reduced correlation matrix) into a \emph{right} and \emph{left} set depending on whether $\expval{\hat{X}}_i > L/4$ or $\expval{\hat{X}}_i < L/4$ (region $A$ ranges from $0$ to $L/2$) and we assign the charges $q_r = +1/2$ and $q_l = -1/2$, respectively.
+ Expression \ref{eqplusminus} then generalizes to
 \begin{equation}
-\chi = \frac{1}{2} \left( \sum_{i\in r}\xi_i-\sum_{i\in l}\xi_i \pm \sum_{i\in b}\xi_i  + N_A \right) \quad {\rm mod} \quad 1,
-\label{loceq}
+\chi = \frac{1}{2} \left( \sum_{i\in r}\xi_i-\sum_{i\in l}\xi_i+ N_A \right) \quad {\rm mod} \quad 1.
+\label{eq_loc}
 \end{equation}
-by defining $ \chi$ modulo $1$ we make the charge of the bulk modes irrelevant but in the next section we will have to take a closer look at it.
-We find the charge of the edge modes by computing $\expval{\hat{X}}_i$ and seeing if it is above or below $L/4$.
-In the case the ES is double degenerate at most, left and right edge states might mix but one will always classify one state of the pair as left and the other one as right, this is however not the case when the degeneracies are higher than two-fold.
-In order to ensure that we obtain edge modes localized at one edge, instead of $C_A$ we diagonalize $C_A + \lambda C_A\hat{X}C_A$ (\carlos{Explain why it does not work in general}), where $\hat{X}$ is the position operator and $\lambda$ is a small parameter that vanishes in the thermodynamic limit.
 
-The reason expression~\eqref{loceq} simplifies so much for systems with only nearest-neighbor hopping is that it is known that in this case the eigenstates of $C_A$ always come with an alternating edge localization (left-right-left...), so that the signs assigned are always $ \pm(+-+-+-...)$.
-Their eigenvalues are also related, so that the modes can only cross each other if all of them do so two by two, so that it just ammounts to a global minus sign.
-
-This relation between the Zak phase and the entanglement spectrum is a consequence of an identity between the Zak phase and the many-body entanglement spectrum previously found for infinite chains \cite{Zaletel2014}.
-In Appendix A we modified this derivation to account for the periodic boundary conditions we use and show how their result reduces to Eq.~\eqref{loceq} when expressing it in terms of the ES.
+In order for this expression to be well-defined however we need localized eigenstates. 
+Non-degenerate eigenstates of the EOS are automatically localized, but degenerate or nearly-degenerate eigenstates (i.e. those exponentially close to $\xi=0,1$) are not necessarily localized. 
+It turns out that we need not worry about the latter set of states, because changing their charge from say $1/2$ to $-1/2$ only changes $1/2(\sum_{i\in r} \xi_i - \sum_{i\in l}\xi_i)$ in Eq.~\eqref{eq_loc} by an integer, which is canceled \mariac{better expression?} by mod 1.
+Also in the case where eigenstates are double degenerate we need not worry about localizing them, because computing the expectation value of the position operator will automatically assign the correct number of eigenstates to the left/right half of $A$.
+This is no longer true for four-fold, or higher-fold, degenerate states.
+%In Fig.\ref{huang}.a the doubly-degenerate states have contributions from both edges.
+Performing the above classification will always lead to assigning one state as \emph{left} and the other one as \emph{right} so there is no issue, but this is no longer true for four-fold, or higher, degenerate states.
+%The eigenstates close to $1$ and $0$ are also highly degenerate and thus delocalized over the whole chain.
+However in this case this is irrelevant since the difference in $\chi$ from changing the charge in one of these modes is either $0$ or $1$, which is cancelled anyway by taking the modulo $1$, so the expression \ref{loceq} is well-defined \carlos{up to exponentially small errors?}.
+In order to ensure that we obtain localized entanglement eigenstates,  we diagonalize $C_A + \lambda C_A\hat{X}C_A$ instead of $C_A$. 
+\mariac{do not commute 
+relation to wannier states
+how to choose lambda }
+where $\lambda$ is a parameter that must be high enough to split all degenerate states into localized ones but small enough to not perturbe the EOS. 
+%Note that this procedure is not necessary when the only degeneracy is at eigenvalues $1$ and $0$, like in Fig.
+\ref{huang}.b, as discussed above.
+\carlos{I need to explain a bit more how this works with the symmetries  ect.}
 
@@
 This happens because of how the two symmetry breaking terms split the four-fold degenerate midgap states.
-While $\kappa'$ splits them into two two-fold degenerate pairs whose contribution cancels, $\kappa$ split thems into two $left-left$ and $right-right$ pairs, with a finite contribution to the Zak phase.
-We also show the cut through the dotted line that shows all possible phases to illustrate how our results also apply to more complex ES.
+While $\kappa'$ splits them into two two-fold degenerate pairs at opposite edges whose contribution cancels, $\kappa$ split thems into two $left-left$ and $right-right$ pairs, with a finite contribution to the Zak phase.
+We also show the cut through the dotted line that shows all possible phases in the BDI class to illustrate how our results also apply to more complex EOS where the observations made in references \cite{Huang2012,Huang2012-2} do not apply anymore.
 
@@
 
+\mariac{issue about  the ambiguity about right - left vs left - right, connect with discussion about Zaletel results and their convention?}
+This relation between the Zak phase and the EOS is a consequence of an identity between the Zak phase and the many-body entanglement spectrum previously found for infinite chains \cite{Zaletel2014}.
+In Appendix A we modified this derivation to account for the periodic boundary conditions we use and show how their result reduces to Eq.~\eqref{eq_loc} when expressing it in terms of the EOS.
+
+\mariac{alternative way of computing Zak phase}
 Using the fact that $N_A = \Tr C_A$ we can rewrite $\chi$ as
 \begin{equation}
-\chi = \left(\sum_{i\ in r} \xi_i+ \sum_{i\in b}\xi_i\right) \, {\rm mod} \, 1.
+\chi = \left(\sum_{i\in r} \xi_i\right) \, {\rm mod} \, 1.
 \end{equation}
@@
 \end{equation}
-This method however will introduce a stronger finite size effect and as we will discuss below it only works for homogeneous systems.
-\carlos{I don't know if its worth it to introduce in this paper.
-Then I dont know what to do for the ES = winding.
-}
+This method however will introduce a stronger finite size effect and as we will discuss below it only works for homogeneous systems. \carlos{I don't know if its worth it to introduce in this paper. }
+
+Before proceeding with the inhomogeneous systems, let us explain why the expression~\eqref{eqplusminus} simplifies so much for systems with an equidistant ES. 
+First note that for translation invariant systems, the ES of the right and the left edge are related by an overall '-' sign, i.e.
+for each eigenvalue $\epsilon$ on the right edge, there is a corresponding value $-\epsilon$ on the left edge.
+Since the spectra are equidistant, this implies that the two edge spectra are merely shifted with respect to each other.
+Note that  the shift could be zero, e.g.  in presence of a symmetry. 
+Independent of the shift, the simple assignment of $ (...+-+-+-...)$ will automatically assign charges $\pm 1/2$ to each edge. 
+The only remaining issue is whether this simple recipe assigns charge $+1/2$ or $-1/2$ to the right edge.
+In presence of a symmetry, this question is irrelevant \mariac{better expression?} because of the degeneracy. 
+In general, it implies that $\chi$ as defined in Eq.~\eqref{eqplusminus} only gives the Zak phase up to an overall sign. 
+This issue can be fixed by determining the localization of a single eigenstate.
+
+%The reason is that, unless they are degenerate, the ordered eigenstates always come with alternating edge localization (...left-right-left-right...)\mariac{ref to peschel}, so that the signs assigned are always .
+%\mariac{expand}
+%\mariac{issue about crossing spectra?}
+%If the eigenvalues cross each other this only adds a global minus sign, however this unlikely and usually only happens when all eigenvalues are close to $0$ and $1$ and $\chi$ is therefore small.
 
@@
 
-\section{Appendix A}
+\section{Appendix A}\label{app:pollmann}
 

### 2020-04-29 draft.tex
..

@@
 The eigenvalues $\epsilon_j$ of $\mathcal{H}$, referred to as entanglement energies,  are related to the EOS by 
-\begin{align}
+\begin{align}\label{eq:xi_eps}
 \xi_j &=\left(e^{\epsilon_j}+1\right)^{-1}, 
@@
 the corresponding eigenstates are the same. 
-XXXX
-Thus, the full information of the \emph{many-particle} ES is fully contained in the single-particle spectrum of the reduced correlation matrix. 
+Eq.~\eqref{eq:xi_eps} implies that the full information of the \emph{many-particle} ES is contained in the single-particle spectrum of the reduced correlation matrix. 
 The latter is much simpler to interpret. 
@@
 We first discuss systems, for which the entanglement energies of a \emph{single} virtual edge are equidistant. 
-This is a common feature for e.g. nearest neighbor hopping models\mariac{ref to Peschel}, such as the SSH chain, and applies to the Hamiltonian~\ref{bdi_model} when $t'=0$. 
+This is a common feature for e.g. nearest neighbor hopping models\mariac{ref to Peschel}, such as the SSH chain, and applies to the Hamiltonian~\eqref{bdi_model} when $t'=0$. 
 In this case, one can obtain the Zak phase in a very simple fashion: We reorder the eigenvalues of the ES by magnitude, $\xi_1 < \xi_2 < ...$, and compute
@@
 Consequently, there is one less mode at $\xi=1$, which yields a difference of $\pi$ in the Zak phase. 
-Note that the full sum~\ref{eqplusminus}  evaluates to zero in the trivial regime, up to finite-size corrections, despite the presence of modes with entanglement occupancy $0<\xi<1$.
+Note that the full sum~\eqref{eqplusminus}  evaluates to zero in the trivial regime, up to finite-size corrections, despite the presence of modes with entanglement occupancy $0<\xi<1$.
 
@@
 }
-\caption{\mariac{system size?}Entanglement spectrum (black), $\chi$ (green) and $\frac{\gamma}{2\pi}$ (red) for a cut in the phase diagram through the $\nu = 1 \rightarrow \nu = 0$ transition (dashed line in 	\ref{fig:bdi_phase_diagram}
+\caption{\mariac{system size?}Entanglement spectrum (black), $\chi$ (green) and $\frac{\gamma}{2\pi}$ (red) for a cut in the phase diagram through the $\nu = 1 \rightarrow \nu = 0$ transition (dashed line in~\ref{fig:bdi_phase_diagram}
 	) for $t'=0$ and different values of $\kappa$.}
@@
 
-In expression 	\eqref{eqplusminus} we assigned a `charge'  of $\pm1/2$ to each entanglement eigenstate depending on the eigenvalue ordering, see also appendix~\ref{app:pollmann} and the related discussion in Ref.~\cite{Zaletel2014}.
+In expression~\eqref{eqplusminus} we assigned a `charge'  of $\pm1/2$ to each entanglement eigenstate depending on the eigenvalue ordering, see also appendix~\ref{app:pollmann} and the related discussion in Ref.~\cite{Zaletel2014}.
 In order to generalize this to systems for which the ES is not equidistant  one needs to know how to properly assign this $\pm 1/2$ charge to each entanglement mode.
@@
 We separate the eigenstates of the entanglement Hamiltonian (or alternatively the reduced correlation matrix) into a \emph{right} and \emph{left} set depending on whether $\expval{\hat{X}}_i > L/4$ or $\expval{\hat{X}}_i < L/4$ (region $A$ ranges from $0$ to $L/2$) and we assign the charges $q_r = +1/2$ and $q_l = -1/2$, respectively.
- Expression \ref{eqplusminus} then generalizes to
+ Expression \eqref{eqplusminus} then generalizes to
 \begin{equation}
@@
 
-\section{Appendix A}\label{app:pollmann}
+\section{Computing the Zak phase from the Schmidt decomposition}\label{app:pollmann}
+\maria{In this appendix, we review  how to compute the Zak phase from the Schmidt decomposition~\cite{Zaletel2014}and relate it to our own results of Eq.s~\eqref{eqplusminus} and \eqref{eq_loc}.}
+
+\mariac{notation ES/EOS consistent with main text?}
+
+\mariac{add some explanatory sentences?}
 
@@
 \end{align}
-And the charge can be assigned so that $Q_p = \sum_{\alpha = b,l,r} q_\alpha\sum_{i \in \alpha} n_i$ with the constrain that $q_r,\pm q_b = q_l-1$ \textcolor{red}{Why?}.
+And the charge can be assigned so that $Q_p = \sum_{\alpha = b,l,r} q_\alpha\sum_{i \in \alpha} n_i$ with the constrain that $q_r\pm q_b = q_l-1$ \textcolor{red}{Why?}.
 As the main result of \cite{Zaletel2014} the Zak phase is then found to be
@@
 
-\section{Correlation matrix}\label{app:CM}
+\section{Alternative expression for the correlation matrix}\label{app:CM}
+In this appendix, we show how to express the correlation matrix in terms of the Hamiltonian, used in Eq.~\eqref{eq:corr_mat2}. 
 We consider the generic, quadratic Hamiltonian in one dimension of Eq.~\ref{eq:quadr_Ham}, which is diagonalized by a unitary matrix $U$ with 
@@
 The second term can be rewritten, using Eq.~\ref{eq:D}, as
-\begin{align}\label{eq:C_rewritten}
+\begin{align}
 UD(\abs{D})^{-1}U^\dagger =& UDU^\dagger U(D^2)^{-1/2}U^\dagger
 \end{align}
-Consider
-\begin{align*}
-[U (D^2)^{-1/2} U^\dagger]^2 =& U (D^2)^{-1/2} U^\dagger U  (D^2)^{-1/2} U^\dagger \\
-=&  U (D^2)^{-1/2}  (D^2)^{-1/2} U^\dagger \\
-=&  U (D^2)^{-1} U^\dagger \\
-=&  (U D^2 U^\dagger)^{-1} \\
-\end{align*}
-Therefore
+We can rewrite this expression further by noting that 
+\begin{align}
+[U (D^2)^{-1/2} U^\dagger]^2 =& U (D^2)^{-1/2} U^\dagger U  (D^2)^{-1/2} U^\dagger \nonumber\\
+=&  U (D^2)^{-1/2}  (D^2)^{-1/2} U^\dagger \nonumber\\
+=&  U (D^2)^{-1} U^\dagger \nonumber\\
+=&  (U D^2 U^\dagger)^{-1}.
+\end{align}
+Therefore,
 \begin{align*}
@@
 \end{align*}
-and we have
-\begin{align}\label{eq:C_rewritten}
-UD(\abs{D})^{-1}U^\dagger =& UDU^\dagger(U D^2 U^\dagger)^{-1/2} \\ 
+and we conclude that
+\begin{align}\label{eq:UDDU}
+UD(\abs{D})^{-1}U^\dagger =& UDU^\dagger(U D^2 U^\dagger)^{-1/2} \nonumber\\ 
 =& H(H^2)^{-1/2}.
 \end{align}
+Combining Eq.s~\eqref{eq:corr_mat},  and \eqref{eq:UDDU}, we arrive at the final expression of the correlation matrix in Eq.~\eqref{eq:corr_mat2}.
 

### 2020-05-07 draft.tex
..

@@
 \subsubsection{Systems with equidistant entanglement energies}
+\mariac{add scaling plot for $\chi$ versus $\gamma/(2\pi)$. }
 We first discuss systems, for which the entanglement energies of a \emph{single} virtual edge are equidistant. 
@@
 However in this case this is irrelevant since the difference in $\chi$ from changing the charge in one of these modes is either $0$ or $1$, which is cancelled anyway by taking the modulo $1$, so the expression \ref{loceq} is well-defined \carlos{up to exponentially small errors?}.
+
 In order to ensure that we obtain localized entanglement eigenstates,  we diagonalize $C_A + \lambda C_A\hat{X}C_A$ instead of $C_A$.
-\mariac{do not commute 
-relation to wannier states
-how to choose lambda }
-where $\lambda$ is a parameter that must be high enough to split all degenerate states into localized ones but small enough to not perturbe the EOS. 
-%Note that this procedure is not necessary when the only degeneracy is at eigenvalues $1$ and $0$, like in Fig.
-\ref{huang}.b, as discussed above.
-\carlos{I need to explain a bit more how this works with the symmetries  ect.}
+For the whole system, one way of obtaining localized eigenstates of the full correlation matrix is to diagonalize $C\hat{X}C$, which results in localized Wannier states. 
+$C$ commutes with this operator because it is a projector, but this is not necessarily the case for $C_A$ and $C_A\hat{X}C_A$. 
+Far from the edges, however, $C_A$ and $C$ are equivalent, so that the states at $\xi=1$ will localize into Wannier states by adding the $\lambda C_A\hat{X}C_A$ term.
+The rest of the eigenstates will also localize because this term, as it splits the degeneracies. 
+But as it does not commute with $C_A$ close to the virtual edges,  it will modify the eigenstates. 
+For that reason $\lambda$ must be a small parameter. \mariac{how to scale $\lambda$ with system size}. 
+%\mariac{do not commute 
+%	relation to wannier states
+%	how to choose lambda }
+%where $\lambda$ is a parameter that must be high enough to split all degenerate states into localized ones but small enough to not perturbe the EOS.
 
-In Fig.\ref{2} we show an example that highlights the importance of the localization structure of the ES.
-For the solid line cut in the phase-diagram, with the $\nu = 2 \rightarrow 0 $ transition, two ES that seem at first very similar result in very different Zak phases.
-This happens because of how the two symmetry breaking terms split the four-fold degenerate midgap states.
-While $\kappa'$ splits them into two two-fold degenerate pairs at opposite edges whose contribution cancels, $\kappa$ split thems into two $left-left$ and $right-right$ pairs, with a finite contribution to the Zak phase.
-We also show the cut through the dotted line that shows all possible phases in the BDI class to illustrate how our results also apply to more complex EOS where the observations made in references \cite{Huang2012,Huang2012-2} do not apply anymore.
+In Fig.\ref{2} we show two examples that highlight the importance of the localization structure of the ES. 
+In subfigures (a) and (b), we compute the EOS for varying $m$ at  $t=1$, $t'=2$, indicated by the solid line  in the phase-diagram~\ref{fig:bdi_phase_diagram}, in the presence of two different symmetry breaking terms: (i) $\kappa =0.3$, $\kappa'=0$ and (ii) $\kappa=0$, $\kappa'=0.3$.
+For $\kappa=\kappa'=0$, there is a phase transition $\nu = 2 \rightarrow 0 $ at $m=2$, whereas both choices of symmetry breaking render the system trivial  for all $m$.
+However, the two symmetry breaking terms split the four-fold degenerate $\xi=1/2$ midgap states very differently, thus resulting in completely different Zak phases despite the superficial resemblance of their respective EOS.
+While $\kappa'$ splits the four-fold degenerate states into two two-fold degenerate pairs at opposite edges whose contribution cancels, $\kappa$ split them into two $left-left$ and $right-right$ pairs, with a finite contribution to the Zak phase.
+In subfigures (c) and (d), we consider the same symmetry breaking terms, but now for $t=1$ and $t'=-2$, where the BDI systems shows a phase transition  $\nu=2\rightarrow 1\rightarrow 0$. 
+Note that the $\kappa'$ term does not render the $\nu=1$ phase trivial, and the Zak phase will remain quantized in that region, see Fig.~\ref{2}(d). 
+Also here, Eq.~\ref{eq_loc} correctly reproduces the Zak phase for all values of $m$. 
+Note that in all examples shown in this subsection, the observations made in references \cite{Huang2012,Huang2012-2}, i.e.
+that the behavior of the Zak phase resembles that of the entanglement occupancy midgap states, do not apply anymore.
 
@@
 
-\mariac{alternative way of computing Zak phase}
-Using the fact that $N_A = \Tr C_A$ we can rewrite $\chi$ as
+Let us also discuss an alternative way of computing Zak phase. 
+Using the fact that $N_A = \Tr C_A$, we can rewrite $\chi$ as
 \begin{equation}
@@
 \end{equation}
-A simpler way one can compute this quantity is by obtaining $C_A$ for an open chain, labeled as $C_A^{ o}$.
-For a long enough chain the right eigenvalues will remain the same, as we still place a virtual cut in the middle of the chain but the left eigenvalues will no longer be present, so that one can simply compute
+A simpler way to compute this quantity is by obtaining $C_A$ for an open chain, labeled as $C_A^{ o}$. 
+For a sufficiently long  chain the right eigenvalues will remain the same, %as we still place a virtual cut in the middle of the chain 
+but the left eigenvalues will no longer be present. 
+This allows one to simply compute
 \begin{equation}
@@
 \end{equation}
-This method however will introduce a stronger finite size effect and as we will discuss below it only works for homogeneous systems. \carlos{I don't know if its worth it to introduce in this paper. }
+This method, however, will introduce a stronger finite size effect and requires substantially larger systems for the same accuracy. 
+In addition, it only works for homogeneous systems as explained below.
 
@@
 In reference \cite{Zaletel2014} translational invariance was used to derive the relation between the many-body entanglement spectrum and the Zak phase.
+If this symmetry is broken by, for example, by adding weak disorder to the system, this will modify slightly the eigenvalues of the reduced density matrix but not the charge assignment in Eq.~\eqref{zaletel} in the Appendix.
+The left side of Eq.~\eqref{zaletel} will be cut independent while the right side is not, and therefore cannot be correct.
+
+We are interested therefore in finding a cut-independent quantity from the EOS that might still result in the Zak phase.
+The first thing one can do is to do an average over the different cuts.
+Naively one would cut-average the expression in Eq.~\eqref{eq_loc}, however the average of a circular quantity is not well defined.
+We defined $\chi$ modulo 1 because it is a phase, but it also served the purpose of getting rid of arbitrary charge assignment of some of the modes close to 1, which is problematic.
+We define
+\begin{equation}
+\chi_s = \frac{1}{2} \left( \sum_{i\in r}\xi^s_i-\sum_{i\in l}\xi^s_i  + N_A \right),
+\end{equation}
+obtained for cuts at sites $s$ and $s+L/2$. The cut-average is then given by
+\begin{equation}
+\bar{\chi} = \left(\frac{1}{L}\sum_{s=1}^L \chi_s \right)\, \rm{mod} \, 1.
+\label{batcutavg}
+\end{equation}
+As we move through the different cuts the average position of  the eigenstates at 1 can move slightly, so that the number of states at each half of the system depends on the cut, which will affect the cut-average.
+In order to obtain the correct result of  the contribution of the states at $\xi=1$ should be constant across cuts, so that it can be correctly cancelled by the modulo 1.
+Therefore we have to find a way of correcting for jumps of $1$ in $\chi_s$ when a mode at one crosses the $L/4$.
+
+%
+%In reference \cite{Zaletel2014} translational invariance was used to derive the relation between the many-body entanglement spectrum and the Zak phase.
 If this symmetry is broken by, for example, adding weak disorder to the system, this will modify slightly the eigenvalues of the reduced density matrix but not the charge assignment in Eq.~\eqref{zaletel} in the Appendix.
 The left side of Eq.~\eqref{zaletel} will be cut independent while the right side is not, and therefore cannot be correct.
-
-We are interested therefore in finding a cut-independent quantity from the ES that might still result in the Zak phase.
+%
+%We are interested therefore in finding a cut-independent quantity from the ES that might still result in the Zak phase.
 The first thing one can do is to do an average over the different cuts.
@@
 As we do for the edge state, we can choose a sign for the bulk modes in the following way
-\begin{equation}
-g_i=
-\begin{cases}
- +1 &,\rm{if} \expval{\hat{X}}_i > L/4  \\
- -1 &,\rm{if} \expval{\hat{X}}_i < L/4.
-\end{cases}
-\end{equation}
-So that now we define
-\begin{equation}
-\chi_s = \frac{1}{2} \left( \sum_{i\in r}\xi^s_i-\sum_{i\in l}\xi^s_i + \sum_{i\in b}g_i^s \xi^s_i  + N_A \right),
-\end{equation}
-obtained for cuts at $s$ and $s+L/2$ and the cut-average is given by
-\begin{equation}
-\bar{\chi} = \frac{1}{L}\sum_{s=1}^L \chi_s.
-\label{batcutavg}
-\end{equation}
-By adding the perturbation proportional to $C_A\hat{X} C_A$ the bulk modes also localize.
+%\begin{equation}
+%g_i=
+%\begin{cases}
+% +1 &,\rm{if} \expval{\hat{X}}_i > L/4  \\
+% -1 &,\rm{if} \expval{\hat{X}}_i < L/4.
+%\end{cases}
+%\end{equation}
+%So that now we define
+%\begin{equation}
+%\chi_s = \frac{1}{2} \left( \sum_{i\in r}\xi^s_i-\sum_{i\in l}\xi^s_i + \sum_{i\in b}g_i^s \xi^s_i  + N_A \right),
+%\end{equation}
+%obtained for cuts at $s$ and $s+L/2$ and the cut-average is given by
+%\begin{equation}
+%\bar{\chi} = \frac{1}{L}\sum_{s=1}^L \chi_s.
+%\label{batcutavg}
+%\end{equation}
+%By adding the perturbation proportional to $C_A\hat{X} C_A$ the bulk modes also localize.
 As we move through the different cuts the average position of the bulk modes moves slightly.
@@
 \subsubsection{Continuous inhomogeneity}
+$\chi$ does not have to be continuous, but jumps of 'integers' can be compensated by hand.
 
@@
 \end{align}
-And the charge can be assigned so that $Q_p = \sum_{\alpha = b,l,r} q_\alpha\sum_{i \in \alpha} n_i$ with the constrain that $q_r\pm q_b = q_l-1$ \textcolor{red}{Why?}.
+And the charge can be assigned so that $Q_p = \sum_{\alpha = l,r} q_\alpha\sum_{i \in \alpha} n_i$ with the constrain that $q_r= q_l-1$ \textcolor{red}{Why?}.
 As the main result of \cite{Zaletel2014} the Zak phase is then found to be

### 2020-05-27 draft.tex
Update 27-5
Updated Fig. 5 with a case where the simple method does not work and added a bit of text to go with it. Update Fig.6 and rewrote parts of Appendix C with added new figures that we should discuss.

@@
 \chi = \left(\sum_{i\in r} \xi_i\right) \, {\rm mod} \, 1.
+\label{eq:chihom}
 \end{equation}
@@
 In reference \cite{Zaletel2014} translational invariance was used to derive the relation between the many-body entanglement spectrum and the Zak phase.
-If this symmetry is broken by, for example, by adding weak disorder to the system, this will modify slightly the eigenvalues of the reduced density matrix but not the charge assignment in Eq.~\eqref{zaletel} in the Appendix.
-The left side of Eq.~\eqref{zaletel} will be cut independent while the right side is not, and therefore cannot be correct.
-
-We are interested therefore in finding a cut-independent quantity from the EOS that might still result in the Zak phase.
-The first thing one can do is to do an average over the different cuts.
-Naively one would cut-average the expression in Eq.~\eqref{eq_loc}, however the average of a circular quantity is not well defined.
-We defined $\chi$ modulo 1 because it is a phase, but it also served the purpose of getting rid of arbitrary charge assignment of some of the modes close to 1, which is problematic.
-We define
+If this symmetry is broken by adding, for example, weak disorder, this result is no longer true.
+However we found, as our main result, that one can still obtain the Zak phase as the cut-average of the relation derived for homogeneous systems.
+For some systems this can be written as (\carlos{not sure about this phrasing})
+\begin{align}
+\gamma =\left(\frac{1}{L}\sum_{s=1}^L (\chi_s \, \rm{mod} \, 1)\right)\, \rm{mod} \, 1,
+\label{eq:inh_simple}
+\end{align}
+where
 \begin{equation}
-\chi_s = \frac{1}{2} \left( \sum_{i\in r}\xi^s_i-\sum_{i\in l}\xi^s_i  + N_A \right),
+\chi_s = \frac{1}{2} \left( \sum_{i\in r}\xi^s_i-\sum_{i\in l}\xi^s_i  + N_A \right)
 \end{equation}
-obtained for cuts at sites $s$ and $s+L/2$. The cut-average is then given by
-\begin{equation}
-\bar{\chi} = \left(\frac{1}{L}\sum_{s=1}^L \chi_s \right)\, \rm{mod} \, 1.
-\label{batcutavg}
-\end{equation}
-As we move through the different cuts the average position of  the eigenstates at 1 can move slightly, so that the number of states at each half of the system depends on the cut, which will affect the cut-average.
-In order to obtain the correct result of  the contribution of the states at $\xi=1$ should be constant across cuts, so that it can be correctly cancelled by the modulo 1.
-Therefore we have to find a way of correcting for jumps of $1$ in $\chi_s$ when a mode at one crosses the $L/4$.
-
-%
-%In reference \cite{Zaletel2014} translational invariance was used to derive the relation between the many-body entanglement spectrum and the Zak phase.
-If this symmetry is broken by, for example, adding weak disorder to the system, this will modify slightly the eigenvalues of the reduced density matrix but not the charge assignment in Eq.~\eqref{zaletel} in the Appendix.
-The left side of Eq.~\eqref{zaletel} will be cut independent while the right side is not, and therefore cannot be correct.
-%
-%We are interested therefore in finding a cut-independent quantity from the ES that might still result in the Zak phase.
-The first thing one can do is to do an average over the different cuts.
-Naively one would cut-average the expression in Eq.~\eqref{loceq}, however the average of a circular quantity is not well defined.
-We defined $\chi$ modulo 1 because the Zak phase is also defined modulo 1, but it also served the purpose of getting rid of the sign choice for the bulk modes, which is problematic.
-As we do for the edge state, we can choose a sign for the bulk modes in the following way
-%\begin{equation}
-%g_i=
-%\begin{cases}
-% +1 &,\rm{if} \expval{\hat{X}}_i > L/4  \\
-% -1 &,\rm{if} \expval{\hat{X}}_i < L/4.
-%\end{cases}
-%\end{equation}
-%So that now we define
-%\begin{equation}
-%\chi_s = \frac{1}{2} \left( \sum_{i\in r}\xi^s_i-\sum_{i\in l}\xi^s_i + \sum_{i\in b}g_i^s \xi^s_i  + N_A \right),
-%\end{equation}
-%obtained for cuts at $s$ and $s+L/2$ and the cut-average is given by
-%\begin{equation}
-%\bar{\chi} = \frac{1}{L}\sum_{s=1}^L \chi_s.
-%\label{batcutavg}
-%\end{equation}
-%By adding the perturbation proportional to $C_A\hat{X} C_A$ the bulk modes also localize.
-As we move through the different cuts the average position of the bulk modes moves slightly.
-For most modes this will not affect the assigned sign, but for the modes near the middle of subsystem $A$ it might be that for strong enough inhomogeneity that the mode crosses $L/4$ and the sign changes.
-We find that in order for us to get the correct result the contribution from the bulk modes should be constant across cuts, so we have to find a way of correcting for these jumps of $1$ in $\chi_s$.
-
-\subsubsection{Continuous inhomogeneity}
-$\chi$ does not have to be continuous, but jumps of 'integers' can be compensated by hand.
-
-In the case we have a continuous inhomogeneity one can obtain a new quantity $\chi'_s$ with a constant bulk contribution, by assuming that the discontinuities of $\chi_s$ above some threshold, $h$, are due to these unphisical changes of sign for the bulk states and eliminating them by shifting it by an integer number so that
-\begin{equation}
-\chi_{s+1}'-\chi'_s < h
-\end{equation}
-and it is continuous.
-$\chi'_s$ can now be cut-averaged to obtain the Zak phase.
-In figure \ref{inh} we show this for a continuous profile for $\kappa_s$.
-We see that $\chi_s$ in fact presents jumps of $1$ but once we correct them in $\chi'_s$, its cut average results in the Zak phase.
-We also see the same for a disordered $\kappa_s$.
-Even though the disorder is relatively strong we see that $\chi_s'$, after the artificial jumps are eliminated, is fairly continuous, which allows us to use this method even for strong inhomogeneity, as long as we can identify this jumps.
+is obtained for cuts at sites $s$ and $s+L/2$.
+In Fig.\ref{fig:simple_inh}.b we show the same system as in Fig.\ref{2}.c with an added moderate disorder in the $\kappa$ parameter shown in \ref{fig:simple_inh}.a .
+We show the EOS of one cut with that particular $\chi_s$, which differs from the Zak phase.
+The Zak phase is shown to be recovered by the cut-average of $\chi$.
+Next we show the results for the same system with $m=-2$, and a continuous profile for $\kappa$ shown in \ref{fig:simple_inh}.c.
+In Fig.
+\ref{fig:simple_inh}.d is shown how the EOS and $\chi_s$ depends on the cuts, and how the Zak phase is again recovered by the cut-averaged $\chi$.
 
 \begin{figure}[h!]
-
 \centering
 \makebox[0pt]{
-\subfloat[]{
-  \includegraphics[width=40mm]{4_a.pdf}
-  \label{inha}
+\subfloat[$\kappa_i$]{
+  \includegraphics[width=35mm]{4cdisorderkap.pdf}
 }
-\subfloat[ ]{
-  \includegraphics[width=40mm]{4_b.pdf}
+\subfloat[$t=1,t'=-2,s=0$]{\includegraphics[width=35mm]{4cdisorderES.pdf}
 }
@@
 \makebox[0pt]{
-\subfloat[]{
-  \includegraphics[width=40mm]{4_c.pdf}
+\subfloat[$\kappa_i$]{
+  \includegraphics[width=35mm]{simplecutkap.pdf}
 }
-\subfloat[]{
-  \includegraphics[width=40mm]{4_d.pdf}
+\subfloat[$t=1,t'=-2,m=2$]{\includegraphics[width=35mm]{simplecutES.pdf}
 }
@@
 \makebox[0pt]{
-\subfloat[]{
-  \includegraphics[width=40mm]{4_e.pdf}
+\subfloat[$\kappa_i$]{
+  \includegraphics[width=35mm]{cutavgwrongkap.pdf}
 }
-\subfloat[]{
-  \includegraphics[width=40mm]{4_f.pdf}
+\subfloat[$t=1,t'=-2,m=1$]{\includegraphics[width=35mm]{cutavgwrong.pdf}
 }
 }
-
-\caption{Using the parameters \label{inh}$t=1,t'=2,m =1$, we show our results for two different profiles for $\kappa_s$, one continuous (a) and one with disorder (b).
-In (c) and (d) we show the ES (black) and $ \chi_s \, {\rm mod} \, 1$ (green).
-In (e) and (f) we show $\chi_s$ (green), $\chi'_s$ (blue dots), $\gamma$ (red) and $\expval{\chi'_s}_{\rm cuts}$ (blue dashed) }
-
+\caption{In a) and b) we show the system in Fig.
+4.c with an added disorder.
+we show the ES (black) and $ \chi_s \, {\rm mod} \, 1$ (green) for one particular cut,$\gamma$ (red) and $\expval{\chi_s}_{\rm cuts}$ (blue dashed).
+In c) and d) we show the same system at $m=2$ and a continuous homogeneous profile for $\kappa_i$.
+In e) and f) we show a case where using \eqref{eq:inh_simple} fails.
+We show the ES and $\chi_s$ for all cuts.
+}
+\label{fig:simple_inh}
 \end{figure}
 
-In some cases, this calculation simplifies a bit more.
-Note that, in the case where one can tell a priori that the width of $\chi'_s$ is lower than 1, as in the example with disorder, to recover $\chi_s'$ one can simply take the modulo 1 and adding an offset, so for instance in the example with disorder we can obtain
-\begin{equation}
-\chi'_s = ([\chi_s-1/2] \, {\rm mod} \, 1)+1/2.
-\end{equation}
-One can then cut-average the resulting $\chi_s \, \rm{mod} \,1$ in the same way to obtain the Zak phase.
-This cannot be done in the first example because as we already see in Fig.\ref{inh}.c, $\chi_s \, \rm{mod} \,1$ wraps aroud the full range, so that there is no offset that will allow us to average it in a well-defined way.
+We have shown two cases were using expression \ref{eq:inh_simple} leads to the correct result.
+The average of a circular quantity is in fact not uniquely defined, and the average we do in Eq.\ref{eq:inh_simple}, which equates to averaging the valus in $(0,1)$ around $1/2$ (\carlos{is this clear?}), is just one choice that works in this cases.
+However in Fig.\ref{fig:simple_inh}.f this cut-average fails to give the Zak phase.
+This happens because by taking $\chi$ modulo 1 we have thrown out relevant information.
+To avoid this issue we can instead take the cut-average 
+\begin{align}
+\gamma =\left(\frac{1}{L}\sum_{s=1}^L \chi_s \right)\, \rm{mod} \, 1,
+\label{eq:inh_fixed}
+\end{align}
 
-In Fig.\ref{disorder} we show $1-\expval{\chi_s}_{\rm cuts}/\gamma$ for different system lengths for system with weak disorder, as well as the disorder average of $1-\chi_0/\gamma$.
-For both cases the error seems to vanish in the thermodynamic limit.
-For small systems both errors are very similar, which suggests that it comes strictly from a systematic finite size error in the computation of either the Zak phase or the ES which is system-independent.
-For larger systems the disoder average is more dispersed, probably due to the number of configurations considered not being high enough.
+but this however has to be done carefully.
+So far we have naively separated our eigenstates into \emph{left} and \emph{rigth} sectors, assigned the charges acordingly, and took the modulo 1 to eliminate the contribution of the bulk modes that localize at the center of A.
+The problem resides in the classification of these bulk modes.
+Even if the rest of the EOS remains constant, due to the inhomogeneity a bulk mode can cross the middle of region A, as shown in Fig.\ref{fig:loc_fix}, which will change $\chi$ by $1/L$, even though choosing $L/4$ as the separation between the \emph{left} and \emph{rigth} sectors is not physical.
+In order for \ref{eq:inh_fixed} to work properly the contribution from the bulk modes, which should vanish in the end, must remain constant across the different cuts.
+In order to achieve this we choose a cut-dependent threshold that the bulk modes do not cross, $l_s$, as the mean position of two of the eigenstates close to this middle point of A, as shown in Fig.\ref{fig:loc_fix}.
+This is done in practice for the Wannier states (\carlos{Do I explain exactly how do I do it numerically?}).
+Note that in the thermodynamic limit it does not matter if the position of these eigenstates cross as they are degenerate, the only important thing is that the number of bulk modes at each sector remains constant (\carlos{Not sure about this phrasing either}).
 
-\begin{figure}[h]
+\begin{figure}[h!]
 \centering
-\makebox[0pt]{
-\subfloat[$t = 1, t' = 2, m=1, \kappa_s \in (-1,1)$]{
-  \includegraphics[width=65mm]{disvscut.pdf}
-}
-}
-\caption{We show $1-\expval{\chi_s'}_{\rm cuts}/\gamma$ (red) and $1-\expval{\chi_s'}_{\rm dis}/\expval{\gamma}_{\rm dis}$.
-The cut average seems to do as good or better than the disorder average.
-For small systems it seems that both errors are similar, pointing to a systematic finite size error in the computation of either the Zak phase or the ES.
-For larger systems the error in the disorder average is more spread, which suggests that one should average more configurations.}
-\label{disorder}
+\includegraphics[width=70mm]{loc_fix.pdf}
+\caption{Average position of the Wannier states in red and, slightly shifted, the localized entanglement eigenstates (\carlos{rename $C_A+\lambda C_A X C_A$?} in blue.
+In black we show the threshold $l_s$ obtained as the mean value of the two Wannier states close de $19$ and $20$) $L=80,\lambda=10^{-6}, t = 1, t'= 2, m=0$}
+\label{fig:loc_fix}
 \end{figure}
-
-\subsubsection{Wannier states, L/4 crossing fix}
-
-\begin{figure}[t]
-\centering
-\makebox[0pt]{
-\subfloat[$\expval{\hat{X}}_{i,s}$ for different cuts, for the entanglement eigenstates such that $\xi_i < 1/2$ (black) and the Wannier states (red)]{
-  \includegraphics[width=90mm]{loca.pdf}
-}}\hspace{0mm}
-
-\makebox[0pt]{
-\subfloat[$\chi_s$]{
-  \includegraphics[width=45mm]{wan1.pdf}
-}
-\subfloat[$\chi_s'$ ]{
-  \includegraphics[width=45mm]{wan2.pdf}
-}
-}
-
-\caption{For the same parameters as used in Fig.5.a,c,e we show the expectation value of the position for the ES eigenstates (black) and the Wannier states (red), showing that the latter are a good approximation of the former in the bulk.
-We highlight by a continuous red line the bulk mode that crosses $L/4$ which causes the jumps in $\chi_s$.
-In (b) we show again $\chi_s$ (black) and the correction for the sign switch (red).
-In (c) $\chi_s'$ is reconstructed via this method, showing the same result as we had before.}
-\label{wan}
-\end{figure}
-
-In a general case where the discontinuities are too strong and we do not know how to correct for these jumps as explained before one has to take a closer look into the bulk states to ensure a consistent sign assignment.
-As we mentioned above the bulk states farthest away from the edges are the problematic ones.
-They are also the ones best approximated by the Wannier states obtained from the chain with periodic boundary conditions, as the eigenstates of
-\begin{equation}
-C\hat{X}_{\rm PBC}C
-\end{equation}
-where $\hat{X}_{\rm PBC} = \exp\{i 2\pi \hat{X}/L\}$.
-Similar to when we go through different cuts labeled by $s$, we now consider the Wannier states dependent on where we define our lattice site 1, shifted by $s$.
-The obtained Wannier states for $s$ and $s+1$ will be therefore exactly the same, but shifted one lattice site.
-By considering their average position $\expval{\hat{X}_{PBC}}_{i,s}$ we can define a set of $2L$ Wannier states that depend continuously on $s$, provided that the Hamiltonian is continuous.
-These Wannier states are then a good approximation for the Bulk states of $C_A$ we find for the different cuts.
-Consider now the quantity $\sum_i\expval{\hat{X}}_{i,s}$.
-So far everything was $s$ independent as our system was periodic, but by considering the operator $\hat{X}$ we introduce an unphysical boundary so that it will depend on how the lattice sites are defined.
-If the states do not cross this unphysical boundary then $\sum_i\expval{\hat{X}}_{i,s} = \sum_i\expval{\hat{X}_{\rm PBC}}_{i,s} $ and it will be a constant with $s$.
-However, when a Wannier state crosses the boundary we obtain the same constant plus $L$.
-If we put this boundary at $L/4$ the quantity
-\begin{equation}
-\delta_s = \frac{1}{L}\left(\sum_i\expval{\hat{X}}_{i,s}-\sum_i\expval{\hat{X}}_{i,0}\right)
-\end{equation}
-will be 1 (0) when a states has crossed the boundary or 0 (-1) when it has not, depending on the number of states in A at $s=0$, which we take as reference.
-Depending which state we take as reference will introduce a global integer shift for all cuts, so it is not relevant as it will go away after we average and take the modulo 1.
-By adding this term to our cut-average we will compensate for when our bulk state in the middle of A crosses $L/4$ and its charge changes, leading to a continuous $\chi_s'$ which can be averaged,
-\begin{equation}
-\chi_s' = \chi_s + \delta_s
-\end{equation}
-
-In Fig.\ref{wan} we show how this works for the same continuous example as in \ref{inh}.
-In Fig.\ref{wan}.a we show the expectation value of the position operator for both the entanglement eigenstates (in black) and the Wannier states (in red) for different values of $s$.
-In Fig.\ref{wan}.b and c we show $\chi_s$, which displays jumps due to a bulk state crossing $L/4$, and $\delta_s$, showing that it properly corrects the sign switch.
-
-This method conceptually works perfectly whenever the bulk states of the ES are equal to the Wannier states, i.e.
-in the thermodynamic limit, but it is not clear what the small parameter is.
-It depends on the $\lambda$ used (it should be small enough to not affect ES but large enough to localize the ES eigenstates) but also on the system.
-Systems with many ES edge modes close to the bulk tend to have a higher finite size effect.
-
-We also find that whenever there is an interphase in our system, and therefore a zero mode in the energy spectrum, the Wannier states are not necesseraly a good approximation of the bulk of finite systems.
-However, in the cases where we have an interphase but $\chi_s$ is fairly continuous, so that we can correct these jumps, cut-averaging results again in the Zak phase.
 
@@
 By adiabatically tuning down the coupling between the $L$ and $1$ sites we see that the eigenvalues of the $left$ modes flow to either $0$ or $1$.
-In the case where this eigenvalues oscillate around $1/2$ for different cuts this leads to discontinuities in the number of bulk modes at $\xi=1$ which again result in an incorrect cut-average and we found no practical way of compensating for these discontinuitiesyfgh.
+This can lead to discontinuities in the number of bulk modes at $\xi=1$ which again result in an incorrect cut-average.
+\carlos{Works in some cases.}
 
@@
 
-The proof in Ref.\cite{Zaletel2014} also breaks down for superconducting systems.
 However, as we did for systems without translational invariance, one can extend this identity for systems with pairing terms as well.
@@
 
+\section{$\lambda$ term}
+
+In order to ensure that we obtain localized entanglement eigenstates,  we diagonalize $C_A^\lambda = C_A + \lambda/L^2 C_A\hat{X}C_A$ instead of $C_A$, where $\lambda$ is our small parameter and the reason we include the $L^2$ scaling will be explained below.
+For the whole system one way of obtaining localized eigenstates of the full correlation function (\carlos{also the Hamiltonian?}) is to diagonalize $C\hat{X}C$, which results in localized Wannier states.
+$C$ commutes with this operator because it is a projector, but this is no longer the case for $C_A$.
+Far from the edges however, $C_A$ and $C$ are equivalent, so that the states close to $\xi=1$ will perfectly localize into Wannier states by adding the $\lambda$ term.
+Note that in the thermodynamic limit the operator $\hat{X}$ acts as proportional to an identity for localized states.
+Therefore in this limit the localized eigenstates of $C_A$ far from $\xi = 0,1$ are also eigenstates of $C_A^\lambda$ while their mixture is not.
+Threfore this operator breaks all degeneracies of $C_A$ and results in localized eigenstates.
+
+One has to be careful when choosing the correct $\lambda$.
+A parameter $\lambda$ which is too small will fail to break the degeneracies and localize the eigenstates.
+On the other hand the $\lambda$ term will  modify the eigenvalues and introduce an additional error to $\chi$.
+Assuming localized eigenstates we have $\expval{\hat{X}}_{i\alpha} \simeq i$.
+Therefore the error it introduces in $\chi$ is roughly
+\begin{equation}
+\frac{\lambda}{L^2} \sum_{\alpha=1,2}\left( \sum_{i=L/4+1}^{L/2}i-\sum_{i=1}^{L/4}i \right) = \frac{\lambda }{8},
+\end{equation}
+it is proportional to $\lambda$ and independent on system size.
+In Fig.\ref{lambdaL} we show this error in $\chi$ together with the maximum width of the eigenstates in the top half of the spectrum (\carlos{as we only care about the midgap states and the bulk modes at 1}), for different system sizes.
+We see that indeed the error remains constant with system size.
+In this case most of the bulk modes are above $\xi = 1-10^{-10}$, so that the error coming from the $\lambda$ term is bigger than the one we make by failing to localize the bulk modes.
+However, by looking at the maximum width of the eigenstates we see three different regions.
+One region at low system size where the width increases linearly.
+A second region where the system localizes and the width seems constant, and finally for large systems a third region where the width increases again because the $\lambda$ term is too small to break the degeneracies and localize the eigenstates.
+
+\begin{figure*}[h!]
+\centering
+\makebox[0pt]{
+\subfloat[${\rm Log}_{10}(\chi - \gamma)$ (\carlos{There are errors at $-5,-9,-12$ because that's the order of the eigenvalue of the bulk modes that should cancel pairwise but dont due to failure of localization.
+Furthermore, the error in yellow around $L=150$ is due to a mode close  to $\xi = 0$ so that it does not appear in Fig.
+b.})]{
+  \includegraphics[width=55mm]{lambdaL1.pdf}
+}
+\subfloat[${\rm Max}\bra{\psi_i}\sigma_x\ket{\psi_i}$ for $\xi_i >1/2$.
+\carlos{I don't know why the green line localizes so bad.
+I have to look further into it}]{\includegraphics[width=55mm]{lambdaL2.pdf}
+}
+}
+\caption{$t = 1, t'= 2, m=0$ for $-{\rm Log}_{10}[\lambda] = 4,5,6,7,8,9,10$.}
+\label{lambdaL}
+\end{figure*}
+
+In Fig.\ref{linearRegion} we take a closer look into the first region with a maximum width proportional to the system size.
+What happens in this region is that due to finite size the midgap modes that localize in the edges overlap so the resulting eigenstates are not localized and are split from $\xi = 1/2$.
+The boundary where this region ends changes with $\lambda$ and coincides with the point where the splitting from the $\lambda$ term is bigger than the splitting due to finite size.
+The gap between these midgap modes that mix starts decreasing with larger system size up to a point, where the system starts to localize, where this splitting is roughly constant with system size and proportional to $\lambda$.
+In the $\nu = 1$ region the two midgap states are degenerate and we do not observe this first linear region, the eigenstates localize for any finite $\lambda$.
+
+\begin{figure*}[h!]
+\centering
+\makebox[0pt]{
+\subfloat[Width of the eigenstate closest to $\xi=1/2$]{
+  \includegraphics[width=55mm]{linearBoundary.pdf}
+}
+\subfloat[Gap between the two eigenstates closest to $\xi = 1/2$]{\includegraphics[width=55mm]{linearBoundarysplit.pdf}
+}
+}
+\caption{$t = 1, t'= 2, m=0, \nu=0 , -{\rm Log}_{10}[\gamma] = 4,5,6,7,8,9,10$ for the different colors. Vertical lines are addded to better show how the features of both plots match.}
+\label{linearRegion}
+\end{figure*}
+
+For inhomogeneous systems the story is slightly different.
+Due to the cut-average a wrong assignment of the bulk modes leads to an error of order $\mathcal{O}(1/L)$.
+In Fig.
+\ref{lambdainh} we have the error made in $\chi$ and the maximum width of the eigenstates for different values of system size $L$.
+The disorder does not change anything discussed above and we observe the same behavior in the localization as in the homogeneous case.
+
+\begin{figure*}[h!]
+\centering
+\makebox[0pt]{
+\subfloat[${\rm Log}_{10}(\chi - \gamma)$ ]{
+  \includegraphics[width=55mm]{lambdaLinherr.pdf}
+}
+\subfloat[${\rm Max}\bra{\psi_i}\sigma_x\ket{\psi_i}$ for $\xi_i >1/2$.]{\includegraphics[width=55mm]{lambdaLinhloc.pdf}
+}
+}
+\caption{$t = 1, t'= 2, m=0, \nu=2$. }
+\label{lambdainh}
+\end{figure*}
+
 \end{document}

### 2020-05-28 draft.tex
..

@@
 }
-}\hspace{0mm}
-
-\makebox[0pt]{
-\subfloat[$\kappa_i$]{
-  \includegraphics[width=35mm]{cutavgwrongkap.pdf}
-}
-\subfloat[$t=1,t'=-2,m=1$]{\includegraphics[width=35mm]{cutavgwrong.pdf}
-}
 }
@@
 In c) and d) we show the same system at $m=2$ and a continuous homogeneous profile for $\kappa_i$.
-In e) and f) we show a case where using \eqref{eq:inh_simple} fails.
 We show the ES and $\chi_s$ for all cuts.
@@
 
-We have shown two cases were using expression \ref{eq:inh_simple} leads to the correct result.
-The average of a circular quantity is in fact not uniquely defined, and the average we do in Eq.\ref{eq:inh_simple}, which equates to averaging the valus in $(0,1)$ around $1/2$ (\carlos{is this clear?}), is just one choice that works in this cases.
-However in Fig.\ref{fig:simple_inh}.f this cut-average fails to give the Zak phase.
-This happens because by taking $\chi$ modulo 1 we have thrown out relevant information.
-To avoid this issue we can instead take the cut-average 
+We have shown two cases were using expression \ref{eq:inh_simple} leads to the correct result but one has to be careful as the average of a circular quantity is not uniquely defined.
+We can instead take the cut-average
 \begin{align}
@@
 
+The proof in Ref.\cite{Zaletel2014} also breaks down for superconducting systems.
 However, as we did for systems without translational invariance, one can extend this identity for systems with pairing terms as well.
@@
 
-In order to ensure that we obtain localized entanglement eigenstates,  we diagonalize $C_A^\lambda = C_A + \lambda/L^2 C_A\hat{X}C_A$ instead of $C_A$, where $\lambda$ is our small parameter and the reason we include the $L^2$ scaling will be explained below.
-For the whole system one way of obtaining localized eigenstates of the full correlation function (\carlos{also the Hamiltonian?}) is to diagonalize $C\hat{X}C$, which results in localized Wannier states.
+In order to ensure that we obtain localized entanglement eigenstates,  we diagonalize $C_A + \lambda C_A\hat{X}C_A$ instead of $C_A$.
+For the whole system, one way of obtaining localized eigenstates of the full correlation function (\carlos{Hamiltonian?}) is to diagonalize $C\hat{X}C$, which results in localized Wannier states.
 $C$ commutes with this operator because it is a projector, but this is no longer the case for $C_A$.
-Far from the edges however, $C_A$ and $C$ are equivalent, so that the states close to $\xi=1$ will perfectly localize into Wannier states by adding the $\lambda$ term.
+Far from the edges however, $C_A$ and $C$ are equivalent, so that the states at 1 will perfectly localize into Wannier states by adding the $\lambda C_A\hat{X}C_A/L$ term (\carlos{Note that the states at 0 or close to 0 don't localize ,but this is not an issue}).
 Note that in the thermodynamic limit the operator $\hat{X}$ acts as proportional to an identity for localized states.
-Therefore in this limit the localized eigenstates of $C_A$ far from $\xi = 0,1$ are also eigenstates of $C_A^\lambda$ while their mixture is not.
+Therefore in this limit the localized eigenstates of $C_A$ with $\xi \neq 0,1$ are also eigenstates of this operator while their mixture is not.
 Threfore this operator breaks all degeneracies of $C_A$ and results in localized eigenstates.
@@
 A parameter $\lambda$ which is too small will fail to break the degeneracies and localize the eigenstates.
-On the other hand the $\lambda$ term will  modify the eigenvalues and introduce an additional error to $\chi$.
-Assuming localized eigenstates we have $\expval{\hat{X}}_{i\alpha} \simeq i$.
+On the other hand the $\lambda$ term will modify the eigenvalues and introduce an additional error.
+The eigenstates roughly spread over the whole chain so that we have $\expval{\hat{X}}_{i\alpha} \simeq i$.
 Therefore the error it introduces in $\chi$ is roughly
 \begin{equation}
-\frac{\lambda}{L^2} \sum_{\alpha=1,2}\left( \sum_{i=L/4+1}^{L/2}i-\sum_{i=1}^{L/4}i \right) = \frac{\lambda }{8},
+\lambda \sum_{\alpha=1,2}\left( \sum_{i=L/4+1}^{L/2}i-\sum_{i=1}^{L/4}i \right) = \frac{\lambda L^2}{8}
 \end{equation}
-it is proportional to $\lambda$ and independent on system size.
-In Fig.\ref{lambdaL} we show this error in $\chi$ together with the maximum width of the eigenstates in the top half of the spectrum (\carlos{as we only care about the midgap states and the bulk modes at 1}), for different system sizes.
+Reescaling $\lambda$ to $\lambda/L^2$ we ensure that this error is roughly independent of system size and controlled only by $\lambda$.
+In Fig.\ref{lambdaL} we show this error in $\chi$ together with the average width for the eigenstates of the top half of the spectrum, for different system sizes.
 We see that indeed the error remains constant with system size.
-In this case most of the bulk modes are above $\xi = 1-10^{-10}$, so that the error coming from the $\lambda$ term is bigger than the one we make by failing to localize the bulk modes.
-However, by looking at the maximum width of the eigenstates we see three different regions.
-One region at low system size where the width increases linearly.
-A second region where the system localizes and the width seems constant, and finally for large systems a third region where the width increases again because the $\lambda$ term is too small to break the degeneracies and localize the eigenstates.
+In this particualr case the degeneracies do not seem to induce a big error.
+However when we look at the average width we see three different regions.
+One region at low system size where the width is bounded by the system size and therefore the $\lambda$ term strongly modifies the eigenstates, which fail to localize.
+For incresing system size we see a second region where the system seems to localize and the width seems constant.
+And finally for further increasing system size a third region where the width increases again because the $\lambda$ term is too small to break the degeneracies and localize the eigenstates.
 
@@
 \makebox[0pt]{
-\subfloat[${\rm Log}_{10}(\chi - \gamma)$ (\carlos{There are errors at $-5,-9,-12$ because that's the order of the eigenvalue of the bulk modes that should cancel pairwise but dont due to failure of localization.
-Furthermore, the error in yellow around $L=150$ is due to a mode close  to $\xi = 0$ so that it does not appear in Fig.
-b.})]{
+\subfloat[${\rm Log}_{10}(\chi - \gamma)$]{
   \includegraphics[width=55mm]{lambdaL1.pdf}
 }
-\subfloat[${\rm Max}\bra{\psi_i}\sigma_x\ket{\psi_i}$ for $\xi_i >1/2$.
-\carlos{I don't know why the green line localizes so bad.
-I have to look further into it}]{\includegraphics[width=55mm]{lambdaL2.pdf}
+\subfloat[$\sum_{i/\xi_i>1/2} \bra{\psi_i}\sigma_x\ket{\psi_i}/L$]{\includegraphics[width=55mm]{lambdaL2.pdf}
 }
 }
-\caption{$t = 1, t'= 2, m=0$ for $-{\rm Log}_{10}[\lambda] = 4,5,6,7,8,9,10$.}
+\caption{$\lambda = 10^{-9}, t = 1, t'= 2, m=0$}
 \label{lambdaL}
@@
 
-In Fig.\ref{linearRegion} we take a closer look into the first region with a maximum width proportional to the system size.
-What happens in this region is that due to finite size the midgap modes that localize in the edges overlap so the resulting eigenstates are not localized and are split from $\xi = 1/2$.
-The boundary where this region ends changes with $\lambda$ and coincides with the point where the splitting from the $\lambda$ term is bigger than the splitting due to finite size.
-The gap between these midgap modes that mix starts decreasing with larger system size up to a point, where the system starts to localize, where this splitting is roughly constant with system size and proportional to $\lambda$.
-In the $\nu = 1$ region the two midgap states are degenerate and we do not observe this first linear region, the eigenstates localize for any finite $\lambda$.
+In Fig.\ref{lambdalambda} we keep the system size constant, and big enough so that we do not introduce a strong finite size error as seen above, and sweep $\lambda$ (from big to small).
+We see two regions.
+For big $\lambda$ the error made decreases proportional to $\lambda$ and the eigenstates remain localize, until some point were $\lambda$ is so small that similar to what hapened in the third region above the eigenstates fail to localize.
+In this case this leads to a wrong classification of the midgap modes which results in a completely wrong $\chi$.
+
+Note that there is a lower bound to $\lambda$ given by the numerical accuracy, $\lambda_{\rm min} \sim 10^{-13} L^2$.
+Since the system has to be much greater than the correlation length in order for it to localize into Wannier states, we have $\lambda_{\rm min} \sim 10^{-13} l_{\rm corr}^2$, which will set the minimum error we can expect, as it scales only with the $\lambda$ parameter.
+(\carlos{Not 100\% sure about this statement.
+Specially how it works in relation to the system with interphase and zero-modes.})
+
+For inhomogeneous systems the story is slightly different.
+Due to the cut-average a wrong assignment of the bulk modes leads to an error of order $\mathcal{O}(1/L)$.
+In Fig.
+\ref{lambdainhlambda} we have the error made in $\chi$ and the maximum width of the eigenstates for different values of $\lambda$ and system size $L$.
+For each system size we observe, similar as before, a region were the system localizes followed by a region where the effect of the $\lambda$ term is too small and the system fails to localize.
+The only difference being that due to the cut-average a wrong assignment of the bulk modes leads to an error of order $\mathcal{O}(1/L)$.
+Note that there is an optimal value for the system size that allows to choose the lowest $\lambda$.
+In Fig \ref{lambdainhL} we see the same results for a system with $\lambda=10^{-8}$, near the minimum allowed where the system still localizes for some $L$.
+We see, as in the homogeneous case, that there are three regions.
+One with a finite-size effect, a second one were the system localizes and a third one where $\lambda/L^2$ is too small and the system delocalizes again.
 
@@
 \makebox[0pt]{
-\subfloat[Width of the eigenstate closest to $\xi=1/2$]{
-  \includegraphics[width=55mm]{linearBoundary.pdf}
+\subfloat[${\rm Log}_{10}(\chi - \gamma)$]{
+  \includegraphics[width=55mm]{lambdalambda1.pdf}
 }
-\subfloat[Gap between the two eigenstates closest to $\xi = 1/2$]{\includegraphics[width=55mm]{linearBoundarysplit.pdf}
+\subfloat[$\sum_{i/\xi_i>1/2} \bra{\psi_i}\sigma_x\ket{\psi_i}/L$]{\includegraphics[width=55mm]{lambdalambda2.pdf}
 }
 }
-\caption{$t = 1, t'= 2, m=0, \nu=0 , -{\rm Log}_{10}[\gamma] = 4,5,6,7,8,9,10$ for the different colors. Vertical lines are addded to better show how the features of both plots match.}
-\label{linearRegion}
+\caption{$L=80, t = 1, t'= 2, m=0$}
+\label{lambdalambda}
 \end{figure*}
-
-For inhomogeneous systems the story is slightly different.
-Due to the cut-average a wrong assignment of the bulk modes leads to an error of order $\mathcal{O}(1/L)$.
-In Fig.
-\ref{lambdainh} we have the error made in $\chi$ and the maximum width of the eigenstates for different values of system size $L$.
-The disorder does not change anything discussed above and we observe the same behavior in the localization as in the homogeneous case.
 
@@
 \makebox[0pt]{
-\subfloat[${\rm Log}_{10}(\chi - \gamma)$ ]{
+\subfloat[${\rm Log}_{10}(\chi - \gamma)$]{
+  \includegraphics[width=55mm]{lambdainherr.pdf}
+}
+\subfloat[$\sum_{i/\xi_i>1/2} \bra{\psi_i}\sigma_x\ket{\psi_i}/L$]{\includegraphics[width=55mm]{lambdainhloc.pdf}
+}
+}
+\caption{$L=40,80,120,160,200; t = 1, t'= 2, m=0$}
+\label{lambdainhlambda}
+\end{figure*}
+
+\begin{figure*}[h!]
+\centering
+\makebox[0pt]{
+\subfloat[${\rm Log}_{10}(\chi - \gamma)$]{
   \includegraphics[width=55mm]{lambdaLinherr.pdf}
 }
-\subfloat[${\rm Max}\bra{\psi_i}\sigma_x\ket{\psi_i}$ for $\xi_i >1/2$.]{\includegraphics[width=55mm]{lambdaLinhloc.pdf}
+\subfloat[$\sum_{i/\xi_i>1/2} \bra{\psi_i}\sigma_x\ket{\psi_i}/L$]{\includegraphics[width=55mm]{lambdaLinhloc.pdf}
 }
 }
-\caption{$t = 1, t'= 2, m=0, \nu=2$. }
-\label{lambdainh}
+\caption{$\lambda=10^{-8}, t = 1, t'= 2, m=0$}
+\label{lambdainhL}
 \end{figure*}

### 2020-06-03 draft.tex

--- 2020-05-27 draft.tex
@@
 \begin{align}
-\gamma =\left(\frac{1}{L}\sum_{s=1}^L (\chi_s \, \rm{mod} \, 1)\right)\, \rm{mod} \, 1,
+\bar{\chi} =\left(\frac{1}{L}\sum_{s=1}^L (\chi_s \, \rm{mod} \, 1)\right)\, \rm{mod} \, 1,
 \label{eq:inh_simple}
@@
 \subfloat[$\kappa_i$]{
-  \includegraphics[width=35mm]{cutavgwrongkap.pdf}
+  \includegraphics[width=35mm]{simplecutavgaround0kap.pdf}
 }
-\subfloat[$t=1,t'=-2,m=1$]{\includegraphics[width=35mm]{cutavgwrong.pdf}
+\subfloat[$t=1,t'=-2,m=1$]{\includegraphics[width=35mm]{simplecutavgaround0.pdf}
 }
@@
 
-We have shown two cases were using expression \ref{eq:inh_simple} leads to the correct result.
-The average of a circular quantity is in fact not uniquely defined, and the average we do in Eq.\ref{eq:inh_simple}, which equates to averaging the valus in $(0,1)$ around $1/2$ (\carlos{is this clear?}), is just one choice that works in this cases.
-However in Fig.\ref{fig:simple_inh}.f this cut-average fails to give the Zak phase.
-This happens because by taking $\chi$ modulo 1 we have thrown out relevant information.
-To avoid this issue we can instead take the cut-average 
+We have shown two cases were using expression \ref{eq:inh_simple} leads to the correct result, but this is in fact not the general case, as we can see in Fig.\ref{fig:simple_inh}.f .
+The average of a circular quantity is not uniquely defined and the choice we made in Eq.\ref{eq:inh_simple}, which equates to averaging the values in $(0,1)$ around $1/2$ (\carlos{is this clear?}) turns out to work in certain cases.
+How to perform this circular average is a priory not known, we can only judge after we compute the Zak phase independetly if we used the correct choice or not.
+Therefore in order to proceed we need to get rid of the circular average which leaves us with
 \begin{align}
-\gamma =\left(\frac{1}{L}\sum_{s=1}^L \chi_s \right)\, \rm{mod} \, 1,
+\bar{\chi} =\left(\frac{1}{L}\sum_{s=1}^L \chi_s \right)\, \rm{mod} \, 1,
 \label{eq:inh_fixed}
 \end{align}
-
 but this however has to be done carefully.
@@
 In order for \ref{eq:inh_fixed} to work properly the contribution from the bulk modes, which should vanish in the end, must remain constant across the different cuts.
-In order to achieve this we choose a cut-dependent threshold that the bulk modes do not cross, $l_s$, as the mean position of two of the eigenstates close to this middle point of A, as shown in Fig.\ref{fig:loc_fix}.
-This is done in practice for the Wannier states (\carlos{Do I explain exactly how do I do it numerically?}).
+In order to achieve this we have to choose a cut-dependent threshold that the bulk modes do not cross.
+We choose this threshold, $l_s$, as the mean position of two of the eigenstates close to this middle point of A, as shown in Fig.\ref{fig:loc_fix}.
+This is done in practice using the Wannier states (\carlos{Do I explain exactly how do I do it numerically?}).
 Note that in the thermodynamic limit it does not matter if the position of these eigenstates cross as they are degenerate, the only important thing is that the number of bulk modes at each sector remains constant (\carlos{Not sure about this phrasing either}).
+
+The reason Eq.\ref{eq:inh_simple} works in some cases is that if, assuming the correct constant bulk contribution, $\chi_s \in [n,n+1)$, wher $n$ is an integer, then taking the modulo 1 before or after averaging does not affect the final result.
+The same argument can be generalize for a $\chi_s \in [a+a,1)$, for any real $a$, given that we offset the modulo 1 in Eq.\ref{eq:inh_simple} appropiately.
 
@@
 \includegraphics[width=70mm]{loc_fix.pdf}
-\caption{Average position of the Wannier states in red and, slightly shifted, the localized entanglement eigenstates (\carlos{rename $C_A+\lambda C_A X C_A$?} in blue.
-In black we show the threshold $l_s$ obtained as the mean value of the two Wannier states close de $19$ and $20$) $L=80,\lambda=10^{-6}, t = 1, t'= 2, m=0$}
+\caption{Average position of the Wannier states in red and, slightly shifted, the localized entanglement eigenstates 
+in blue. In black we show the threshold $l_s$ obtained as the mean value of the two Wannier states close de $19$ and $20$) $L=80,\lambda=10^{-6}, t = 1, t'= 2, m=0$}
 \label{fig:loc_fix}
@@
 
-As the sign assigned to the bulk modes becomes relevant in the inhomogeneous case one might think that it is better to use the open chain, $\chi_s = \Tr C_A^{o,s}$, as one avoids the issue of having to assign any signs at all.
+As the sign assigned to the bulk modes becomes relevant in the inhomogeneous case one might think that it is better to use the open chain, $\chi_s = \Tr C_A^{{\rm open},s}$, as one avoids the issue of having to assign any signs at all.
 Although this is true we find another problem.
@@
 This can lead to discontinuities in the number of bulk modes at $\xi=1$ which again result in an incorrect cut-average.
-\carlos{Works in some cases.}
+\carlos{Do we want to include is?
+Or leave it for later?}
 

### 2020-06-04 draft.tex
..

@@
 }
-\caption{\mariac{increase system size and colorcode edge modes.}Edge spectrum along the dotted path (i.e.
+\caption{\mariac{Figures still need to be replaced}Edge spectrum along the dotted path (i.e.
 $t=1,t'=-2$) of Fig.~\ref{fig:bdi_phase_diagram} for different values of $\kappa$ and $\kappa'$.\mariac{note on numerical system size}}
@@
 Interpreting this as a one-dimensional system with a parameter, we can explain this behavior by Eq.~\eqref{eqplusminus}, noting that there is a single pair of eigenvalues that dominates the sum. 
-Note that the observation of Ref.s~\cite{Huang2012,Huang2012-2} is particular to a system with Chern number 0 or $\pm 1$, and it generically fails for systems with higher Chern numbers.
+Note that the observation of Ref.s~\cite{Huang2012,Huang2012-2} is particular to systems with Chern number 0 or $\pm 1$, and usually fails for more general systems, see for instance Fig.~\ref{2}.  
 In the latter case, there are several terms in Eq.~\eqref{eqplusminus} with comparably large contributions. 
@@
 This is a common feature for e.g. nearest neighbor hopping models\mariac{ref to Peschel}, such as the SSH chain, and applies to the Hamiltonian~\eqref{bdi_model} when $t'=0$. 
-In this case, one can obtain the Zak phase in a very simple fashion: We reorder the eigenvalues of the ES by magnitude, $\xi_1 < \xi_2 < ...$, and compute
+In this case, one can obtain the Zak phase in a very simple fashion: We reorder the eigenvalues of the EOS by magnitude, $\xi_1 < \xi_2 < ...$, and compute
 \begin{equation}
@@
 In the thermodynamic limit, $\chi$ becomes identical to the Zak phase divided by $2\pi$:
-\begin{equation}
+\begin{equation}\label{eq:chi_gamma_simple}
 \lim_{L \rightarrow \infty} \frac{\gamma}{2\pi} = \lim_{L \rightarrow \infty} \chi.
 \end{equation}
+In the case where there is only one eigenvalue per edge that is $\neq 0,1$ in  the EOS, like in the fully dimerized limit of the SSH chain, we recover the equality between the midgap eigenvalue and the Zak phase~\cite{Ryu2006}.
+We illustrate the general behavior of Eq.~\eqref{eq:chi_gamma_simple} in Fig.~\ref{huang}, both in the presence, \ref{huang}(a),  and absence, \ref{huang}(b),  of chiral symmetry. 
+We show the EOS (black), the Zak phase$/2\pi$ (red) and $\chi$ (green), setting the next-nearest neighbor hopping $t'=0$, i.e.
+along the dashed cut in the phase diagram of Fig.~\ref{fig:bdi_phase_diagram}.
+Along this line, our model is equivalent to an SSH chain.
 
-In the case where there is only one eigenvalue per edge that is $\neq 0,1$ in  the EOS, like in the fully dimerized limit of the SSH chain, we recover the equality between the midgap eigenvalue and the Zak phase~\cite{Ryu2006}.
-
-In Fig.~\ref{huang} we show  the EOS along the dashed cut in the phase diagram of Fig.~\ref{fig:bdi_phase_diagram}. 
-Along this line, our model is equivalent to an SSH chain. 
 In the presence of chiral and translation symmetry, all eigenvalues of the EOS are doubly degenerate --- one eigenvalue from each edge. 
@@
 Note that the full sum~\eqref{eqplusminus}  evaluates to zero in the trivial regime, up to finite-size corrections, despite the presence of modes with entanglement occupancy $0<\xi<1$. 
+When breaking chiral symmetry, the EOS will no longer be doubly degenerate. 
+Still, eq.~\eqref{eqplusminus} faithfully reproduces the Zak phase, see Fig.~\ref{huang}(b). 
+\paragraph*{Finite-size scaling}
 
-When breaking chiral symmetry, the EOS will no longer be doubly degenerate. 
-For small $m$, there are two dominant midgap eigenvalues, and the Zak phase follows the lower of them very closely, see Fig.~\ref{huang}(b).
-When this midgap eigenvalue approaches  0, the contribution from the other modes becomes more relevant and the Zak phase starts to deviate from the midgap eigenvalue, \mariac{see inset}.
+% \mariac{still missing}
+%
+%
+%For small $m$, there are two dominant midgap eigenvalues, and the Zak phase follows the lower of them very closely, see Fig.~\ref{huang}(b).
+%When this midgap eigenvalue approaches  0, the contribution from the other modes becomes more relevant and the Zak phase starts to deviate from the midgap eigenvalue, \mariac{see inset}.
 \begin{figure}[h!]
@@
 \end{equation}
+Note that the details on how to define \emph{right} versus \emph{left} are actually not important. 
+We could have chosen any other boundary than $L/4$ as long as it is sufficiently far away from the boundary to ensure that the boundary modes are correctly assigned. 
+The underlying reason is that  bulk modes have eigenvalues $\xi=0,1$.
+Thus, re-assigning, say, a \emph{right} to  \emph{left} bulk mode changes the expression within the parenthesis in eq.~\eqref{eq_loc} by either $0$ or $1$, and the full expression remains unchanged due to the modulo.
 
-In order for this expression to be well-defined however we need localized eigenstates. 
+However, in order for  \eqref{eq_loc} to be well-defined,  we need localized eigenstates. 
 Non-degenerate eigenstates of the EOS are automatically localized, but degenerate or nearly-degenerate eigenstates (i.e. those exponentially close to $\xi=0,1$) are not necessarily localized. 
-It turns out that we need not worry about the latter set of states, because changing their charge from say $1/2$ to $-1/2$ only changes $1/2(\sum_{i\in r} \xi_i - \sum_{i\in l}\xi_i)$ in Eq.~\eqref{eq_loc} by an integer, which is canceled \mariac{better expression?} by mod 1.
+As explained above, we need not worry about the latter set of states.
+%, because changing their charge from say $1/2$ to $-1/2$ only changes $1/2(\sum_{i\in r} \xi_i - \sum_{i\in l}\xi_i)$ in Eq.~\eqref{eq_loc} by an integer, which is canceled \mariac{better expression?} by mod 1.
 Also in the case where eigenstates are double degenerate we need not worry about localizing them, because computing the expectation value of the position operator will automatically assign the correct number of eigenstates to the left/right half of $A$.
 This is no longer true for four-fold, or higher-fold, degenerate states.
+
 %In Fig.\ref{huang}.a the doubly-degenerate states have contributions from both edges.
@@
 
-In order to ensure that we obtain localized entanglement eigenstates,  we diagonalize $C_A + \lambda C_A\hat{X}C_A$ instead of $C_A$.
+{\color{red} Localization section needs to be rewritten}
+we diagonalize $C_A + \lambda C_A\hat{X}C_A$ instead of $C_A$.
 For the whole system, one way of obtaining localized eigenstates of the full correlation matrix is to diagonalize $C\hat{X}C$, which results in localized Wannier states. 
@@
 But as it does not commute with $C_A$ close to the virtual edges,  it will modify the eigenstates. 
-For that reason $\lambda$ must be a small parameter. \mariac{how to scale $\lambda$ with system size}. 
+For that reason $\lambda$ must be a small parameter. 
 %\mariac{do not commute 
@@
 In Fig.\ref{2} we show two examples that highlight the importance of the localization structure of the ES. 
-In subfigures (a) and (b), we compute the EOS for varying $m$ at  $t=1$, $t'=2$, indicated by the solid line  in the phase-diagram~\ref{fig:bdi_phase_diagram}, in the presence of two different symmetry breaking terms: (i) $\kappa =0.3$, $\kappa'=0$ and (ii) $\kappa=0$, $\kappa'=0.3$.
+In subfigures (a) and (b), we compute the EOS for varying $m$ at  $t=1$, $t'=2$ --- indicated by the solid line  in the phase-diagram~\ref{fig:bdi_phase_diagram} --- in the presence of two different symmetry breaking terms: (i) $\kappa =0.3$, $\kappa'=0$ and (ii) $\kappa=0$, $\kappa'=0.3$.
 For $\kappa=\kappa'=0$, there is a phase transition $\nu = 2 \rightarrow 0 $ at $m=2$, whereas both choices of symmetry breaking render the system trivial  for all $m$.
 However, the two symmetry breaking terms split the four-fold degenerate $\xi=1/2$ midgap states very differently, thus resulting in completely different Zak phases despite the superficial resemblance of their respective EOS.
-While $\kappa'$ splits the four-fold degenerate states into two two-fold degenerate pairs at opposite edges whose contribution cancels, $\kappa$ split them into two $left-left$ and $right-right$ pairs, with a finite contribution to the Zak phase.
+While $\kappa'$ splits the four-fold degenerate states into two two-fold degenerate pairs at opposite edges whose contribution cancels, $\kappa$ split them into two \emph{left}-\emph{left} and \emph{right}-\emph{right} pairs, with a finite contribution to the Zak phase.
 In subfigures (c) and (d), we consider the same symmetry breaking terms, but now for $t=1$ and $t'=-2$, where the BDI systems shows a phase transition  $\nu=2\rightarrow 1\rightarrow 0$. 
@@
 \end{figure}
-
-\mariac{issue about  the ambiguity about right - left vs left - right, connect with discussion about Zaletel results and their convention?}
 This relation between the Zak phase and the EOS is a consequence of an identity between the Zak phase and the many-body entanglement spectrum previously found for infinite chains \cite{Zaletel2014}.
@@
 This method, however, will introduce a stronger finite size effect and requires substantially larger systems for the same accuracy. 
-In addition, it only works for homogeneous systems as explained below.
+In addition, it only works for homogeneous systems as explained in the next section.
 
@@
 The only remaining issue is whether this simple recipe assigns charge $+1/2$ or $-1/2$ to the right edge.
-In presence of a symmetry, this question is irrelevant \mariac{better expression?} because of the degeneracy. 
+In presence of a symmetry, this question is irrelevant because of the degeneracy. 
 In general, it implies that $\chi$ as defined in Eq.~\eqref{eqplusminus} only gives the Zak phase up to an overall sign. 
 This issue can be fixed by determining the localization of a single eigenstate.
+
+This is closely related to another issue, namely \mariac{issue about  the ambiguity about right - left vs left - right, connect with discussion about Zaletel results and their convention}
 
@@
 \section{Inhomogeneous systems}
+
+In the last section, we have argued that the Zak phase is encoded in the EOS. 
+This result could, in fact, already been inferred from Ref.~\cite{Zaletel2014}, where the relation between the many-body entanglement spectrum and the Zak phase was derived. 
+However, for the latter  translation invariance was crucial for the derivation, and it was unclear if (and how) it can be generalized to inhomogeneous systems.
+In this section, we demonstrate that the EOS contains the full information about the Zak phase even for inhomogeneous systems and present a well-defined procedure for extracting the Zak phase.
 

### 2020-06-04 draft.tex
started on changes in chapter V

@@
 In the last section, we have argued that the Zak phase is encoded in the EOS. 
-This result could, in fact, already been inferred from Ref.~\cite{Zaletel2014}, where the relation between the many-body entanglement spectrum and the Zak phase was derived. 
-However, for the latter  translation invariance was crucial for the derivation, and it was unclear if (and how) it can be generalized to inhomogeneous systems.
-In this section, we demonstrate that the EOS contains the full information about the Zak phase even for inhomogeneous systems and present a well-defined procedure for extracting the Zak phase.
+This result could, in fact, already have been inferred from Ref.~\cite{Zaletel2014}, where the relation between the singular values of the Schmidt decomposition and the Zak phase was derived. 
+However, for the latter translation invariance was crucial in the derivation, and it was unclear if (and how) it can be generalized to inhomogeneous systems.
+Note that both the entanglement spectrum --- single-body or many-body --- and the Schmidt decomposition depend on where the virtual boundaries are located in a system that is not translationally invariant.
+Despite this obvious concern, we will show that the EOS still contains the full information about the Zak phase even for inhomogeneous systems, and present a well-defined procedure for extracting it.
 
-In reference \cite{Zaletel2014} translational invariance was used to derive the relation between the many-body entanglement spectrum and the Zak phase.
+%In reference \cite{Zaletel2014} translational invariance was used to derive the relation between the many-body entanglement spectrum and the Zak phase.
 If this symmetry is broken by adding, for example, weak disorder, this result is no longer true.
@@
 For some systems this can be written as (\carlos{not sure about this phrasing})
+The main idea is that even though $\chi$ depends on the details of the location of the virtual boundaries, by averaging over different such `cuts' we can define a well-defined quantity. 
+The most obvious generalization of eq.~\eqref{eq_loc} is given by 
 \begin{align}
@@
 \begin{equation}
-\chi_s = \frac{1}{2} \left( \sum_{i\in r}\xi^s_i-\sum_{i\in l}\xi^s_i  + N_A \right)
+\chi_s = \frac{1}{2} \left( \sum_{i\in r}\xi^s_i-\sum_{i\in l}\xi^s_i  + N_A \right). 
 \end{equation}
-is obtained for cuts at sites $s$ and $s+L/2$.
+Thus, we average $\chi$ over cuts that partition the system into two equal halves with the virtual boundaries located between sites sites $s$, $s+1$ and $s+L/2$,$s+L/2+1$ respectively. 
+This generalization indeed reproduces the Zak phase for some systems, but is unfortunately too simple to work in general. 
+Remember that $\chi$ refers to an angular variable,
+
+XXXXXXX
+
 In Fig.\ref{fig:simple_inh}.b we show the same system as in Fig.\ref{2}.c with an added moderate disorder in the $\kappa$ parameter shown in \ref{fig:simple_inh}.a .

### 2020-06-18 draft.tex
Updated Restas formula and inhomogeneous plots

--- 2020-06-03 draft.tex
@@
 
-The polarization can be obtained as the many-body average of the position operator, but the latter has to be regularized for periodic boundary conditions \cite{Resta1997}. 
+The polarization can be obtained from the many-body average of the position operator of electrons and ions, but the position operator has to be regularized for periodic boundary conditions \cite{Resta1997}.
 We can define a position operator for PBC as
@@
 and  ${\rm det }\!' $ indicates that the determinant is restricted to the space of \emph{occupied} single-particle states.
+Note that for an open chain the many-body average position of the ions will sit at a site when the chain length is odd and at a link when the length is even.
+The same result applies in the system with periodic boundary conditions, i.e.
+$\expval{\hat{X}_{PBC,+}}_{0}=(1+(-1)^{L})/4$.
 The Zak phase can then be obtained as
-\begin{equation}
-\gamma = \expval{\hat{X}_{\rm PBC}}_0\frac{2\pi}{L} + \frac{1}{2}.
-\end{equation}
-\carlos{There is an extra 1/2 factor compared to Restas's formula which I don't know where it is coming from.}
+\begin{align}
+\gamma =& \left( \expval{\hat{X}_{\rm PBC}}_0-\expval{\hat{X}_{\rm PBC,+}}_0 \right)\frac{2\pi}{L} \\
+=& \expval{\hat{X}_{\rm PBC}}_0\frac{2\pi}{L} -(1+(-1)^{L})/4.
+\end{align}
 
@@
 The problem resides in the classification of these bulk modes.
-Even if the rest of the EOS remains constant, due to the inhomogeneity a bulk mode can cross the middle of region A, as shown in Fig.\ref{fig:loc_fix}, which will change $\chi$ by $1/L$, even though choosing $L/4$ as the separation between the \emph{left} and \emph{rigth} sectors is not physical.
-In order for \ref{eq:inh_fixed} to work properly the contribution from the bulk modes, which should vanish in the end, must remain constant across the different cuts.
-In order to achieve this we have to choose a cut-dependent threshold that the bulk modes do not cross.
-We choose this threshold, $l_s$, as the mean position of two of the eigenstates close to this middle point of A, as shown in Fig.\ref{fig:loc_fix}.
-This is done in practice using the Wannier states (\carlos{Do I explain exactly how do I do it numerically?}).
-Note that in the thermodynamic limit it does not matter if the position of these eigenstates cross as they are degenerate, the only important thing is that the number of bulk modes at each sector remains constant (\carlos{Not sure about this phrasing either}).
+Even if the rest of the EOS remains constant, due to the inhomogeneity a bulk mode can cross the middle of region A, as shown in Fig.\ref{fig:wan1}.c, which will change $\chi$ by $1/L$, even though choosing $L/4$ as the separation between the \emph{left} and \emph{rigth} sectors is not physical.
+In order for Eq.\ref{eq:inh_fixed} to work properly the contribution from the bulk modes, which should vanish in the end, must remain constant across the different cuts.
+In order to achieve this we have to choose a cut-dependent threshold such that it is not crossed by the bulk modes, as shown in figure \ref{fig:wan1}.c.
+This is done in practice using the Wannier states.
+Note that in the thermodynamic limit it does not matter if the position of these eigenstates cross as they are degenerate, the only important thing is that the number of bulk modes at each sector remains constant.
 
@@
 \centering
-\includegraphics[width=70mm]{loc_fix.pdf}
-\caption{Average position of the Wannier states in red and, slightly shifted, the localized entanglement eigenstates 
-in blue. In black we show the threshold $l_s$ obtained as the mean value of the two Wannier states close de $19$ and $20$) $L=80,\lambda=10^{-6}, t = 1, t'= 2, m=0$}
-\label{fig:loc_fix}
+\makebox[0pt]{
+\subfloat[$\kappa_i$]{
+  \includegraphics[width=35mm]{wan2_kap.pdf}
+}
+\subfloat[EOS (black) and $\chi_s$ mod $1$ (green)]{\includegraphics[width=35mm]{wan2_EOS.pdf}
+}
+}\hspace{0mm}
+
+\makebox[0pt]{
+\subfloat[Average position of the EOS eigenstates (blue) and Wannier states (red) around the threshold that separates the left and right sector (black).]{
+  \includegraphics[width=35mm]{wan2_wan.pdf}
+}
+\subfloat[$\chi_s$ obtained using the appropiate left-right threshold (black), Zak phase (red) and $\bar{\chi}$ (dashed blue).]{\includegraphics[width=35mm]{wan2_essum.pdf}
+}
+}
+\caption{Parameters used are: $t=1,t'=-2,m=0,\kappa'=0$. With the profile of $\kappa_i$ shown in (a) the width of $\chi_s$ is larger than one and therefore Eq.\ref{eq:inh_fixed} cannot be employed.}
+\label{fig:wan1}
 \end{figure}
 
-As the sign assigned to the bulk modes becomes relevant in the inhomogeneous case one might think that it is better to use the open chain, $\chi_s = \Tr C_A^{{\rm open},s}$, as one avoids the issue of having to assign any signs at all.
-Although this is true we find another problem.
+\begin{figure}[h!]
+\centering
+\makebox[0pt]{
+\subfloat[$\kappa_i$]{
+  \includegraphics[width=35mm]{wan_kap.pdf}
+}
+\subfloat[EOS (black) and $\chi_s$ mod $1$ (green)]{\includegraphics[width=35mm]{wan_EOS.pdf}
+}
+}\hspace{0mm}
+
+\makebox[0pt]{
+\subfloat[Average position of the EOS eigenstates (blue) and Wannier states (red) around the threshold that separates the left and right sector (black).]{
+  \includegraphics[width=35mm]{wan_wan.pdf}
+}
+\subfloat[$\chi_s$ obtained using the appropiate left-right threshold (black), Zak phase (red) and $\bar{\chi}$ (dashed blue).]{\includegraphics[width=35mm]{wan_essum.pdf}
+}
+}
+\caption{Parameters used are: $t=1,t'=-2,m=0,\kappa'=0$.
+Adding disorder to an oscillating profile similar to the one in Fig.\ref{fig:wan1} shows that in general there is not enough information in $\chi_s$ mod $1$ to recover the Zak phase.
+}
+\label{fig:wan2}
+\end{figure}
+
+In Fig.\ref{fig:wan1} we have shown an example where an oscillating profile for $\kappa_i$ results in a $\chi_s$ with a width which is larger than one, meaning that using Eq.\ref{eq:inh_fixed} is not posible in any case.
+Since we have chosen a continuous profile one might argue that one can recover the correct contribution of $\chi_s$, shown in Fig.\ref{fig:wan1}.c from $\chi_s$ mod $1$, there is however not possible a priori to know if that is the correct result.
+Furthermore one cannot do this in general, as shown in figure \ref{fig:wan2}.
+In this example we have added disorder to a similar oscillating profile for $\kappa_i$ that there are cases in general where clearly there is not enough information in $\chi_s$ mod $1$ to recover the Zak phase.
+\carlos{The profile in Fig.\ref{fig:wan2} is a really good example where there is no way one can recover the Zak phase from $\chi_s$ mod $1$.
+However I think the plot of average positions is too messy to understand what I mean by the bulk modes crossing $L/4$ and understand how the threshold is chosen, that's why I added the example in Fig.\ref{fig:wan1} but I dont know if we want to keep both.}
+
+As the sign assigned to the bulk modes becomes relevant in the inhomogeneous case and leads to problems one might think that it is better to use the open chain, $\chi_s = \Tr C_A^{{\rm open},s}$, as one avoids the issue of having to assign any signs at all.
+Although this is true we find another issue.
 By adiabatically tuning down the coupling between the $L$ and $1$ sites we see that the eigenvalues of the $left$ modes flow to either $0$ or $1$.
 This can lead to discontinuities in the number of bulk modes at $\xi=1$ which again result in an incorrect cut-average.
-\carlos{Do we want to include is?
-Or leave it for later?}
+\carlos{Do we want to include is?}
 
@@
 
+\carlos{I added this thinking it was generalizing on what Zaletel did but that was wrong, they also allow for interactions.
+Therefore I think it's too trivial for a section.
+It might be enough to mention that the above is valid for any quadratic Hamiltonian.}
 However, as we did for systems without translational invariance, one can extend this identity for systems with pairing terms as well.
@@
 
+\begin{figure*}[h!]
+\centering
+\makebox[0pt]{
+\subfloat[${\rm Max}\bra{\psi_i}\sigma_x\ket{\psi_i}$ for $\xi_i >1/2$.]{
+  \includegraphics[width=55mm]{bigsystemloc.pdf}
+}
+\subfloat[${\rm Log}_{10}(\chi - \gamma)$ ]{\includegraphics[width=55mm]{bigsystemerr.pdf}
+}
+}
+\caption{$t = 1, t'= 2, m=0, \nu=2$. }
+\label{lambdainh}
+\end{figure*}
+
+\begin{figure*}[h!]
+\centering
+\makebox[0pt]{
+\subfloat[ ${\rm Max}\bra{\psi_i}\sigma_x\ket{\psi_i}$ for $\xi_i >1/2$]{
+  \includegraphics[width=55mm]{smallsystemloc.pdf}
+}
+\subfloat[${\rm Log}_{10}(\chi - \gamma)$.]{\includegraphics[width=55mm]{smallsystemerr.pdf}
+}
+}
+\caption{$t = 1, t'= 2, m=0, \nu=2$. }
+\label{lambdainh}
+\end{figure*}
+
+\begin{figure*}[h!]
+\centering
+\makebox[0pt]{
+\subfloat[${\rm Max}\bra{\psi_i}\sigma_x\ket{\psi_i}$ for $\xi_i >1/2$. ]{
+  \includegraphics[width=55mm]{lambdaloc.pdf}
+}
+\subfloat[spectrum of $H_{ent}$]{\includegraphics[width=55mm]{lambdalocES.pdf}
+}
+}
+\caption{$t = 1, t'= 2, m=0, \nu=2$. }
+\label{lambdainh}
+\end{figure*}
+
+\begin{figure*}[h!]
+\centering
+\makebox[0pt]{
+\subfloat[$\gamma$ (red) and $\chi$ (blue) ]{
+  \includegraphics[width=55mm]{nonquanterror2.pdf}
+}
+\subfloat[${\rm Log}_{10}(\chi - \gamma)$]{\includegraphics[width=55mm]{nonquanterror3.pdf}
+}
+}
+\caption{$t = 1, t'= 2, m=2, \kappa = 0.2 $. }
+\label{lambdainh}
+\end{figure*}
+
+\section{Resta's formula}
+
+Consider a 1D chain with L unit cells. For a half-filled two-orbital system the many-body ground-state is given by
+\begin{equation}
+\Psi_0(\vec{x}) = A\prod_{\mu} \psi_\mu (x_\mu),
+\end{equation}
+where $A$ is an antisymmetrizer and $\mu$ is restricted to the occupied states. The polarization is give by the expectation value of the operator
+\begin{equation}
+\hat{P}=\frac{-e}{L}\left(\sum_j X_j - \sum_j  x_j\right),
+\end{equation}
+where $X_j$ stands for the position operator of the nuclei, and $x_j$ the one for the electrons. Assuming a tight-binding approximation $X_j$ corresponds simply to the lattice positions.
+
+Given periodic boundary conditions the polarization is obtained as
+\begin{align*}
+P =& \frac{-e}{2\pi}{\rm Im}\,{\rm Ln}\bra{\Psi_0}e^{i\frac{2\pi}{L} \left(\sum_j X_j - \sum_j  x_j\right)}\ket{\Psi_0} \\
+=&\frac{-e}{2\pi}{\rm Im}\,{\rm Ln}\bra{\Psi_0}e^{-i\frac{2\pi}{L} \sum_j  x_j}\ket{\Psi_0}+\frac{-e(1+L)}{2}.
+\end{align*}
+Note that it is very important to include the contribution from the nuclei whenever we are working with periodic boundary conditions, as a new term dependent on system size appears.
+This can now be expressed as
+\begin{align*}
+P =& \frac{-e}{2\pi}{\rm Im}\,{\rm Ln} \, {\rm det}[S]+\frac{-e(1+L)}{2},
+\end{align*}
+where
+\begin{equation}
+S_{\mu \nu} = \int_0^L dx \, \psi^*_{\mu}(x)e^{i\frac{2\pi}{L}x}\psi_\nu(x).
+\end{equation}
+Consider now an spetial case where the system is translational invariant with only one occupied band. The indeces $\mu,\nu$ are now given by the momentum on the occupied band and we have
+\begin{equation}
+S_{q,q+1} = \int_0^L dx \, \psi^*_{q}(x)e^{i\frac{2\pi}{L}x}\psi_{q+1}(x),
+\end{equation}
+since all other elements of $S$ will vanish due to orthogonality proeprties of the Bloch functions. Now the determinant can be obtained as
+\begin{equation}
+{\rm det[S]} = \varepsilon_{2,3,...L,1}\prod_q S(q,q+1)
+\end{equation}
+Permuting two by two the $1$ back to the first position gives for the antisymmetric tensor $+1$ for odd $L$ and $-1$ for even $L$. The polarization is given therefore by
+\begin{align*}
+P =& \frac{-e}{2\pi}{\rm Im}\,{\rm Ln} \, \left\{e^{-i \pi (L+1)}\prod_q S(q,q+1)\right\}+\frac{-e(1+L)}{2} \\
+=& \frac{-e}{2\pi}{\rm Im}\,{\rm Ln} \, \left\{\prod_q S(q,q+1)\right\}
+\end{align*} 
+which coincides with the known result for the discretized Zak phase.
+Both the term due to the nuclei and this sign coming from the antisymmetric tensor are missing from the Resta paper, but since they cancel each other its final result was ok.
+It's only when you compute things without translational invariance that the issue arised.
+
 \end{document}

### 2020-08-05 version2/draft.tex
Add files via upload

\documentclass[twocolumn,amsmath,longbibliography,amssymb,superscriptaddress]{revtex4-1}
\usepackage[pdftex]{graphics}
\usepackage{graphicx}
\graphicspath{{figures/}}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{physics}
\usepackage{subfig}
\usepackage{bm}
\usepackage{caption}
\usepackage{amsmath} 
\usepackage{mathtools}

\newcommand{\carlos}[1]{{\color{red} #1}}
\newcommand{\maria}[1]{{\color{blue} #1}}
\newcommand{\mariac}[1]{{\it\color{cyan}#1}}

\begin{document}

\title{On the relation of the entanglement spectrum to the Zak-phase}
\author{Carlos Ortega Taberner}
\affiliation{Department of Physics, Stockholm University, AlbaNova University Center, SE-106 91 Stockholm, Sweden}
\affiliation{Nordita, KTH Royal Institute of Technology and Stockholm University, SE-106 91 Stockholm, Sweden}

\author{Maria Hermanns}
\affiliation{Department of Physics, Stockholm University, AlbaNova University Center, SE-106 91 Stockholm, Sweden}
\affiliation{Nordita, KTH Royal Institute of Technology and Stockholm University, SE-106 91 Stockholm, Sweden}
\date{\today}

\maketitle

\section{Introduction}
Topological phases of matter have attracted a lot of attention during the last decades, not the least because a large variety of relevant systems have been realized experimentally. 
The early focus was mainly on \emph{topologically ordered} systems \cite{wenbook}, where interaction effects are crucial for stabilizing the phase. 
The most notable examples are the fractional quantum Hall effect~\cite{Tsui1982} and quantum spin liquids~\cite{Balents2010spin}. 
However, since 2005~\cite{kane2005quantum, roy2009topological} the focus has shifted to non-interacting topological phases, which can be characterized in terms of topological invariants~\cite{ryu2010topological}. 
Symmetries are often necessary to protect these topological phases, and determine which distinct topological phases can be realized for a given dimensionality.

There are few tools to identify topological phases. 
A very efficient one is the entanglement entropy~\cite{Kitaev2006topological, Levin2006detecting}, which allows one to identify the total quantum dimension of the underlying topological quantum field theory. 
As such, the entanglement entropy can only be used for interacting systems. 
Its main drawback is the fact that it can only provide one of the pertinent quantum numbers that characterizes the system, and is, thus,  not able to uniquely determine the topologically ordered phase at hand. 
In addition, it requires a careful scaling analysis, which can be challenging for strongly correlated systems. 
Another, closely related tool, is the entanglement spectrum (ES), originally introduced for fractional quantum Hall systems~\cite{Li2008entanglement}. 
It was conjectured to provide information about the edge spectrum, which was later shown to be a general feature of ground states with an effective topological quantum field theory description~\cite{Qi2012general}. 
It also proved to be an effective tool for fractional Chern insulators~\cite{Regnault2011fractional} and certain quantum spin liquids~\cite{yao2010entanglement}. \mariac{Others?}

Generically, any of the proposed tools provides some information about the phase naturally, while others seem inaccessible. 
It is not known, whether there exists a method to obtain the full topological data from the ground state. 
For the simplest quantum Hall states, the Laughlin states,  it was shown that the ES does, in fact, contain the full information about the topological phase~\cite{hermanns2011haldane}.
However, it is far from clear if this holds for more complicated, in particular nonabelian, states as well.

Here, we address a much simpler question: what information can be extracted from the ES of \emph{non-interacting} systems?
The latter can be computed very efficiently, using methods developed by Peschel and others~\cite{Peschel2003}. 
It was subsequently shown for topological insulators/superconductors that the ES for (gapped) periodic systems is equivalent to the flat-band energy spectrum of the corresponding system with open boundaries~\cite{Fidkowski2010entanglement}. 
The same correspondence was also found for closely related gapless systems~\cite{Matern2018entanglement}
However, even for non-interacting systems, it is unclear which information (beyond the  `edge' spectrum) is encoded in the entanglement spectrum.

In this manuscript, we show that the information about the Zak phase, or polarization, can be completely recovered from the entanglement spectrum and provide a simple formula to do so. 
The relation between the entanglement spectrum and various topological invariants has already been studied in the literature. 
In Ref.~\cite{Zaletel2014}, the authors show that the Zak phase can be computed from the Schmidt decomposition of a translation invariant, infinite chain. 
This demonstrates that there should be a direct relation between the single-particle entanglement energies and the Zak phase, although deriving that relation turned out to be non-trivial. 
An added benefit of our approach is that it can be generalized to systems that are not translationally invariant in a numerically efficient way \carlos{Does using the MBES imply this is no longer correct?}. 
A particular limit of our result for homogeneous systems~\eqref{eq:chi_integrable} was derived by Ryu and Hatsugai for a fully dimerized SSH chain~\cite{Ryu2006}. 
The similarity between the behavior of the Zak phase and the 'virtual edge modes' in the entanglement spectrum of a Chern insulator, which was observed in Ref.~\cite{Huang2012,Huang2012-2}, can also be explained by our results.

\carlos{I added this last part}
Although the polarization obtained via this method for infinite systems or systems with periodic boundary conditions is only well-defined modulo 1 one can compute it in the equivalent open chain to obtain a polarization defined as a real number \carlos{Not sure about this expression. I mean it's not defined on a circle, not that is not a complex number, which it trivially isnt.}. We use this method to compute Chern numbers, which can also serve as topological invariants in one-dimensional systems, and for the system studied we show that it is equivalent to the winding number.

\emph{Outline of the paper}: In section II we introduce the model used throughout the paper and we discuss its phase diagram. In section III we introduce the polarization, the different ways one can define it and its relation to the Zak phase. In section IV we introduce the entanglement spectrum and set the notation that will be used. In section V we introduce the method and show how it reproduces the Zak phase for both systems with and without translational invariance. In section VI we explain how the method applied to the opened chain allows us to compute Chern numbers and winding numbers and finally in section VII we present the conclusions and possible outlook.

% - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % 
\section{Model.}
To showcase our results we consider as a starting point the model in the BDI class used in reference~\cite{Song2014}. 
This is a very simple model which nevertheless supports a rich phase diagram containing topological phases with winding numbers up to $\nu =2$. 
In order to show that our results are not limited to any symmetry class~\cite{ryu2010topological}, we also include two symmetry breaking terms. 
The Hamiltonian is then given by
\begin{align}
H =& \sum_{i\alpha,j\beta} c_{i\alpha}^\dagger H_{ij,\alpha \beta} c_{j\beta} \\
H_{ij} =& (m \sigma_x + \kappa \sigma_z)\delta_{ij}  + \frac{1}{2i}\kappa'\sigma_z (\delta_{i-j,1}-\delta_{i-j,-1})\nonumber\\
&+ \frac{1}{2} t \left[(\sigma_x + i \sigma_y)\delta_{i,j+1} + (\sigma_x - i \sigma_y) \delta_{i,j-1} \right] \nonumber\\
&+  \frac{1}{2} t' \left[(\sigma_x + i \sigma_y)\delta_{i,j+2} + (\sigma_x - i \sigma_y) \delta_{i,j-2} \right],
\label{bdi_model}
\end{align}
with the corresponding Bloch Hamiltonian being
\begin{align}
H(k)=&\mqty( \kappa + \kappa' \sin(k) & t' e^{i2k} + t e^{ik}+m \nonumber\\t' e^{-i2k} + t e^{-ik}+m & -\kappa-\kappa' \sin(k)  )\nonumber \\
=& (t' \cos(2k)+t\cos(k)+m)\sigma_x \\
&+ (-t' \sin(2k)-t\sin(k))\sigma_y + (\kappa+\kappa'\sin(k)) \sigma_z.\nonumber
\end{align}

For $\kappa = \kappa' = 0$ this Hamiltonian is in the BDI class, i.e. it has time-reversal ($T$), particle-hole ($C$), and chiral symmetry ($S$):
\begin{alignat}{2}
&T = \mathcal{K} ; \quad &&T H(-k) T^{-1} = H(k) \nonumber\\
&C = \sigma_z\mathcal{K} ; \quad &&C H(-k) C^{-1} = -H(k) \nonumber\\
&S = \sigma_z ; \quad &&S H(k)S^{-1} = -H(k) .
\end{alignat}
Using the results of Ref.~\cite{Song2014}, the corresponding phase diagram can be calculated and is shown in Fig.\ref{fig:bdi_phase_diagram} for $t=1$. 
\begin{figure}[t]
	\centering
	\includegraphics[width=45mm]{phase_diagram2.pdf}
	\caption{Phase diagram for the system in the $BDI$ class showing the winding number of each region. Plotted for parameters $t = 1,\kappa =\kappa'=0.$}
\label{fig:bdi_phase_diagram}
\end{figure}
The BDI class  is characterized by a $\mathbb{Z}$ invariant, the winding number,  in one spatial dimension.
When imposing open boundary conditions on the system, the number of symmetry-protected zero energy edge modes is equal to the winding number. 
The behavior of the  edge modes  is  shown in Fig.~\ref{fig:zero_E_modes}, using the dotted path ($t=1,t'=-2$) marked in the phase diagram. 
Subfigure (a) shows the edge spectrum in the BDI class, with 4 (resp. 2) symmetry-protected zero modes for winding number 2 (1). 
%We also a cut through the three phases (dotted line in the phase diagram) for different values of $\kappa,\kappa'$.

For $\kappa' \neq 0$ only particle-hole symmetry $C$ is preserved and the system belongs to symmetry class $D$. 
In one spatial dimension, the latter is characterized by a $\mathbb{Z}_2$ invariant.  
Consequently, adding such a term causes the zero modes in the $\nu=2$ phase to split  pairwise and the $\nu = 2$ and $\nu=0$ phases become equivalent. 
This is shown in subfigure (b), where one clearly sees that the zero modes below $m=1$ are split.

For $\kappa \neq 0$ only time-reversal $T$ is preserved and the system is in the AI class. 
In one dimensions, this class is trivial and, thus,  all zero-energy modes split. 
With both $\kappa \neq 0$ and $\kappa' \neq0$ the system has no local symmetry, thus belonging to the $A$ class.  
Also this symmetry class is topologically trivial in one dimension. 
Subfigures (c) and (d) show the edge spectrum for class AI and A, respectively. In both cases, the edge modes are split from zero for all values of $m$.

%
%In subfigure (b) and (d), where the system is in a topologically trivial phase, all edge modes are gapped. 
%Subfigure (c) shows the edge spectrum of the system being in class D, where the $\nu=2 $ phase becomes trivial. 
%Consequently, symmetry-protected zero modes only exist for $1<m<2.5$\mariac{?} (corresponding to $\nu=1$ for $\kappa'=0$) , whereas the modes are split for $m<1$ (corresponding to $\nu=2$ for $\kappa'=0$).

\begin{figure}[h!]
\centering
%\makebox[0pt]{
%\subfloat[$t = 1,\kappa =\kappa'=0$]{
%  \includegraphics[width=45mm]{phase_diagram2.pdf}
%}}\hspace{0mm}

\makebox[0pt]{
\subfloat[$\kappa =\kappa'=0$]{
  \includegraphics[width=35mm]{2a.pdf}
}
\subfloat[$\kappa = 0, \kappa'=0.3$]{\includegraphics[width=35mm]{2b.pdf}
}
}\hspace{0mm}

\makebox[0pt]{
\subfloat[$\kappa = .3,\kappa'=0$]{
  \includegraphics[width=35mm]{2c.pdf}
}
\subfloat[$\kappa = 0.3,\kappa'=0.3$]{\includegraphics[width=35mm]{2d.pdf}
}
}
\caption{\carlos{Use another colour for the blue edge mode.}Edge spectrum along the dotted path (i.e. $t=1,t'=-2$) of Fig.~\ref{fig:bdi_phase_diagram} for different values of $\kappa$ and $\kappa'$.}
\label{fig:zero_E_modes}
\end{figure}

% - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % 
\section{Geometric phases and polarization}

In this section we review some results of the modern theory of polarization, which gave a geometrical description of the polarization \cite{Resta1992,KingSmith1993,Vanderbilt1993,Resta1997} , as well as a recent article by Watanabe and Oshikawa \cite{Watanabe2018} that clarified the meanings of the different polarizations one can define this way and the relations between them. For clarity we will use the same notation they use.

Consider a generic, quadratic Hamiltonian in one dimension
\begin{equation}\label{eq:quadr_Ham}
\mathcal{H} = \sum_{ij,\alpha\beta} c_{i\alpha}^\dagger H_{ij,\alpha \beta}c_{j\beta}.
\end{equation}
We can obtain its single-particle eigenstates as
\begin{align}
\sum_{j\beta}H_{ij,\alpha\beta} u_{p\mu}^{j\beta} = E_{p\mu} u^{i\alpha}_{p\mu},
\end{align}
where $[U]_{j\beta,p\mu} = u_{p\mu}^{j\beta}$ is the unitary matrix that diagonalizes $H$. In translational invariant systems $u_{p\mu}^{j\beta} = e^{ipj}\tilde{u}_{p \mu}^{\beta}$, where $\tilde{u}_{p \mu}^{\beta}$ are the components of the eigenstates of the Bloch Hamiltonian, $\ket{\tilde{u}_{k\mu}}$. One can now compute the Zak phase \cite{Zak1989}, which for a two-band model is defined as the geometric phase acquired by the occupied state $\ket{\tilde{u}_k}$ as it winds around the Brillouin zone,
\begin{equation}
\gamma = \int_{0}^{2\pi} dk\, i A_k, 
\label{eq:zak_phase}
\end{equation}
where $A_k = \bra{\tilde{u}_k}\partial_k \ket{\tilde{u}_k}$ is the Berry connection \cite{Berry1984}. The Zak phase is typically regarded as just a phase because it is only gauge invariant modulo $2\pi$. However, for two different states defined by a parameter $\lambda$, as long as $A_k(\lambda)$ is smooth in the path connecting them the change in the Zak phase can be computed as
\begin{equation}
\Delta {\gamma_{\lambda_i \lambda_f}} = \int_{\lambda_i}^{\lambda_f}\int_{0}^{2\pi} dk \, \Omega_{\lambda k},
\label{eq:change_zak_phase}
\end{equation}
where $\Omega_{\lambda k} = \partial_\lambda A_k(\lambda,k) - \partial_k A_\lambda(\lambda,k)$ is the Berry curvature. The Berry curvature is fully gauge invariant and therefore the change in the Zak phase as defined in \eqref{eq:change_zak_phase} is defined as a real number. The Zak phase itself, as defined above, can be shown to be proportional to the winding number for a certain gauge (see Appendix \ref{appendix:winding_ssh}), which means that it carries physical information beyond modulo $2\pi$. Note that other ways of computing the Zak phase rely on computing $e^{i\gamma}$ instead, such that the result is always defined only modulo $2\pi$.

The derivative of the polarization with respect to some parameter $\lambda$ was obtained \cite{KingSmith1993} as
\begin{equation}
\partial_\lambda \tilde{\mathcal{P}}^{\rm Bloch} = \int_{0}^{2\pi} \frac{dk}{2\pi} \, \Omega_{\lambda k},
\end{equation}
which allows us to make the identification $ \tilde{\mathcal{P}}^{\rm Bloch} = \gamma/2\pi$. As discussed above this polarization is defined as an absolute value only modulo 1, however its changes can be defined as a real number. This is no surprise as the derivative of the polarization is related to the current which is a physical observable, unlike $\tilde{\mathcal{P}}^{\rm Bloch}$.

One can also compute the geometric phase obtained by threading a $U(1)$ flux through the ring \cite{Watanabe2018}, which allows us to compute a polarization in the absence of translational invariance. If the flux is introduced homogeneously through a vector potential $A_x = \Phi/L$ an equivalent polarization can be defined as
\begin{equation}
\mathcal{P} = \int_0^{2\pi}\frac{d\Phi}{2\pi} i\bra{\Psi_0^\Phi}\partial_\Phi \ket{\Psi_0^\Phi} + {\rm Im}{\rm Ln} \bra{\Psi_0^0}e^{2\pi i \hat{P}} \ket{\Psi_0^{2\pi}}, 
\end{equation}
where $\ket{\Psi_0^{\Phi}}$ is the ground state in the presence of flux and $\hat{P}=\frac{1}{L}\sum_{j\alpha} j\hat{n}_{j\alpha}$ is the polarization operator. This is equivalent to the polarization obtained by Resta \cite{Resta1997} as
\begin{align}
\mathcal{P} = \frac{1}{2\pi} \rm{ Im }\,{\rm Ln}\bra{\Psi_0}e^{2\pi i \hat{P}}\ket{\Psi_0}.
\end{align}
This expression is particularly useful because it can be easily expressed in terms of single-particle eigenstates as
\begin{align}
\mathcal{P} = \frac{1}{2\pi} \rm{Im \,Ln} \, {\rm det }\!' \, S,
\label{eq:polarization_resta}
\end{align}
where the matrix $S$ is given by
\begin{equation}
S_{p\mu,q\nu} = \sum_{j\alpha} u_{p\mu}^{j \alpha \, \ast} e^{i\frac{2\pi}{L}j}u_{q\nu}^{j \alpha},
\end{equation}
\carlos{I have to check overall sign and sign of exponential.}and  ${\rm det }\!' $ indicates that the determinant is restricted to the space of \emph{occupied} single-particle states. Below we will employ this expression to compare our results with.

There is one last polarization we must consider and it is the one obtained when the flux is introduced in the bond between sites $i=1$ and $i=L$ via twisted boundary conditions. This is equivalent to performing the gauge transformation $\ket{\tilde{\Psi}_0^\Phi}=e^{i\Phi \hat{P}}\ket{\Psi_0^\Phi}$, which makes $\ket{\tilde{\Psi}_0^\Phi}$ fully periodic in $\Phi$. We can now define the polarization
\begin{equation}
\tilde{\mathcal{P}} = \int_0^{2\pi}\frac{d\Phi}{2\pi} i\bra{\tilde{\Psi_0}^\Phi}\partial_\Phi \ket{\tilde{\Psi_0}^\Phi},
\label{eq:p_tilde}
\end{equation}
which is related to the current flowing through the bond \cite{Watanabe2018}. This polarization is related to the homogeneous polarization by 
\begin{equation}
\mathcal{P}= \tilde{\mathcal{P}}+\tilde{\mathcal{P}}_0,
\end{equation} 
where
\begin{equation}
\mathcal{P}_0= \int_0^{2\pi}\frac{d\Phi}{2\pi} \bra{\tilde{\Psi_0}^\Phi}\hat{P}\ket{\tilde{\Psi_0}^\Phi}.
\end{equation} 
The previous discussion about the Zak phase being defined only modulo $2\pi$ although its changes can be defined as a real number applies in the same way to the polarization obtained in \eqref{eq:p_tilde}.

\section{Entanglement spectrum}
\label{section:ES}
Let us now proceed to discuss the entanglement spectrum. 
An important quantity is the correlation matrix, which in position space is defined by
\begin{align}
C_{ij}^{\alpha \beta} = \expval{c_{i\alpha}^\dagger c_{j\beta}}.
\end{align}
For a system where we occupy all single-particle eigenstates with energies $E<0$ the correlation matrix can be written  as
\begin{align}\label{eq:corr_mat2}
C_{ij}^{\alpha \beta} =& \frac{1}{2}\left[I - H/ (H^2)^{-1/2} \right]_{ij, \alpha \beta}.
\end{align}
This form is particularly useful for all our subsequent calculations, since it does not require diagonalizing the Hamiltonian.

In this manuscript, we are mainly interested in the spectrum of the correlation matrix, when restricted to a spatial sub-region $A$ (its complement will be denoted by $B$ in the following). 
Following Ref.s~\cite{Huang2012,Huang2012-2}, we refer to this spectrum as the entanglement occupancy spectrum (EOS). 
The resulting eigenvalues  are denoted by $\xi_j$. 
%We will differentiate between two types of eigenvalues: those (exponentially) close to $\xi_j=0 (1)$ correspond to ``bulk'' states, by which we mean that the corresponding eigenstates lie (up to exponentially small corrections) in  $A$ ($B$). 
%Values $0<\xi_{l,r}<1$ indicate states that are located at the virtual boundaries -- labeled `left' and `right' in the following-- between $A$ and $B$~\cite{Peschel2008}. 
%There is, of course, an ambiguity here on dividing eigenstates of the EOS into ``bulk'' and ``edge'' states and `how close' the eigenvalue has to be to $0,1$ in order to qualify as a bulk state. 
%XXXX
Topological phases are characterized by (symmetry-protected) zero-energy modes in the edge spectrum and $\xi=\frac 1 2$ modes in the EOS~\cite{Fidkowski2010entanglement}.

The EOS is in one-to-one correspondence to the entanglement spectrum (ES). 
Following Ref.~\cite{Peschel2003}, one can write the  reduced density matrix as 
\begin{align}\label{eq:red_dens_mat}
\rho_A&=\mathcal{K} \exp(-\mathcal H),
\end{align}
where $\mathcal{K}$ is a normalization constant and $\mathcal{H}$ is a quadratic Hamiltonian, referred to as entanglement Hamiltonian. 
The eigenvalues $\epsilon_j$ of $\mathcal{H}$, referred to as entanglement energies,  are related to the EOS by 
\begin{align}\label{eq:xi_eps}
\xi_j &=\left(e^{\epsilon_j}+1\right)^{-1}, 
\end{align}
the corresponding eigenstates are the same. 
Eq.~\eqref{eq:xi_eps} implies that the full information of the \emph{many-particle} ES is contained in the single-particle spectrum of the reduced correlation matrix. 
The latter is much simpler to interpret. 
Consequently, we will focus on the EOS in the remainder of the manuscript and only mention the ES when necessary.

One property of the subsystem correlation matrix we will make use of later is that, for sufficiently large systems, its eigenstates are found to be localize on either virtual edge if $\xi \neq 0,1$, or they are found to be bulk modes if the correspondent eigenvalues are exponentially close to $\xi = 0,1$ \cite{Peschel2008}.

%The single-particle entanglement spectrum (ES) is then obtained as the spectrum of the subsystem correlation function obtained after bisecting the system into two equal parts.

%In the ES we can differentiate between two types of eigenvalues, which we denote by $\xi$. The ones at $\xi_b = 0,1$ correspond to bulk states while for the ones in between, $0<\xi_{l,r}<1$, we find their correspondent eigenstate localized in either the left, $\xi_l$, or right, $\xi_r$, virtual edge \cite{Peschel2008}. 
%\mariac{introduce entanglement energy, relation between C and energies}

A few comments are needed regarding  bipartition. 
In the following sections, we show how the Zak phase can be recovered from the EOS.
We also give a simple formula that is identical to the Zak phase in the thermodyamic limit.  
However, for a generic, finite size system, there will be finite-size discrepancies between our formula and the Zak phase.
In order to reduce these, we are going to choose $A$ as half the system in the remainder of the manuscript.
The results (in the thermodynamic limit) do not depend on this choice. 
% - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % 
\section{Zak phase in the EOS}

Let us first review previous results on the relation between the entanglement spectrum and the Zak phase. 
For symmetry-protected topological systems in one dimension, the quantized Zak phase is zero whenever there is an even number  of eigenvalues per edge at $\xi = 1/2$, and $\pi$ when the number is odd~\cite{Peschel2008}. 
Ryu and Hatsugai considered the fully dimerized SSH chain with broken chiral symmetry~\cite{Ryu2006}.  
This model is rather special in that there is only a single pair of eigenvalues, related by symmetry, in the EOS that is not identical to $0$ or $1$. 
In fact, the authors could show that the value of one of these `midgap states' is identical to the Zak phase divided by $2\pi$, or $\tilde{\mathcal{P}}^{\rm Bloch}$. 
Their result is a particular limit of Eq.~\eqref{eq:chi_integrable}. 
%
%Beyond symmetry protected systems, there was observed that in the fully dimerized SSH chain, with only a single mode present in the ES, the Zak phase is equal to the midgap eigenvalue \cite{Ryu2006}. 
Huang and Arovas considered a two-dimensional chern insulator and noted that the Zak phase qualitatively follows the `virtual edge mode' that connects the EOS values at 0 to those at 1~\cite{Huang2012,Huang2012-2}. 
Interpreting this as a one-dimensional system with a parameter, we can explain this behavior by Eq.~\eqref{eq:chi_integrable}, noting that there is a single pair of eigenvalues that dominates the sum. 
Note that the observation of Ref.s~\cite{Huang2012,Huang2012-2} is particular to a system with Chern number 0 or $\pm 1$, and it generically fails for systems with higher Chern numbers. 
In the latter case, there are several terms in Eq.~\eqref{eq:chi_integrable} with comparably large contributions. 
%For cases with more edge modes present in the ES it has been shown that the Zak phase follows the midgap eigenvalues qualitatively \cite{Huang2012,Huang2012-2}.

\subsubsection{Systems with equidistant entanglement energies}
\mariac{add scaling plot for $\chi$ versus $\gamma/(2\pi)$. }
We first discuss systems, for which the entanglement energies of a \emph{single} virtual edge are equidistant. This is a feature of integrable systems \cite{Peschel1999} for e.g. nearest neighbor hopping models such as the SSH chain, and applies to the Hamiltonian~\eqref{bdi_model} when $t'=0$. 
In this case, one can obtain the Zak phase in a very simple fashion: We reorder the eigenvalues of the ES by magnitude, $\xi_1 < \xi_2 < ...< \xi_{L_AM}$, and compute
\begin{equation}
\chi = \sum_{j=1}^{L_A M/2} \xi_{2j-1} \, {\rm mod} \, 1.
\label{eq:chi_integrable}
\end{equation}
In the thermodynamic limit, $\chi$ becomes identical to either $\tilde{\mathcal{P}}^{\rm Bloch}$ or $1-\tilde{\mathcal{P}}^{\rm Bloch}$
\begin{equation}
\lim_{L \rightarrow \infty} \abs{\tilde{\mathcal{P}}^{\rm Bloch}} = \lim_{L \rightarrow \infty} \abs{\chi}.
\end{equation}
For translational invariant systems the EOS is symmetric such that a crossing of the eigenvalues $\xi$ results in $\chi$ changing sign. This is particularly irrelevant in the case of symmetry-protected states where the Zak phase is quantized to $0$ or $\pi$.

In the case where there is only one eigenvalue per edge that is $\neq 0,1$ in  the EOS, like in the fully dimerized limit of the SSH chain, we recover the equality between the midgap eigenvalue and the Zak phase~\cite{Ryu2006}.

In Fig.~\ref{huang} we show  the EOS along the dashed cut in the phase diagram of Fig.~\ref{fig:bdi_phase_diagram}. 
Along this line, our model is equivalent to an SSH chain. 
In the presence of chiral and translation symmetry, all eigenvalues of the EOS are doubly degenerate --- one eigenvalue from each edge, such that $\chi$ from \eqref{eq:chi_integrable} will pick up one eigenvalue from each pair. Because of chiral symmetry the sum of the two eigenvalues related by the symmetry gives always $1$ so they do not contribute to $\chi$. The only contribution to $\chi$ is the one from the eigenstates with $\xi=1/2$ that do not have a chiral partner, see Fig.~\ref{huang}(a). Here we see explicitely the known relation between the Zak phase and the number of $\xi = 1/2$ modes.

Note that the full sum~\eqref{eq:chi_integrable}  evaluates to zero in the trivial regime %up to finite-size corrections% \carlos{It is actually exact due to symmetry}.
despite the presence of modes with entanglement occupancy $0<\xi<1$.

When breaking chiral symmetry, the EOS will no longer be doubly degenerate. 
For small $m$, there are two dominant midgap eigenvalues, and the Zak phase follows the lower one very closely, see Fig.~\ref{huang}(b).
When this midgap eigenvalue approaches  0, the contribution from the other modes becomes more relevant and the Zak phase starts to deviate from the midgap eigenvalue, \mariac{see inset}.
\begin{figure}[t]
\centering
\makebox[0pt]{
\subfloat[$t = 1,t'=0$, \, \,$\kappa =\kappa'=0$]{
  \includegraphics[width=35mm]{2_a.pdf}
}
\subfloat[$t=1,t'=0$, $\kappa = 0.3, \kappa'=0$]{\includegraphics[width=35mm]{2_b.pdf}
}
}
\caption{Entanglement spectrum (black), $\chi$ (green) and $\frac{\gamma}{2\pi}$ (red) for a cut in the phase diagram through the $\nu = 1 \rightarrow \nu = 0$ transition (dashed line in~\ref{fig:bdi_phase_diagram}
	) for $t'=0$ and different values of $\kappa$. Computed for $L=40$.}
\label{huang}
\end{figure}

\subsubsection{General case}

In general we find that it is only the eigenstates that localize on the left edge that should contribute to $\chi$. The inclusion of bulk modes is irrelevant since they have $\xi = 0,1$. We can then compute the average position of each eigenstate and construct $\chi$ from the eigenstates with $\expval{\hat{x}}<L/4$, with $L/4$ being the middle point of region $A$,
\begin{equation}
\chi = \sum_{\mu \in A_L} \xi_\mu \, {\rm mod} \, 1. 
\label{eq:chi_left}
\end{equation}
In the thermodynamic limit it is $\tilde{\mathcal{P}}$ that we actually identify with $\chi$
\begin{equation}
\lim_{L \rightarrow \infty} \tilde{\mathcal{P}} = \lim_{L \rightarrow \infty} \chi,
\label{eq:ptilde_eq_chi}
\end{equation}
however for translationally invariant systems $\tilde{\mathcal{P}} = \tilde{\mathcal{P}}^{\rm Bloch}$ \cite{Watanabe2018}. Note that in order to compute $\chi$ correctly the eigenstates with $\xi \neq 0,1$ need to be localized in order for us to select the correct eigenvalues. It might be, like in Fig.\ref{huang}(a) that the spectrum has degenerate pairs of left and right eigenstates and the numerical diagonalization mixes them. This case is trivial as we know that there is one left and one right eigenstate but if the degeneracy is four-fold or higher one needs to obtain the localized eigenstates. Localizing the eigenstates that are exponentially close to $\xi = 0,1$ is more difficult but not doing so only results in an exponentially small error.

Before proceeding lets review again the result in the previous section. First note that for translation invariant systems, the ES of the right and the left edge are related by an overall '-' sign, i.e. for each eigenvalue $\epsilon$ on the right edge, there is a corresponding value $-\epsilon$ on the left edge. 
Since the spectra are equidistant, this implies that the two edge spectra are merely shifted with respect to each other.
If the shift is zero, e.g.  in presence of a symmetry, summing every other eigenvalue is equivalent, modulo 1, to summing all the left eigenvalues. With a finite shift summing all odd eigenvalues corresponds to either summing all left eigenvalues or all right eigenvalues. In general, it implies that $\chi$ as defined in Eq.~\eqref{eq:chi_integrable} only gives the Zak phase up to an overall sign. This issue can be fixed by determining the localization of a single eigenstate. 
\begin{figure}[t]
	\centering
	\makebox[0pt]{
		\subfloat[$t = 1, t' = 2,$   $\kappa =0.3, \kappa' = 0$]{
			\includegraphics[width=35mm]{3_a.pdf}
		}
		\subfloat[$t = 1, t' = 2,$ $\kappa = 0, \kappa' = 0.3$]{
			\includegraphics[width=35mm]{3_b.pdf}
		}
	}\hspace{0mm}

	\makebox[0pt]{
		\subfloat[$t = 1, t' = -2,$ $\kappa = 0.3, \kappa' = 0$]{
			\includegraphics[width=35mm]{3_c.pdf}
		}
		\subfloat[$t = 1, t' = -2,$ $\kappa = 0, \kappa' = 0.3$]{
			\includegraphics[width=35mm]{3_d.pdf}
		}
	}\hspace{0mm}

	\makebox[0pt]{
		\subfloat[$t = 1, t' = -2,$ $\kappa_i \neq 0 , \kappa' = 0$]{
			\includegraphics[width=35mm]{3_e.pdf}
		}
	}
	\caption{(a,b,c,d) Entanglement spectrum (black), $\chi$ (green) and $\tilde{\mathcal{P}}$ (red) for two different cuts with both $\kappa$ and $\kappa'$ symmetry breaking terms. Note that in this case where there is translational invariance $\tilde{\mathcal{P}}=\tilde{\mathcal{P}}^{\rm Bloch}$, which is nothing else than the Zak phase over $2\pi$. (e) Same as (c) but with a position dependent $\kappa_i=(i+L/4 \, {\rm mod}\, L)/L$. We also plot $\mathcal{P}+1/2$ (blue) to showcase that $\chi$ is indeed equal to $\tilde{\mathcal{P}}$ and not $\mathcal{P}$ as in the translationally invariant case $\tilde{\mathcal{P}}=\mathcal{P}+(L+1)/2$. }
	\label{2}
\end{figure}

In Fig.\ref{2} we show two examples that highlight the importance of the localization structure of the EOS. 
In subfigures (a) and (b), we compute the EOS for varying $m$ at  $t=1$, $t'=2$, indicated by the solid line  in the phase-diagram in Fig.~\ref{fig:bdi_phase_diagram}, in the presence of two different symmetry breaking terms: (i) $\kappa =0.3$, $\kappa'=0$ and (ii) $\kappa=0$, $\kappa'=0.3$.  
For $\kappa=\kappa'=0$, there is a phase transition $\nu = 2 \rightarrow 0 $ at $m=2$, whereas both choices of symmetry breaking render the system trivial  for all $m$.
However, the two symmetry breaking terms split the four-fold degenerate $\xi=1/2$ midgap states very differently, thus resulting in completely different Zak phases despite the superficial resemblance of their respective EOS.  
While $\kappa'$ splits the four-fold degenerate states into two two-fold degenerate pairs at opposite edges whose contribution cancels (modulo 1), $\kappa$ split them into two $left-left$ and $right-right$ pairs, with a finite contribution to the Zak phase. 
In subfigures (c) and (d), we consider the same symmetry breaking terms, but now for $t=1$ and $t'=-2$, where the BDI systems shows a phase transition  $\nu=2\rightarrow 1\rightarrow 0$. 
Note that the $\kappa'$ term does not render the $\nu=1$ phase trivial, and the Zak phase will remain quantized in that region, see Fig.~\ref{2}(d). 
Also here, Eq.~\ref{eq:chi_left} correctly reproduces the Zak phase for all values of $m$. 
Note that in all examples shown in this subsection, the observations made in references \cite{Huang2012,Huang2012-2}, i.e. that the behavior of the Zak phase resembles that of the entanglement occupancy midgap states, do not apply anymore.

Since $\chi$ is naturally defined in position space it is also valid for systems without translationally invariance. We show this in Fig.\ref{2}(e) where we show the EOS, $\chi$, $\mathcal{P}$ and $\tilde{\mathcal{P}}$ along the dotted cut in the phase diagram with an added position dependent symmetry breaking term $\kappa_i = (i+L/4 \, {\rm mod}\, L)/L$ (\carlos{I added this profile because with disorder $\tilde{\mathcal{P}}$ and $\mathcal{P}$ are difficult to distinguish}). We obtain that $\chi$ is indeed identified with $\tilde{\mathcal{P}}$ as we hinted at before, and not $\mathcal{P}$, although these two polarization only differ by $(L+1)/2$ in the translationally invariant case.

This relation between the Zak phase and the EOS is a consequence of an identity between the Zak phase and the many-body entanglement spectrum previously found for infinite chains \cite{Zaletel2014}. In Appendix A we modified this derivation to account for the periodic boundary conditions we use and show how their result simplifies greatly when expressing it in terms of the EOS, resulting in.~\eqref{eq:ptilde_eq_chi}.

% - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - %

\section{Chern number and Winding number}

In Appendix A we obtain that $\chi$ should be the sum of all the eigenvalues of eigenstates related to the cut between $i=1$ and $i=L$ that make a subspace $A_L$. The dimension of $A_L$ is however not well-defined as one can always include bulk modes into it. Since the bulk modes have $\xi = 0,1$ this makes $\chi$ well-defined only modulo 1. This can be avoided by computing instead
\begin{equation}
\tilde{\mathcal{P}}_{\rm open} = \int_0^{2\pi}\frac{d\Phi}{2\pi} i\bra{\tilde{\Psi}_{0,{\rm open}}^\Phi}\partial_\Phi \ket{\tilde{\Psi}_{0,{\rm open}}^\Phi}.
\end{equation}
where $\ket{\tilde{\Psi}_{0,{\rm open}}^0}$ is the ground state of the chain opened between sites $i=L/2$ and $i=L/2+1$ and 
\begin{equation}
\ket{\tilde{\Psi}_{0,{\rm open}}^\Phi} =e^{-i \Phi \hat{N}_A} \ket{\tilde{\Psi}_{0,{\rm open}}^0}.
\label{eq:gauge}
\end{equation}
The difference with the closed chain being that now there is no ambiguity on when an electron crosses the boundary with the twisted boundary condition. Note that since the chain is open $\Phi$ is not a flux but should now be regarded as a parameter and therefore $\tilde{\mathcal{P}}_{\rm open}$ does not need to be a physical polarization.  $\tilde{\mathcal{P}}_{\rm open}$ can be obtained in an even simpler way than $\tilde{\mathcal{P}}$ as
\begin{align*}
\tilde{\mathcal{P}}_{\rm open} =&  \int_0^{2\pi}\frac{d\Phi}{2\pi} \bra{\tilde{\Psi}_{0,{\rm open}}^0}  \hat{N}_A \ket{\tilde{\Psi}_{0,{\rm open}}^0} \\
=&\sum_{\alpha,j\in A}\bra{\tilde{\Psi}_{0,{\rm open}}^0} c_{j\alpha}^\dagger c_{j\alpha} \ket{\tilde{\Psi}_{0,{\rm open}}^0} \\
=&\sum_{\alpha,j\in A} C_{\rm open,j\alpha,j\alpha} \\
=& {\rm Tr}[C_{A,\rm open}].
\end{align*}
Which is nothing else than the generalization of $\chi$ to the open chain. The idea is that in the process of oppening the chain, which can be done adiabatically, all eigenvalues that are not in $A_L$ should get pushed to $\xi=0$ or $\xi=1$. If for a given path in parameter space $\boldsymbol{\lambda}_i \rightarrow \boldsymbol{\lambda}_f$ their contribution remains constant, that is, there are no jumps between $\xi=0$ and $\xi=1$ and $\tilde{\mathcal{P}}_{\rm open} (\boldsymbol{\lambda} )$ is smooth, then changes in $\tilde{\mathcal{P}}_{\rm open}(\boldsymbol{\lambda}_i ) $ along the path will be defined as a real number and not only modulo 1. Since $\tilde{\mathcal{P}}$ is a local property of the bond between $i=1$ and $i=L$ one would expect it to not depend on what happens far from it such that $\tilde{\mathcal{P}}_{\rm open}$ and $\tilde{\mathcal{P}}$ are equivalent.

\subsubsection{SSH chain}

Consider the simple case of the SSH chain, with $t'=0$ in the model. In the chiral symmetric case, $\kappa = 0$, the open chain presents zero-energy modes for $m<t$ that seem problematic. $\tilde{\mathcal{P}}_{\rm open}$  will depend on the occupation of these zero modes although we just argued that $\tilde{\mathcal{P}}$ should not, since the zero-energy modes are localized far from the virtual cut at $i=1$. We can break the chiral symmetry in order to split the zero-energy modes so we do not have to make this choice, but the resulting $\tilde{\mathcal{P}}_{\rm open}$ will be discontinuous at $\kappa = 0$, suggesting that the limit does not exist. This can be seen from the fact that the Hamiltonian now has the symmetry
\begin{equation}
SH(\kappa)S = -H(-\kappa), \quad S C_A(\kappa) S = \mathbb{I}-C_A(-\kappa)
\end{equation}
so $\tilde{\mathcal{P}}_{\rm open}(-\kappa) = L-\tilde{\mathcal{P}}_{\rm open}(\kappa)$. Unless $\tilde{\mathcal{P}}_{\rm open}(0) = L/2$, which only happens in the trivial case, the result will be discontinuous at $\kappa = 0$. We can see this in Fig.\ref{ssh_chern}.(a) where we show $\tilde{\mathcal{P}}_{\rm open}$ in the space $(m,\kappa)$. This branch cut is in fact not an artifice of the open chain trick, but the result of the degenerate point at $(m=t,\kappa=0)$ having an associated Chern number $C=1$ \cite{Asboth2016}.

Consider a loop around the singularity at $(m=t,\kappa=0)$ parametrized by $\theta$. The Chern number is then defined as
\begin{equation}
C = \int_0^{2\pi}d\theta\int_0^{2\pi}\frac{d\Phi}{2\pi} \Omega_{\theta\Phi}.
\label{eq:chern}
\end{equation} 
The Berry connection for the flux is simply $A_\Phi(\theta,\Phi,) = -i\tilde{\mathcal{P}}(\theta)$, so we now need to compute $A_\theta(\theta,\Phi)$. (\carlos{Please double check this calculation}) Expanding $\ket{\tilde{\Psi}_{0,{\rm open}}^0}$ in the eigenbasis of $\hat{N_A}$, $\{\ket{j}\}$ we can compute its derivative with respect to the parameter $\theta$ as
\begin{align*}
\partial_\theta \ket{\tilde{\Psi}_{0,{\rm open}}^0} =& \sum_j [\partial_\theta c_j (\theta)]e^{-i\Phi N_A^j}\ket{j}.
\end{align*}
where $c_j(\theta) = \bra{j}\ket{\tilde{\Psi}_{0,{\rm open}}^0}$. The correspondent Berry connection then gives
\begin{align*}
\bra{\tilde{\Psi}_{0,{\rm open}}^\Phi}\partial_\theta \ket{\tilde{\Psi}_{0,{\rm open}}^\Phi} =& \sum_{lj} c^*_l(\theta) [\partial_\theta c_j(\theta)] \bra{l} e^{i\Phi (N_A^lN_A^j)}\ket{j}\\
=& \sum_j  c^*_j(\theta) \partial_\theta c_j(\theta) \\
=&\bra{\tilde{\Psi}_{0,{\rm open}}^0}\partial_\theta \ket{\tilde{\Psi}_{0,{\rm open}}^0}
\end{align*}
and is therefore independent of $\Phi$.

The integral in \eqref{eq:chern} is performed on the surface of the torus defined by $(\theta,\Phi)$. Because the Berry connection $A_\Phi(\theta)$ is not smooth one cannot directly apply Stokes theorem. Instead, consider that we split the torus in two cylinders, $S_1$ with $\theta\in[\theta_0,2\pi-\theta_0]$ and  $S_2$ with $\theta\in[2\pi-\theta_0,2\pi+\theta_0]$ and we change the gauge in $\boldsymbol{A}(\theta)$ such that it is smooth in each cylinder. We can now apply Stokes theorem to the two cylinders independently
\begin{align*}
C_1 =& \int_{S_1} d\boldsymbol{S}\cdot (\nabla \times \boldsymbol{A}) \\
=& \int_{\partial S_1} d\boldsymbol{l}\cdot \boldsymbol{A}(\theta,\Phi) \\
=& \int_a^{2\pi-a} d\theta A_\theta(\theta,0) +\int_0^{2\pi} \frac{d\Phi}{2\pi} A_\Phi (2\pi-a,\Phi) \\
&-\int_a^{2\pi-a} d\theta A_\theta(\theta,2\pi) -\int_0^{2\pi} \frac{d\Phi}{2\pi} A_\Phi (a,\Phi).
\end{align*}
In our case, $A_\theta$ is independent of $\Phi$ so the terms involving it cancel and we have
\begin{align*}
C_1 =& \int_0^{2\pi} \frac{d\Phi}{2\pi} [A_\Phi (2\pi-a,\Phi) -A_\Phi (a,\Phi)].
\end{align*}
Similarly computing the integral for the other cylinder we obtain
\begin{align*}
C_2 =& \int_0^{2\pi} \frac{d\Phi}{2\pi} [A'_\Phi (2\pi+a,\Phi) -A'_\Phi (2\pi-a,\Phi)].
\end{align*}
The Chern number is then $C=C_1+C_2$. Noting again that $A'_\Phi(\theta)$ is continuous in $S_2$ taking the limit of $a\rightarrow 0^+$ makes $C_2$ vanish and the Chern number is given by
\begin{align*}
C =& \int_0^{2\pi} \frac{d\Phi}{2\pi} [A_\Phi (2\pi^-,\Phi) -A_\Phi (0^+,\Phi)].
\end{align*}
If we consider the loop starting at $\kappa = 0$ this is exactly the case shown in Fig.\ref{ssh_chern}.(a), where $A_\Phi (\theta,\Phi) = \tilde{\mathcal{P}}_{\rm open}(\theta)$ is smooth everywhere but in $\theta = 0$, and therefore we can compute the Chern number simply as 
\begin{align*}
C= \tilde{\mathcal{P}}_{\rm open}(\kappa=0^+)-\tilde{\mathcal{P}}_{\rm open}(\kappa=0^-).
\end{align*}
Using the symmetry of the polarization described above this results in 
\begin{align}
C= 2\tilde{\mathcal{P}}_{\rm open}(0^+)-L.
\label{eq:Chern_number}
\end{align}
The discontinuity in $\tilde{\mathcal{P}}_{\rm open}$ is therefore related to the Chern number and it is not merely continuous modulo 1.

One can also obtain a polarization continuous in $\kappa$ by doing a gauge transformation $\ket{\tilde{\psi}^{\prime \,\Phi}_{0}}=e^{ i \Phi \Theta(-\kappa)\Theta(t-m)}\ket{\tilde{\psi}^{\Phi}_{0}}$ that places the branch cut at $m=t,\kappa < 0$ as we see in Fig.\ref{ssh_chern}.(b)  (\carlos{need to double check signs}). This is in fact the result one gets when using the gauge that allows one to compute the winding number from the Zak phase.

\begin{figure}[t]
	\centering
	\makebox[0pt]{
		\subfloat[$\tilde{\mathcal{P}}_{\rm open}-L/2$]{
			\includegraphics[width=50mm]{windingSSH1.pdf}
		}
		\subfloat[$\tilde{\mathcal{P}}_{\rm open,+}-L/2$]{
			\includegraphics[width=45mm]{windingSSH2.pdf}
		}
	}\hspace{0mm}

	\makebox[0pt]{
		\subfloat[$\tilde{\mathcal{P}}_{\rm open}-L/2$]{
			\includegraphics[width=45mm]{windingtp1.pdf}
		}
		\subfloat[$\tilde{\mathcal{P}}_{\rm open,+}-L/2$]{
			\includegraphics[width=45mm]{windingtp2.pdf}
		}
	}
	\caption{Polarization $\tilde{\mathcal{P}}_{\rm open}$ computed for three different gauge choices, for $t=1,\kappa'=0$ and system size $L=160$. (a) and (b) with $t'=0$ and (c) and (d) for $t'=-2$. }
	\label{ssh_chern}
\end{figure}

\begin{figure}[t]
	\centering
	\makebox[0pt]{
		\subfloat[$\tilde{\mathcal{P}}_{\rm open}$]{
			\includegraphics[width=50mm]{windingtpdis.pdf}
		}
		\subfloat[$C $]{
			\includegraphics[width=45mm]{plotwinding.pdf}
		}		
	}\hspace{0mm}

	\makebox[0pt]{
		\subfloat[${\rm Log}_{10} (C \, {\rm mod} \, 1)$ ]{
			\includegraphics[width=45mm]{plotwindingerror.pdf}
		}
	}
	\caption{We compute (a)$\tilde{\mathcal{P}}_{\rm open}$, (b) $\nu$ and (c) error in the quantization of $\nu$ for the continuous line in the phase diagram, $t=1,t'=2$ and system size $L=400$. (a) is computed for $W=6$ showing that the branch cut is not affected by the disorder. (b) and (c) are computed for increasing disorder strength, showing a good quantization of the resulting winding number far from the transitions. }
	\label{winding}
\end{figure}

\subsubsection{$t' \neq 0$}

The analysis we did above of $\tilde{\mathcal{P}}_{\rm open}$ for the SSH chain seems rather trivial as $\tilde{\mathcal{P}}_{\rm open}$ does not seem to have more information than $\tilde{\mathcal{P}}$ modulo 1. This is no longer true when $t' \neq 0$ allowing for a $\nu = 2$ phase. In Fig.\ref{ssh_chern}.(c) we see $\tilde{\mathcal{P}}_{\rm open}$ for the dotted cut in the phase diagram with an added symmetry breaking $\kappa$. We see clearly that $\tilde{\mathcal{P}}_{\rm open}$ distinguises between the $\nu=0$ and $\nu = 2$ phases. Similar to before there is a gauge choice that makes $\tilde{\mathcal{P}}_{\rm open}$ continuous in $\kappa$, however which gauge transformation does this is a priori not known, so in practice we can only compute the limit $\lim_{\kappa \rightarrow 0} \tilde{\mathcal{P}}_{\rm open}$ in the original gauge.

Our expression for the Chern number \eqref{eq:Chern_number} is also valid without translational invariance. Consider now that we add disorder in the $t$ and $m$ parameters
\begin{align}
&t_i = t + \frac{W}{2} \omega_i \\
&m_i = m + W \omega'_i
\end{align}
where $\omega_i$ and $\omega_i$ are randomly selected from $[-1/2,1/2]$. For very strong disorder, where the gap fills with states, the limit $\lim_{\kappa \rightarrow 0} \tilde{\mathcal{P}}_{\rm open}$ might become problematic as bulk states can cross $E=0$ as we tune $\kappa$ down. In order to avoid it we break chiral symmetry locally in the edge of the open chain so it only affects the zero modes. If any of the other states cross $E=0$ they do it in pairs such that $\tilde{\mathcal{P}}_{\rm open}$ is not affected by it. We show $\tilde{\mathcal{P}}_{\rm open}$ in Fig.\ref{winding}.(a) using the local symmetry breaking term for a strong disorder, $W=6t$, where the gap is indeed filled with states. In Fig.\ref{winding}.(b) we show $C$ for the continuous line in the phase diagram and increasing disorder strength. The Chern number as defined in \eqref{eq:Chern_number} indeed provides a topological invariant which can be compared with the winding number obtained in reference \cite{Song2014} for the same system. Note that for large enough systems the spectrum of $C_A$ is not affected by the local chiral symmetry breaking term other than the occupation of the zero-energy modes such that $C$ is still quantized, however near the transitions it is smoothen by the disorder average We show the error in the quantization of $C$ in figure \ref{winding}.(c) .

\carlos{In Appendix D I give an expression for the winding number for flux insertion. However this expression seems problematic with a filled gap (Prodan refers to this). If we found an inhomogeneous term that induce a topological transition without filling the gap we can use this formula instead of comparing with the Prodan results, so we dont have to worry about the filled gap. 
Furthermore I worked an expression for this winding number in terms of the correlation function, which I believe we might be able to relate to our $\chi$. However I can only do this on the open chain, where the winding number itself is not well-defined because the Hamiltonian is not invertible. We can do the trick of splitting the edge-modes and arguing that the winding number should not change by doing that because its a  bulk property, but I'm not too sure about that, so it would be better if we found a way of computing the winding number for open chains.   }

% - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - %

\section{Conclusion}

It was previously known that the Zak phase, or the geometric phase for $U(1)$ flux insertion, was encoded in the many-body entanglement spectrum. Here we have shown how it is encoded in the single-body entanglement spectrum. This relation is simpler than the one involving the MBES and it provides more insight into the ES. Based on this method we show that one can obtain a polarization that is defined in the real axis. Although it is not gauge invariant it provides a method for computing Chern numbers for systems without translational invariance. We have also shown that the Chern number obtained this way is, for the system studied here, equivalent to the winding number.

\bibliography{references}

\appendix

% - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - %

\section{Computing the Zak phase from the Schmidt decomposition}\label{app:pollmann}

\maria{In this appendix, we review  how to compute the Zak phase from the Schmidt decomposition~\cite{Zaletel2014}and relate it to our own results}

Consider a closed chain of length $L$ and a bipartition into regions $A$, for $i\in [1,L/2]$, and $B$, for $i \in [L/2+1,L]$. Using a Schmidt decomposition the ground state gives
\begin{equation}
\ket{\Psi_C} = \sum_{p,q} s_p s'_q \ket{p,q}_A \ket{q,p}_B,
\end{equation}
The Schmidt indeces $p$ and $q$ label the fluctuations at the cuts in $i=1$ and $i=L/2$, respectively. The convention used for the states $\ket{p,q}_{R}$ is that the first (second) index labels the state near the left-most (right-most) region of $R$, where $R$ can be $A$ or $B$. Assuming the system is large enough compared with the correlation length the fluctuations across the two cuts are independent of each other. The reduced density matrix can be computed as
\begin{equation}
\rho_A = \sum_{p,q} s_p^2 s^{\prime \, 2}_q \ket{p,q}_{A}\bra{p,q}_A .
\end{equation}
As mentioned in section \ref{section:ES} it is fully determined by the correlation matrix $C_A$. In the state $\ket{p,q}_A$, the Schmidt index $p$ ($q$) labels a set of occupation numbers of the eigenstates of $C_A$ related to the cut at $i=1$ ($i=L/2$). we will refer to these subspaces of the Hilbert space as $A_L$ ($A_R$). Note that the remaining subspace $A_{\rm bulk}$ is composed of eigenstates that have eigenvalues $\xi = 0,1$. The eigenvalues of $\rho_A$, $\lambda_{pq}=s_p^2 s_q^{\prime \, 2}$ can be obtained as
\begin{align}
\lambda_{pq} =& \prod_{\mu \in A} (1-\xi_\mu)\left(\frac{\xi_\mu}{1-\xi_\mu}\right)^{n_\mu^{pq}} \\
=&\left(\prod_{\mu\in A_L} (1-\xi_\mu)\left(\frac{\xi_\mu}{1-\xi_\mu}\right)^{n_\mu^{p}} \right) \\
&\cross \left(\prod_{\nu \in A_R} (1-\xi_\nu)\left(\frac{\xi_\nu}{1-\xi_\nu}\right)^{n_\nu^{q}} \right) ,
\label{eq:Alexandrinata}
\end{align}
from which we identify 
\begin{align*}
&s_p^2 = \prod_{\mu\in A_L} (1-\xi_\mu)\left(\frac{\xi_\mu}{1-\xi_\mu}\right)^{n_\mu^{p}} \\
&s_q^{\prime 2} = \prod_{\nu \in A_R} (1-\xi_\nu)\left(\frac{\xi_\nu}{1-\xi_\nu}\right)^{n_\nu^{q}}.
\label{eq:spsq_A}
\end{align*}

We are now in position to introduce the $U(1)$ flux via a twisted boundary condition as
\begin{equation}
\ket{\Psi_0^\Phi} = \sum_{pq} s_p s_q' e^{-i\Phi N_{A_L}^p} \ket{p,q}_A \ket{q,p}_B,
\label{eq:Ground_state_flux}
\end{equation}
where $N_{A_L}^p = \sum_{\mu \in A_L}n_\mu^p$. As mentioned above, the label $p$ describes the set of occupation numbers $\{n_\mu^p\}$. Equation \eqref{eq:Ground_state_flux} means that an electron crossing the virtual cut at $i=1$ between region $B$ and $A_L$ will acquire a phase $-\Phi$, the convention on the sign is the same as the one used in reference \cite{Watanabe2018}.

The polarization can be now computed as
\begin{align}
\tilde{\mathcal{P}} =& \int_0^{2\pi} \frac{d\Phi}{2\pi}i\bra{\Psi_0^\Phi}\partial_\Phi\ket{\Psi_0^\Phi} \\
=& \int_0^{2\pi} \frac{d\Phi}{2\pi}i\sum_{p,q} s_p^2 s_q^{\prime \, 2} e^{i\Phi N_{A_L}^p}\partial_{\Phi}e^{-i\Phi  N_{A_L}^p}\\
=&\int_0^{2\pi} \frac{d\Phi}{2\pi}\sum_{p,q} s_p^2 s_q^{\prime \, 2}  N_{A_L}^p\\
=&\sum_{p} s_p^2 \sum_{\mu \in A_L}n^p_\mu,
\end{align}
where we used that $\sum_q s_q^{\prime \,2} = 1$. Inserting now the expression for $s_p^2$ we obtain
\begin{align}
\tilde{\mathcal{P}}=&\sum_{\{n_{i \in A_L}\} = 0,1} \left(\sum_{j\in A_{L}} n_j \right) \left(\prod_{k\in A_L} \lambda_k \right),
\end{align} 
where we defined
\begin{equation}
\lambda_k = (1- \xi_k)\left(\frac{\xi_k}{1-\xi_k} \right)^{n_k}.
\end{equation}

If we expand the sum for one particular occupation number $n_p$ we have
\begin{align}
\tilde{\mathcal{P}} =& \sum_{\substack{\{n_{i \neq p \in A_L}\} = 0,1 \\ n_p=0}} \left(\sum_{j\neq p \in A_{L}} n_j \right) (1-\xi_p)\left(\prod_{k \neq p\in A_L} \lambda_k \right) \\
&+ \sum_{\substack{\{n_{i \neq p \in A_L}\} = 0,1 \\ n_p=0}} \left(\sum_{j\neq p \in A_{L}} n_j+1 \right)\xi_p \left(\prod_{k \neq p\in A_L} \lambda_k \right) \\
=& \sum_{\substack{\{n_{i \neq p \in A_L}\} = 0,1 }} \left(\sum_{j\neq p \in A_{L}} n_j \right) \left(\prod_{k \neq p\in A_L} \lambda_k \right) \\
&+ \xi_p \sum_{\substack{\{n_{i \neq p \in A_L}\} = 0,1 }} \left(\prod_{k \neq p\in A_L} \lambda_k \right) 
\end{align}
If we continue expanding the sum in the first term we will get terms like the second one for all the other $k\neq p$ eigenvalues. If we further expand the sum in the factor accompanying $\xi_p$ we have
\begin{align}
&(1-\xi_{p'})\sum_{\substack{\{n_{i \neq p,p' \in A_L}\} = 0,1 \\ n_{p'}=0}} \left(\prod_{k \neq p,p'\in A_L} \lambda_k\right) \\
&+ \xi_{p'}\sum_{\substack{\{n_{i \neq p,p' \in A_L}\} = 0,1 \\ n_{p'}=1}} \left(\prod_{k \neq p,p'\in A_L}\lambda_k\right) =1,
\end{align}
which we can see is $1$ if we continue this procedure for all other occupation numbers. Therefore in terms of the spectrum of $C_A$ the first term of the polarization greatly simplifies to 
\begin{align}
\tilde{\mathcal{P}} =& \sum_{p\in A_L}\xi_p  \quad {\rm mod}\, 1.
\end{align}
The subspace $A_L$ is not well-defined, as one can always include bulk modes. As a result the polarization obtained this way is defined only modulo 1. In practice we extend $A_L$ to include all eigenstates whose average position lies in the left half of $A$ and we finally get
\begin{align}
\tilde{\mathcal{P}} =& \sum_{i\in L}\xi_i \quad {\rm mod}\, 1.
\label{eq:polarization_left}
\end{align}

\section{Alternative expression for the correlation matrix}\label{app:CM}
In this appendix, we show how to express the correlation matrix in terms of the Hamiltonian, used in Eq.~\eqref{eq:corr_mat2}. 
We consider the generic, quadratic Hamiltonian in one dimension of Eq.~\ref{eq:quadr_Ham}, which is diagonalized by a unitary matrix $U$ with 
	\begin{align}\label{eq:D}
	H&=UDU^\dagger& \mbox{with } D=\mbox{diag}(E_{p\mu}).
	\end{align} 
In terms of the fermionic operators that diagonalize the Hamiltonian, 
\begin{align}
& \gamma_{i\alpha} = \sum_{j\beta}u_{j\beta}^{i\alpha \, \ast} c_{j\beta},
%& c_{i\alpha} = \sum_{j\beta} u^{j\beta}_{i\alpha} \gamma_{j\beta}
\end{align}
we find that the correlation matrix can be written as 
\begin{align}\label{eq:corr_mat}
C_{ij}^{\alpha \beta} =& \sum_{pq,\mu\nu} u^{q\nu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} \expval{\gamma^\dagger_{p\mu} \gamma_{q\nu} }\nonumber \\
=&  \sum_{p\mu} u^{p\mu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} \expval{\gamma^\dagger_{p\mu} \gamma_{p\mu} }\nonumber \\
=&  \sum_{p\mu} u^{p\mu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} [1-{\rm sign}(E_{p\mu})]/2.
\end{align}
The first term of the last line is simply a kronecker delta between both sets of indices.

The second term can be rewritten, using Eq.~\ref{eq:D}, as
\begin{align}
UD(\abs{D})^{-1}U^\dagger =& UDU^\dagger U(D^2)^{-1/2}U^\dagger
\end{align}
We can rewrite this expression further by noting that 
\begin{align}
[U (D^2)^{-1/2} U^\dagger]^2 =& U (D^2)^{-1/2} U^\dagger U  (D^2)^{-1/2} U^\dagger \nonumber\\
=&  U (D^2)^{-1/2}  (D^2)^{-1/2} U^\dagger \nonumber\\
=&  U (D^2)^{-1} U^\dagger \nonumber\\
=&  (U D^2 U^\dagger)^{-1}.
\end{align}
Therefore,
\begin{align*}
U (D^2)^{-1/2} U^\dagger =&(U D^2 U^\dagger)^{-1/2} 
\end{align*}
and we conclude that
\begin{align}\label{eq:UDDU}
UD(\abs{D})^{-1}U^\dagger =& UDU^\dagger(U D^2 U^\dagger)^{-1/2} \nonumber\\ 
=& H(H^2)^{-1/2}.
\end{align}
Combining Eq.s~\eqref{eq:corr_mat},  and \eqref{eq:UDDU}, we arrive at the final expression of the correlation matrix in Eq.~\eqref{eq:corr_mat2}.

\section{Winding number in the SSH chain}
\label{appendix:winding_ssh}

Consider the Bloch Hamiltonian of the Rice-Mele model
\begin{equation}
H=\mqty(\kappa & f(k) \\ f^\dagger(k) & -\kappa),
\end{equation}
which can also be written as $H = \boldsymbol{h}\cdot \boldsymbol{\sigma}$
where 
\begin{align*}
\boldsymbol{h}=& ({\rm Re}[f(k)],-{\rm Im}[f(k)],\kappa) \\
=& \sqrt{\abs{f(k)}^2 +\kappa^2}(\sin(\theta)\cos(\phi),\sin(\theta)\sin(\phi),\cos(\theta))
\end{align*}
In the gauge where the second component of the eigenstates remains real they are given by
\begin{align}
&\ket{+}  = \frac{1}{N_+}\mqty(\cot(\theta/2)e^{-i\phi}\\ 1 )\\
&\ket{-}  =\frac{1}{N_-} \mqty(-\tan(\theta/2)e^{-i\phi}\\ 1),
\end{align}
where
\begin{align*}
&N_+ = \sqrt{1+\cot^2(\theta/2)} \\
&N_- = \sqrt{1+\tan^2(\theta/2)} .
\end{align*}
The Berry connections of the occupied state, $A_\lambda = \bra{-}\partial_\lambda \ket{-}$, can be obtained as
\begin{align*}
A_{\lambda}=& i\frac{\cos(\theta)-1}{2}\partial_q \phi \\
=&i\frac{\cos(\theta)-1}{2} \cos(\phi)^2 \partial_q  \tan(\phi) \\
=&\frac{i}{2}\left(\frac{\kappa}{\sqrt{\abs{f(k)}^2+\kappa^2}}-1\right) \left( \frac{\Re [f(k)]}{\abs{f(k)}}\right)^2 \partial_\lambda  \left( -\frac{\Im [f(k)]}{\Re [f(k)]} \right) \\
=&\frac{-1}{2}\left(\frac{\kappa}{\sqrt{\abs{f(k)}^2+\kappa^2}}-1\right) \frac{f(k)^\dagger \partial_\lambda f(k)-f(k)\partial_\lambda f(k)^\dagger}{2\abs{f(k)}^2}
\end{align*}
The Berry connection with respect to momentum gives
\begin{equation}
A_k = \frac{it [t+m \cos(k)][\kappa - \sqrt{\kappa^2+m^2+t^2+2mt \cos(k)}]}{2[m^2+t^2+2mt\cos(k)]\sqrt{\kappa^2+m^2+t^2+2mt \cos(k)}},
\end{equation}
and the resulting Zak phase can be seen in Fig.\ref{fig:ssh_zak}, which presents a branch cut for $m=0,\kappa<0$ for this particular gauge.

\begin{figure}[t]
	\centering
	\includegraphics[width=50mm]{sshwindinganalytical.pdf}
	\caption{$\gamma/2\pi$ obtained in the smooth gauge that provides the winding number in the chiral symmetric limit.}
	\label{fig:ssh_zak}
\end{figure}

Note that in the chiral symmetric case, for $\kappa = 0$, the Zak phase is
\begin{align*}
\gamma =& \int_0^{2\pi} dk\,\frac{1}{2}\frac{f(k)^\dagger \partial_k f(k)-f(k)\partial_k f(k)^\dagger}{2\abs{f(k)}^2} \\
=& \int_0^{2\pi} dk\,\frac{1}{2}q(k)^\dagger \partial_k q(k),
\end{align*}
where $q(k) = f(k)/\abs{f(k)}$. This is nothing else than the winding number \cite{ryu2010topological}.

Consider now that we parametrize a loop in parameter space as $m=1-\cos(\theta),\kappa = \sin(\theta)$.

\section{Twisted boundary condition Winding number}
\carlos{Work in progress}
For a system with chiral symmetry the Hamiltonian can be written in some basis as
\begin{equation}
H = \mqty(0 & h \\ h^\dagger & 0),
\end{equation}
where $h$ is in general a matrix. In translationally invariant systems the winding number is defined as 
\begin{equation}
\nu = -i\int_0^{2\pi} \frac{dk}{2\pi} {\rm Tr}[h(k)^{-1} \partial_k h(k)],
\label{eq:winding_k}
\end{equation}
where $h(k)$ is the off-diagonal block of the Bloch Hamiltonian. In systems without translational invariance, one can similarly define the winding number
\begin{equation}
\nu = -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[h(\Phi)^{-1} \partial_\Phi h(\Phi)],
\label{eq:winding_flux}
\end{equation}
for a $U(1)$ flux introduced (\carlos{I have only found a similar expression in the context of non-hermitian systems \cite{Gong2018}, but I assume this is the expression Prodan refers to. It gives the correct results in translationally invariant systems, however it is very unpractical.}). Note that a generalization of \eqref{eq:winding_k} to position space has also been developed which not necesseraly gives the same results as \eqref{eq:winding_flux} (\carlos{The only comparison is in the Prodan paper where they talk about practical numerical implementations but don't adress if both formulations are equivalent.}) Both formulations however reduce to \eqref{eq:winding_k} for translationally invariant systems \cite{Gong2018}.  In the following we will omit the flux parameter for simplicity.

Note that the winding number can also be expressed in terms of the flat-band Hamiltonian
\begin{align*}
Q =& H (H^2)^{-1/2} \\
=& \mqty(0 & q \\ q^\dagger & 0).
\end{align*}
Note that since $Q$ only has eigenvalues $\pm 1 $ , $Q^2 = \mathbb{I}$ and therefore $q^\dagger = q^{-1}$. We can therefore write
\begin{equation}
\nu = -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[q^\dagger \partial_\Phi q].
\end{equation}
In this basis the chiral symmetry operator is given by $S = \sigma_z \otimes \mathbb{I}$ such that the winding number can be expressed in terms of the full flat-band Hamiltonian as
\begin{align}
\nu =& \frac{i}{2}\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S Q^\dagger \partial_\Phi Q] \\
=& \frac{i}{2}\int_0^{2\pi} \frac{d\Phi}{2\pi} \left({\rm Tr}[q \partial_\Phi q^\dagger]-{\rm Tr}[q^\dagger \partial_\Phi q] \right)\\
=& -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[q^\dagger \partial_\Phi q] 
\end{align}
where in the last step we use the fact that the winding number is a real number. We can now write the winding number in terms of the correlation matrix
\begin{align}
\nu =& \frac{i}{2}\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S (1-2C) \partial_\Phi (1-2C)] \\
=& -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S (1-2C) \partial_\Phi C] .
\end{align}
Introducing the flux via twisted boundary conditions allows us to easily construct the ground state with flux in the case of the open chain \eqref{eq:gauge}, therefore we will employ this gauge to compute the correlation matrix. The correlation matrix can be written in position space as
\begin{equation}
C = \mqty(C_A & C_{AB} \\ C_{AB}^\dagger & C_B),
\end{equation}
where in this basis, assuming $S$ is a local symmetry it is also diagonal. Note that the flux will only appear in the off-diagonal blocks $C_{AB}$ and therefore the first term in the trace above will vanish, leaving the winding number as
\begin{align}
\nu =&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S C \partial_\Phi C] \\
=&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} \left( {\rm Tr}[S C_{AB} \partial_\Phi C_{AB}^\dagger]+{\rm Tr}[S C_{AB}^\dagger \partial_\Phi C_{AB}] \right)
\end{align}
We can not explicitely construct the correlation matrix with flux as
\begin{align*}
C_{ij}^\Phi =& \bra{\tilde{\Psi}_{0,{\rm open}}^\Phi}c_i^\dagger c_j \ket{\tilde{\Psi}_{0,{\rm open}}^\Phi} \\
=&\bra{\tilde{\Psi}_{0,{\rm open}}^0}e^{i\Phi \hat{N}_A} c_i^\dagger c_j e^{-i\Phi \hat{N}_A}\ket{\tilde{\Psi}_{0,{\rm open}}^0} .
\end{align*}
Depending on if $i,j$ belong to $A$ or not, commuting the exponentials will leave additional factors or not, resulting in
\begin{align*}
C^\Phi = \mqty(C_A^0 & e^{i\Phi} C_{AB}^0 \\e^{-i\Phi} C_{AB}^{0\,\dagger} &C_B^{0}),
\end{align*}
where we have made explicit the flux dependence. The winding number is then
\begin{align}
\nu =&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S C^\Phi \partial_\Phi C^\Phi] \\
=&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} \left( {\rm Tr}[-iS C_{AB}^0 C_{AB}^{0\,\dagger}]+{\rm Tr}[iS C_{AB}^{0\,^\dagger} C_{AB}^0] \right) \\
=&  {\rm Tr}[S C_B^0(1-C_B^0)]-{\rm Tr}[S C_A^0(1-C_A^0)] 
\end{align}
where in the last step we use one of the properties of the correlation function stemming from $C^2 = C$. (\carlos{So far this expression gives the correct result for translationally invariant systems however one needs to do the trick of breaking chiral symmetry slightly. However, unlike with $\tilde{\mathcal{P}}$ this expression is not discontinuous at $\kappa = 0$. })

\end{document}

### 2020-08-23 version2/draft.tex
small changes

@@
 \newcommand{\mariac}[1]{{\it\color{cyan}#1}}
+\newcommand{\changed}[1]{{\color{orange}#1}}
+\newcommand{\tpo}{\tilde{\mathcal{P}}_{\rm o}}
+\newcommand{\brapsio}[1]{\bra{\tilde{\Psi}_{{\rm o}}^#1}}
+\newcommand{\ketpsio}[1]{\ket{\tilde{\Psi}_{{\rm o}}^#1}}
 
@@
 It was subsequently shown for topological insulators/superconductors that the ES for (gapped) periodic systems is equivalent to the flat-band energy spectrum of the corresponding system with open boundaries~\cite{Fidkowski2010entanglement}.
-The same correspondence was also found for closely related gapless systems~\cite{Matern2018entanglement}
+The same correspondence was also found for closely related gapless systems~\cite{Matern2018entanglement}.
 However, even for non-interacting systems, it is unclear which information (beyond the  `edge' spectrum) is encoded in the entanglement spectrum. 
+In this manuscript, we show that the information about the Zak phase, or polarization, can be completely recovered from the entanglement spectrum and provide a simple formula to do so. 
+We also show how to extend this method to compute the Chern number directly from the entanglement spectrum.
 
-In this manuscript, we show that the information about the Zak phase, or polarization, can be completely recovered from the entanglement spectrum and provide a simple formula to do so. 
 The relation between the entanglement spectrum and various topological invariants has already been studied in the literature. 
@@
 A particular limit of our result for homogeneous systems~\eqref{eq:chi_integrable} was derived by Ryu and Hatsugai for a fully dimerized SSH chain~\cite{Ryu2006}. 
-The similarity between the behavior of the Zak phase and the 'virtual edge modes' in the entanglement spectrum of a Chern insulator, which was observed in Ref.~\cite{Huang2012,Huang2012-2}, can also be explained by our results.
-
-\carlos{I added this last part}
-Although the polarization obtained via this method for infinite systems or systems with periodic boundary conditions is only well-defined modulo 1 one can compute it in the equivalent open chain to obtain a polarization defined as a real number \carlos{Not sure about this expression.
+The similarity between the behavior of the Zak phase and the 'virtual edge modes' in the entanglement spectrum of certain Chern insulators, which was observed in Ref.~\cite{Huang2012,Huang2012-2}, can also be explained by our results.
+The polarization obtained via this method for infinite systems or systems with periodic boundary conditions is only well-defined modulo 1, i.e. in the range $[0,1]$. 
+One can, however, compute it in the equivalent open chain to obtain a polarization defined on the full real axis.
+%\carlos{Not sure about this expression.
 I mean it's not defined on a circle, not that is not a complex number, which it trivially isnt.}.
-We use this method to compute Chern numbers, which can also serve as topological invariants in one-dimensional systems, and for the system studied we show that it is equivalent to the winding number.
+We use the latter method to compute Chern numbers, which can also serve as topological invariants in one-dimensional systems.
+For the system studied, the Chern number is equivalent to the winding number, but it is unclear whether this result continues to hold for generic systems.
 
@@
 In section V we introduce the method and show how it reproduces the Zak phase for both systems with and without translational invariance.
-In section VI we explain how the method applied to the opened chain allows us to compute Chern numbers and winding numbers and finally in section VII we present the conclusions and possible outlook.
+In section VI we explain how the method applied to the opened chain allows us to compute Chern numbers %and winding numbers
+and finally in section VII we present the conclusions and possible outlook.
 
 % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % 
-\section{Model.}
+\section{Model}
 To showcase our results we consider as a starting point the model in the BDI class used in reference~\cite{Song2014}. 
@@
 In translational invariant systems $u_{p\mu}^{j\beta} = e^{ipj}\tilde{u}_{p \mu}^{\beta}$, where $\tilde{u}_{p \mu}^{\beta}$ are the components of the eigenstates of the Bloch Hamiltonian, $\ket{\tilde{u}_{k\mu}}$.
-One can now compute the Zak phase \cite{Zak1989}, which for a two-band model is defined as the geometric phase acquired by the occupied state $\ket{\tilde{u}_k}$ as it winds around the Brillouin zone,
+One can now compute the Zak phase \cite{Zak1989}, which for a two-band \mariac{do we need the restriction to two-band models?} model is defined as the geometric phase acquired by the occupied state $\ket{\tilde{u}_k}$ as it winds around the Brillouin zone,
 \begin{equation}
@@
 \end{equation}
-where $\Omega_{\lambda k} = \partial_\lambda A_k(\lambda,k) - \partial_k A_\lambda(\lambda,k)$ is the Berry curvature.
-The Berry curvature is fully gauge invariant and therefore the change in the Zak phase as defined in \eqref{eq:change_zak_phase} is defined as a real number.
+where 
+\begin{align}\label{eq:BerryCurvature}
+\Omega_{\lambda k} = \partial_\lambda A_k(\lambda,k) - \partial_k A_\lambda(\lambda,k)
+\end{align}
+ is the Berry curvature.
+The Berry curvature is fully gauge invariant and, therefore, the change in the Zak phase as defined in \eqref{eq:change_zak_phase} is defined along the full real line.
 The Zak phase itself, as defined above, can be shown to be proportional to the winding number for a certain gauge (see Appendix \ref{appendix:winding_ssh}), which means that it carries physical information beyond modulo $2\pi$.
@@
 The derivative of the polarization with respect to some parameter $\lambda$ was obtained \cite{KingSmith1993} as
-\begin{equation}
+\begin{equation}\label{eq:Bloch_polarization}
 \partial_\lambda \tilde{\mathcal{P}}^{\rm Bloch} = \int_{0}^{2\pi} \frac{dk}{2\pi} \, \Omega_{\lambda k},
@@
 which allows us to make the identification $ \tilde{\mathcal{P}}^{\rm Bloch} = \gamma/2\pi$.
-As discussed above this polarization is defined as an absolute value only modulo 1, however its changes can be defined as a real number.
+As discussed above,  this polarization is defined as an absolute value only modulo 1, however its changes can be defined as a real number.
 This is no surprise as the derivative of the polarization is related to the current which is a physical observable, unlike $\tilde{\mathcal{P}}^{\rm Bloch}$.
@@
 \begin{equation}
-\mathcal{P} = \int_0^{2\pi}\frac{d\Phi}{2\pi} i\bra{\Psi_0^\Phi}\partial_\Phi \ket{\Psi_0^\Phi} + {\rm Im}{\rm Ln} \bra{\Psi_0^0}e^{2\pi i \hat{P}} \ket{\Psi_0^{2\pi}}, 
+\mathcal{P} = \int_0^{2\pi}\frac{d\Phi}{2\pi} i\bra{\Psi^\Phi}\partial_\Phi \ket{\Psi^\Phi} + \Im\ln \bra{\Psi^0}e^{2\pi i \hat{P}} \ket{\Psi^{2\pi}}, 
 \end{equation}
-where $\ket{\Psi_0^{\Phi}}$ is the ground state in the presence of flux and $\hat{P}=\frac{1}{L}\sum_{j\alpha} j\hat{n}_{j\alpha}$ is the polarization operator.
+where $\ket{\Psi^{\Phi}}$ is the ground state in the presence of flux and $\hat{P}=\frac{1}{L}\sum_{j\alpha} j\hat{n}_{j\alpha}$ is the polarization operator.
 This is equivalent to the polarization obtained by Resta \cite{Resta1997} as
+\maria{should I interpret the following line as $\Phi=0$?}
 \begin{align}
-\mathcal{P} = \frac{1}{2\pi} \rm{ Im }\,{\rm Ln}\bra{\Psi_0}e^{2\pi i \hat{P}}\ket{\Psi_0}.
+\mathcal{P} = \frac{1}{2\pi} \rm{ Im }\,\ln\bra{\Psi}e^{2\pi i \hat{P}}\ket{\Psi}.
 \end{align}
@@
 \begin{align}
-\mathcal{P} = \frac{1}{2\pi} \rm{Im \,Ln} \, {\rm det }\!' \, S,
+\mathcal{P} = \frac{1}{2\pi}\Im \ln \, {\rm det }\!' \, S,
 \label{eq:polarization_resta}
@@
 There is one last polarization we must consider and it is the one obtained when the flux is introduced in the bond between sites $i=1$ and $i=L$ via twisted boundary conditions.
-This is equivalent to performing the gauge transformation $\ket{\tilde{\Psi}_0^\Phi}=e^{i\Phi \hat{P}}\ket{\Psi_0^\Phi}$, which makes $\ket{\tilde{\Psi}_0^\Phi}$ fully periodic in $\Phi$.
-We can now define the polarization
+This is equivalent to performing the gauge transformation
+\begin{align}\label{eq:Psi_tilde}
+\ket{\tilde{\Psi}^\Phi}=e^{i\Phi \hat{P}}\ket{\Psi^\Phi},
+\end{align}
+which makes $\ket{\tilde{\Psi}_0^\Phi}$ fully periodic in $\Phi$. We can now define the polarization
 \begin{equation}
-\tilde{\mathcal{P}} = \int_0^{2\pi}\frac{d\Phi}{2\pi} i\bra{\tilde{\Psi_0}^\Phi}\partial_\Phi \ket{\tilde{\Psi_0}^\Phi},
+\tilde{\mathcal{P}} = \int_0^{2\pi}\frac{d\Phi}{2\pi} i\bra{\tilde{\Psi}^\Phi}\partial_\Phi \ket{\tilde{\Psi}^\Phi},
 \label{eq:p_tilde}
@@
 \begin{equation}
-\mathcal{P}_0= \int_0^{2\pi}\frac{d\Phi}{2\pi} \bra{\tilde{\Psi_0}^\Phi}\hat{P}\ket{\tilde{\Psi_0}^\Phi}.
+\tilde {\mathcal{P}}_0= \int_0^{2\pi}\frac{d\Phi}{2\pi} \bra{\tilde{\Psi}^\Phi}\hat{P}\ket{\tilde{\Psi}^\Phi}.
 \end{equation} 
-The previous discussion about the Zak phase being defined only modulo $2\pi$ although its changes can be defined as a real number applies in the same way to the polarization obtained in \eqref{eq:p_tilde}.
+\maria{Ref to Oshikawa paper?}
+%The previous discussion about the Zak phase being defined only modulo $2\pi$ although its changes can be defined as a real number applies in the same way to the polarization obtained in \eqref{eq:p_tilde}.
+Note that while $\tilde{\mathcal{P}}$ itself is only defined modulo 1, its changes are again well-defined on the full real axis, similar to the discussion about the Bloch polarization below Eq.~\eqref{eq:Bloch_polarization}.
+In fact, for translationally invariant systems one finds that $\tilde{\mathcal{P}} = \tilde{\mathcal{P}}^{\rm Bloch}$ \cite{Watanabe2018}.
 
@@
 
-In the case where there is only one eigenvalue per edge that is $\neq 0,1$ in  the EOS, like in the fully dimerized limit of the SSH chain, we recover the equality between the midgap eigenvalue and the Zak phase~\cite{Ryu2006}.
+In the case where there is only one eigenvalue per edge that is $\neq 0,1$ in  the EOS, like in the fully dimerized limit of the SSH chain, we recover the equality between the midgap eigenvalue and the Zak phase found in Ref.~\cite{Ryu2006}.
 
@@
 \end{equation}
-In the thermodynamic limit it is $\tilde{\mathcal{P}}$ that we actually identify with $\chi$
+In the thermodynamic limit, it is $\tilde{\mathcal{P}}$ that we actually identify with $\chi$
 \begin{equation}
-\lim_{L \rightarrow \infty} \tilde{\mathcal{P}} = \lim_{L \rightarrow \infty} \chi,
+\lim_{L \rightarrow \infty} \tilde{\mathcal{P}} = \lim_{L \rightarrow \infty} \chi.
 \label{eq:ptilde_eq_chi}
 \end{equation}
-however for translationally invariant systems $\tilde{\mathcal{P}} = \tilde{\mathcal{P}}^{\rm Bloch}$ \cite{Watanabe2018}.
 Note that in order to compute $\chi$ correctly the eigenstates with $\xi \neq 0,1$ need to be localized in order for us to select the correct eigenvalues.
-It might be, like in Fig.\ref{huang}(a) that the spectrum has degenerate pairs of left and right eigenstates and the numerical diagonalization mixes them.
-This case is trivial as we know that there is one left and one right eigenstate but if the degeneracy is four-fold or higher one needs to obtain the localized eigenstates.
+It might be --- like in Fig.\ref{huang}(a) --- that the spectrum has degenerate pairs of left and right eigenstates and the numerical diagonalization mixes them.
+This case is trivial, as we know that there is one left and one right eigenstate but if the degeneracy is four-fold or higher one needs to obtain the localized eigenstates.
 Localizing the eigenstates that are exponentially close to $\xi = 0,1$ is more difficult but not doing so only results in an exponentially small error.
@@
 This can be avoided by computing instead
+\begin{equation}\label{eq:tpo}
+\tpo = \int_0^{2\pi}\frac{d\Phi}{2\pi} i\brapsio{\Phi}\partial_\Phi \ketpsio{\Phi}.
+\end{equation}
+where $\ketpsio{{\Phi=0}}$ is the ground state of the chain opened between sites $i=L/2$ and $i=L/2+1$ and 
 \begin{equation}
-\tilde{\mathcal{P}}_{\rm open} = \int_0^{2\pi}\frac{d\Phi}{2\pi} i\bra{\tilde{\Psi}_{0,{\rm open}}^\Phi}\partial_\Phi \ket{\tilde{\Psi}_{0,{\rm open}}^\Phi}.
-\end{equation}
-where $\ket{\tilde{\Psi}_{0,{\rm open}}^0}$ is the ground state of the chain opened between sites $i=L/2$ and $i=L/2+1$ and 
-\begin{equation}
-\ket{\tilde{\Psi}_{0,{\rm open}}^\Phi} =e^{-i \Phi \hat{N}_A} \ket{\tilde{\Psi}_{0,{\rm open}}^0}.
+\ketpsio{\Phi} =e^{-i \Phi \hat{N}_A} \ketpsio{0}.
 \label{eq:gauge}
@@
 The difference with the closed chain being that now there is no ambiguity on when an electron crosses the boundary with the twisted boundary condition.
-Note that since the chain is open $\Phi$ is not a flux but should now be regarded as a parameter and therefore $\tilde{\mathcal{P}}_{\rm open}$ does not need to be a physical polarization.
-$\tilde{\mathcal{P}}_{\rm open}$ can be obtained in an even simpler way than $\tilde{\mathcal{P}}$ as
-\begin{align*}
-\tilde{\mathcal{P}}_{\rm open} =&  \int_0^{2\pi}\frac{d\Phi}{2\pi} \bra{\tilde{\Psi}_{0,{\rm open}}^0}  \hat{N}_A \ket{\tilde{\Psi}_{0,{\rm open}}^0} \\
-=&\sum_{\alpha,j\in A}\bra{\tilde{\Psi}_{0,{\rm open}}^0} c_{j\alpha}^\dagger c_{j\alpha} \ket{\tilde{\Psi}_{0,{\rm open}}^0} \\
-=&\sum_{\alpha,j\in A} C_{\rm open,j\alpha,j\alpha} \\
-=& {\rm Tr}[C_{A,\rm open}].
-\end{align*}
-Which is nothing else than the generalization of $\chi$ to the open chain.
-The idea is that in the process of oppening the chain, which can be done adiabatically, all eigenvalues that are not in $A_L$ should get pushed to $\xi=0$ or $\xi=1$.
-If for a given path in parameter space $\boldsymbol{\lambda}_i \rightarrow \boldsymbol{\lambda}_f$ their contribution remains constant, that is, there are no jumps between $\xi=0$ and $\xi=1$ and $\tilde{\mathcal{P}}_{\rm open} (\boldsymbol{\lambda} )$ is smooth, then changes in $\tilde{\mathcal{P}}_{\rm open}(\boldsymbol{\lambda}_i ) $ along the path will be defined as a real number and not only modulo 1.
-Since $\tilde{\mathcal{P}}$ is a local property of the bond between $i=1$ and $i=L$ one would expect it to not depend on what happens far from it such that $\tilde{\mathcal{P}}_{\rm open}$ and $\tilde{\mathcal{P}}$ are equivalent.
+Note that since the chain is open, $\Phi$ is not a flux but should now be regarded as a parameter and therefore $\tpo$ does not need to be a physical polarization.
+$\tpo$ can be obtained in an even simpler way than $\tilde{\mathcal{P}}$ as
+\begin{align}\label{eq:phi_independence}
+\tpo =&  \int_0^{2\pi}\frac{d\Phi}{2\pi} \brapsio{0} \hat{N}_A \ketpsio{0} \nonumber \\
+=&\sum_{\alpha,j\in A}\brapsio{0} c_{j\alpha}^\dagger c_{j\alpha} \ketpsio{0} \nonumber\\
+=&\sum_{\alpha,j\in A} C_{\rm o,j\alpha,j\alpha}\nonumber \\
+=& {\rm Tr}[C_{A,\rm o}],
+\end{align}
+where we used Eq.~\eqref{eq:gauge} and $C_{\rm o}$($C_{A,\rm o}$) denotes the (reduced) density matrix of the open chain. 
+Note that Eq.~\eqref{eq:phi_independence} is nothing else than the generalization of $\chi$ to the open chain. 
+\maria{I am not entirely happy with the following argument, as it goes wrong in the topological phase. We mention this later, but we should already mention it here. }
+The idea is that in the process of opening the chain, which can be done adiabatically, all eigenvalues that are not in $A_L$ should get pushed to $\xi=0$ or $\xi=1$. 
+If for a given path in parameter space $\boldsymbol{\lambda}_i \rightarrow \boldsymbol{\lambda}_f$ their contribution remains constant, that is, there are no jumps between $\xi=0$ and $\xi=1$ and $\tpo (\boldsymbol{\lambda} )$ is smooth, then changes in $\tpo(\boldsymbol{\lambda}_i ) $ along the path will be defined as a real number and not only modulo 1.
+Since $\tilde{\mathcal{P}}$ is a local property of the bond between $i=1$ and $i=L$ one would expect it to not depend on what happens far from it such that $\tpo$ and $\tilde{\mathcal{P}}$ are equivalent.
 
@@
 \end{equation}
-so $\tilde{\mathcal{P}}_{\rm open}(-\kappa) = L-\tilde{\mathcal{P}}_{\rm open}(\kappa)$.
-Unless $\tilde{\mathcal{P}}_{\rm open}(0) = L/2$, which only happens in the trivial case, the result will be discontinuous at $\kappa = 0$.
-We can see this in Fig.\ref{ssh_chern}.(a) where we show $\tilde{\mathcal{P}}_{\rm open}$ in the space $(m,\kappa)$.
+so $\tpo(-\kappa) = L-\tpo(\kappa)$.
+Unless $\tpo(0) = L/2$, which only happens in the trivial case, the result will be discontinuous at $\kappa = 0$.
+We can see this in Fig.\ref{ssh_chern}.(a) where we show $\tpo$ in the space $(m,\kappa)$.
 This branch cut is in fact not an artifice of the open chain trick, but the result of the degenerate point at $(m=t,\kappa=0)$ having an associated Chern number $C=1$ \cite{Asboth2016}.
@@
 \begin{equation}
-C = \int_0^{2\pi}d\theta\int_0^{2\pi}\frac{d\Phi}{2\pi} \Omega_{\theta\Phi}.
+C = \int_0^{2\pi}d\theta\int_0^{2\pi}\frac{d\Phi}{2\pi} \Omega_{\theta\Phi},
 \label{eq:chern}
 \end{equation} 
-The Berry connection for the flux is simply $A_\Phi(\theta,\Phi,) = -i\tilde{\mathcal{P}}(\theta)$, so we now need to compute $A_\theta(\theta,\Phi)$.
-(\carlos{Please double check this calculation}) Expanding $\ket{\tilde{\Psi}_{0,{\rm open}}^0}$ in the eigenbasis of $\hat{N_A}$, $\{\ket{j}\}$ we can compute its derivative with respect to the parameter $\theta$ as
-\begin{align*}
-\partial_\theta \ket{\tilde{\Psi}_{0,{\rm open}}^0} =& \sum_j [\partial_\theta c_j (\theta)]e^{-i\Phi N_A^j}\ket{j}.
-\end{align*}
-where $c_j(\theta) = \bra{j}\ket{\tilde{\Psi}_{0,{\rm open}}^0}$. The correspondent Berry connection then gives
-\begin{align*}
-\bra{\tilde{\Psi}_{0,{\rm open}}^\Phi}\partial_\theta \ket{\tilde{\Psi}_{0,{\rm open}}^\Phi} =& \sum_{lj} c^*_l(\theta) [\partial_\theta c_j(\theta)] \bra{l} e^{i\Phi (N_A^lN_A^j)}\ket{j}\\
-=& \sum_j  c^*_j(\theta) \partial_\theta c_j(\theta) \\
-=&\bra{\tilde{\Psi}_{0,{\rm open}}^0}\partial_\theta \ket{\tilde{\Psi}_{0,{\rm open}}^0}
-\end{align*}
-and is therefore independent of $\Phi$.
+with the Berry curvature $\Omega_{\theta\Phi}$ defined in Eq.~\ref{eq:BerryCurvature}. 
+From  Eq.~\eqref{eq:tpo}, it follows that the Berry connection for the flux is simply $A_\Phi(\theta,\Phi,) = -i\tpo(\theta)$, so we only need to compute $A_\theta(\theta,\Phi)$.
+(\carlos{Please double check this calculation}) Expanding $\ketpsio{\Phi}$ in the eigenbasis of the number operator $\hat{N_A}$ -- denoted by $\{\ket{j}\}$ -- we can compute its derivative with respect to the parameter $\theta$ as
+\changed{changed}
+\begin{align}
+\partial_\theta \ketpsio{\Phi} =& \sum_j [\partial_\theta c_j (\theta)]e^{-i\Phi N_A^j}\ket{j}.
+\end{align}
+where $c_j(\theta) = \Big\langle j\ketpsio{0}$ does not depend on $\Phi$. The correspondent Berry connection then gives
+\begin{align}
+\brapsio{\Phi}\partial_\theta \ketpsio{\Phi} =& \sum_{lj} c^*_l(\theta) [\partial_\theta c_j(\theta)] \bra{l} e^{i\Phi (N_A^lN_A^j)}\ket{j}\nonumber\\
+=& \sum_j  c^*_j(\theta) \partial_\theta c_j(\theta) \nonumber\\
+=&\brapsio{0}\partial_\theta \ketpsio{0}, 
+\end{align}
+i.e. it is independent of $\Phi$.
 
@@
 We can now apply Stokes theorem to the two cylinders independently
-\begin{align*}
-C_1 =& \int_{S_1} d\boldsymbol{S}\cdot (\nabla \times \boldsymbol{A}) \\
-=& \int_{\partial S_1} d\boldsymbol{l}\cdot \boldsymbol{A}(\theta,\Phi) \\
-=& \int_a^{2\pi-a} d\theta A_\theta(\theta,0) +\int_0^{2\pi} \frac{d\Phi}{2\pi} A_\Phi (2\pi-a,\Phi) \\
+\begin{align}
+C_1 =& \int_{S_1} d\boldsymbol{S}\cdot (\nabla \times \boldsymbol{A}) \nonumber\\
+=& \int_{\partial S_1} d\boldsymbol{l}\cdot \boldsymbol{A}(\theta,\Phi) \nonumber\\
+=& \int_a^{2\pi-a} d\theta A_\theta(\theta,0) +\int_0^{2\pi} \frac{d\Phi}{2\pi} A_\Phi (2\pi-a,\Phi) \nonumber\\
 &-\int_a^{2\pi-a} d\theta A_\theta(\theta,2\pi) -\int_0^{2\pi} \frac{d\Phi}{2\pi} A_\Phi (a,\Phi).
-\end{align*}
+\end{align}
 In our case, $A_\theta$ is independent of $\Phi$ so the terms involving it cancel and we have
-\begin{align*}
+\begin{align}
 C_1 =& \int_0^{2\pi} \frac{d\Phi}{2\pi} [A_\Phi (2\pi-a,\Phi) -A_\Phi (a,\Phi)].
-\end{align*}
+\end{align}
 Similarly computing the integral for the other cylinder we obtain
-\begin{align*}
+\begin{align}
 C_2 =& \int_0^{2\pi} \frac{d\Phi}{2\pi} [A'_\Phi (2\pi+a,\Phi) -A'_\Phi (2\pi-a,\Phi)].
-\end{align*}
+\end{align}
 The Chern number is then $C=C_1+C_2$. Noting again that $A'_\Phi(\theta)$ is continuous in $S_2$ taking the limit of $a\rightarrow 0^+$ makes $C_2$ vanish and the Chern number is given by
-\begin{align*}
+\begin{align}
 C =& \int_0^{2\pi} \frac{d\Phi}{2\pi} [A_\Phi (2\pi^-,\Phi) -A_\Phi (0^+,\Phi)].
-\end{align*}
+\end{align}
 If we consider the loop starting at $\kappa = 0$ this is exactly the case shown in Fig.\ref{ssh_chern}.(a), where $A_\Phi (\theta,\Phi) = \tilde{\mathcal{P}}_{\rm open}(\theta)$ is smooth everywhere but in $\theta = 0$, and therefore we can compute the Chern number simply as
-\begin{align*}
-C= \tilde{\mathcal{P}}_{\rm open}(\kappa=0^+)-\tilde{\mathcal{P}}_{\rm open}(\kappa=0^-).
-\end{align*}
+\begin{align}
+C= \tpo(\kappa=0^+)-\tpo(\kappa=0^-).
+\end{align}
 Using the symmetry of the polarization described above this results in 
 \begin{align}
-C= 2\tilde{\mathcal{P}}_{\rm open}(0^+)-L.
+C= 2\tpo(0^+)-L.
 \label{eq:Chern_number}
 \end{align}
-The discontinuity in $\tilde{\mathcal{P}}_{\rm open}$ is therefore related to the Chern number and it is not merely continuous modulo 1.
+The discontinuity in $\tpo$ is therefore related to the Chern number.% and it is not merely continuous modulo 1.
 
-One can also obtain a polarization continuous in $\kappa$ by doing a gauge transformation $\ket{\tilde{\psi}^{\prime \,\Phi}_{0}}=e^{ i \Phi \Theta(-\kappa)\Theta(t-m)}\ket{\tilde{\psi}^{\Phi}_{0}}$ that places the branch cut at $m=t,\kappa < 0$ as we see in Fig.\ref{ssh_chern}.(b)  (\carlos{need to double check signs}).
-This is in fact the result one gets when using the gauge that allows one to compute the winding number from the Zak phase.
+One can also obtain a polarization continuous in $\kappa$ by doing a gauge transformation 
+\begin{align}
+\ket{\tilde{\psi}^{\prime \,\Phi}}=e^{ i \Phi \Theta(-\kappa)\Theta(t-m)}\ket{\tilde{\psi}^{\Phi}}
+\end{align}
+\maria{where $\Theta$ is the Heaviside theta function (?)}. 
+This gauge transformation places the branch cut at $m=t,\kappa < 0$, as can be seen in Fig.\ref{ssh_chern}(b)  (\carlos{need to double check signs}). 
+This is in fact the result one gets when using the gauge that allows one to compute the winding number from the Zak phase. \maria{???}
 
-\begin{figure}[t]
-	\centering
-	\makebox[0pt]{
-		\subfloat[$\tilde{\mathcal{P}}_{\rm open}-L/2$]{
-			\includegraphics[width=50mm]{windingSSH1.pdf}
-		}
-		\subfloat[$\tilde{\mathcal{P}}_{\rm open,+}-L/2$]{
-			\includegraphics[width=45mm]{windingSSH2.pdf}
-		}
-	}\hspace{0mm}
+\changed{undo figure uncomment}
+%\begin{figure}[t]
+%	\centering
+%	\makebox[0pt]{
+%		\subfloat[$\tilde{\mathcal{P}}_{\rm open}-L/2$]{
+%			\includegraphics[width=50mm]{windingSSH1.pdf}
+%		}
+%		\subfloat[$\tilde{\mathcal{P}}_{\rm open,+}-L/2$]{
+%			\includegraphics[width=45mm]{windingSSH2.pdf}
+%		}
+%	}\hspace{0mm}
+%	
+%	\makebox[0pt]{
+%		\subfloat[$\tilde{\mathcal{P}}_{\rm open}-L/2$]{
+%			\includegraphics[width=45mm]{windingtp1.pdf}
+%		}
+%		\subfloat[$\tilde{\mathcal{P}}_{\rm open,+}-L/2$]{
+%			\includegraphics[width=45mm]{windingtp2.pdf}
+%		}
+%	}
+%	\caption{Polarization $\tilde{\mathcal{P}}_{\rm open}$ computed for three different gauge choices, for $t=1,\kappa'=0$ and system size $L=160$. (a) and (b) with $t'=0$ and (c) and (d) for $t'=-2$. }
+%	\label{ssh_chern}
+%\end{figure}
 
-	\makebox[0pt]{
-		\subfloat[$\tilde{\mathcal{P}}_{\rm open}-L/2$]{
-			\includegraphics[width=45mm]{windingtp1.pdf}
-		}
-		\subfloat[$\tilde{\mathcal{P}}_{\rm open,+}-L/2$]{
-			\includegraphics[width=45mm]{windingtp2.pdf}
-		}
-	}
-	\caption{Polarization $\tilde{\mathcal{P}}_{\rm open}$ computed for three different gauge choices, for $t=1,\kappa'=0$ and system size $L=160$. (a) and (b) with $t'=0$ and (c) and (d) for $t'=-2$. }
-	\label{ssh_chern}
-\end{figure}
-
-\begin{figure}[t]
-	\centering
-	\makebox[0pt]{
-		\subfloat[$\tilde{\mathcal{P}}_{\rm open}$]{
-			\includegraphics[width=50mm]{windingtpdis.pdf}
-		}
-		\subfloat[$C $]{
-			\includegraphics[width=45mm]{plotwinding.pdf}
-		}		
-	}\hspace{0mm}
-
-	\makebox[0pt]{
-		\subfloat[${\rm Log}_{10} (C \, {\rm mod} \, 1)$ ]{
-			\includegraphics[width=45mm]{plotwindingerror.pdf}
-		}
-	}
-	\caption{We compute (a)$\tilde{\mathcal{P}}_{\rm open}$, (b) $\nu$ and (c) error in the quantization of $\nu$ for the continuous line in the phase diagram, $t=1,t'=2$ and system size $L=400$.
+\changed{undo figure uncomment}
+%
+%\begin{figure}[t]
+%	\centering
+%	\makebox[0pt]{
+%		\subfloat[$\tilde{\mathcal{P}}_{\rm open}$]{
+%			\includegraphics[width=50mm]{windingtpdis.pdf}
+%		}
+%		\subfloat[$C $]{
+%			\includegraphics[width=45mm]{plotwinding.pdf}
+%		}		
+%	}\hspace{0mm}
+%	
+%	\makebox[0pt]{
+%		\subfloat[${\rm Log}_{10} (C \, {\rm mod} \, 1)$ ]{
+%			\includegraphics[width=45mm]{plotwindingerror.pdf}
+%		}
+%	}
+%	\caption{We compute (a)$\tilde{\mathcal{P}}_{\rm open}$, (b) $\nu$ and (c) error in the quantization of $\nu$ for the continuous line in the phase diagram, $t=1,t'=2$ and system size $L=400$.
 (a) is computed for $W=6$ showing that the branch cut is not affected by the disorder.
@@
 }
-	\label{winding}
-\end{figure}
+%	\label{winding}
+%\end{figure}
 
@@
 
-The analysis we did above of $\tilde{\mathcal{P}}_{\rm open}$ for the SSH chain seems rather trivial as $\tilde{\mathcal{P}}_{\rm open}$ does not seem to have more information than $\tilde{\mathcal{P}}$ modulo 1.
+The analysis we did above of $\tpo$ for the SSH chain seems rather trivial as $\tpo$ does not seem to have more information than $\tilde{\mathcal{P}}$ modulo 1.
 This is no longer true when $t' \neq 0$ allowing for a $\nu = 2$ phase.
-In Fig.\ref{ssh_chern}.(c) we see $\tilde{\mathcal{P}}_{\rm open}$ for the dotted cut in the phase diagram with an added symmetry breaking $\kappa$.
-We see clearly that $\tilde{\mathcal{P}}_{\rm open}$ distinguises between the $\nu=0$ and $\nu = 2$ phases.
-Similar to before there is a gauge choice that makes $\tilde{\mathcal{P}}_{\rm open}$ continuous in $\kappa$, however which gauge transformation does this is a priori not known, so in practice we can only compute the limit $\lim_{\kappa \rightarrow 0} \tilde{\mathcal{P}}_{\rm open}$ in the original gauge.
+In Fig.\ref{ssh_chern}.(c) we see $\tpo$ for the dotted cut in the phase diagram with an added symmetry breaking $\kappa$.
+We see clearly that $\tpo$ distinguishes between the $\nu=0$ and $\nu = 2$ phases.
+Similar to before, there is a gauge choice that makes $\tpo$ continuous in $\kappa$.
+However, which gauge transformation does this is a priori not known, so in practice we can only compute the limit $\lim_{\kappa \rightarrow 0} \tpo$ in the original gauge.
 
@@
 \begin{align}
-&t_i = t + \frac{W}{2} \omega_i \\
+&t_i = t + \frac{W}{2} \omega_i \nonumber\\
 &m_i = m + W \omega'_i
@@
 
-\carlos{In Appendix D I give an expression for the winding number for flux insertion.
+\mariac{removed the paragraph about appendix D}
+%\carlos{In Appendix D I give an expression for the winding number for flux insertion.
 However this expression seems problematic with a filled gap (Prodan refers to this).
 If we found an inhomogeneous term that induce a topological transition without filling the gap we can use this formula instead of comparing with the Prodan results, so we dont have to worry about the filled gap.
-Furthermore I worked an expression for this winding number in terms of the correlation function, which I believe we might be able to relate to our $\chi$.
+%Furthermore I worked an expression for this winding number in terms of the correlation function, which I believe we might be able to relate to our $\chi$.
 However I can only do this on the open chain, where the winding number itself is not well-defined because the Hamiltonian is not invertible.
@@
 
-\section{Twisted boundary condition Winding number}
-\carlos{Work in progress}
-For a system with chiral symmetry the Hamiltonian can be written in some basis as
-\begin{equation}
-H = \mqty(0 & h \\ h^\dagger & 0),
-\end{equation}
-where $h$ is in general a matrix. In translationally invariant systems the winding number is defined as 
-\begin{equation}
-\nu = -i\int_0^{2\pi} \frac{dk}{2\pi} {\rm Tr}[h(k)^{-1} \partial_k h(k)],
-\label{eq:winding_k}
-\end{equation}
-where $h(k)$ is the off-diagonal block of the Bloch Hamiltonian. In systems without translational invariance, one can similarly define the winding number
-\begin{equation}
-\nu = -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[h(\Phi)^{-1} \partial_\Phi h(\Phi)],
-\label{eq:winding_flux}
-\end{equation}
-for a $U(1)$ flux introduced (\carlos{I have only found a similar expression in the context of non-hermitian systems \cite{Gong2018}, but I assume this is the expression Prodan refers to.
+\mariac{removed appendix D}
+%\section{Twisted boundary condition Winding number}
+%\carlos{Work in progress}
+%For a system with chiral symmetry the Hamiltonian can be written in some basis as
+%\begin{equation}
+%H = \mqty(0 & h \\ h^\dagger & 0),
+%\end{equation}
+%where $h$ is in general a matrix. In translationally invariant systems the winding number is defined as 
+%\begin{equation}
+%\nu = -i\int_0^{2\pi} \frac{dk}{2\pi} {\rm Tr}[h(k)^{-1} \partial_k h(k)],
+%\label{eq:winding_k}
+%\end{equation}
+%where $h(k)$ is the off-diagonal block of the Bloch Hamiltonian. In systems without translational invariance, one can similarly define the winding number
+%\begin{equation}
+%\nu = -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[h(\Phi)^{-1} \partial_\Phi h(\Phi)],
+%\label{eq:winding_flux}
+%\end{equation}
+%for a $U(1)$ flux introduced (\carlos{I have only found a similar expression in the context of non-hermitian systems \cite{Gong2018}, but I assume this is the expression Prodan refers to.
 It gives the correct results in translationally invariant systems, however it is very unpractical.}).
@@
 In the following we will omit the flux parameter for simplicity.
-
-Note that the winding number can also be expressed in terms of the flat-band Hamiltonian
-\begin{align*}
-Q =& H (H^2)^{-1/2} \\
-=& \mqty(0 & q \\ q^\dagger & 0).
-\end{align*}
-Note that since $Q$ only has eigenvalues $\pm 1 $ , $Q^2 = \mathbb{I}$ and therefore $q^\dagger = q^{-1}$. We can therefore write
-\begin{equation}
-\nu = -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[q^\dagger \partial_\Phi q].
-\end{equation}
-In this basis the chiral symmetry operator is given by $S = \sigma_z \otimes \mathbb{I}$ such that the winding number can be expressed in terms of the full flat-band Hamiltonian as
-\begin{align}
-\nu =& \frac{i}{2}\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S Q^\dagger \partial_\Phi Q] \\
-=& \frac{i}{2}\int_0^{2\pi} \frac{d\Phi}{2\pi} \left({\rm Tr}[q \partial_\Phi q^\dagger]-{\rm Tr}[q^\dagger \partial_\Phi q] \right)\\
-=& -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[q^\dagger \partial_\Phi q] 
-\end{align}
-where in the last step we use the fact that the winding number is a real number. We can now write the winding number in terms of the correlation matrix
-\begin{align}
-\nu =& \frac{i}{2}\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S (1-2C) \partial_\Phi (1-2C)] \\
-=& -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S (1-2C) \partial_\Phi C] .
-\end{align}
-Introducing the flux via twisted boundary conditions allows us to easily construct the ground state with flux in the case of the open chain \eqref{eq:gauge}, therefore we will employ this gauge to compute the correlation matrix.
+%
+%Note that the winding number can also be expressed in terms of the flat-band Hamiltonian
+%\begin{align*}
+%Q =& H (H^2)^{-1/2} \\
+%=& \mqty(0 & q \\ q^\dagger & 0).
+%\end{align*}
+%Note that since $Q$ only has eigenvalues $\pm 1 $ , $Q^2 = \mathbb{I}$ and therefore $q^\dagger = q^{-1}$. We can therefore write
+%\begin{equation}
+%\nu = -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[q^\dagger \partial_\Phi q].
+%\end{equation}
+%In this basis the chiral symmetry operator is given by $S = \sigma_z \otimes \mathbb{I}$ such that the winding number can be expressed in terms of the full flat-band Hamiltonian as
+%\begin{align}
+%\nu =& \frac{i}{2}\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S Q^\dagger \partial_\Phi Q] \\
+%=& \frac{i}{2}\int_0^{2\pi} \frac{d\Phi}{2\pi} \left({\rm Tr}[q \partial_\Phi q^\dagger]-{\rm Tr}[q^\dagger \partial_\Phi q] \right)\\
+%=& -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[q^\dagger \partial_\Phi q] 
+%\end{align}
+%where in the last step we use the fact that the winding number is a real number. We can now write the winding number in terms of the correlation matrix
+%\begin{align}
+%\nu =& \frac{i}{2}\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S (1-2C) \partial_\Phi (1-2C)] \\
+%=& -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S (1-2C) \partial_\Phi C] .
+%\end{align}
+%Introducing the flux via twisted boundary conditions allows us to easily construct the ground state with flux in the case of the open chain \eqref{eq:gauge}, therefore we will employ this gauge to compute the correlation matrix.
 The correlation matrix can be written in position space as
-\begin{equation}
-C = \mqty(C_A & C_{AB} \\ C_{AB}^\dagger & C_B),
-\end{equation}
-where in this basis, assuming $S$ is a local symmetry it is also diagonal.
+%\begin{equation}
+%C = \mqty(C_A & C_{AB} \\ C_{AB}^\dagger & C_B),
+%\end{equation}
+%where in this basis, assuming $S$ is a local symmetry it is also diagonal.
 Note that the flux will only appear in the off-diagonal blocks $C_{AB}$ and therefore the first term in the trace above will vanish, leaving the winding number as
-\begin{align}
-\nu =&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S C \partial_\Phi C] \\
-=&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} \left( {\rm Tr}[S C_{AB} \partial_\Phi C_{AB}^\dagger]+{\rm Tr}[S C_{AB}^\dagger \partial_\Phi C_{AB}] \right)
-\end{align}
-We can not explicitely construct the correlation matrix with flux as
-\begin{align*}
-C_{ij}^\Phi =& \bra{\tilde{\Psi}_{0,{\rm open}}^\Phi}c_i^\dagger c_j \ket{\tilde{\Psi}_{0,{\rm open}}^\Phi} \\
-=&\bra{\tilde{\Psi}_{0,{\rm open}}^0}e^{i\Phi \hat{N}_A} c_i^\dagger c_j e^{-i\Phi \hat{N}_A}\ket{\tilde{\Psi}_{0,{\rm open}}^0} .
-\end{align*}
-Depending on if $i,j$ belong to $A$ or not, commuting the exponentials will leave additional factors or not, resulting in
-\begin{align*}
-C^\Phi = \mqty(C_A^0 & e^{i\Phi} C_{AB}^0 \\e^{-i\Phi} C_{AB}^{0\,\dagger} &C_B^{0}),
-\end{align*}
-where we have made explicit the flux dependence. The winding number is then
-\begin{align}
-\nu =&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S C^\Phi \partial_\Phi C^\Phi] \\
-=&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} \left( {\rm Tr}[-iS C_{AB}^0 C_{AB}^{0\,\dagger}]+{\rm Tr}[iS C_{AB}^{0\,^\dagger} C_{AB}^0] \right) \\
-=&  {\rm Tr}[S C_B^0(1-C_B^0)]-{\rm Tr}[S C_A^0(1-C_A^0)] 
-\end{align}
-where in the last step we use one of the properties of the correlation function stemming from $C^2 = C$.
+%\begin{align}
+%\nu =&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S C \partial_\Phi C] \\
+%=&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} \left( {\rm Tr}[S C_{AB} \partial_\Phi C_{AB}^\dagger]+{\rm Tr}[S C_{AB}^\dagger \partial_\Phi C_{AB}] \right)
+%\end{align}
+%We can not explicitely construct the correlation matrix with flux as
+%\begin{align*}
+%C_{ij}^\Phi =& \bra{\tilde{\Psi}_{0,{\rm open}}^\Phi}c_i^\dagger c_j \ket{\tilde{\Psi}_{0,{\rm open}}^\Phi} \\
+%=&\bra{\tilde{\Psi}_{0,{\rm open}}^0}e^{i\Phi \hat{N}_A} c_i^\dagger c_j e^{-i\Phi \hat{N}_A}\ket{\tilde{\Psi}_{0,{\rm open}}^0} .
+%\end{align*}
+%Depending on if $i,j$ belong to $A$ or not, commuting the exponentials will leave additional factors or not, resulting in
+%\begin{align*}
+%C^\Phi = \mqty(C_A^0 & e^{i\Phi} C_{AB}^0 \\e^{-i\Phi} C_{AB}^{0\,\dagger} &C_B^{0}),
+%\end{align*}
+%where we have made explicit the flux dependence. The winding number is then
+%\begin{align}
+%\nu =&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S C^\Phi \partial_\Phi C^\Phi] \\
+%=&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} \left( {\rm Tr}[-iS C_{AB}^0 C_{AB}^{0\,\dagger}]+{\rm Tr}[iS C_{AB}^{0\,^\dagger} C_{AB}^0] \right) \\
+%=&  {\rm Tr}[S C_B^0(1-C_B^0)]-{\rm Tr}[S C_A^0(1-C_A^0)] 
+%\end{align}
+%where in the last step we use one of the properties of the correlation function stemming from $C^2 = C$.
 (\carlos{So far this expression gives the correct result for translationally invariant systems however one needs to do the trick of breaking chiral symmetry slightly.
@@
 })
+%
 

### 2020-08-31 version2/draft.tex
..

@@
 This is a very simple model which nevertheless supports a rich phase diagram containing topological phases with winding numbers up to $\nu =2$. 
-In order to show that our results are not limited to any symmetry class~\cite{ryu2010topological}, we also include two symmetry breaking terms. 
+In order to show that our results are not limited to any particular symmetry class~\cite{ryu2010topological}, we also include two symmetry breaking terms. 
 The Hamiltonian is then given by
 \begin{align}
-H =& \sum_{i\alpha,j\beta} c_{i\alpha}^\dagger H_{ij,\alpha \beta} c_{j\beta} \\
+H =& \sum_{i\alpha,j\beta} c_{i\alpha}^\dagger H_{ij,\alpha \beta} c_{j\beta},
+\end{align}
+where $H_{ij} $ is given by 
+\begin{align}
 H_{ij} =& (m \sigma_x + \kappa \sigma_z)\delta_{ij}  + \frac{1}{2i}\kappa'\sigma_z (\delta_{i-j,1}-\delta_{i-j,-1})\nonumber\\
@@
 \end{align}
-with the corresponding Bloch Hamiltonian being
+and the corresponding Bloch Hamiltonian is given by
 \begin{align}
@@
 
-For $\kappa' \neq 0$ only particle-hole symmetry $C$ is preserved and the system belongs to symmetry class $D$. 
+For $\kappa' \neq 0$, only particle-hole symmetry $C$ is preserved and the system belongs to symmetry class $D$. 
 In one spatial dimension, the latter is characterized by a $\mathbb{Z}_2$ invariant.  
@@
 \makebox[0pt]{
-\subfloat[$\kappa = .3,\kappa'=0$]{
+\subfloat[$\kappa = 0.3,\kappa'=0$]{
   \includegraphics[width=35mm]{2c.pdf}
@@
 
-In this section we review some results of the modern theory of polarization, which gave a geometrical description of the polarization \cite{Resta1992,KingSmith1993,Vanderbilt1993,Resta1997} , as well as a recent article by Watanabe and Oshikawa \cite{Watanabe2018} that clarified the meanings of the different polarizations one can define this way and the relations between them.
+In this section we review some aspects of the modern theory of polarization, which gave a geometrical description of the polarization \cite{Resta1992,KingSmith1993,Vanderbilt1993,Resta1997}. 
+We also discuss the results of a recent article by Watanabe and Oshikawa \cite{Watanabe2018} that clarified the meanings of the different polarizations one can define this way, as well as the relations between them.
 For clarity we will use the same notation they use.
@@
 In translational invariant systems $u_{p\mu}^{j\beta} = e^{ipj}\tilde{u}_{p \mu}^{\beta}$, where $\tilde{u}_{p \mu}^{\beta}$ are the components of the eigenstates of the Bloch Hamiltonian, $\ket{\tilde{u}_{k\mu}}$.
-One can now compute the Zak phase \cite{Zak1989}, which for a two-band \mariac{do we need the restriction to two-band models?} model is defined as the geometric phase acquired by the occupied state $\ket{\tilde{u}_k}$ as it winds around the Brillouin zone,
+One can now compute the Zak phase \cite{Zak1989}, which for a two-band  model is defined as the geometric phase acquired by the occupied state $\ket{\tilde{u}_k}$ as it winds around the Brillouin zone,
 \begin{equation}
@@
 where $A_k = \bra{\tilde{u}_k}\partial_k \ket{\tilde{u}_k}$ is the Berry connection \cite{Berry1984}. 
+The generalization to higher-band models is straightforward.
+
 The Zak phase is typically regarded as just a phase because it is only gauge invariant modulo $2\pi$.
-However, for two different states defined by a parameter $\lambda$, as long as $A_k(\lambda)$ is smooth in the path connecting them the change in the Zak phase can be computed as
+However, for two different states defined by a parameter $\lambda$ --- assuming $A_k(\lambda)$ is smooth in the path connecting them --- the change in the Zak phase can be computed as
 \begin{equation}
@@
  is the Berry curvature.
-The Berry curvature is fully gauge invariant and, therefore, the change in the Zak phase as defined in \eqref{eq:change_zak_phase} is defined along the full real line.
-The Zak phase itself, as defined above, can be shown to be proportional to the winding number for a certain gauge (see Appendix \ref{appendix:winding_ssh}), which means that it carries physical information beyond modulo $2\pi$.
+The Berry curvature is fully gauge invariant and, therefore, the change in the Zak phase \eqref{eq:change_zak_phase} is defined along the full real line.
+The Zak phase itself, as defined in Eq.~\eqref{eq:zak_phase}, can be shown to be proportional to the winding number for a certain gauge (see Appendix \ref{appendix:winding_ssh}), which means that it carries physical information beyond modulo $2\pi$.
 Note that other ways of computing the Zak phase rely on computing $e^{i\gamma}$ instead, such that the result is always defined only modulo $2\pi$.
@@
 which allows us to make the identification $ \tilde{\mathcal{P}}^{\rm Bloch} = \gamma/2\pi$. 
-As discussed above,  this polarization is defined as an absolute value only modulo 1, however its changes can be defined as a real number.
-This is no surprise as the derivative of the polarization is related to the current which is a physical observable, unlike $\tilde{\mathcal{P}}^{\rm Bloch}$.
+As discussed above,  this polarization is defined only modulo 1.
+Its changes, however, can be defined as a real number. 
+This is not surprising,  as the derivative of the polarization is related to the current which is a physical observable, unlike $\tilde{\mathcal{P}}^{\rm Bloch}$.
 
@@
 \carlos{I have to check overall sign and sign of exponential.}and  ${\rm det }\!' $ indicates that the determinant is restricted to the space of \emph{occupied} single-particle states.
-Below we will employ this expression to compare our results with.
+Below we will employ this expression to compare our results with \mariac{XXX}.
 
-There is one last polarization we must consider and it is the one obtained when the flux is introduced in the bond between sites $i=1$ and $i=L$ via twisted boundary conditions.
-This is equivalent to performing the gauge transformation
+There is one last polarization we must consider. 
+It is obtained when the flux is introduced in the bond between sites $i=1$ and $i=L$ via twisted boundary conditions. This is equivalent to performing the gauge transformation
 \begin{align}\label{eq:Psi_tilde}
@@
 \end{equation} 
-\maria{Ref to Oshikawa paper?}
 %The previous discussion about the Zak phase being defined only modulo $2\pi$ although its changes can be defined as a real number applies in the same way to the polarization obtained in \eqref{eq:p_tilde}.
@@
 \end{align}
-For a system where we occupy all single-particle eigenstates with energies $E<0$ the correlation matrix can be written  as
+For a system where all single-particle eigenstates with energies $E<0$ are occupied, the correlation matrix can be written  in terms of the Hamiltonian as
 \begin{align}\label{eq:corr_mat2}
@@
 where $\mathcal{K}$ is a normalization constant and $\mathcal{H}$ is a quadratic Hamiltonian, referred to as entanglement Hamiltonian. 
-The eigenvalues $\epsilon_j$ of $\mathcal{H}$, referred to as entanglement energies,  are related to the EOS by 
+The eigenvalues $\epsilon_j$ of $\mathcal{H}$, referred to as entanglement energies,  are related to those of the EOS by 
 \begin{align}\label{eq:xi_eps}
@@
 
-One property of the subsystem correlation matrix we will make use of later is that, for sufficiently large systems, its eigenstates are found to be localize on either virtual edge if $\xi \neq 0,1$, or they are found to be bulk modes if the correspondent eigenvalues are exponentially close to $\xi = 0,1$ \cite{Peschel2008}.
+One property of the subsystem correlation matrix we will make use of later is that, for sufficiently large systems, its eigenstates are found to be localized on either virtual edge if $\xi \neq 0,1$, or they are found to be bulk modes if the correspondent eigenvalues are exponentially close to $\xi = 0,1$ \cite{Peschel2008}.
 
@@
 In the following sections, we show how the Zak phase can be recovered from the EOS.
-We also give a simple formula that is identical to the Zak phase in the thermodyamic limit.  
-However, for a generic, finite size system, there will be finite-size discrepancies between our formula and the Zak phase.
-In order to reduce these, we are going to choose $A$ as half the system in the remainder of the manuscript.
+We also provide a simple formula in terms of the EOS eigenvalues that is identical to the Zak phase in the thermodyamic limit.  
+However, for a generic, finite size system, there will be finite-size discrepancies between our formula and the Zak phase. \mariac{see scaling plot}
+In order to reduce these, we are going to choose subsystem $A$ as half the system in the remainder of the manuscript.
 The results (in the thermodynamic limit) do not depend on this choice. 
@@
 For symmetry-protected topological systems in one dimension, the quantized Zak phase is zero whenever there is an even number  of eigenvalues per edge at $\xi = 1/2$, and $\pi$ when the number is odd~\cite{Peschel2008}.
+In the absence of symmetry-protection, much less is known. 
 Ryu and Hatsugai considered the fully dimerized SSH chain with broken chiral symmetry~\cite{Ryu2006}.  
-This model is rather special in that there is only a single pair of eigenvalues, related by symmetry, in the EOS that is not identical to $0$ or $1$. 
-In fact, the authors could show that the value of one of these `midgap states' is identical to the Zak phase divided by $2\pi$, or $\tilde{\mathcal{P}}^{\rm Bloch}$. 
+This model is rather special in that there is only a single pair of eigenvalues, related by symmetry \mariac{which symmetry?}, in the EOS that is not strictly identical to $0$ or $1$. 
+The authors could show that the value of one of these `midgap states' is identical to the Zak phase divided by $2\pi$ or, alternatively, $\tilde{\mathcal{P}}^{\rm Bloch}$. 
 Their result is a particular limit of Eq.~\eqref{eq:chi_integrable}. 
@@
 %Beyond symmetry protected systems, there was observed that in the fully dimerized SSH chain, with only a single mode present in the ES, the Zak phase is equal to the midgap eigenvalue \cite{Ryu2006}.
-Huang and Arovas considered a two-dimensional chern insulator and noted that the Zak phase qualitatively follows the `virtual edge mode' that connects the EOS values at 0 to those at 1~\cite{Huang2012,Huang2012-2}.
+
+Huang and Arovas considered a two-dimensional Chern insulator and noted that the Zak phase qualitatively follows the `virtual edge mode' that connects the EOS values at 0 to those at 1~\cite{Huang2012,Huang2012-2}.
 Interpreting this as a one-dimensional system with a parameter, we can explain this behavior by Eq.~\eqref{eq:chi_integrable}, noting that there is a single pair of eigenvalues that dominates the sum.
-Note that the observation of Ref.s~\cite{Huang2012,Huang2012-2} is particular to a system with Chern number 0 or $\pm 1$, and it generically fails for systems with higher Chern numbers. 
-In the latter case, there are several terms in Eq.~\eqref{eq:chi_integrable} with comparably large contributions. 
+Note that the observation of Ref.s~\cite{Huang2012,Huang2012-2} is particular to systems with Chern number 0 or $\pm 1$, and it generically fails for systems with higher Chern numbers. 
+In the latter case, there are several terms in Eq.~\eqref{eq:chi_integrable} with comparably large contributions and, consequently, the Zak phase deviates considerably from the 'virtual edge modes'. 
 %For cases with more edge modes present in the ES it has been shown that the Zak phase follows the midgap eigenvalues qualitatively \cite{Huang2012,Huang2012-2}.
@@
 nearest neighbor hopping models such as the SSH chain, and applies to the Hamiltonian~\eqref{bdi_model} when $t'=0$.
-In this case, one can obtain the Zak phase in a very simple fashion: We reorder the eigenvalues of the ES by magnitude, $\xi_1 < \xi_2 < ...< \xi_{L_AM}$, and compute
+In this case, one can obtain the Zak phase in a very simple fashion: We reorder the eigenvalues of the EOS by magnitude, $\xi_1 < \xi_2 < ...< \xi_{L_AM}$, and compute
 \begin{equation}
@@
 \end{equation}
-In the thermodynamic limit, $\chi$ becomes identical to either $\tilde{\mathcal{P}}^{\rm Bloch}$ or $1-\tilde{\mathcal{P}}^{\rm Bloch}$
+In the thermodynamic limit, $\chi$ becomes identical to either $\tilde{\mathcal{P}}^{\rm Bloch}$ or $1-\tilde{\mathcal{P}}^{\rm Bloch}$: \mariac{comment on the ambiguity?}
 \begin{equation}
@@
 \end{equation}
-For translational invariant systems the EOS is symmetric such that a crossing of the eigenvalues $\xi$ results in $\chi$ changing sign.
+\mariac{I don't understand what we want to say here.} For translational invariant systems the EOS is symmetric such that a crossing of the eigenvalues $\xi$ results in $\chi$ changing sign.
 This is particularly irrelevant in the case of symmetry-protected states where the Zak phase is quantized to $0$ or $\pi$.
-
-In the case where there is only one eigenvalue per edge that is $\neq 0,1$ in  the EOS, like in the fully dimerized limit of the SSH chain, we recover the equality between the midgap eigenvalue and the Zak phase found in Ref.~\cite{Ryu2006}.
+In the special case where there is only one eigenvalue per edge that is $\neq 0,1$ in  the EOS, like in the fully dimerized limit of the SSH chain, we recover the equality between the midgap eigenvalue and the Zak phase found in Ref.~\cite{Ryu2006}.
 

### 2020-08-31 version2/draft.tex
Add files via upload

@@
 
-In Appendix A we obtain that $\chi$ should be the sum of all the eigenvalues of eigenstates related to the cut between $i=1$ and $i=L$ that make a subspace $A_L$.
+In Appendix A we obtain that $\tilde{\mathcal{P}}$ is given by the sum of all the eigenvalues of $C_A$ with eigenstates related to the cut between $i=1$ and $i=L$ that make a subspace $A_L$.
 The dimension of $A_L$ is however not well-defined as one can always include bulk modes into it.
-Since the bulk modes have $\xi = 0,1$ this makes $\chi$ well-defined only modulo 1.
+Since the bulk modes have $\xi = 0,1$ this makes $\tilde{\mathcal{P}}$ well-defined only modulo 1.
 This can be avoided by computing instead
@@
 \end{equation}
-The difference with the closed chain being that now there is no ambiguity on when an electron crosses the boundary with the twisted boundary condition.
-Note that since the chain is open, $\Phi$ is not a flux but should now be regarded as a parameter and therefore $\tpo$ does not need to be a physical polarization.
+There is now no ambiguity in how the flux is introduced, as oppossed to the case with PBC.
+Note that $\Phi$ is not a magnetic flux but should be treated as an additional parameter, and therefore $\tpo$ is not necessarily the physical polarization of the open chain.
 $\tpo$ can be obtained in an even simpler way than $\tilde{\mathcal{P}}$ as
@@
 where we used Eq.~\eqref{eq:gauge} and $C_{\rm o}$($C_{A,\rm o}$) denotes the (reduced) density matrix of the open chain.
-Note that Eq.~\eqref{eq:phi_independence} is nothing else than the generalization of $\chi$ to the open chain. 
-\maria{I am not entirely happy with the following argument, as it goes wrong in the topological phase. We mention this later, but we should already mention it here. }
-The idea is that in the process of opening the chain, which can be done adiabatically, all eigenvalues that are not in $A_L$ should get pushed to $\xi=0$ or $\xi=1$. 
-If for a given path in parameter space $\boldsymbol{\lambda}_i \rightarrow \boldsymbol{\lambda}_f$ their contribution remains constant, that is, there are no jumps between $\xi=0$ and $\xi=1$ and $\tpo (\boldsymbol{\lambda} )$ is smooth, then changes in $\tpo(\boldsymbol{\lambda}_i ) $ along the path will be defined as a real number and not only modulo 1.
-Since $\tilde{\mathcal{P}}$ is a local property of the bond between $i=1$ and $i=L$ one would expect it to not depend on what happens far from it such that $\tpo$ and $\tilde{\mathcal{P}}$ are equivalent.
+
+The idea is that in the process of opening the chain, which can be done adiabatically, all eigenvalues that are not in $A_L$ should get pushed to $\xi=0$ or $\xi=1$, by getting rid of the subspace $A_R$ by this process we can finally obtain $\tpo$ by summing all eigenvalues of $C_{A,\rm o}$.
+The advantage of doing this is that, if for a given path in parameter space $\boldsymbol{\lambda}_i \rightarrow \boldsymbol{\lambda}_f$ their contribution remains constant, that is, there are no jumps between $\xi=0$ and $\xi=1$ such that $\tpo (\boldsymbol{\lambda} )$ is smooth, then changes in $\tpo(\boldsymbol{\lambda}_i ) $ along the path will be defined as a "real number" and not only modulo 1.
+There are, however, two caveats.
+When we open the chain, if the system is in a topological non-trivial state, zero-energy modes will appear.
+The ground state of the closed chain has several equivalent degenerate ground states for the open chain, so that $C_{A,\rm o}$ is not well-defined at these points.
+The second issue is that, even if we allow for symmetry-breaking terms so that we can compute $\tpo$, there might be integer jumps in $\tpo$ where the edge-modes cross zero-energy.
+However, as we will see below these jumps in $\tpo$ due to the zero modes are not simple jumps but branch cuts that provide the correct description of the closed system.
 
@@
 
-Consider the simple case of the SSH chain, with $t'=0$ in the model.
-In the chiral symmetric case, $\kappa = 0$, the open chain presents zero-energy modes for $m<t$ that seem problematic.
-$\tilde{\mathcal{P}}_{\rm open}$  will depend on the occupation of these zero modes although we just argued that $\tilde{\mathcal{P}}$ should not, since the zero-energy modes are localized far from the virtual cut at $i=1$.
-We can break the chiral symmetry in order to split the zero-energy modes so we do not have to make this choice, but the resulting $\tilde{\mathcal{P}}_{\rm open}$ will be discontinuous at $\kappa = 0$, suggesting that the limit does not exist.
-This can be seen from the fact that the Hamiltonian now has the symmetry
+Consider first the simple case of the SSH chain, with $t'=0$ in the model. When we include the chiral-symmetry braking parameter $\kappa$ the Hamiltonian now has the symmetry
 \begin{equation}
-SH(\kappa)S = -H(-\kappa), \quad S C_A(\kappa) S = \mathbb{I}-C_A(-\kappa)
+SH(\kappa)S = -H(-\kappa), \quad S C_A(\kappa) S = \mathbb{I}-C_A(-\kappa),
 \end{equation}
 so $\tpo(-\kappa) = L-\tpo(\kappa)$.
-Unless $\tpo(0) = L/2$, which only happens in the trivial case, the result will be discontinuous at $\kappa = 0$.
-We can see this in Fig.\ref{ssh_chern}.(a) where we show $\tpo$ in the space $(m,\kappa)$.
-This branch cut is in fact not an artifice of the open chain trick, but the result of the degenerate point at $(m=t,\kappa=0)$ having an associated Chern number $C=1$ \cite{Asboth2016}.
+Unless $\tpo(0) = L/2$, which is the result expected for a trivial insulator, the result will be discontinuous at $\kappa = 0$.
+We can see this in Fig.\ref{ssh_chern}.(a) where we show $\tpo$ in the parameter space $(m,\kappa)$.
+The presence of the zero-mode for $m<1$ appears in $\tpo$ as a branch cut.
+This branch cut is in fact not an artifice of the open chain trick, but the result of the degenerate point at $(m=1,\kappa=0)$ having an associated Chern number $C=1$ \cite{Asboth2016}.
 
-Consider a loop around the singularity at $(m=t,\kappa=0)$ parametrized by $\theta$. The Chern number is then defined as
+Consider a counterclockwise loop around the degenerate point at $(m=t,\kappa=0)$ parametrized by $\theta$, starting from the branch cut. The Chern number is then defined as
 \begin{equation}
@@
 \end{equation} 
-with the Berry curvature $\Omega_{\theta\Phi}$ defined in Eq.~\ref{eq:BerryCurvature}. 
-From  Eq.~\eqref{eq:tpo}, it follows that the Berry connection for the flux is simply $A_\Phi(\theta,\Phi,) = -i\tpo(\theta)$, so we only need to compute $A_\theta(\theta,\Phi)$.
-(\carlos{Please double check this calculation}) Expanding $\ketpsio{\Phi}$ in the eigenbasis of the number operator $\hat{N_A}$ -- denoted by $\{\ket{j}\}$ -- we can compute its derivative with respect to the parameter $\theta$ as
+with the Berry curvature $\Omega_{\theta\Phi}$ defined in equation~\eqref{eq:BerryCurvature}. 
+From  Eq.~\eqref{eq:tpo}, it follows that the Berry connection for the flux is simply $A_\Phi(\theta,\Phi,) = \tpo(\theta)$, so we now need to compute $A_\theta(\theta,\Phi)$.
+Expanding $\ketpsio{\Phi}$ in the eigenbasis of the number operator $\hat{N_A}$ -- denoted by $\{\ket{j}\}$ -- we can compute its derivative with respect to the parameter $\theta$ as
 \changed{changed}
@@
 The integral in \eqref{eq:chern} is performed on the surface of the torus defined by $(\theta,\Phi)$.
-Because the Berry connection $A_\Phi(\theta)$ is not smooth one cannot directly apply Stokes theorem.
+Because of the branch cut the Berry connection $A_\Phi(\theta)$ cannot be made smooth for the whole torus and directly apply Stokes theorem.
 Instead, consider that we split the torus in two cylinders, $S_1$ with $\theta\in[\theta_0,2\pi-\theta_0]$ and  $S_2$ with $\theta\in[2\pi-\theta_0,2\pi+\theta_0]$ and we change the gauge in $\boldsymbol{A}(\theta)$ such that it is smooth in each cylinder.
-We can now apply Stokes theorem to the two cylinders independently
+We can now apply Stokes theorem to the two cylinders independently \cite{Kohmoto1985}.
+For $S_1$ we have
 \begin{align}
@@
 \end{align}
-The Chern number is then $C=C_1+C_2$. Noting again that $A'_\Phi(\theta)$ is continuous in $S_2$ taking the limit of $a\rightarrow 0^+$ makes $C_2$ vanish and the Chern number is given by
+The Chern number is then $C=C_1+C_2$. Noting again that $A'_\Phi(\theta)$ is continuous in $S_2$ taking the limit of $\theta_0\rightarrow 0^+$ makes $C_2$ vanish and the Chern number is given by
 \begin{align}
-C =& \int_0^{2\pi} \frac{d\Phi}{2\pi} [A_\Phi (2\pi^-,\Phi) -A_\Phi (0^+,\Phi)].
+C =& \int_0^{2\pi} \frac{d\Phi}{2\pi} [A_\Phi (2\pi^-,\Phi) -A_\Phi (0^+,\Phi)],
 \end{align}
-If we consider the loop starting at $\kappa = 0$ this is exactly the case shown in Fig.\ref{ssh_chern}.(a), where $A_\Phi (\theta,\Phi) = \tilde{\mathcal{P}}_{\rm open}(\theta)$ is smooth everywhere but in $\theta = 0$, and therefore we can compute the Chern number simply as
+or
 \begin{align}
-C= \tpo(\kappa=0^+)-\tpo(\kappa=0^-).
+C(m)= \tpo(m,\kappa=0^+)-\tpo(m,\kappa=0^-),
 \end{align}
+where $C(m)$ is the Chern number of any loop around the degenerate point crossing $\kappa=0$ at $m$.
 Using the symmetry of the polarization described above this results in 
 \begin{align}
-C= 2\tpo(0^+)-L.
+C(m)= 2\tpo(m,\kappa=0^+)-L,
 \label{eq:Chern_number}
 \end{align}
-The discontinuity in $\tpo$ is therefore related to the Chern number.% and it is not merely continuous modulo 1.
+which results in $+1$ when $m<1$ and $0$ when $m>1$, as it is known for the SSH chain. The discontinuity in $\tpo$ is therefore related to the Chern number.% and it is not merely continuous modulo 1.
 
-One can also obtain a polarization continuous in $\kappa$ by doing a gauge transformation 
-\begin{align}
-\ket{\tilde{\psi}^{\prime \,\Phi}}=e^{ i \Phi \Theta(-\kappa)\Theta(t-m)}\ket{\tilde{\psi}^{\Phi}}
-\end{align}
-\maria{where $\Theta$ is the Heaviside theta function (?)}. 
-This gauge transformation places the branch cut at $m=t,\kappa < 0$, as can be seen in Fig.\ref{ssh_chern}(b)  (\carlos{need to double check signs}). 
-This is in fact the result one gets when using the gauge that allows one to compute the winding number from the Zak phase. \maria{???}
+%\changed{undo figure uncomment}
+\begin{figure}[t]
+	\centering
+	\makebox[0pt]{
+		\subfloat[$\tpo-L/2$]{
+			\includegraphics[width=50mm]{windingSSH1.pdf}
+		}
+		\subfloat[$\tpo-L/2$]{
+			\includegraphics[width=45mm]{windingtp1.pdf}
+		}
+	}
+	\caption{Polarization $\tpo$ computed for $t=1,\kappa'=0$ and system size $L=160$, for the SSH chain $t'=0$ in (a) and for $t'=-2$ in (b) }
+	\label{ssh_chern}
+\end{figure}
 
-\changed{undo figure uncomment}
-%\begin{figure}[t]
-%	\centering
-%	\makebox[0pt]{
-%		\subfloat[$\tilde{\mathcal{P}}_{\rm open}-L/2$]{
-%			\includegraphics[width=50mm]{windingSSH1.pdf}
-%		}
-%		\subfloat[$\tilde{\mathcal{P}}_{\rm open,+}-L/2$]{
-%			\includegraphics[width=45mm]{windingSSH2.pdf}
-%		}
-%	}\hspace{0mm}
-%	
-%	\makebox[0pt]{
-%		\subfloat[$\tilde{\mathcal{P}}_{\rm open}-L/2$]{
-%			\includegraphics[width=45mm]{windingtp1.pdf}
-%		}
-%		\subfloat[$\tilde{\mathcal{P}}_{\rm open,+}-L/2$]{
-%			\includegraphics[width=45mm]{windingtp2.pdf}
-%		}
-%	}
-%	\caption{Polarization $\tilde{\mathcal{P}}_{\rm open}$ computed for three different gauge choices, for $t=1,\kappa'=0$ and system size $L=160$. (a) and (b) with $t'=0$ and (c) and (d) for $t'=-2$. }
-%	\label{ssh_chern}
-%\end{figure}
+%\changed{undo figure uncomment}
 
-\changed{undo figure uncomment}
-%
-%\begin{figure}[t]
-%	\centering
-%	\makebox[0pt]{
-%		\subfloat[$\tilde{\mathcal{P}}_{\rm open}$]{
-%			\includegraphics[width=50mm]{windingtpdis.pdf}
-%		}
-%		\subfloat[$C $]{
-%			\includegraphics[width=45mm]{plotwinding.pdf}
-%		}		
-%	}\hspace{0mm}
-%	
-%	\makebox[0pt]{
-%		\subfloat[${\rm Log}_{10} (C \, {\rm mod} \, 1)$ ]{
-%			\includegraphics[width=45mm]{plotwindingerror.pdf}
-%		}
-%	}
-%	\caption{We compute (a)$\tilde{\mathcal{P}}_{\rm open}$, (b) $\nu$ and (c) error in the quantization of $\nu$ for the continuous line in the phase diagram, $t=1,t'=2$ and system size $L=400$.
+\begin{figure}[t]
+	\centering
+	\makebox[0pt]{
+		\subfloat[$2\tilde{\mathcal{P}}_{\rm open}-L$($m$ should go from 0 to 3)]{
+			\includegraphics[width=45mm]{windingtpdis.pdf}
+		}
+		\subfloat[$C $]{
+			\includegraphics[width=45mm]{plotwinding.pdf}
+		}		
+	}\hspace{0mm}
+
+	\makebox[0pt]{
+		\subfloat[${\rm Log}_{10} (C \, {\rm mod} \, 1)$ ]{
+			\includegraphics[width=45mm]{plotwindingerror.pdf}
+		}
+	}
+	\caption{We compute (a)$\tilde{\mathcal{P}}_{\rm open}$, (b) $\nu$ and (c) error in the quantization of $\nu$ for the continuous line in the phase diagram, $t=1,t'=2$ and system size $L=400$.
 (a) is computed for $W=6$ showing that the branch cut is not affected by the disorder.
@@
 }
-%	\label{winding}
-%\end{figure}
+	\label{winding}
+\end{figure}
 
@@
 This is no longer true when $t' \neq 0$ allowing for a $\nu = 2$ phase.
-In Fig.\ref{ssh_chern}.(c) we see $\tpo$ for the dotted cut in the phase diagram with an added symmetry breaking $\kappa$.
-We see clearly that $\tpo$ distinguishes between the $\nu=0$ and $\nu = 2$ phases.
-Similar to before, there is a gauge choice that makes $\tpo$ continuous in $\kappa$.
-However, which gauge transformation does this is a priori not known, so in practice we can only compute the limit $\lim_{\kappa \rightarrow 0} \tpo$ in the original gauge.
+In Fig.\ref{ssh_chern}.(b) we see $\tpo$ for the dotted cut in the phase diagram with an added symmetry breaking $\kappa$ term.
+We can see that a loop around the two degenerate points will have a Chern number of $C=2$.
 
-Our expression for the Chern number \eqref{eq:Chern_number} is also valid without translational invariance. Consider now that we add disorder in the $t$ and $m$ parameters
+A Chern number C calculated for a gapped loop crossing the trivial phase and some point at $(m,\kappa=0)$ requires the presence of at least C zero-modes.
+In this case we can see that the equivalence between the Chern number and the number of zero-modes, or the winding number, is exact.
+It is also interesting to point out that even far away from the symmetric point $\tpo$ still shows plateaus that point to a nearby topological transition.
+Note , however, that there are many ways of breaking chiral symmetry and one can find terms where a loop crossing two points in the $\nu=0$ and $\nu=2$ has Chern number $C=0$ such that these two phases would appear equivalent.
+\carlos{write term}
+
+Since $\tpo$ is naturally described in position space we can also use this method for a disordered system. Consider now that we add disorder in the $t$ and $m$ parameters
 \begin{align}
@@
 In order to avoid it we break chiral symmetry locally in the edge of the open chain so it only affects the zero modes.
-If any of the other states cross $E=0$ they do it in pairs such that $\tilde{\mathcal{P}}_{\rm open}$ is not affected by it.
-We show $\tilde{\mathcal{P}}_{\rm open}$ in Fig.\ref{winding}.(a) using the local symmetry breaking term for a strong disorder, $W=6t$, where the gap is indeed filled with states.
+If any of the other states cross $E=0$ they do it in pairs such that $\tpo$ is not affected by it.
+We show $\tpo$ in Fig.\ref{winding}.(a) using the local symmetry breaking term for a strong disorder, $W=6t$, where the gap is indeed filled with states.
+For the chain without disorder this local symmetry-breaking term would result in the same Chern number.
+The main difference being that since it only affects the zero-modes, $\tpo$ is quantized to multiples of $1/2$.
+
 In Fig.\ref{winding}.(b) we show $C$ for the continuous line in the phase diagram and increasing disorder strength.
 The Chern number as defined in \eqref{eq:Chern_number} indeed provides a topological invariant which can be compared with the winding number obtained in reference \cite{Song2014} for the same system.
-Note that for large enough systems the spectrum of $C_A$ is not affected by the local chiral symmetry breaking term other than the occupation of the zero-energy modes such that $C$ is still quantized, however near the transitions it is smoothen by the disorder average We show the error in the quantization of $C$ in figure \ref{winding}.(c) .
+Note that for large enough systems the spectrum of $C_A$ is not affected by the local chiral symmetry breaking term other than the occupation of the zero-energy modes such that $C$ is still quantized, however near the transitions it is smoothen by the disorder average.
+We show the error in the quantization of $C$ in figure \ref{winding}.(c) .
+\carlos{Fig.6.a seems important to discuss how with local symmetry braking term the relation between the chern number and the number of zero modes still remains.
+Maybe it would be enough to show a cut of 6.b }
 
@@
 }
+
+% - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - %
+
+\section{Chern Number in 2D}
+
+We have shown how $\tpo$ allows us to compute Chern numbers of 1D chains in a simpler way. This can be generalized to 2D systems. Consider the following Chern insulator
+\begin{align}
+H =& \sum_{im\alpha,jn\beta} c_{im\alpha}^\dagger H_{im\alpha,jn\beta} c_{jn\beta}, \\
+H_{im\alpha,jn\beta}=& \frac{1}{2}(i\sigma_x-\sigma_z)\delta_{i,j+1}\delta_{mn}+\frac{1}{2}(-i\sigma_x-\sigma_z)\delta_{i,j-1}\delta_{mn} \nonumber\\
++&\frac{1}{2}(i\sigma_y -\sigma_z)\delta_{ij}\delta_{m,n+1} +\frac{1}{2}(-i\sigma_y -\sigma_z)\delta_{ij}\delta_{m,n-1} \nonumber\\
++&(2-m)\sigma_z\delta_{ij}\delta_{mn},
+\end{align}
+where $i,j$ describe the $x$ direction and $m,n$ the $y$ direction.
+The system with periodic boundary conditions is defined on a torus.
+In order to apply the method of computing $\tpo$ described in the last section we will open the torus into a cylinder periodic in $y$.
+We then fourier transform only the $y$ direction to obtain
+\begin{align}
+H =& \sum_{i\alpha,j\beta}\sum_{k_y} c_{i\alpha}^\dagger(k_y) H_{i\alpha,j\beta}(k_y) c_{j\beta}(k_y), \\
+H_{i\alpha,j\beta}(k_y)=& \frac{1}{2}(i\sigma_x-\sigma_z)\delta_{i,j+1}+\frac{1}{2}(-i\sigma_x-\sigma_z)\delta_{i,j-1} \nonumber\\
++&\sin(k_y)\sigma_y+[2-m-\cos(k_y)]\sigma_z.
+\end{align}
+We can treat the Hamiltonian with elements $H_{i\alpha,j\beta}(k_y)$ as a 1D Hamiltonian with an extra parameter $k_y$ and compute $\tpo$, or equivalently $A_\Phi(\Phi,k_y)$, as in last section.
+But first lets analyze the entanglement occupancy spectrum of this system, shown in Fig.\ref{chern_insulator}(a) and (b) for $m=1,3$ respectively.
+For $0<m<2$ the EOS presents two edge modes that connect the occupied and empty bands that cross $\xi=1/2$ at $k_y=0$, in this region the system has a Chern number $C=+1$.
+For $2<m<4$ there are also edge modes but they cross $\xi=1/2$ at $k_y=\pi$ and the system has $C=-1$.
+For $m>4$ the system is a trivial insulator with $C=0$.
+When we open the chain one of the edge modes will be pushed to the bulk and will present a jump between $\xi=1,0$ whenever the edge-modes crossed zero energy.
+In this model, due to the symmetries, this is only possible at $k_y=0,\pi$.
+
+We can now take a look at $\tpo$ in the parameter space $(m,k_y)$, shown in figure \ref{chern_insulator}.(c).
+We see, again, that $\tpo$ remains continous except for the branch cuts that appear at the points where the physical edge-modes cross zero energy.
+The Chern number can be obtained as
+\begin{equation}
+C(m) = \int_{-\pi}^{\pi}\frac{dk_y}{2\pi} \partial_{k_y} \tpo(m,k_y),
+\label{eq:Zaletel_chern}
+\end{equation}
+However, since we know the position of the branch cuts we can integrate avoiding them and obtain
+\begin{align}
+C(m) =& [\tpo(m,k_y=0^-)-\tpo(m,k_y=-\pi^+)]\\
+&+[\tpo(m,k_y=\pi^-)-\tpo(m,k_y=0^+)].
+\end{align}
+
+\begin{figure}[t]
+	\centering
+	\makebox[0pt]{
+		\subfloat[Spectrum of $C_{A}$ for $m=1$]{
+			\includegraphics[width=45mm]{chernES0.pdf}
+		}
+		\subfloat[Spectrum of $C_{A}$ for $m=3$]{
+			\includegraphics[width=45mm]{chernESpi.pdf}
+		}
+	}\hspace{0mm}
+
+	\makebox[0pt]{
+		\subfloat[$\tpo-L/2$]{
+			\includegraphics[width=45mm]{cherntpo.pdf}
+		}
+		\subfloat[$\tilde{\mathcal{P}}$ mod 1]{
+			\includegraphics[width=45mm]{cherntp.pdf}
+		}
+	}\hspace{0mm}
+
+	\makebox[0pt]{
+		\subfloat[$\tpo-L/2$]{
+			\includegraphics[width=45mm]{cherntpo4.pdf}
+		}
+		\subfloat[$\tilde{\mathcal{P}}$ mod 1]{
+			\includegraphics[width=45mm]{cherntp4.pdf}
+		}
+	}
+
+	\caption{Polarization $\tpo$ computed for $t=1,\kappa'=0$ and system size $L=160$, for the SSH chain $t'=0$ in (a) and for $t'=-2$ in (b) CHANGE}
+	\label{chern_insulator}
+\end{figure}
+
+To ilustrate the advantage of using $\tpo$ instead of $\tilde{\mathcal{P}}$ we also show the latter in the same parameter space.
+In order to obtain the Chern number there is no way of avoiding having to compute the derivative and subsequent integration of
+\begin{align}
+C(m) =& [\tpo(m,k_y=0^-)-\tilde{\mathcal{P}}(m,k_y=-\pi^+)]\\
+&+[\tpo(m,k_y=\pi^-)-\tpo(m,k_y=0^+)],
+\end{align}
+as we no longer know in advance where the branch cuts will appear, due to it being defined modulo 1.
+This example seems a bit trivial as the Chern number is also easy to identify from the plot of $\tilde{\mathcal{P}}$ as well but in systems with higher Chern numbers $\tilde{\mathcal{P}}$ will wind around the circle several times which will make this more difficult.
+\carlos{Figures 7.e and f are computed for the same system where instead of nearest-neighbor hopping in the x direction we use a 4th neighbor hopping ($\delta_{i,j+1} \rightarrow \delta_{i,j+4}$), to illustrate the challenges of computing the Chern number when the integrand is only defined modulo 1.
+This is similar to the issues we had when doing the cut-average.
+One needs to identify whenever $\tilde{\mathcal{P}}$ (or $\chi$) crosses 1 or 0 to compute the derivative and integral properly.
+
+Note also that we are computing the Chern number in the torus defined by $(\Phi,k_y)$.
+It seems a bit weird to mix both flux and momentum but it works in the end.
+If we were to introduce another flux in the other direction the system would no longer be effectively 1D so I don't know if everything we do would still apply in the same way or if it would provide any advantage.´ }
 

### 2020-08-31 version2/draft.tex
Add files via upload

@@
 It seems a bit weird to mix both flux and momentum but it works in the end.
-If we were to introduce another flux in the other direction the system would no longer be effectively 1D so I don't know if everything we do would still apply in the same way or if it would provide any advantage.´ }
+If we were to introduce another flux in the other direction the system would no longer be effectively 1D so I don't know if everything we do would still apply in the same way or if it would provide any advantage.
+}
 

### 2020-10-28 version2/draft.tex
Add files via upload

@@
 
-\title{On the relation of the entanglement spectrum to the Zak-phase}
+\title{On the relation of the entanglement spectrum to the bulk polarization \\ DRAFT}
 \author{Carlos Ortega Taberner}
@@
 \date{\today}
+\begin{abstract}
+The bulk polarization is a $\mathbb{Z}_2$ topological invariant for systems in one dimension with chiral or particle-hole symmetries.
+It is therefore related to the number of topological edge modes by the bulk-boundary correspondence.
+We show that the bulk polarization can always be decoded from the single-particle entanglement spectrum, even in the absence of symmetries that quantize it.
+From this relation the bulk-boundary correspondence can be derived in the presence of chiral and particle-hole symmetries.
+We use the bulk polarization to compute Chern numbers in 1D and 2D systems and show their relation to the entanglement spectrum.
+Furthermore we propose an alternative bulk polarization that can carry more information about the surface spectrum than the conventional one and can simplify the calculation of Chern numbers.
+\end{abstract}
 
@@
 As such, the entanglement entropy can only be used for interacting systems. 
-Its main drawback is the fact that it can only provide one of the pertinent quantum numbers that characterizes the system, and is, thus,  not able to uniquely determine the topologically ordered phase at hand.
+Its main drawback is the fact that it can only provide one of the pertinent quantum numbers that characterizes the system, and is, in general,  not able to uniquely determine the topologically ordered phase at hand.
 In addition, it requires a careful scaling analysis, which can be challenging for strongly correlated systems. 
@@
 It was conjectured to provide information about the edge spectrum, which was later shown to be a general feature of ground states with an effective topological quantum field theory description~\cite{Qi2012general}.
-It also proved to be an effective tool for fractional Chern insulators~\cite{Regnault2011fractional} and certain quantum spin liquids~\cite{yao2010entanglement}. \mariac{Others?}
+It also proved to be an effective tool for fractional Chern insulators~\cite{Regnault2011fractional} and certain quantum spin liquids~\cite{yao2010entanglement}. %\mariac{Others?}
 
 Generically, any of the proposed tools provides some information about the phase naturally, while others seem inaccessible. 
-It is not known, whether there exists a method to obtain the full topological data from the ground state. 
+It is not known, whether there exists a method to obtain the full topological content from the ground state. 
 For the simplest quantum Hall states, the Laughlin states,  it was shown that the ES does, in fact, contain the full information about the topological phase~\cite{hermanns2011haldane}.
@@
 
-Here, we address a much simpler question: what information can be extracted from the ES of \emph{non-interacting} systems?
-The latter can be computed very efficiently, using methods developed by Peschel and others~\cite{Peschel2003}. 
-It was subsequently shown for topological insulators/superconductors that the ES for (gapped) periodic systems is equivalent to the flat-band energy spectrum of the corresponding system with open boundaries~\cite{Fidkowski2010entanglement}.
+For \emph{non-interacting} topological insulators/superconductors the ES for (gapped) periodic systems can be computed very efficiently, using methods developed by Peschel and others~\cite{Peschel2003}.
+The ES in these systems is equivalent to the band-flattened energy spectrum of the corresponding system with open boundaries~\cite{Fidkowski2010entanglement}.
 The same correspondence was also found for closely related gapless systems~\cite{Matern2018entanglement}.
 However, even for non-interacting systems, it is unclear which information (beyond the  `edge' spectrum) is encoded in the entanglement spectrum.
-In this manuscript, we show that the information about the Zak phase, or polarization, can be completely recovered from the entanglement spectrum and provide a simple formula to do so. 
+In this manuscript, we show that the information about the Zak phase, or the bulk polarization, can be completely decoded from the entanglement spectrum and provide a simple formula to do so.
 We also show how to extend this method to compute the Chern number directly from the entanglement spectrum.
+In obtaining these topological invariants, the Zak phase and the Chern number, directly from the entanglement spectrum the \emph{bulk-boundary correspondence} is seen in an explicit way.
 
@@
 This demonstrates that there should be a direct relation between the single-particle entanglement energies and the Zak phase, although deriving that relation turned out to be non-trivial.
-An added benefit of our approach is that it can be generalized to systems that are not translationally invariant in a numerically efficient way \carlos{Does using the MBES imply this is no longer correct?}.
-A particular limit of our result for homogeneous systems~\eqref{eq:chi_integrable} was derived by Ryu and Hatsugai for a fully dimerized SSH chain~\cite{Ryu2006}. 
+A particular limit of our result for homogeneous (translationally invariant) systems~\eqref{eq:chi_integrable} was derived by Ryu and Hatsugai for a fully dimerized SSH chain~\cite{Ryu2006}.
 The similarity between the behavior of the Zak phase and the 'virtual edge modes' in the entanglement spectrum of certain Chern insulators, which was observed in Ref.~\cite{Huang2012,Huang2012-2}, can also be explained by our results.
-The polarization obtained via this method for infinite systems or systems with periodic boundary conditions is only well-defined modulo 1, i.e. in the range $[0,1]$. 
-One can, however, compute it in the equivalent open chain to obtain a polarization defined on the full real axis.
-%\carlos{Not sure about this expression.
-I mean it's not defined on a circle, not that is not a complex number, which it trivially isnt.}.
-We use the latter method to compute Chern numbers, which can also serve as topological invariants in one-dimensional systems.
-For the system studied, the Chern number is equivalent to the winding number, but it is unclear whether this result continues to hold for generic systems.
+The bulk polarization is defined in $\mathbb{R}$ mod 1, which presents challenges when computing changes in the polarization, which are defined  in $\mathbb{R}$.
+We show how, using our method, one can define an alternative bulk polarization computed in the equivalent open chain that is defined in $\mathbb{R}$.
+The bulk polarization obtained this way can give more information about the edge spectrum than the usual one, and can be used to simplify the computation of Chern numbers.
+We also discuss the relation between this alternative bulk polarization and another topological invariant known as the trace index \cite{Alexandrinata2011}.
 
@@
 In section V we introduce the method and show how it reproduces the Zak phase for both systems with and without translational invariance.
-In section VI we explain how the method applied to the opened chain allows us to compute Chern numbers %and winding numbers
+In section VI we explain how the method applied to the opened chain allows us to compute Chern numbers for systems both in 1D and 2D.%and winding numbers
 and finally in section VII we present the conclusions and possible outlook.
@@
 \section{Model}
-To showcase our results we consider as a starting point the model in the BDI class used in reference~\cite{Song2014}. 
-This is a very simple model which nevertheless supports a rich phase diagram containing topological phases with winding numbers up to $\nu =2$. 
+To showcase our results we consider as a starting point a model in the BDI class~\cite{Song2014}. 
+This is a simple model which nevertheless supports a rich phase diagram containing topological phases with winding numbers up to $\nu =2$. 
 In order to show that our results are not limited to any particular symmetry class~\cite{ryu2010topological}, we also include two symmetry breaking terms. 
@@
 \end{alignat}
-Using the results of Ref.~\cite{Song2014}, the corresponding phase diagram can be calculated and is shown in Fig.\ref{fig:bdi_phase_diagram} for $t=1$. 
+The corresponding phase diagram \cite{Song2014} is shown in Fig.\ref{fig:bdi_phase_diagram} for $t=1$. 
 \begin{figure}[t]
 	\centering
-	\includegraphics[width=45mm]{phase_diagram2.pdf}
-	\caption{Phase diagram for the system in the $BDI$ class showing the winding number of each region. Plotted for parameters $t = 1,\kappa =\kappa'=0.$}
+	\includegraphics[width=66mm]{fig1.pdf}
+	\caption{Phase diagram for the system in the $BDI$ class showing the winding number of each region.
+Plotted for parameters $t = 1,\kappa =\kappa'=0$.
+Three cuts of the phase diagram are shown with continuous, dashed and dotted lines which will be used in the rest of the figures.}
 \label{fig:bdi_phase_diagram}
@@
 The BDI class  is characterized by a $\mathbb{Z}$ invariant, the winding number,  in one spatial dimension.
-When imposing open boundary conditions on the system, the number of symmetry-protected zero energy edge modes is equal to the winding number. 
+When open boundary conditions are imposed on the system, the number of symmetry-protected zero energy edge modes is equal to the winding number. 
 The behavior of the  edge modes  is  shown in Fig.~\ref{fig:zero_E_modes}, using the dotted path ($t=1,t'=-2$) marked in the phase diagram. 
-Subfigure (a) shows the edge spectrum in the BDI class, with 4 (resp. 2) symmetry-protected zero modes for winding number 2 (1). 
+Figure~\ref{fig:zero_E_modes}.a shows the edge spectrum in the BDI class, with 4 (resp. 2) symmetry-protected zero modes for winding number 2 (1). 
 %We also a cut through the three phases (dotted line in the phase diagram) for different values of $\kappa,\kappa'$.
@@
 Consequently, adding such a term causes the zero modes in the $\nu=2$ phase to split  pairwise and the $\nu = 2$ and $\nu=0$ phases become equivalent. 
-This is shown in subfigure (b), where one clearly sees that the zero modes below $m=1$ are split.
+This is shown in figure~\ref{fig:zero_E_modes}.b, where one clearly sees that the zero modes below $m=1$ are split.
 
@@
 Also this symmetry class is topologically trivial in one dimension. 
-Subfigures (c) and (d) show the edge spectrum for class AI and A, respectively. In both cases, the edge modes are split from zero for all values of $m$.
-
-%
-%In subfigure (b) and (d), where the system is in a topologically trivial phase, all edge modes are gapped. 
-%Subfigure (c) shows the edge spectrum of the system being in class D, where the $\nu=2 $ phase becomes trivial. 
-%Consequently, symmetry-protected zero modes only exist for $1<m<2.5$\mariac{?} (corresponding to $\nu=1$ for $\kappa'=0$) , whereas the modes are split for $m<1$ (corresponding to $\nu=2$ for $\kappa'=0$).
-
+Figure~\ref{fig:zero_E_modes}.c and d show the edge spectrum for class AI and A, respectively. In both cases, the edge modes are split from zero for all values of $m$.
+
+\iffalse
 \begin{figure}[h!]
 \centering
-%\makebox[0pt]{
-%\subfloat[$t = 1,\kappa =\kappa'=0$]{
-%  \includegraphics[width=45mm]{phase_diagram2.pdf}
-%}}\hspace{0mm}
-
 \makebox[0pt]{
 \subfloat[$\kappa =\kappa'=0$]{
-  \includegraphics[width=35mm]{2a.pdf}
-}
-\subfloat[$\kappa = 0, \kappa'=0.3$]{\includegraphics[width=35mm]{2b.pdf}
+  \includegraphics[width=43mm]{fig21.pdf}
+}
+\subfloat[$\kappa = 0, \kappa'=0.3$]{\includegraphics[width=43mm]{fig22.pdf}
 }
@@
 \subfloat[$\kappa = 0.3,\kappa'=0$]{
-  \includegraphics[width=35mm]{2c.pdf}
-}
-\subfloat[$\kappa = 0.3,\kappa'=0.3$]{\includegraphics[width=35mm]{2d.pdf}
-}
-}
-\caption{\carlos{Use another colour for the blue edge mode.}Edge spectrum along the dotted path (i.e. $t=1,t'=-2$) of Fig.~\ref{fig:bdi_phase_diagram} for different values of $\kappa$ and $\kappa'$.}
+  \includegraphics[width=43mm]{fig23.pdf}
+}
+\subfloat[$\kappa = 0.3,\kappa'=0.3$]{\includegraphics[width=43mm]{fig24.pdf}
+}
+}
+\caption{Edge spectrum along the dotted path (i.e. $t=1,t'=-2$) of Fig.~\ref{fig:bdi_phase_diagram} for different values of $\kappa$ and $\kappa'$.}
 \label{fig:zero_E_modes}
 \end{figure}
+\fi
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=86mm]{fig2comp.pdf}
+\caption{Energy spectrum with open boundary conditions along the dotted path (i.e.
+$t=1,t'=-2$) of Fig.~\ref{fig:bdi_phase_diagram} plotted for different values of the symmetry-breaking terms (a) $\kappa =\kappa'=0$ , (b) $\kappa = 0, \kappa'=0.3$, (c) $\kappa = 0.3,\kappa'=0$ and (d) $\kappa = 0.3,\kappa'=0.3$}
+\label{fig:zero_E_modes}
+\end{figure}
+
+In Table \ref{fig:class_table} we summarize the different classes that are hosted by this model together with the correspondent symmetries and topological invariants in 1D.
+
+\begin{table}[h]
+\centering % centering table
+\begin{tabular}{|c |rrr| c|c|} % creating eight columns
+\hline\hline %inserting double-line
+Model &  $T$ & $C$ & $S$ & Class & inv.  \\  [1ex] 
+\hline
+$\kappa =0,\kappa'=0$ &  1 & 1 & 1 & BDI  & $\mathbb{Z}$\\  [1ex] 
+$\kappa =0,\kappa' \neq0$ &  0 & 1 & 0 & D  & $\mathbb{Z}_2$ \\  [1ex] 
+$\kappa \neq 0,\kappa'=0$ &  1 & 0 & 0 & AI & 0 \\  [1ex] 
+$\kappa \neq 0,\kappa' \neq 0$ &  0 & 0 & 0 & A & 0  \\  [1ex] 
+\hline \hline%title of the table
+\end{tabular}
+\caption{Different Cartan classes hosted by the model in Eq.\ref{bdi_model}, with the correspondent symmetries and topological invariants in 1D.} 
+\label{fig:class_table}
+\end{table}
 
@@
 where $[U]_{j\beta,p\mu} = u_{p\mu}^{j\beta}$ is the unitary matrix that diagonalizes $H$.
-In translational invariant systems $u_{p\mu}^{j\beta} = e^{ipj}\tilde{u}_{p \mu}^{\beta}$, where $\tilde{u}_{p \mu}^{\beta}$ are the components of the eigenstates of the Bloch Hamiltonian, $\ket{\tilde{u}_{k\mu}}$.
+In translational invariant systems $u_{k\mu}^{j\beta} = e^{ikj}\tilde{u}_{k \mu}^{\beta}$, where $\tilde{u}_{k \mu}^{\beta}$ are the components of the eigenstates of the Bloch Hamiltonian, $\ket{\tilde{u}_{k\mu}}$ and $k$ is the momentum.
 One can now compute the Zak phase \cite{Zak1989}, which for a two-band  model is defined as the geometric phase acquired by the occupied state $\ket{\tilde{u}_k}$ as it winds around the Brillouin zone,
 \begin{equation}
-\gamma = \int_{0}^{2\pi} dk\, i A_k, 
+\gamma = \int_{0}^{2\pi} {\rm d}k\, i A_k, 
 \label{eq:zak_phase}
@@
 \begin{equation}
-\Delta {\gamma_{\lambda_i \lambda_f}} = \int_{\lambda_i}^{\lambda_f}\int_{0}^{2\pi} dk \, \Omega_{\lambda k},
+\Delta {\gamma_{\lambda_i \lambda_f}} = \int_{\lambda_i}^{\lambda_f} d\lambda\int_{0}^{2\pi} dk \, \Omega_{\lambda k},
 \label{eq:change_zak_phase}
@@
 \end{align}
- is the Berry curvature.
-The Berry curvature is fully gauge invariant and, therefore, the change in the Zak phase \eqref{eq:change_zak_phase} is defined along the full real line.
+ is the Berry curvature. The Berry curvature is fully gauge invariant and, therefore, the change in the Zak phase in Eq.\eqref{eq:change_zak_phase} is defined in $\mathbb{R}$. %?
 The Zak phase itself, as defined in Eq.~\eqref{eq:zak_phase}, can be shown to be proportional to the winding number for a certain gauge (see Appendix \ref{appendix:winding_ssh}), which means that it carries physical information beyond modulo $2\pi$.
@@
 
-The derivative of the polarization with respect to some parameter $\lambda$ was obtained \cite{KingSmith1993} as
+The derivative of the polarization with respect to $\lambda$ was obtained \cite{KingSmith1993} as
 \begin{equation}\label{eq:Bloch_polarization}
@@
 As discussed above,  this polarization is defined only modulo 1.
-Its changes, however, can be defined as a real number. 
-This is not surprising,  as the derivative of the polarization is related to the current which is a physical observable, unlike $\tilde{\mathcal{P}}^{\rm Bloch}$.
+Its changes, however, are defined in $\mathbb{R}$. The polarization itself is not a physical observable. Only its changes, related to the current flowing through the system, are.
 
@@
 \begin{equation}
-\mathcal{P} = \int_0^{2\pi}\frac{d\Phi}{2\pi} i\bra{\Psi^\Phi}\partial_\Phi \ket{\Psi^\Phi} + \Im\ln \bra{\Psi^0}e^{2\pi i \hat{P}} \ket{\Psi^{2\pi}}, 
-\end{equation}
-where $\ket{\Psi^{\Phi}}$ is the ground state in the presence of flux and $\hat{P}=\frac{1}{L}\sum_{j\alpha} j\hat{n}_{j\alpha}$ is the polarization operator.
+\mathcal{P} = \int_0^{2\pi}\frac{d\Phi}{2\pi} i\bra{\Psi^\Phi}\partial_\Phi \ket{\Psi^\Phi} + \frac{1}{2\pi}\Im\ln \bra{\Psi^0}e^{2\pi i \hat{P}} \ket{\Psi^{2\pi}}, 
+\end{equation}
+where $\ket{\Psi^{\Phi}}$ is the ground state in the presence of flux and $\hat{P}=\frac{1}{L}\sum_{j\alpha} j\hat{n}_{j\alpha}$ is the polarization operator, where $\hat{n}_{j\alpha}$ is the number operator.
 This is equivalent to the polarization obtained by Resta \cite{Resta1997} as
-\maria{should I interpret the following line as $\Phi=0$?}
-\begin{align}
-\mathcal{P} = \frac{1}{2\pi} \rm{ Im }\,\ln\bra{\Psi}e^{2\pi i \hat{P}}\ket{\Psi}.
+\begin{align}
+\mathcal{P} = \frac{1}{2\pi} \Im \,\ln\bra{\Psi^{\Phi=0}}e^{2\pi i \hat{P}}\ket{\Psi^{\Phi=0}}.
 \end{align}
@@
 \end{equation}
-\carlos{I have to check overall sign and sign of exponential.}and  ${\rm det }\!' $ indicates that the determinant is restricted to the space of \emph{occupied} single-particle states.
-Below we will employ this expression to compare our results with \mariac{XXX}.
+and  ${\rm det }\!' $ indicates that the determinant is restricted to the space of \emph{occupied} single-particle states.
+Below we will compare the bulk polarization obtained using our method to the one obtained using equation~\eqref{eq:polarization_resta}.
 
 There is one last polarization we must consider. 
-It is obtained when the flux is introduced in the bond between sites $i=1$ and $i=L$ via twisted boundary conditions. This is equivalent to performing the gauge transformation
+It is obtained when the flux is introduced in the bond between sites $j=1$ and $j=L$ via twisted boundary conditions. This is equivalent to performing the gauge transformation
 \begin{align}\label{eq:Psi_tilde}
@@
 \end{align}
-which makes $\ket{\tilde{\Psi}_0^\Phi}$ fully periodic in $\Phi$. We can now define the polarization
+which makes $\ket{\tilde{\Psi}^\Phi}$ fully periodic in $\Phi$. We can now define the polarization
 \begin{equation}
@@
 \end{equation}
-which is related to the current flowing through the bond \cite{Watanabe2018}. This polarization is related to the homogeneous polarization by 
+which is related to the current flowing through this bond \cite{Watanabe2018}. This polarization is related to the homogeneous polarization by 
 \begin{equation}
@@
 %The previous discussion about the Zak phase being defined only modulo $2\pi$ although its changes can be defined as a real number applies in the same way to the polarization obtained in \eqref{eq:p_tilde}.
-Note that while $\tilde{\mathcal{P}}$ itself is only defined modulo 1, its changes are again well-defined on the full real axis, similar to the discussion about the Bloch polarization below Eq.~\eqref{eq:Bloch_polarization}.
-In fact, for translationally invariant systems one finds that $\tilde{\mathcal{P}} = \tilde{\mathcal{P}}^{\rm Bloch}$ \cite{Watanabe2018}.
+Similar to the discussion about the Bloch polarization below Eq.~\eqref{eq:Bloch_polarization}, both $\mathcal{P}$ and $\tilde{\mathcal{P}}$ are defined modulo 1, while their changes are defiend in $\mathbb{R}$.
+For translationally invariant systems one finds that $\tilde{\mathcal{P}} = \tilde{\mathcal{P}}^{\rm Bloch}$ \cite{Watanabe2018}.
 
@@
 \label{section:ES}
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=85mm]{chain.pdf}
+	\caption{Schematic of the chain with periodic boundary conditions considered, where the different sites are labeled by their position.
+The bipartition into regions $A$ and $B$ is shown.
+Other subregions of $A$ considered in the text are also shown.
+}
+\label{fig:chain}
+\end{figure}
+
 Let us now proceed to discuss the entanglement spectrum. 
@@
 \begin{align}\label{eq:corr_mat2}
-C_{ij}^{\alpha \beta} =& \frac{1}{2}\left[I - H/ (H^2)^{-1/2} \right]_{ij, \alpha \beta}.
+C_{ij}^{\alpha \beta} =& \frac{1}{2}\left[I - H/ (H^2)^{1/2} \right]_{ij, \alpha \beta}.
 \end{align}
@@
 Following Ref.s~\cite{Huang2012,Huang2012-2}, we refer to this spectrum as the entanglement occupancy spectrum (EOS).
-The resulting eigenvalues  are denoted by $\xi_j$. 
-%We will differentiate between two types of eigenvalues: those (exponentially) close to $\xi_j=0 (1)$ correspond to ``bulk'' states, by which we mean that the corresponding eigenstates lie (up to exponentially small corrections) in  $A$ ($B$).
-%Values $0<\xi_{l,r}<1$ indicate states that are located at the virtual boundaries -- labeled `left' and `right' in the following-- between $A$ and $B$~\cite{Peschel2008}. 
-%There is, of course, an ambiguity here on dividing eigenstates of the EOS into ``bulk'' and ``edge'' states and `how close' the eigenvalue has to be to $0,1$ in order to qualify as a bulk state. 
-%XXXX
+We denote the eigenvalues of the EOS by $\xi_j$. 
 Topological phases are characterized by (symmetry-protected) zero-energy modes in the edge spectrum and $\xi=\frac 1 2$ modes in the EOS~\cite{Fidkowski2010entanglement}.
@@
 the corresponding eigenstates are the same. 
-Eq.~\eqref{eq:xi_eps} implies that the full information of the \emph{many-particle} ES is contained in the single-particle spectrum of the reduced correlation matrix. 
+Eq.~\eqref{eq:xi_eps} implies that the full information of the \emph{many-body} ES (MBES) is contained in the single-particle spectrum of the reduced correlation matrix. 
 The latter is much simpler to interpret. 
@@
 
-One property of the subsystem correlation matrix we will make use of later is that, for sufficiently large systems, its eigenstates are found to be localized on either virtual edge if $\xi \neq 0,1$, or they are found to be bulk modes if the correspondent eigenvalues are exponentially close to $\xi = 0,1$ \cite{Peschel2008}.
+One property of the subsystem correlation matrix we will make use of later is that, for sufficiently large systems, its eigenstates are found to be localized on either virtual edge if $\xi$ is away from $0$ and $1$, or they are found to be bulk modes if the correspondent eigenvalues are exponentially close to $\xi = 0,1$ \cite{Peschel2008}.
 
@@
 We also provide a simple formula in terms of the EOS eigenvalues that is identical to the Zak phase in the thermodyamic limit.  
-However, for a generic, finite size system, there will be finite-size discrepancies between our formula and the Zak phase. \mariac{see scaling plot}
+However, for a generic, finite size system, there will be finite-size discrepancies between our formula and the Zak phase, see figure \ref{huang}.d. 
 In order to reduce these, we are going to choose subsystem $A$ as half the system in the remainder of the manuscript.
@@
 
-Let us first review previous results on the relation between the entanglement spectrum and the Zak phase. 
-For symmetry-protected topological systems in one dimension, the quantized Zak phase is zero whenever there is an even number  of eigenvalues per edge at $\xi = 1/2$, and $\pi$ when the number is odd~\cite{Peschel2008}.
+Let us first review previous results on the relation between the entanglement occupancy spectrum and the Zak phase. 
+For symmetry-protected topological systems in one dimension, the quantized Zak phase is zero whenever there is an even number  of eigenvalues per vitual edge at $\xi = 1/2$, and $\pi$ when the number is odd~\cite{Peschel2008}.
 In the absence of symmetry-protection, much less is known. 
 Ryu and Hatsugai considered the fully dimerized SSH chain with broken chiral symmetry~\cite{Ryu2006}.  
-This model is rather special in that there is only a single pair of eigenvalues, related by symmetry \mariac{which symmetry?}, in the EOS that is not strictly identical to $0$ or $1$. 
+This model is rather special in that there is only a single pair of eigenvalues in the EOS that is not strictly identical to $0$ or $1$, they are $\xi$ and $1-\xi$, related due to translational symmetry.
 The authors could show that the value of one of these `midgap states' is identical to the Zak phase divided by $2\pi$ or, alternatively, $\tilde{\mathcal{P}}^{\rm Bloch}$.
+Although they do not specify to which of the two eigenvalues it corresponds.
 Their result is a particular limit of Eq.~\eqref{eq:chi_integrable}. 
@@
 
-Huang and Arovas considered a two-dimensional Chern insulator and noted that the Zak phase qualitatively follows the `virtual edge mode' that connects the EOS values at 0 to those at 1~\cite{Huang2012,Huang2012-2}.
+Huang and Arovas considered a two-dimensional Chern insulator and noted that there was a similarity between the Wannier centers (and therefore the Zak phase) and the `virtual edge mode' that connects the EOS values at 0 to those at 1~\cite{Huang2012,Huang2012-2}.
 Interpreting this as a one-dimensional system with a parameter, we can explain this behavior by Eq.~\eqref{eq:chi_integrable}, noting that there is a single pair of eigenvalues that dominates the sum.
@@
 In the latter case, there are several terms in Eq.~\eqref{eq:chi_integrable} with comparably large contributions and, consequently, the Zak phase deviates considerably from the 'virtual edge modes'.
-%For cases with more edge modes present in the ES it has been shown that the Zak phase follows the midgap eigenvalues qualitatively \cite{Huang2012,Huang2012-2}.
-
-\subsubsection{Systems with equidistant entanglement energies}
-\mariac{add scaling plot for $\chi$ versus $\gamma/(2\pi)$. }
-We first discuss systems, for which the entanglement energies of a \emph{single} virtual edge are equidistant.
+% This last sentece is not entirely ok.
+They observe flow in the zak phase that correspond s to flow in the EOS.
+
+\subsubsection{Systems with equispaced entanglement energies}
+
+We first discuss systems for which the entanglement energies of a \emph{single} virtual edge are equispaced.
+That is, they are given in the thermodynamic limit by $\varepsilon_{n\alpha} = \varepsilon_{0\alpha}+n\delta_\alpha$, where $\alpha = L,R$.
 This is a feature of integrable systems \cite{Peschel1999} for e.g.
-nearest neighbor hopping models such as the SSH chain, and applies to the Hamiltonian~\eqref{bdi_model} when $t'=0$.
+nearest neighbor hopping models such as the SSH chain, and applies to the Hamiltonian in Eq.~\eqref{bdi_model} when $t'=0$.
 In this case, one can obtain the Zak phase in a very simple fashion: We reorder the eigenvalues of the EOS by magnitude, $\xi_1 < \xi_2 < ...< \xi_{L_AM}$, and compute
 \begin{equation}
-\chi = \sum_{j=1}^{L_A M/2} \xi_{2j-1} \, {\rm mod} \, 1.
+\chi = \sum_{j=1}^{L_A M} \xi_{2j-1} \, {\rm mod} \, 1,
 \label{eq:chi_integrable}
 \end{equation}
-In the thermodynamic limit, $\chi$ becomes identical to either $\tilde{\mathcal{P}}^{\rm Bloch}$ or $1-\tilde{\mathcal{P}}^{\rm Bloch}$: \mariac{comment on the ambiguity?}
-\begin{equation}
-\lim_{L \rightarrow \infty} \abs{\tilde{\mathcal{P}}^{\rm Bloch}} = \lim_{L \rightarrow \infty} \abs{\chi}.
-\end{equation}
-\mariac{I don't understand what we want to say here.} For translational invariant systems the EOS is symmetric such that a crossing of the eigenvalues $\xi$ results in $\chi$ changing sign.
-This is particularly irrelevant in the case of symmetry-protected states where the Zak phase is quantized to $0$ or $\pi$.
-In the special case where there is only one eigenvalue per edge that is $\neq 0,1$ in  the EOS, like in the fully dimerized limit of the SSH chain, we recover the equality between the midgap eigenvalue and the Zak phase found in Ref.~\cite{Ryu2006}.
+where $L_A$ is the length of subsystem $A$ and $M$ is the number of orbitals per site.
+In the thermodynamic limit, $\chi$ becomes identical to either $\tilde{\mathcal{P}}^{\rm Bloch}$ or $1-\tilde{\mathcal{P}}^{\rm Bloch}$,
+\begin{equation}
+\lim_{L \rightarrow \infty} \abs{\tilde{\mathcal{P}}^{\rm Bloch}} = \lim_{L \rightarrow \infty} \chi \, {\rm mod} \, 1.
+\end{equation}
+This generalizes the observation made by Ryu and Hatsugai \cite{Ryu2006} beyond the fully dimerized limit, $m=0$, but the ambiguity remains. This issue will be resolved in the next section. 
+For translational invariant systems the EOS is symmetric such that a crossing of the eigenvalues $\xi$ results in $\chi$ changing sign.
+This is irrelevant in the case of symmetry-protected states where the Zak phase is quantized to $0$ or $\pi$.
+In the special case where there is only one eigenvalue per edge that is $\neq 0,1$ in  the EOS, like in the fully dimerized limit of the SSH chain, we recover the equality between one of the two midgap eigenvalues and the Zak phase found in Ref.~\cite{Ryu2006}.
 
@@
 Along this line, our model is equivalent to an SSH chain. 
-In the presence of chiral and translation symmetry, all eigenvalues of the EOS are doubly degenerate --- one eigenvalue from each edge, such that $\chi$ from \eqref{eq:chi_integrable} will pick up one eigenvalue from each pair.
+In the presence of chiral and translation symmetry, all eigenvalues of the EOS are doubly degenerate --- one eigenvalue from each virtual edge, such that $\chi$ from Eq.\eqref{eq:chi_integrable} will pick up one eigenvalue from each pair.
 Because of chiral symmetry the sum of the two eigenvalues related by the symmetry gives always $1$ so they do not contribute to $\chi$.
-The only contribution to $\chi$ is the one from the eigenstates with $\xi=1/2$ that do not have a chiral partner, see Fig.~\ref{huang}(a).
+The only contribution to $\chi$ is the one from the eigenstates with $\xi=1/2$ that do not have a chiral partner, see figure~\ref{huang}.a.
 Here we see explicitely the known relation between the Zak phase and the number of $\xi = 1/2$ modes.
 
-Note that the full sum~\eqref{eq:chi_integrable}  evaluates to zero in the trivial regime %up to finite-size corrections% \carlos{It is actually exact due to symmetry}.
+Note that the full sum  in Eq.~\eqref{eq:chi_integrable}  evaluates to zero in the trivial regime %up to finite-size corrections% \carlos{It is actually exact due to symmetry}.
 despite the presence of modes with entanglement occupancy $0<\xi<1$.
@@
 When breaking chiral symmetry, the EOS will no longer be doubly degenerate. 
-For small $m$, there are two dominant midgap eigenvalues, and the Zak phase follows the lower one very closely, see Fig.~\ref{huang}(b).
-When this midgap eigenvalue approaches  0, the contribution from the other modes becomes more relevant and the Zak phase starts to deviate from the midgap eigenvalue, \mariac{see inset}.
+For small $m$, there are two dominant midgap eigenvalues, and the Zak phase follows the lower one very closely, see figure~\ref{huang}.b.
+When this midgap eigenvalue approaches  0, the contribution from the other modes becomes more relevant and the Zak phase starts to deviate from the midgap eigenvalue, see figure \ref{huang}.c.
+
 \begin{figure}[t]
 \centering
-\makebox[0pt]{
-\subfloat[$t = 1,t'=0$, \, \,$\kappa =\kappa'=0$]{
-  \includegraphics[width=35mm]{2_a.pdf}
-}
-\subfloat[$t=1,t'=0$, $\kappa = 0.3, \kappa'=0$]{\includegraphics[width=35mm]{2_b.pdf}
-}
-}
-\caption{Entanglement spectrum (black), $\chi$ (green) and $\frac{\gamma}{2\pi}$ (red) for a cut in the phase diagram through the $\nu = 1 \rightarrow \nu = 0$ transition (dashed line in~\ref{fig:bdi_phase_diagram}
-	) for $t'=0$ and different values of $\kappa$. Computed for $L=40$.}
+\includegraphics[width=86mm]{fig3comp.pdf}
+\caption{In (a) and (b)  we show the EOS (black), $\chi$ (green) and $\tilde{\mathcal{P}}^{\rm Bloch}$(red) for a cut in the phase diagram through the $\nu = 1 \rightarrow \nu = 0$ transition (dashed line in~\ref{fig:bdi_phase_diagram}), $t = 1,t'=0$, for parameters  (a) $\kappa =\kappa'=0$ and (b) $\kappa = 0.3, \kappa'=0$.
+Computed for $L=40$.
+In (c) we zoom in a region of the plot in (b) to show the splitting between $\tilde{\mathcal{P}}^{\rm Bloch}$ and the eigenvalue closest to $\xi=1/2$.
+In (d) we show a scaling plot of the difference between $\chi$ and $\tilde{\mathcal{P}}^{\rm Bloch}$ for parameters $t=1,t'=0,m=1$, $\kappa = 0.3, \kappa'=0$ and increasing system size to show that the difference vanishes in the thermodynamic limit.
+}
 \label{huang}
@@
 The inclusion of bulk modes is irrelevant since they have $\xi = 0,1$.
-We can then compute the average position of each eigenstate and construct $\chi$ from the eigenstates with $\expval{\hat{x}}<L/4$, with $L/4$ being the middle point of region $A$,
-\begin{equation}
-\chi = \sum_{\mu \in A_L} \xi_\mu \, {\rm mod} \, 1. 
+We can then compute
+\begin{equation}
+\chi = \sum_{\mu \in L} \xi_\mu \, {\rm mod} \, 1. 
 \label{eq:chi_left}
 \end{equation}
+the sum is performed over the subspace $L$ formed by the eigenstates with $\expval{\hat{x}}<L/4$, where $L/4$ is the middle point of region $A$.
+It does not matter how we choose the eigenvalues we sum over as long as we do not include the ones related to the right edge with $\xi$ away from $1$.
+In practice only a few eigenvalues give a significant contribution to $\chi$.
+
 In the thermodynamic limit, it is $\tilde{\mathcal{P}}$ that we actually identify with $\chi$
@@
 Before proceeding lets review again the result in the previous section.
-First note that for translation invariant systems, the ES of the right and the left edge are related by an overall '-' sign, i.e.
-for each eigenvalue $\epsilon$ on the right edge, there is a corresponding value $-\epsilon$ on the left edge.
-Since the spectra are equidistant, this implies that the two edge spectra are merely shifted with respect to each other.
+First note that for translation invariant systems, the ES of the right and the left virtual edge are related by an overall '-' sign, i.e.
+for each eigenvalue $\epsilon$ on the right virutal edge, there is a corresponding value $-\varepsilon$ on the left one.
+Since the spectra is equispaced, this implies that the two virtual edge spectra are merely shifted with respect to each other.
 If the shift is zero, e.g.
@@
 This issue can be fixed by determining the localization of a single eigenstate.
+MinMax{]
+
 \begin{figure}[t]
 \centering
-	\makebox[0pt]{
-		\subfloat[$t = 1, t' = 2,$   $\kappa =0.3, \kappa' = 0$]{
-			\includegraphics[width=35mm]{3_a.pdf}
-		}
-		\subfloat[$t = 1, t' = 2,$ $\kappa = 0, \kappa' = 0.3$]{
-			\includegraphics[width=35mm]{3_b.pdf}
-		}
-	}\hspace{0mm}
-
-	\makebox[0pt]{
-		\subfloat[$t = 1, t' = -2,$ $\kappa = 0.3, \kappa' = 0$]{
-			\includegraphics[width=35mm]{3_c.pdf}
-		}
-		\subfloat[$t = 1, t' = -2,$ $\kappa = 0, \kappa' = 0.3$]{
-			\includegraphics[width=35mm]{3_d.pdf}
-		}
-	}\hspace{0mm}
-
-	\makebox[0pt]{
-		\subfloat[$t = 1, t' = -2,$ $\kappa_i \neq 0 , \kappa' = 0$]{
-			\includegraphics[width=35mm]{3_e.pdf}
-		}
-	}
-	\caption{(a,b,c,d) Entanglement spectrum (black), $\chi$ (green) and $\tilde{\mathcal{P}}$ (red) for two different cuts with both $\kappa$ and $\kappa'$ symmetry breaking terms.
-Note that in this case where there is translational invariance $\tilde{\mathcal{P}}=\tilde{\mathcal{P}}^{\rm Bloch}$, which is nothing else than the Zak phase over $2\pi$.
-(e) Same as (c) but with a position dependent $\kappa_i=(i+L/4 \, {\rm mod}\, L)/L$.
-We also plot $\mathcal{P}+1/2$ (blue) to showcase that $\chi$ is indeed equal to $\tilde{\mathcal{P}}$ and not $\mathcal{P}$ as in the translationally invariant case $\tilde{\mathcal{P}}=\mathcal{P}+(L+1)/2$.
+\includegraphics[width=86mm]{fig4comp.pdf}
+\caption{We show the Entanglement spectrum (black), $\chi$ (green) and $\tilde{\mathcal{P}}$ (red) computed for $L=40$ sites.
+(a) and (b) for the continuous line in the phase diagram, $t=1,t'=2$  with the symmetry-breaking parameters (a) $\kappa=0.3,\kappa'=0$ and (b) $\kappa=0,\kappa'=0.3$.
+(c) and (d) for the dotted line in the phase diagram, $t=1,t'=-2$  with the symmetry-breaking parameters (c) $\kappa=0.3,\kappa'=0$ and (d) $\kappa=0,\kappa'=0.3$.
+(d) is the same as (c) but with a position dependent $\kappa_i=(i+L/4 \, {\rm mod}\, L)/L$.
+We also plot $\mathcal{P}+1/2$ (blue) to showcase that $\chi$ is indeed equal to $\tilde{\mathcal{P}}$ and not $\mathcal{P}$.
 }
@@
 In Fig.\ref{2} we show two examples that highlight the importance of the localization structure of the EOS. 
-In subfigures (a) and (b), we compute the EOS for varying $m$ at  $t=1$, $t'=2$, indicated by the solid line  in the phase-diagram in Fig.~\ref{fig:bdi_phase_diagram}, in the presence of two different symmetry breaking terms: (i) $\kappa =0.3$, $\kappa'=0$ and (ii) $\kappa=0$, $\kappa'=0.3$.
+In figures \ref{2}.a and \ref{2}.b, we compute the EOS for varying $m$ at  $t=1$, $t'=2$, indicated by the solid line  in the phase-diagram in Fig.~\ref{fig:bdi_phase_diagram}, in the presence of two different symmetry breaking terms: (a) $\kappa =0.3$, $\kappa'=0$ and (b) $\kappa=0$, $\kappa'=0.3$.
 For $\kappa=\kappa'=0$, there is a phase transition $\nu = 2 \rightarrow 0 $ at $m=2$, whereas both choices of symmetry breaking render the system trivial  for all $m$.
@@
 While $\kappa'$ splits the four-fold degenerate states into two two-fold degenerate pairs at opposite edges whose contribution cancels (modulo 1), $\kappa$ split them into two $left-left$ and $right-right$ pairs, with a finite contribution to the Zak phase.
-In subfigures (c) and (d), we consider the same symmetry breaking terms, but now for $t=1$ and $t'=-2$, where the BDI systems shows a phase transition  $\nu=2\rightarrow 1\rightarrow 0$. 
-Note that the $\kappa'$ term does not render the $\nu=1$ phase trivial, and the Zak phase will remain quantized in that region, see Fig.~\ref{2}(d). 
+In figures \ref{2}.c and d, we consider the same symmetry breaking terms, but now for $t=1$ and $t'=-2$, where the BDI systems shows a phase transition  $\nu=2\rightarrow 1\rightarrow 0$. 
+Note that the $\kappa'$ term does not render the $\nu=1$ phase trivial, and the Zak phase will remain quantized in that region, see Fig.~\ref{2}.d. 
 Also here, Eq.~\ref{eq:chi_left} correctly reproduces the Zak phase for all values of $m$. 
@@
 
-Since $\chi$ is naturally defined in position space it is also valid for systems without translationally invariance.
-We show this in Fig.\ref{2}(e) where we show the EOS, $\chi$, $\mathcal{P}$ and $\tilde{\mathcal{P}}$ along the dotted cut in the phase diagram with an added position dependent symmetry breaking term $\kappa_i = (i+L/4 \, {\rm mod}\, L)/L$ (\carlos{I added this profile because with disorder $\tilde{\mathcal{P}}$ and $\mathcal{P}$ are difficult to distinguish}).
+Since $\chi$ is naturally defined in position space our results are also valid for systems without translationally invariance.
+We show this in Fig.\ref{2}.e where we show the EOS, $\chi$, $\mathcal{P}$ and $\tilde{\mathcal{P}}$ along the dotted cut in the phase diagram with an added position dependent symmetry breaking term $\kappa_i = (i+L/4 \, {\rm mod}\, L)/L$.
+We use this profile for $\kappa_i$ instead of simple disorder because with the disorder the difference between $\mathcal{P}$ and $\tilde{\mathcal{P}}$ is very small.
 We obtain that $\chi$ is indeed identified with $\tilde{\mathcal{P}}$ as we hinted at before, and not $\mathcal{P}$, although these two polarization only differ by $(L+1)/2$ in the translationally invariant case.
@@
 This relation between the Zak phase and the EOS is a consequence of an identity between the Zak phase and the many-body entanglement spectrum previously found for infinite chains \cite{Zaletel2014}.
-In Appendix A we modified this derivation to account for the periodic boundary conditions we use and show how their result simplifies greatly when expressing it in terms of the EOS, resulting in.~\eqref{eq:ptilde_eq_chi}.
+In Appendix A we modified this derivation to account for the periodic boundary conditions we use and show how their result simplifies greatly when expressing it in terms of the EOS, resulting in equation~\eqref{eq:ptilde_eq_chi}.
 
@@
 
-In Appendix A we obtain that $\tilde{\mathcal{P}}$ is given by the sum of all the eigenvalues of $C_A$ with eigenstates related to the cut between $i=1$ and $i=L$ that make a subspace $A_L$.
-The dimension of $A_L$ is however not well-defined as one can always include bulk modes into it.
-Since the bulk modes have $\xi = 0,1$ this makes $\tilde{\mathcal{P}}$ well-defined only modulo 1.
-This can be avoided by computing instead
+As mentioned above the bulk polarization is only defined modulo 1.
+This is obvious when we express it in terms of a geometric phase, which is only defined modulo $2\pi$.
+In our formulation (See Eq.~\eqref{eq:ptilde_eq_chi}) this is reflected in an ambiguity in the number of bulk modes with $\xi = 1$ that are included in $\chi$.
+Being defined only modulo $1$, the bulk polarization cannot differentiate the phases $\nu=0$ and $\nu=2$, even though the EOS of both phases is distinct.
+Another issue due to it being defined in $\mathbb{R}$ mod $1$ appears when we try to follow its evolution over a path $C_\lambda$.
+Between any two points $\lambda$ and $\lambda'$ in this path it is not possible to tell if the bulk polarization has increased or decreased by only looking at $\tilde{\mathcal{P}}(\lambda)$, this becomes relevant when computing Chern numbers.
+
+In an attempt to solve this issues we introduce an alternative bulk polarization given by
 \begin{equation}\label{eq:tpo}
@@
 \end{equation}
-where $\ketpsio{{\Phi=0}}$ is the ground state of the chain opened between sites $i=L/2$ and $i=L/2+1$ and 
+where $\ketpsio{{\Phi=0}}$ is the ground state of the periodic chain opened between sites $j=L/2$ and $j=L/2+1$ and 
 \begin{equation}
@@
 \end{equation}
-There is now no ambiguity in how the flux is introduced, as oppossed to the case with PBC.
+When introducing the flux in the system with PBC (See Eq.~\eqref{eq:Ground_state_flux}) there is an ambiguity due to the bulk modes.
+This ambiguity dissapears when we introduce the flux as above in the open chain.
 Note that $\Phi$ is not a magnetic flux but should be treated as an additional parameter, and therefore $\tpo$ is not necessarily the physical polarization of the open chain.
@@
 \end{align}
-where we used Eq.~\eqref{eq:gauge} and $C_{\rm o}$($C_{A,\rm o}$) denotes the (reduced) density matrix of the open chain.
-
-The idea is that in the process of opening the chain, which can be done adiabatically, all eigenvalues that are not in $A_L$ should get pushed to $\xi=0$ or $\xi=1$, by getting rid of the subspace $A_R$ by this process we can finally obtain $\tpo$ by summing all eigenvalues of $C_{A,\rm o}$.
-The advantage of doing this is that, if for a given path in parameter space $\boldsymbol{\lambda}_i \rightarrow \boldsymbol{\lambda}_f$ their contribution remains constant, that is, there are no jumps between $\xi=0$ and $\xi=1$ such that $\tpo (\boldsymbol{\lambda} )$ is smooth, then changes in $\tpo(\boldsymbol{\lambda}_i ) $ along the path will be defined as a "real number" and not only modulo 1.
-There are, however, two caveats.
+where we used Eq.~\eqref{eq:gauge} and ($C_{A,\rm o}$)$C_{\rm o}$ denotes the (subsystem) correlation matrix of the open chain.
+
+The idea is that in the process of opening the chain, which can be done adiabatically, all eigenvalues that are related to the right virtual edge should get pushed to $\xi=0,1$.
+As a result, the sum of all eigenvalues of $C_{A,o}$ is equal, modulo 1 and in the thermodynamic limit, to $\chi$ (See Eq.~\eqref{eq:chi_left}).
+The advantage of using $\tpo$ is that the contribution from the eigenvalues far from the virtual edge is always constant, as opposed to the case of $\chi$ where this is only true when taken modulo 1.
+The consequence is that for paths in parameter space, $C_\lambda$, which are gapped, i.e.
+no states cross the Fermi energy, $\tpo (\lambda)$ is defined in $\mathbb{R}$, such that the total change in $\tpo$ along this path can be obtained knowing $\tpo$ only in the initial and final states, $\Delta \tpo ({C_\lambda}) = \tpo(\lambda_f) - \tpo(\lambda_i)$.
+Note, however, that $\tpo$ is still gauge-invariant only when taken modulo 1.
+
+There are, however, two caveats to this method.
 When we open the chain, if the system is in a topological non-trivial state, zero-energy modes will appear.
@@
 The second issue is that, even if we allow for symmetry-breaking terms so that we can compute $\tpo$, there might be integer jumps in $\tpo$ where the edge-modes cross zero-energy.
-However, as we will see below these jumps in $\tpo$ due to the zero modes are not simple jumps but branch cuts that provide the correct description of the closed system.
+However, as we will see below these jumps in $\tpo$ due to the zero modes are branch cuts that provide the correct description of the system.
 
@@
 \begin{equation}
-SH(\kappa)S = -H(-\kappa), \quad S C_A(\kappa) S = \mathbb{I}-C_A(-\kappa),
+SH(\kappa)S^\dagger = -H(-\kappa), \quad S C_A(\kappa) S^\dagger = \mathbb{I}-C_A(-\kappa),
 \end{equation}
@@
 Unless $\tpo(0) = L/2$, which is the result expected for a trivial insulator, the result will be discontinuous at $\kappa = 0$.
-We can see this in Fig.\ref{ssh_chern}.(a) where we show $\tpo$ in the parameter space $(m,\kappa)$.
+We can see this in Fig.\ref{ssh_chern}.a where we show $\tpo$ in the parameter space $(m,\kappa)$.
 The presence of the zero-mode for $m<1$ appears in $\tpo$ as a branch cut.
+The relation between the zero-modes and $\Tr[C_{A,o}]$ was already observed in the context of Chern insulators \cite{Alexandrinata2011}.
 This branch cut is in fact not an artifice of the open chain trick, but the result of the degenerate point at $(m=1,\kappa=0)$ having an associated Chern number $C=1$ \cite{Asboth2016}.
 
-Consider a counterclockwise loop around the degenerate point at $(m=t,\kappa=0)$ parametrized by $\theta$, starting from the branch cut. The Chern number is then defined as
-\begin{equation}
-C = \int_0^{2\pi}d\theta\int_0^{2\pi}\frac{d\Phi}{2\pi} \Omega_{\theta\Phi},
+Consider a counterclockwise loop around the degenerate point at $(m=t,\kappa=0)$ starting from the branch cut parametrized by $\theta$,
+\begin{align*}
+& m(\theta) = 1+\frac{1}{2}\sin(\theta-\frac{\pi}{2}) \\
+& \kappa(\theta) = -\sin(\theta).
+\end{align*}
+The Chern number is then defined as
+\begin{equation}
+C = \frac{1}{2\pi}\oint d\theta\oint d\Phi \, \Omega_{\theta\Phi},
 \label{eq:chern}
@@
 with the Berry curvature $\Omega_{\theta\Phi}$ defined in equation~\eqref{eq:BerryCurvature}. 
-From  Eq.~\eqref{eq:tpo}, it follows that the Berry connection for the flux is simply $A_\Phi(\theta,\Phi,) = \tpo(\theta)$, so we now need to compute $A_\theta(\theta,\Phi)$.
-Expanding $\ketpsio{\Phi}$ in the eigenbasis of the number operator $\hat{N_A}$ -- denoted by $\{\ket{j}\}$ -- we can compute its derivative with respect to the parameter $\theta$ as
-\changed{changed}
-\begin{align}
-\partial_\theta \ketpsio{\Phi} =& \sum_j [\partial_\theta c_j (\theta)]e^{-i\Phi N_A^j}\ket{j}.
-\end{align}
-where $c_j(\theta) = \Big\langle j\ketpsio{0}$ does not depend on $\Phi$. The correspondent Berry connection then gives
-\begin{align}
-\brapsio{\Phi}\partial_\theta \ketpsio{\Phi} =& \sum_{lj} c^*_l(\theta) [\partial_\theta c_j(\theta)] \bra{l} e^{i\Phi (N_A^lN_A^j)}\ket{j}\nonumber\\
+From  Eq.~\eqref{eq:tpo}, it follows that the Berry connection for the flux is simply $A_\Phi(\theta,\Phi) = \tpo(\theta)$, so we now need to compute $A_\theta(\theta,\Phi)$.
+
+We can expand $\ket{\tilde{\Psi}^{\Phi = 0}_o}$ in the eigenbasis of the number operator $\hat{N}_A$, denoted by $\{\ket{j}\}$, as
+\begin{equation}
+\ket{\tilde{\Psi}^{\Phi = 0}_o(\theta)} = \sum_j c_j(\theta)\ket{j},
+\end{equation}
+where $c_j(\theta) = \Big\langle j\ket{\tilde{\Psi}^{\Phi = 0}_o(\theta)}$. The flux can then be introduced as
+\begin{equation}
+\ket{\tilde{\Psi}^{\Phi }_o(\theta)} = \sum_j c_j(\theta)e^{-i \Phi N_A^j}\ket{j}.
+\end{equation}
+Computing the derivative of the ground state with respect to the parameter $\theta$ gives
+\begin{align}
+\partial_\theta \ketpsio{\Phi(\theta)} =& \sum_j [\partial_\theta c_j (\theta)]e^{-i\Phi N_A^j}\ket{j}.
+\end{align}
+The correspondent Berry connection then gives
+\begin{align}
+\brapsio{\Phi(\theta)}\partial_\theta \ketpsio{\Phi(\theta)} =& \sum_{lj} c^*_l(\theta) [\partial_\theta c_j(\theta)] \bra{l} e^{i\Phi (N_A^lN_A^j)}\ket{j}\nonumber\\
 =& \sum_j  c^*_j(\theta) \partial_\theta c_j(\theta) \nonumber\\
-=&\brapsio{0}\partial_\theta \ketpsio{0}, 
+=&\brapsio{0(\theta)}\partial_\theta \ketpsio{0(\theta)}, 
 \end{align}
@@
 
-The integral in \eqref{eq:chern} is performed on the surface of the torus defined by $(\theta,\Phi)$.
-Because of the branch cut the Berry connection $A_\Phi(\theta)$ cannot be made smooth for the whole torus and directly apply Stokes theorem.
-Instead, consider that we split the torus in two cylinders, $S_1$ with $\theta\in[\theta_0,2\pi-\theta_0]$ and  $S_2$ with $\theta\in[2\pi-\theta_0,2\pi+\theta_0]$ and we change the gauge in $\boldsymbol{A}(\theta)$ such that it is smooth in each cylinder.
+The integral in Eq.\eqref{eq:chern} is performed on the surface of the torus defined by $(\theta,\Phi)$.
+Because of the branch cut the Berry connection $A_\Phi(\theta)$ cannot be made smooth for the whole torus and therefore we cannot directly apply Stokes theorem.
+Instead, we split the torus in two cylinders, $S_1$ with $\theta\in[\theta_0,2\pi-\theta_0]$ and  $S_2$ with $\theta\in[2\pi-\theta_0,2\pi+\theta_0]$ and we change the gauge in $\boldsymbol{A}(\theta)$ such that it is smooth in each cylinder.
 We can now apply Stokes theorem to the two cylinders independently \cite{Kohmoto1985}.
@@
 \begin{align}
-C_1 =& \int_{S_1} d\boldsymbol{S}\cdot (\nabla \times \boldsymbol{A}) \nonumber\\
-=& \int_{\partial S_1} d\boldsymbol{l}\cdot \boldsymbol{A}(\theta,\Phi) \nonumber\\
-=& \int_a^{2\pi-a} d\theta A_\theta(\theta,0) +\int_0^{2\pi} \frac{d\Phi}{2\pi} A_\Phi (2\pi-a,\Phi) \nonumber\\
-&-\int_a^{2\pi-a} d\theta A_\theta(\theta,2\pi) -\int_0^{2\pi} \frac{d\Phi}{2\pi} A_\Phi (a,\Phi).
+C_1 =& \frac{1}{2\pi} \int_{S_1} d\boldsymbol{S}\cdot (\nabla \times \boldsymbol{A}) \nonumber\\
+=&\frac{1}{2\pi} \int_{\partial S_1} d\boldsymbol{l}\cdot \boldsymbol{A}(\theta,\Phi) \nonumber\\
+=& \frac{1}{2\pi} \int_{\theta_0}^{2\pi-{\theta_0}} d\theta A_\theta(\theta,0) +\frac{1}{2\pi}\int_0^{2\pi} \frac{d\Phi}{2\pi} A_\Phi (2\pi-{\theta_0},\Phi) \nonumber\\
+&-\frac{1}{2\pi}\int_{\theta_0}^{2\pi-{\theta_0}} d\theta A_\theta(\theta,2\pi) -\frac{1}{2\pi}\int_0^{2\pi} \frac{d\Phi}{2\pi} A_\Phi ({\theta_0},\Phi).
 \end{align}
@@
 \begin{align}
-C_1 =& \int_0^{2\pi} \frac{d\Phi}{2\pi} [A_\Phi (2\pi-a,\Phi) -A_\Phi (a,\Phi)].
+C_1 =& \int_0^{2\pi} \frac{d\Phi}{2\pi} [A_\Phi (2\pi-{\theta_0},\Phi) -A_\Phi ({\theta_0},\Phi)].
 \end{align}
@@
 \begin{align}
-C_2 =& \int_0^{2\pi} \frac{d\Phi}{2\pi} [A'_\Phi (2\pi+a,\Phi) -A'_\Phi (2\pi-a,\Phi)].
+C_2 =& \int_0^{2\pi} \frac{d\Phi}{2\pi} [A'_\Phi (2\pi+{\theta_0},\Phi) -A'_\Phi (2\pi-{\theta_0},\Phi)].
 \end{align}
@@
 \end{align}
-which results in $+1$ when $m<1$ and $0$ when $m>1$, as it is known for the SSH chain. The discontinuity in $\tpo$ is therefore related to the Chern number.% and it is not merely continuous modulo 1.
-
-%\changed{undo figure uncomment}
+which results in $+1$ when $m<1$ and $0$ when $m>1$, as it is known for the SSH chain.
+The only contribution to the Chern number comes from the discontinuity of $\tpo$ at the branch cut.
+In general the Chern number over a loop is equal to the sum of the discontinuities of $\tpo$ across the branch cuts.
+
 \begin{figure}[t]
-	\centering
-	\makebox[0pt]{
-		\subfloat[$\tpo-L/2$]{
-			\includegraphics[width=50mm]{windingSSH1.pdf}
-		}
-		\subfloat[$\tpo-L/2$]{
-			\includegraphics[width=45mm]{windingtp1.pdf}
-		}
-	}
-	\caption{Polarization $\tpo$ computed for $t=1,\kappa'=0$ and system size $L=160$, for the SSH chain $t'=0$ in (a) and for $t'=-2$ in (b) }
+\includegraphics[width=86mm]{fig5comp.pdf}
+\caption{Polarization $\tpo-L/2$ computed for $t=1,\kappa'=0$.
+(a) and (b) are computed for $L=40$ sites and parameters $t'=0$ and $t'=-2$, respectively.
+$\tpo$ presents branch cuts that indicate the presence of zero-energy modes.
+$\tpo$ is able to differentiate between the phases $\nu=2$ and $\nu=0$.
+(c) is computed for $t'=2$, with an added disorder to $t$ and $m$ with strength $W=3$, and $\kappa$ is applied only at the edges.
+We compute it for $L=400$, as the disorder increases the correlation length.
+The disorder opens a region with $\nu=1$ as it has been observed in \cite{Song2014} }
 	\label{ssh_chern}
@@
 
-%\changed{undo figure uncomment}
-
-\begin{figure}[t]
-	\centering
-	\makebox[0pt]{
-		\subfloat[$2\tilde{\mathcal{P}}_{\rm open}-L$($m$ should go from 0 to 3)]{
-			\includegraphics[width=45mm]{windingtpdis.pdf}
-		}
-		\subfloat[$C $]{
-			\includegraphics[width=45mm]{plotwinding.pdf}
-		}		
-	}\hspace{0mm}
-
-	\makebox[0pt]{
-		\subfloat[${\rm Log}_{10} (C \, {\rm mod} \, 1)$ ]{
-			\includegraphics[width=45mm]{plotwindingerror.pdf}
-		}
-	}
-	\caption{We compute (a)$\tilde{\mathcal{P}}_{\rm open}$, (b) $\nu$ and (c) error in the quantization of $\nu$ for the continuous line in the phase diagram, $t=1,t'=2$ and system size $L=400$.
+% \begin{figure}[t]
+%	\centering
+%	\makebox[0pt]{
+%		\subfloat[$2\tilde{\mathcal{P}}_{\rm open}-L$($m$ should go from 0 to 3)]{
+%			\includegraphics[width=45mm]{fig61.pdf}
+%		}
+%		\subfloat[$C $]{
+%			\includegraphics[width=45mm]{fig62.pdf}
+%		}		
+%	}\hspace{0mm}
+%	
+%	\makebox[0pt]{
+%		\subfloat[${\rm Log}_{10} (C \, {\rm mod} \, 1)$ ]{
+%			\includegraphics[width=45mm]{fig63.pdf}
+%		}
+%	}
+%	\caption{We compute (a)$\tilde{\mathcal{P}}_{\rm open}$, (b) $\nu$ and (c) error in the quantization of $\nu$ for the continuous line in the phase diagram, $t=1,t'=2$ and system size $L=400$.
 (a) is computed for $W=6$ showing that the branch cut is not affected by the disorder.
@@
 }
-	\label{winding}
-\end{figure}
-
-\subsubsection{$t' \neq 0$}
+%	\label{winding}
+%\end{figure}
+
+\subsubsection{Second neighbor hopping, $t' \neq 0$}
 
@@
 This is no longer true when $t' \neq 0$ allowing for a $\nu = 2$ phase.
-In Fig.\ref{ssh_chern}.(b) we see $\tpo$ for the dotted cut in the phase diagram with an added symmetry breaking $\kappa$ term.
+In Fig.\ref{ssh_chern}.b we see $\tpo$ for the dotted cut in the phase diagram with an added symmetry breaking $\kappa$ term.
 We can see that a loop around the two degenerate points will have a Chern number of $C=2$.
 
-A Chern number C calculated for a gapped loop crossing the trivial phase and some point at $(m,\kappa=0)$ requires the presence of at least C zero-modes.
-In this case we can see that the equivalence between the Chern number and the number of zero-modes, or the winding number, is exact.
-It is also interesting to point out that even far away from the symmetric point $\tpo$ still shows plateaus that point to a nearby topological transition.
-Note , however, that there are many ways of breaking chiral symmetry and one can find terms where a loop crossing two points in the $\nu=0$ and $\nu=2$ has Chern number $C=0$ such that these two phases would appear equivalent.
-\carlos{write term}
+A jump of $n$ in $\tpo$ implies the existence of \emph{at least} $n$ zero-energy modes at that point.
+For this model this equivalence is exact and the jump of $\tpo$ at $\kappa = 0$ gives the number of topological zero-modes, or the winding number.
+It is also interesting to point out that even far away from the symmetric point $\tpo$ still shows plateaus that point to a possible nearby topological transition.
+Note , however, that there are many ways of breaking chiral symmetry.
+Replacing $\kappa$ in Eq.~\eqref{bdi_model} by $\kappa_i=\kappa (-1)^i$, the zero-energy modes split in such a way that $\tpo$ is continuous at $\kappa=0$.
+It cannot differenciate between the phases $\nu=0$ and $\nu=2$, and a loop like the one discussed above results in a Chern number $C=0$.
 
@@
 \end{align}
-where $\omega_i$ and $\omega_i$ are randomly selected from $[-1/2,1/2]$.
-For very strong disorder, where the gap fills with states, the limit $\lim_{\kappa \rightarrow 0} \tilde{\mathcal{P}}_{\rm open}$ might become problematic as bulk states can cross $E=0$ as we tune $\kappa$ down.
-In order to avoid it we break chiral symmetry locally in the edge of the open chain so it only affects the zero modes.
-If any of the other states cross $E=0$ they do it in pairs such that $\tpo$ is not affected by it.
-We show $\tpo$ in Fig.\ref{winding}.(a) using the local symmetry breaking term for a strong disorder, $W=6t$, where the gap is indeed filled with states.
-For the chain without disorder this local symmetry-breaking term would result in the same Chern number.
-The main difference being that since it only affects the zero-modes, $\tpo$ is quantized to multiples of $1/2$.
-
-In Fig.\ref{winding}.(b) we show $C$ for the continuous line in the phase diagram and increasing disorder strength.
+where $\omega_i$ and $\omega_i'$ are selected from a uniform distribution with range $[-1/2,1/2]$.
+For very strong disorder where the gap fills with states $\tpo$ will present many branch cuts as we approach the limit $\lim_{\kappa \rightarrow 0}$, which makes it unpractical.
+In order to avoid it we break chiral symmetry locally in the edge of the open chain so it only affects the edge modes.
+If any of the other states cross $E=0$ they do it in  chiral-symmetric pairs such that $\tpo$ is not affected by it.
+We show $\tpo$ in Fig.\ref{ssh_chern}.c using the local symmetry breaking term for a strong disorder, $W=3t$, where the gap is indeed filled with states.
+We see that it works just as in the case without disorder, where now due to the disorder a region of $\nu=1$ opens up \cite{Song2014}.
+The main difference is that since we break chiral symmetry only for the edge modes, $\tpo$ is quantized to multiples of $1/2$.
+
+%In Fig.\ref{winding}.(b) we show $C$ for the continuous line in the phase diagram and increasing disorder strength.
 The Chern number as defined in \eqref{eq:Chern_number} indeed provides a topological invariant which can be compared with the winding number obtained in reference \cite{Song2014} for the same system.
-Note that for large enough systems the spectrum of $C_A$ is not affected by the local chiral symmetry breaking term other than the occupation of the zero-energy modes such that $C$ is still quantized, however near the transitions it is smoothen by the disorder average.
+Note that for large enough systems the spectrum of $C_A$ is not affected by the local chiral symmetry breaking term other than the occupation of the zero-energy modes such that $C$ is still quantized for each disorder configuration, however near the transitions it is smoothen by the disorder average.
 We show the error in the quantization of $C$ in figure \ref{winding}.(c) .
@@
 
-\mariac{removed the paragraph about appendix D}
-%\carlos{In Appendix D I give an expression for the winding number for flux insertion.
-However this expression seems problematic with a filled gap (Prodan refers to this).
-If we found an inhomogeneous term that induce a topological transition without filling the gap we can use this formula instead of comparing with the Prodan results, so we dont have to worry about the filled gap.
-%Furthermore I worked an expression for this winding number in terms of the correlation function, which I believe we might be able to relate to our $\chi$.
-However I can only do this on the open chain, where the winding number itself is not well-defined because the Hamiltonian is not invertible.
-We can do the trick of splitting the edge-modes and arguing that the winding number should not change by doing that because its a  bulk property, but I'm not too sure about that, so it would be better if we found a way of computing the winding number for open chains.
-}
-
 % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - %
@@
 
-We have shown how $\tpo$ allows us to compute Chern numbers of 1D chains in a simpler way. This can be generalized to 2D systems. Consider the following Chern insulator
+We have shown how $\tpo$ simplifies the computation of Chern numbers in 1D chains. This can be generalized to 2D systems. Consider the following model for a Chern insulator
 \begin{align}
@@
 +&(2-m)\sigma_z\delta_{ij}\delta_{mn},
+\label{eq:chern_model}
 \end{align}
@@
 H_{i\alpha,j\beta}(k_y)=& \frac{1}{2}(i\sigma_x-\sigma_z)\delta_{i,j+1}+\frac{1}{2}(-i\sigma_x-\sigma_z)\delta_{i,j-1} \nonumber\\
-+&\sin(k_y)\sigma_y+[2-m-\cos(k_y)]\sigma_z.
++&(\sin(k_y)\sigma_y+[2-m-\cos(k_y)]\sigma_z) \delta_{ij}.
 \end{align}
 We can treat the Hamiltonian with elements $H_{i\alpha,j\beta}(k_y)$ as a 1D Hamiltonian with an extra parameter $k_y$ and compute $\tpo$, or equivalently $A_\Phi(\Phi,k_y)$, as in last section.
-But first lets analyze the entanglement occupancy spectrum of this system, shown in Fig.\ref{chern_insulator}(a) and (b) for $m=1,3$ respectively.
-For $0<m<2$ the EOS presents two edge modes that connect the occupied and empty bands that cross $\xi=1/2$ at $k_y=0$, in this region the system has a Chern number $C=+1$.
+But first lets analyze the entanglement occupancy spectrum of the system with PBC, shown in Fig.\ref{chern_insulator}.a and b for $m=1,3$ respectively.
+For $0<m<2$ the EOS presents two edge modes that connect the occupied and empty bands that cross $\xi=1/2$ at $k_y=0$, in this region the occupied bands have a Chern number $C=+1$.
 For $2<m<4$ there are also edge modes but they cross $\xi=1/2$ at $k_y=\pi$ and the system has $C=-1$.
 For $m>4$ the system is a trivial insulator with $C=0$.
-When we open the chain one of the edge modes will be pushed to the bulk and will present a jump between $\xi=1,0$ whenever the edge-modes crossed zero energy.
+When we open the chain the eigenvalue of one of the virtual edge modes will be pushed to $\xi=1,0$ as seen in figure~\ref{chern_insulator}.a and b.
+It will have a jump whenever the correspondent edge-mode crossed zero energy, going between $\xi=1$ and $\xi=0$, i.e.
+the number of eigenvalues at $\xi = 1$ is not constant across $k$.
 In this model, due to the symmetries, this is only possible at $k_y=0,\pi$.
-
-We can now take a look at $\tpo$ in the parameter space $(m,k_y)$, shown in figure \ref{chern_insulator}.(c).
-We see, again, that $\tpo$ remains continous except for the branch cuts that appear at the points where the physical edge-modes cross zero energy.
-The Chern number can be obtained as
-\begin{equation}
-C(m) = \int_{-\pi}^{\pi}\frac{dk_y}{2\pi} \partial_{k_y} \tpo(m,k_y),
-\label{eq:Zaletel_chern}
-\end{equation}
-However, since we know the position of the branch cuts we can integrate avoiding them and obtain
-\begin{align}
-C(m) =& [\tpo(m,k_y=0^-)-\tpo(m,k_y=-\pi^+)]\\
-&+[\tpo(m,k_y=\pi^-)-\tpo(m,k_y=0^+)].
-\end{align}
 
@@
 	\makebox[0pt]{
-		\subfloat[Spectrum of $C_{A}$ for $m=1$]{
-			\includegraphics[width=45mm]{chernES0.pdf}
-		}
-		\subfloat[Spectrum of $C_{A}$ for $m=3$]{
-			\includegraphics[width=45mm]{chernESpi.pdf}
+		\subfloat[EOS for $m=1$]{
+			\includegraphics[width=45mm]{fig71.pdf}
+		}
+		\subfloat[EOS for $m=3$]{
+			\includegraphics[width=45mm]{fig72.pdf}
 		}
@@
 		\subfloat[$\tpo-L/2$]{
-			\includegraphics[width=45mm]{cherntpo.pdf}
+			\includegraphics[width=45mm]{fig73.pdf}
 		}
 		\subfloat[$\tilde{\mathcal{P}}$ mod 1]{
-			\includegraphics[width=45mm]{cherntp.pdf}
-		}
-	}\hspace{0mm}
-
-	\makebox[0pt]{
-		\subfloat[$\tpo-L/2$]{
-			\includegraphics[width=45mm]{cherntpo4.pdf}
-		}
-		\subfloat[$\tilde{\mathcal{P}}$ mod 1]{
-			\includegraphics[width=45mm]{cherntp4.pdf}
-		}
-	}
-
-	\caption{Polarization $\tpo$ computed for $t=1,\kappa'=0$ and system size $L=160$, for the SSH chain $t'=0$ in (a) and for $t'=-2$ in (b) CHANGE}
+			\includegraphics[width=45mm]{fig74.pdf}
+		}
+	}
+%	\hspace{0mm}
+%	
+%	\makebox[0pt]{
+%		\subfloat[$\tpo-L/2$]{
+%			\includegraphics[width=45mm]{fig75.pdf}
+%		}
+%		\subfloat[$\tilde{\mathcal{P}}$ mod 1]{
+%			\includegraphics[width=45mm]{fig76.pdf}
+%		}
+%	}
+%	
+	\caption{(a) and (b) EOS.
+The evolution of the edge state towards $\xi=1,0$ as the chain is opened adiabatically is shown in two intermediate points.
+(c) $\tpo$ and (d) $\tilde{\mathcal{P}}$.
+Computed for the Chern insulator given by the model in Eq.~\eqref{eq:chern_model}.
+}
 	\label{chern_insulator}
@@
 
-To ilustrate the advantage of using $\tpo$ instead of $\tilde{\mathcal{P}}$ we also show the latter in the same parameter space.
-In order to obtain the Chern number there is no way of avoiding having to compute the derivative and subsequent integration of
-\begin{align}
-C(m) =& [\tpo(m,k_y=0^-)-\tilde{\mathcal{P}}(m,k_y=-\pi^+)]\\
-&+[\tpo(m,k_y=\pi^-)-\tpo(m,k_y=0^+)],
-\end{align}
+We can now take a look at $\tpo$ in the parameter space $(m,k_y)$, shown in figure~\ref{chern_insulator}.c .
+We see, again, that $\tpo$ remains continous except for the branch cuts that appear at the points where the physical edge-modes cross zero energy.
+The Chern number can be obtained as
+\begin{equation}
+C(m) = \int_{-\pi}^{\pi}\frac{dk_y}{2\pi} \partial_{k_y} \tpo(m,k_y),
+\label{eq:Zaletel_chern}
+\end{equation}
+However, since we know the position of the branch cuts we can integrate avoiding them and obtain
+\begin{align}
+C(m) =& [\tpo(m,k_y=0^-)-\tpo(m,k_y=-\pi^+)]\\
+&+[\tpo(m,k_y=\pi^-)-\tpo(m,k_y=0^+)].
+\end{align}
+The expression in the right hand side is actually a known topological invariant, the trace index \cite{Alexandrinata2011}, defined in terms of the trace of $C_{A,o}$.
+In reference \cite{Alexandrinata2011} it is shown to be equal to the Chern number by relating it to the Hall current.
+By understanding $\Tr[C_{A,o}]$ as a geometric quantity the connection between the trace index and the Chern number is made more explicit.
+
+To ilustrate the advantage of using $\tpo$ over $\tilde{\mathcal{P}}$ we also show the latter in the same parameter space.
+In order to obtain the Chern number via $\tilde{\mathcal{P}}$ for PBC there is no way of avoiding having to compute the derivative and subsequent integration of
+\begin{equation}
+C(m) = \int_{-\pi}^{\pi}\frac{dk_y}{2\pi} \partial_{k_y} \tilde{\mathcal{P}}(m,k_y),
+\end{equation}
 as we no longer know in advance where the branch cuts will appear, due to it being defined modulo 1.
-This example seems a bit trivial as the Chern number is also easy to identify from the plot of $\tilde{\mathcal{P}}$ as well but in systems with higher Chern numbers $\tilde{\mathcal{P}}$ will wind around the circle several times which will make this more difficult.
-\carlos{Figures 7.e and f are computed for the same system where instead of nearest-neighbor hopping in the x direction we use a 4th neighbor hopping ($\delta_{i,j+1} \rightarrow \delta_{i,j+4}$), to illustrate the challenges of computing the Chern number when the integrand is only defined modulo 1.
+
+% This example might seem insufficient as the Chern number is also easy to identify from the plot of $\tilde{\mathcal{P}}$ as well but in systems with higher Chern numbers $\tilde{\mathcal{P}}$ will wind around the circle several times which will make this more difficult.
+We exemplify this in figures \ref{chern_insulator}.(e) and (f).
+To obtain these we change the nearest neighbor hoping in the $x$ direction into a third neighbor hopping, $\delta_{i,j+1} \rightarrow \delta_{i,j+3}$.
+This results in the same phase diagram with the Chern number, and the number of edge modes, multiplied by 3.
+As we see, while the plot of $\tpo$ remains simple, $\tilde{\mathcal{P}}$ becomes more complex, with the appearance of additional branch cuts.
+
+%
+%   \carlos{Figures 7.e and f are computed for the same system where instead of nearest-neighbor hopping in the x direction we use a 4th neighbor hopping ($\delta_{i,j+1} \rightarrow \delta_{i,j+4}$), to illustrate the challenges of computing the Chern number when the integrand is only defined modulo 1.
 This is similar to the issues we had when doing the cut-average.
 One needs to identify whenever $\tilde{\mathcal{P}}$ (or $\chi$) crosses 1 or 0 to compute the derivative and integral properly.
-
-Note also that we are computing the Chern number in the torus defined by $(\Phi,k_y)$.
+%
+%Note also that we are computing the Chern number in the torus defined by $(\Phi,k_y)$.
 It seems a bit weird to mix both flux and momentum but it works in the end.
@@
 Here we have shown how it is encoded in the single-body entanglement spectrum.
-This relation is simpler than the one involving the MBES and it provides more insight into the ES.
+This relation is simpler than the one involving the MBES and it provides more insight into the ES and EOS, as well as it provides another simple method for computing the Zak phase for systems without translational invariance.
 Based on this method we show that one can obtain a polarization that is defined in the real axis.
-Although it is not gauge invariant it provides a method for computing Chern numbers for systems without translational invariance.
-We have also shown that the Chern number obtained this way is, for the system studied here, equivalent to the winding number.
+Although it is not gauge invariant it provides additional information and also simplifies the method for computing Chern numbers for systems in 1D without translational invariance or systems in 2D with translational invariance in only one direction.
 
@@
 
-\maria{In this appendix, we review  how to compute the Zak phase from the Schmidt decomposition~\cite{Zaletel2014}and relate it to our own results}
+In this appendix, we review  how to compute the Zak phase from the Schmidt decomposition~\cite{Zaletel2014}and relate it to our own results.
 
@@
 \begin{equation}
-\ket{\Psi_C} = \sum_{p,q} s_p s'_q \ket{p,q}_A \ket{q,p}_B,
+\ket{\Psi^0} = \sum_{p,q} s_p s'_q \ket{p,q}_A \ket{q,p}_B,
 \end{equation}
@@
 \begin{equation}
-\ket{\Psi_0^\Phi} = \sum_{pq} s_p s_q' e^{-i\Phi N_{A_L}^p} \ket{p,q}_A \ket{q,p}_B,
+\ket{\Psi^\Phi} = \sum_{pq} s_p s_q' e^{-i\Phi N_{A_L}^p} \ket{p,q}_A \ket{q,p}_B,
 \label{eq:Ground_state_flux}
@@
 \begin{align}
-\tilde{\mathcal{P}} =& \int_0^{2\pi} \frac{d\Phi}{2\pi}i\bra{\Psi_0^\Phi}\partial_\Phi\ket{\Psi_0^\Phi} \\
+\tilde{\mathcal{P}} =& \int_0^{2\pi} \frac{d\Phi}{2\pi}i\bra{\Psi^\Phi}\partial_\Phi\ket{\Psi^\Phi} \\
 =& \int_0^{2\pi} \frac{d\Phi}{2\pi}i\sum_{p,q} s_p^2 s_q^{\prime \, 2} e^{i\Phi N_{A_L}^p}\partial_{\Phi}e^{-i\Phi  N_{A_L}^p}\\
@@
 
-Consider now that we parametrize a loop in parameter space as $m=1-\cos(\theta),\kappa = \sin(\theta)$.
-
+\iffalse
 \mariac{removed appendix D}
-%\section{Twisted boundary condition Winding number}
-%\carlos{Work in progress}
-%For a system with chiral symmetry the Hamiltonian can be written in some basis as
-%\begin{equation}
-%H = \mqty(0 & h \\ h^\dagger & 0),
-%\end{equation}
-%where $h$ is in general a matrix. In translationally invariant systems the winding number is defined as 
-%\begin{equation}
-%\nu = -i\int_0^{2\pi} \frac{dk}{2\pi} {\rm Tr}[h(k)^{-1} \partial_k h(k)],
-%\label{eq:winding_k}
-%\end{equation}
-%where $h(k)$ is the off-diagonal block of the Bloch Hamiltonian. In systems without translational invariance, one can similarly define the winding number
-%\begin{equation}
-%\nu = -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[h(\Phi)^{-1} \partial_\Phi h(\Phi)],
-%\label{eq:winding_flux}
-%\end{equation}
-%for a $U(1)$ flux introduced (\carlos{I have only found a similar expression in the context of non-hermitian systems \cite{Gong2018}, but I assume this is the expression Prodan refers to.
+\section{Twisted boundary condition Winding number}
+\carlos{Work in progress}
+For a system with chiral symmetry the Hamiltonian can be written in some basis as
+\begin{equation}
+H = \mqty(0 & h \\ h^\dagger & 0),
+\end{equation}
+where $h$ is in general a matrix. In translationally invariant systems the winding number is defined as 
+\begin{equation}
+\nu = -i\int_0^{2\pi} \frac{dk}{2\pi} {\rm Tr}[h(k)^{-1} \partial_k h(k)],
+\label{eq:winding_k}
+\end{equation}
+where $h(k)$ is the off-diagonal block of the Bloch Hamiltonian. In systems without translational invariance, one can similarly define the winding number
+\begin{equation}
+\nu = -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[h(\Phi)^{-1} \partial_\Phi h(\Phi)],
+\label{eq:winding_flux}
+\end{equation}
+for a $U(1)$ flux introduced (\carlos{I have only found a similar expression in the context of non-hermitian systems \cite{Gong2018}, but I assume this is the expression Prodan refers to.
 It gives the correct results in translationally invariant systems, however it is very unpractical.}).
@@
 In the following we will omit the flux parameter for simplicity.
-%
-%Note that the winding number can also be expressed in terms of the flat-band Hamiltonian
-%\begin{align*}
-%Q =& H (H^2)^{-1/2} \\
-%=& \mqty(0 & q \\ q^\dagger & 0).
-%\end{align*}
-%Note that since $Q$ only has eigenvalues $\pm 1 $ , $Q^2 = \mathbb{I}$ and therefore $q^\dagger = q^{-1}$. We can therefore write
-%\begin{equation}
-%\nu = -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[q^\dagger \partial_\Phi q].
-%\end{equation}
-%In this basis the chiral symmetry operator is given by $S = \sigma_z \otimes \mathbb{I}$ such that the winding number can be expressed in terms of the full flat-band Hamiltonian as
-%\begin{align}
-%\nu =& \frac{i}{2}\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S Q^\dagger \partial_\Phi Q] \\
-%=& \frac{i}{2}\int_0^{2\pi} \frac{d\Phi}{2\pi} \left({\rm Tr}[q \partial_\Phi q^\dagger]-{\rm Tr}[q^\dagger \partial_\Phi q] \right)\\
-%=& -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[q^\dagger \partial_\Phi q] 
-%\end{align}
-%where in the last step we use the fact that the winding number is a real number. We can now write the winding number in terms of the correlation matrix
-%\begin{align}
-%\nu =& \frac{i}{2}\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S (1-2C) \partial_\Phi (1-2C)] \\
-%=& -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S (1-2C) \partial_\Phi C] .
-%\end{align}
-%Introducing the flux via twisted boundary conditions allows us to easily construct the ground state with flux in the case of the open chain \eqref{eq:gauge}, therefore we will employ this gauge to compute the correlation matrix.
+
+Note that the winding number can also be expressed in terms of the flat-band Hamiltonian
+\begin{align*}
+Q =& H (H^2)^{-1/2} \\
+=& \mqty(0 & q \\ q^\dagger & 0).
+\end{align*}
+Note that since $Q$ only has eigenvalues $\pm 1 $ , $Q^2 = \mathbb{I}$ and therefore $q^\dagger = q^{-1}$. We can therefore write
+\begin{equation}
+\nu = -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[q^\dagger \partial_\Phi q].
+\end{equation}
+In this basis the chiral symmetry operator is given by $S = \sigma_z \otimes \mathbb{I}$ such that the winding number can be expressed in terms of the full flat-band Hamiltonian as
+\begin{align}
+\nu =& \frac{i}{2}\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S Q^\dagger \partial_\Phi Q] \\
+=& \frac{i}{2}\int_0^{2\pi} \frac{d\Phi}{2\pi} \left({\rm Tr}[q \partial_\Phi q^\dagger]-{\rm Tr}[q^\dagger \partial_\Phi q] \right)\\
+=& -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[q^\dagger \partial_\Phi q] 
+\end{align}
+where in the last step we use the fact that the winding number is a real number. We can now write the winding number in terms of the correlation matrix
+\begin{align}
+\nu =& \frac{i}{2}\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S (1-2C) \partial_\Phi (1-2C)] \\
+=& -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S (1-2C) \partial_\Phi C] .
+\end{align}
+Introducing the flux via twisted boundary conditions allows us to easily construct the ground state with flux in the case of the open chain \eqref{eq:gauge}, therefore we will employ this gauge to compute the correlation matrix.
 The correlation matrix can be written in position space as
-%\begin{equation}
-%C = \mqty(C_A & C_{AB} \\ C_{AB}^\dagger & C_B),
-%\end{equation}
-%where in this basis, assuming $S$ is a local symmetry it is also diagonal.
+\begin{equation}
+C = \mqty(C_A & C_{AB} \\ C_{AB}^\dagger & C_B),
+\end{equation}
+where in this basis, assuming $S$ is a local symmetry it is also diagonal.
 Note that the flux will only appear in the off-diagonal blocks $C_{AB}$ and therefore the first term in the trace above will vanish, leaving the winding number as
-%\begin{align}
-%\nu =&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S C \partial_\Phi C] \\
-%=&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} \left( {\rm Tr}[S C_{AB} \partial_\Phi C_{AB}^\dagger]+{\rm Tr}[S C_{AB}^\dagger \partial_\Phi C_{AB}] \right)
-%\end{align}
-%We can not explicitely construct the correlation matrix with flux as
-%\begin{align*}
-%C_{ij}^\Phi =& \bra{\tilde{\Psi}_{0,{\rm open}}^\Phi}c_i^\dagger c_j \ket{\tilde{\Psi}_{0,{\rm open}}^\Phi} \\
-%=&\bra{\tilde{\Psi}_{0,{\rm open}}^0}e^{i\Phi \hat{N}_A} c_i^\dagger c_j e^{-i\Phi \hat{N}_A}\ket{\tilde{\Psi}_{0,{\rm open}}^0} .
-%\end{align*}
-%Depending on if $i,j$ belong to $A$ or not, commuting the exponentials will leave additional factors or not, resulting in
-%\begin{align*}
-%C^\Phi = \mqty(C_A^0 & e^{i\Phi} C_{AB}^0 \\e^{-i\Phi} C_{AB}^{0\,\dagger} &C_B^{0}),
-%\end{align*}
-%where we have made explicit the flux dependence. The winding number is then
-%\begin{align}
-%\nu =&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S C^\Phi \partial_\Phi C^\Phi] \\
-%=&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} \left( {\rm Tr}[-iS C_{AB}^0 C_{AB}^{0\,\dagger}]+{\rm Tr}[iS C_{AB}^{0\,^\dagger} C_{AB}^0] \right) \\
-%=&  {\rm Tr}[S C_B^0(1-C_B^0)]-{\rm Tr}[S C_A^0(1-C_A^0)] 
-%\end{align}
-%where in the last step we use one of the properties of the correlation function stemming from $C^2 = C$.
+\begin{align}
+\nu =&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S C \partial_\Phi C] \\
+=&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} \left( {\rm Tr}[S C_{AB} \partial_\Phi C_{AB}^\dagger]+{\rm Tr}[S C_{AB}^\dagger \partial_\Phi C_{AB}] \right)
+\end{align}
+We can not explicitely construct the correlation matrix with flux as
+\begin{align*}
+C_{ij}^\Phi =& \bra{\tilde{\Psi}_{0,{\rm open}}^\Phi}c_i^\dagger c_j \ket{\tilde{\Psi}_{0,{\rm open}}^\Phi} \\
+=&\bra{\tilde{\Psi}_{0,{\rm open}}^0}e^{i\Phi \hat{N}_A} c_i^\dagger c_j e^{-i\Phi \hat{N}_A}\ket{\tilde{\Psi}_{0,{\rm open}}^0} .
+\end{align*}
+Depending on if $i,j$ belong to $A$ or not, commuting the exponentials will leave additional factors or not, resulting in
+\begin{align*}
+C^\Phi = \mqty(C_A^0 & e^{i\Phi} C_{AB}^0 \\e^{-i\Phi} C_{AB}^{0\,\dagger} &C_B^{0}),
+\end{align*}
+where we have made explicit the flux dependence. The winding number is then
+\begin{align}
+\nu =&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S C^\Phi \partial_\Phi C^\Phi] \\
+=&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} \left( {\rm Tr}[-iS C_{AB}^0 C_{AB}^{0\,\dagger}]+{\rm Tr}[iS C_{AB}^{0\,^\dagger} C_{AB}^0] \right) \\
+=&  {\rm Tr}[S C_B^0(1-C_B^0)]-{\rm Tr}[S C_A^0(1-C_A^0)] 
+\end{align}
+where in the last step we use one of the properties of the correlation function stemming from $C^2 = C$.
 (\carlos{So far this expression gives the correct result for translationally invariant systems however one needs to do the trick of breaking chiral symmetry slightly.
@@
 })
-%
+\fi
 

### 2020-10-28 version2/draft.tex
..

--- 2020-08-31 version2/draft.tex
@@
 \newcommand{\changed}[1]{{\color{orange}#1}}
+
 \newcommand{\tpo}{\tilde{\mathcal{P}}_{\rm o}}
@@
 This is a very simple model which nevertheless supports a rich phase diagram containing topological phases with winding numbers up to $\nu =2$. 
-In order to show that our results are not limited to any particular symmetry class~\cite{ryu2010topological}, we also include two symmetry breaking terms. 
+In order to show that our results are not limited to any symmetry class~\cite{ryu2010topological}, we also include two symmetry breaking terms. 
 The Hamiltonian is then given by
 \begin{align}
-H =& \sum_{i\alpha,j\beta} c_{i\alpha}^\dagger H_{ij,\alpha \beta} c_{j\beta},
-\end{align}
-where $H_{ij} $ is given by 
-\begin{align}
+H =& \sum_{i\alpha,j\beta} c_{i\alpha}^\dagger H_{ij,\alpha \beta} c_{j\beta} \\
 H_{ij} =& (m \sigma_x + \kappa \sigma_z)\delta_{ij}  + \frac{1}{2i}\kappa'\sigma_z (\delta_{i-j,1}-\delta_{i-j,-1})\nonumber\\
@@
 \end{align}
-and the corresponding Bloch Hamiltonian is given by
+with the corresponding Bloch Hamiltonian being
 \begin{align}
@@
 
-For $\kappa' \neq 0$, only particle-hole symmetry $C$ is preserved and the system belongs to symmetry class $D$. 
+For $\kappa' \neq 0$ only particle-hole symmetry $C$ is preserved and the system belongs to symmetry class $D$. 
 In one spatial dimension, the latter is characterized by a $\mathbb{Z}_2$ invariant.  
@@
 \makebox[0pt]{
-\subfloat[$\kappa = 0.3,\kappa'=0$]{
+\subfloat[$\kappa = .3,\kappa'=0$]{
   \includegraphics[width=35mm]{2c.pdf}
@@
 
-In this section we review some aspects of the modern theory of polarization, which gave a geometrical description of the polarization \cite{Resta1992,KingSmith1993,Vanderbilt1993,Resta1997}. 
-We also discuss the results of a recent article by Watanabe and Oshikawa \cite{Watanabe2018} that clarified the meanings of the different polarizations one can define this way, as well as the relations between them.
+In this section we review some results of the modern theory of polarization, which gave a geometrical description of the polarization \cite{Resta1992,KingSmith1993,Vanderbilt1993,Resta1997} , as well as a recent article by Watanabe and Oshikawa \cite{Watanabe2018} that clarified the meanings of the different polarizations one can define this way and the relations between them.
 For clarity we will use the same notation they use.
@@
 In translational invariant systems $u_{p\mu}^{j\beta} = e^{ipj}\tilde{u}_{p \mu}^{\beta}$, where $\tilde{u}_{p \mu}^{\beta}$ are the components of the eigenstates of the Bloch Hamiltonian, $\ket{\tilde{u}_{k\mu}}$.
-One can now compute the Zak phase \cite{Zak1989}, which for a two-band  model is defined as the geometric phase acquired by the occupied state $\ket{\tilde{u}_k}$ as it winds around the Brillouin zone,
+One can now compute the Zak phase \cite{Zak1989}, which for a two-band \mariac{do we need the restriction to two-band models?} model is defined as the geometric phase acquired by the occupied state $\ket{\tilde{u}_k}$ as it winds around the Brillouin zone,
 \begin{equation}
@@
 where $A_k = \bra{\tilde{u}_k}\partial_k \ket{\tilde{u}_k}$ is the Berry connection \cite{Berry1984}.
-The generalization to higher-band models is straightforward.
-
 The Zak phase is typically regarded as just a phase because it is only gauge invariant modulo $2\pi$.
-However, for two different states defined by a parameter $\lambda$ --- assuming $A_k(\lambda)$ is smooth in the path connecting them --- the change in the Zak phase can be computed as
+However, for two different states defined by a parameter $\lambda$, as long as $A_k(\lambda)$ is smooth in the path connecting them the change in the Zak phase can be computed as
 \begin{equation}
@@
  is the Berry curvature.
-The Berry curvature is fully gauge invariant and, therefore, the change in the Zak phase \eqref{eq:change_zak_phase} is defined along the full real line.
-The Zak phase itself, as defined in Eq.~\eqref{eq:zak_phase}, can be shown to be proportional to the winding number for a certain gauge (see Appendix \ref{appendix:winding_ssh}), which means that it carries physical information beyond modulo $2\pi$.
+The Berry curvature is fully gauge invariant and, therefore, the change in the Zak phase as defined in \eqref{eq:change_zak_phase} is defined along the full real line.
+The Zak phase itself, as defined above, can be shown to be proportional to the winding number for a certain gauge (see Appendix \ref{appendix:winding_ssh}), which means that it carries physical information beyond modulo $2\pi$.
 Note that other ways of computing the Zak phase rely on computing $e^{i\gamma}$ instead, such that the result is always defined only modulo $2\pi$.
@@
 which allows us to make the identification $ \tilde{\mathcal{P}}^{\rm Bloch} = \gamma/2\pi$.
-As discussed above,  this polarization is defined only modulo 1.
-Its changes, however, can be defined as a real number. 
-This is not surprising,  as the derivative of the polarization is related to the current which is a physical observable, unlike $\tilde{\mathcal{P}}^{\rm Bloch}$.
+As discussed above,  this polarization is defined as an absolute value only modulo 1, however its changes can be defined as a real number.
+This is no surprise as the derivative of the polarization is related to the current which is a physical observable, unlike $\tilde{\mathcal{P}}^{\rm Bloch}$.
 
@@
 \carlos{I have to check overall sign and sign of exponential.}and  ${\rm det }\!' $ indicates that the determinant is restricted to the space of \emph{occupied} single-particle states.
-Below we will employ this expression to compare our results with \mariac{XXX}.
+Below we will employ this expression to compare our results with.
 
-There is one last polarization we must consider. 
-It is obtained when the flux is introduced in the bond between sites $i=1$ and $i=L$ via twisted boundary conditions. This is equivalent to performing the gauge transformation
+There is one last polarization we must consider and it is the one obtained when the flux is introduced in the bond between sites $i=1$ and $i=L$ via twisted boundary conditions.
+This is equivalent to performing the gauge transformation
 \begin{align}\label{eq:Psi_tilde}
@@
 \end{equation} 
+\maria{Ref to Oshikawa paper?}
 %The previous discussion about the Zak phase being defined only modulo $2\pi$ although its changes can be defined as a real number applies in the same way to the polarization obtained in \eqref{eq:p_tilde}.
@@
 \end{align}
-For a system where all single-particle eigenstates with energies $E<0$ are occupied, the correlation matrix can be written  in terms of the Hamiltonian as
+For a system where we occupy all single-particle eigenstates with energies $E<0$ the correlation matrix can be written  as
 \begin{align}\label{eq:corr_mat2}
@@
 where $\mathcal{K}$ is a normalization constant and $\mathcal{H}$ is a quadratic Hamiltonian, referred to as entanglement Hamiltonian. 
-The eigenvalues $\epsilon_j$ of $\mathcal{H}$, referred to as entanglement energies,  are related to those of the EOS by 
+The eigenvalues $\epsilon_j$ of $\mathcal{H}$, referred to as entanglement energies,  are related to the EOS by 
 \begin{align}\label{eq:xi_eps}
@@
 
-One property of the subsystem correlation matrix we will make use of later is that, for sufficiently large systems, its eigenstates are found to be localized on either virtual edge if $\xi \neq 0,1$, or they are found to be bulk modes if the correspondent eigenvalues are exponentially close to $\xi = 0,1$ \cite{Peschel2008}.
+One property of the subsystem correlation matrix we will make use of later is that, for sufficiently large systems, its eigenstates are found to be localize on either virtual edge if $\xi \neq 0,1$, or they are found to be bulk modes if the correspondent eigenvalues are exponentially close to $\xi = 0,1$ \cite{Peschel2008}.
 
@@
 In the following sections, we show how the Zak phase can be recovered from the EOS.
-We also provide a simple formula in terms of the EOS eigenvalues that is identical to the Zak phase in the thermodyamic limit.  
-However, for a generic, finite size system, there will be finite-size discrepancies between our formula and the Zak phase. \mariac{see scaling plot}
-In order to reduce these, we are going to choose subsystem $A$ as half the system in the remainder of the manuscript.
+We also give a simple formula that is identical to the Zak phase in the thermodyamic limit.  
+However, for a generic, finite size system, there will be finite-size discrepancies between our formula and the Zak phase.
+In order to reduce these, we are going to choose $A$ as half the system in the remainder of the manuscript.
 The results (in the thermodynamic limit) do not depend on this choice. 
@@
 For symmetry-protected topological systems in one dimension, the quantized Zak phase is zero whenever there is an even number  of eigenvalues per edge at $\xi = 1/2$, and $\pi$ when the number is odd~\cite{Peschel2008}.
-In the absence of symmetry-protection, much less is known. 
 Ryu and Hatsugai considered the fully dimerized SSH chain with broken chiral symmetry~\cite{Ryu2006}.  
-This model is rather special in that there is only a single pair of eigenvalues, related by symmetry \mariac{which symmetry?}, in the EOS that is not strictly identical to $0$ or $1$. 
-The authors could show that the value of one of these `midgap states' is identical to the Zak phase divided by $2\pi$ or, alternatively, $\tilde{\mathcal{P}}^{\rm Bloch}$. 
+This model is rather special in that there is only a single pair of eigenvalues, related by symmetry, in the EOS that is not identical to $0$ or $1$. 
+In fact, the authors could show that the value of one of these `midgap states' is identical to the Zak phase divided by $2\pi$, or $\tilde{\mathcal{P}}^{\rm Bloch}$. 
 Their result is a particular limit of Eq.~\eqref{eq:chi_integrable}. 
@@
 %Beyond symmetry protected systems, there was observed that in the fully dimerized SSH chain, with only a single mode present in the ES, the Zak phase is equal to the midgap eigenvalue \cite{Ryu2006}.
-
-Huang and Arovas considered a two-dimensional Chern insulator and noted that the Zak phase qualitatively follows the `virtual edge mode' that connects the EOS values at 0 to those at 1~\cite{Huang2012,Huang2012-2}.
+Huang and Arovas considered a two-dimensional chern insulator and noted that the Zak phase qualitatively follows the `virtual edge mode' that connects the EOS values at 0 to those at 1~\cite{Huang2012,Huang2012-2}.
 Interpreting this as a one-dimensional system with a parameter, we can explain this behavior by Eq.~\eqref{eq:chi_integrable}, noting that there is a single pair of eigenvalues that dominates the sum.
-Note that the observation of Ref.s~\cite{Huang2012,Huang2012-2} is particular to systems with Chern number 0 or $\pm 1$, and it generically fails for systems with higher Chern numbers. 
-In the latter case, there are several terms in Eq.~\eqref{eq:chi_integrable} with comparably large contributions and, consequently, the Zak phase deviates considerably from the 'virtual edge modes'. 
+Note that the observation of Ref.s~\cite{Huang2012,Huang2012-2} is particular to a system with Chern number 0 or $\pm 1$, and it generically fails for systems with higher Chern numbers. 
+In the latter case, there are several terms in Eq.~\eqref{eq:chi_integrable} with comparably large contributions. 
 %For cases with more edge modes present in the ES it has been shown that the Zak phase follows the midgap eigenvalues qualitatively \cite{Huang2012,Huang2012-2}.
@@
 nearest neighbor hopping models such as the SSH chain, and applies to the Hamiltonian~\eqref{bdi_model} when $t'=0$.
-In this case, one can obtain the Zak phase in a very simple fashion: We reorder the eigenvalues of the EOS by magnitude, $\xi_1 < \xi_2 < ...< \xi_{L_AM}$, and compute
+In this case, one can obtain the Zak phase in a very simple fashion: We reorder the eigenvalues of the ES by magnitude, $\xi_1 < \xi_2 < ...< \xi_{L_AM}$, and compute
 \begin{equation}
@@
 \end{equation}
-In the thermodynamic limit, $\chi$ becomes identical to either $\tilde{\mathcal{P}}^{\rm Bloch}$ or $1-\tilde{\mathcal{P}}^{\rm Bloch}$: \mariac{comment on the ambiguity?}
+In the thermodynamic limit, $\chi$ becomes identical to either $\tilde{\mathcal{P}}^{\rm Bloch}$ or $1-\tilde{\mathcal{P}}^{\rm Bloch}$
 \begin{equation}
@@
 \end{equation}
-\mariac{I don't understand what we want to say here.} For translational invariant systems the EOS is symmetric such that a crossing of the eigenvalues $\xi$ results in $\chi$ changing sign.
+For translational invariant systems the EOS is symmetric such that a crossing of the eigenvalues $\xi$ results in $\chi$ changing sign.
 This is particularly irrelevant in the case of symmetry-protected states where the Zak phase is quantized to $0$ or $\pi$.
-In the special case where there is only one eigenvalue per edge that is $\neq 0,1$ in  the EOS, like in the fully dimerized limit of the SSH chain, we recover the equality between the midgap eigenvalue and the Zak phase found in Ref.~\cite{Ryu2006}.
+
+In the case where there is only one eigenvalue per edge that is $\neq 0,1$ in  the EOS, like in the fully dimerized limit of the SSH chain, we recover the equality between the midgap eigenvalue and the Zak phase found in Ref.~\cite{Ryu2006}.
 
@@
 
-In Appendix A we obtain that $\tilde{\mathcal{P}}$ is given by the sum of all the eigenvalues of $C_A$ with eigenstates related to the cut between $i=1$ and $i=L$ that make a subspace $A_L$.
+In Appendix A we obtain that $\chi$ should be the sum of all the eigenvalues of eigenstates related to the cut between $i=1$ and $i=L$ that make a subspace $A_L$.
 The dimension of $A_L$ is however not well-defined as one can always include bulk modes into it.
-Since the bulk modes have $\xi = 0,1$ this makes $\tilde{\mathcal{P}}$ well-defined only modulo 1.
+Since the bulk modes have $\xi = 0,1$ this makes $\chi$ well-defined only modulo 1.
 This can be avoided by computing instead
@@
 \end{equation}
-There is now no ambiguity in how the flux is introduced, as oppossed to the case with PBC.
-Note that $\Phi$ is not a magnetic flux but should be treated as an additional parameter, and therefore $\tpo$ is not necessarily the physical polarization of the open chain.
+The difference with the closed chain being that now there is no ambiguity on when an electron crosses the boundary with the twisted boundary condition.
+Note that since the chain is open, $\Phi$ is not a flux but should now be regarded as a parameter and therefore $\tpo$ does not need to be a physical polarization.
 $\tpo$ can be obtained in an even simpler way than $\tilde{\mathcal{P}}$ as
@@
 where we used Eq.~\eqref{eq:gauge} and $C_{\rm o}$($C_{A,\rm o}$) denotes the (reduced) density matrix of the open chain. 
-
-The idea is that in the process of opening the chain, which can be done adiabatically, all eigenvalues that are not in $A_L$ should get pushed to $\xi=0$ or $\xi=1$, by getting rid of the subspace $A_R$ by this process we can finally obtain $\tpo$ by summing all eigenvalues of $C_{A,\rm o}$.
-The advantage of doing this is that, if for a given path in parameter space $\boldsymbol{\lambda}_i \rightarrow \boldsymbol{\lambda}_f$ their contribution remains constant, that is, there are no jumps between $\xi=0$ and $\xi=1$ such that $\tpo (\boldsymbol{\lambda} )$ is smooth, then changes in $\tpo(\boldsymbol{\lambda}_i ) $ along the path will be defined as a "real number" and not only modulo 1.
-There are, however, two caveats.
-When we open the chain, if the system is in a topological non-trivial state, zero-energy modes will appear.
-The ground state of the closed chain has several equivalent degenerate ground states for the open chain, so that $C_{A,\rm o}$ is not well-defined at these points.
-The second issue is that, even if we allow for symmetry-breaking terms so that we can compute $\tpo$, there might be integer jumps in $\tpo$ where the edge-modes cross zero-energy.
-However, as we will see below these jumps in $\tpo$ due to the zero modes are not simple jumps but branch cuts that provide the correct description of the closed system.
+Note that Eq.~\eqref{eq:phi_independence} is nothing else than the generalization of $\chi$ to the open chain. 
+\maria{I am not entirely happy with the following argument, as it goes wrong in the topological phase. We mention this later, but we should already mention it here. }
+The idea is that in the process of opening the chain, which can be done adiabatically, all eigenvalues that are not in $A_L$ should get pushed to $\xi=0$ or $\xi=1$. 
+If for a given path in parameter space $\boldsymbol{\lambda}_i \rightarrow \boldsymbol{\lambda}_f$ their contribution remains constant, that is, there are no jumps between $\xi=0$ and $\xi=1$ and $\tpo (\boldsymbol{\lambda} )$ is smooth, then changes in $\tpo(\boldsymbol{\lambda}_i ) $ along the path will be defined as a real number and not only modulo 1.
+Since $\tilde{\mathcal{P}}$ is a local property of the bond between $i=1$ and $i=L$ one would expect it to not depend on what happens far from it such that $\tpo$ and $\tilde{\mathcal{P}}$ are equivalent.
 
@@
 
-Consider first the simple case of the SSH chain, with $t'=0$ in the model. When we include the chiral-symmetry braking parameter $\kappa$ the Hamiltonian now has the symmetry
+Consider the simple case of the SSH chain, with $t'=0$ in the model.
+In the chiral symmetric case, $\kappa = 0$, the open chain presents zero-energy modes for $m<t$ that seem problematic.
+$\tilde{\mathcal{P}}_{\rm open}$  will depend on the occupation of these zero modes although we just argued that $\tilde{\mathcal{P}}$ should not, since the zero-energy modes are localized far from the virtual cut at $i=1$.
+We can break the chiral symmetry in order to split the zero-energy modes so we do not have to make this choice, but the resulting $\tilde{\mathcal{P}}_{\rm open}$ will be discontinuous at $\kappa = 0$, suggesting that the limit does not exist.
+This can be seen from the fact that the Hamiltonian now has the symmetry
 \begin{equation}
-SH(\kappa)S = -H(-\kappa), \quad S C_A(\kappa) S = \mathbb{I}-C_A(-\kappa),
+SH(\kappa)S = -H(-\kappa), \quad S C_A(\kappa) S = \mathbb{I}-C_A(-\kappa)
 \end{equation}
 so $\tpo(-\kappa) = L-\tpo(\kappa)$.
-Unless $\tpo(0) = L/2$, which is the result expected for a trivial insulator, the result will be discontinuous at $\kappa = 0$.
-We can see this in Fig.\ref{ssh_chern}.(a) where we show $\tpo$ in the parameter space $(m,\kappa)$.
-The presence of the zero-mode for $m<1$ appears in $\tpo$ as a branch cut.
-This branch cut is in fact not an artifice of the open chain trick, but the result of the degenerate point at $(m=1,\kappa=0)$ having an associated Chern number $C=1$ \cite{Asboth2016}.
+Unless $\tpo(0) = L/2$, which only happens in the trivial case, the result will be discontinuous at $\kappa = 0$.
+We can see this in Fig.\ref{ssh_chern}.(a) where we show $\tpo$ in the space $(m,\kappa)$.
+This branch cut is in fact not an artifice of the open chain trick, but the result of the degenerate point at $(m=t,\kappa=0)$ having an associated Chern number $C=1$ \cite{Asboth2016}.
 
-Consider a counterclockwise loop around the degenerate point at $(m=t,\kappa=0)$ parametrized by $\theta$, starting from the branch cut. The Chern number is then defined as
+Consider a loop around the singularity at $(m=t,\kappa=0)$ parametrized by $\theta$. The Chern number is then defined as
 \begin{equation}
@@
 \end{equation} 
-with the Berry curvature $\Omega_{\theta\Phi}$ defined in equation~\eqref{eq:BerryCurvature}. 
-From  Eq.~\eqref{eq:tpo}, it follows that the Berry connection for the flux is simply $A_\Phi(\theta,\Phi,) = \tpo(\theta)$, so we now need to compute $A_\theta(\theta,\Phi)$.
-Expanding $\ketpsio{\Phi}$ in the eigenbasis of the number operator $\hat{N_A}$ -- denoted by $\{\ket{j}\}$ -- we can compute its derivative with respect to the parameter $\theta$ as
+with the Berry curvature $\Omega_{\theta\Phi}$ defined in Eq.~\ref{eq:BerryCurvature}. 
+From  Eq.~\eqref{eq:tpo}, it follows that the Berry connection for the flux is simply $A_\Phi(\theta,\Phi,) = -i\tpo(\theta)$, so we only need to compute $A_\theta(\theta,\Phi)$.
+(\carlos{Please double check this calculation}) Expanding $\ketpsio{\Phi}$ in the eigenbasis of the number operator $\hat{N_A}$ -- denoted by $\{\ket{j}\}$ -- we can compute its derivative with respect to the parameter $\theta$ as
 \changed{changed}
@@
 The integral in \eqref{eq:chern} is performed on the surface of the torus defined by $(\theta,\Phi)$.
-Because of the branch cut the Berry connection $A_\Phi(\theta)$ cannot be made smooth for the whole torus and directly apply Stokes theorem.
+Because the Berry connection $A_\Phi(\theta)$ is not smooth one cannot directly apply Stokes theorem.
 Instead, consider that we split the torus in two cylinders, $S_1$ with $\theta\in[\theta_0,2\pi-\theta_0]$ and  $S_2$ with $\theta\in[2\pi-\theta_0,2\pi+\theta_0]$ and we change the gauge in $\boldsymbol{A}(\theta)$ such that it is smooth in each cylinder.
-We can now apply Stokes theorem to the two cylinders independently \cite{Kohmoto1985}.
-For $S_1$ we have
+We can now apply Stokes theorem to the two cylinders independently
 \begin{align}
@@
 \end{align}
-The Chern number is then $C=C_1+C_2$. Noting again that $A'_\Phi(\theta)$ is continuous in $S_2$ taking the limit of $\theta_0\rightarrow 0^+$ makes $C_2$ vanish and the Chern number is given by
+The Chern number is then $C=C_1+C_2$. Noting again that $A'_\Phi(\theta)$ is continuous in $S_2$ taking the limit of $a\rightarrow 0^+$ makes $C_2$ vanish and the Chern number is given by
 \begin{align}
-C =& \int_0^{2\pi} \frac{d\Phi}{2\pi} [A_\Phi (2\pi^-,\Phi) -A_\Phi (0^+,\Phi)],
+C =& \int_0^{2\pi} \frac{d\Phi}{2\pi} [A_\Phi (2\pi^-,\Phi) -A_\Phi (0^+,\Phi)].
 \end{align}
-or
+If we consider the loop starting at $\kappa = 0$ this is exactly the case shown in Fig.\ref{ssh_chern}.(a), where $A_\Phi (\theta,\Phi) = \tilde{\mathcal{P}}_{\rm open}(\theta)$ is smooth everywhere but in $\theta = 0$, and therefore we can compute the Chern number simply as
 \begin{align}
-C(m)= \tpo(m,\kappa=0^+)-\tpo(m,\kappa=0^-),
+C= \tpo(\kappa=0^+)-\tpo(\kappa=0^-).
 \end{align}
-where $C(m)$ is the Chern number of any loop around the degenerate point crossing $\kappa=0$ at $m$.
 Using the symmetry of the polarization described above this results in 
 \begin{align}
-C(m)= 2\tpo(m,\kappa=0^+)-L,
+C= 2\tpo(0^+)-L.
 \label{eq:Chern_number}
 \end{align}
-which results in $+1$ when $m<1$ and $0$ when $m>1$, as it is known for the SSH chain. The discontinuity in $\tpo$ is therefore related to the Chern number.% and it is not merely continuous modulo 1.
+The discontinuity in $\tpo$ is therefore related to the Chern number.% and it is not merely continuous modulo 1.
+
+One can also obtain a polarization continuous in $\kappa$ by doing a gauge transformation 
+\begin{align}
+\ket{\tilde{\psi}^{\prime \,\Phi}}=e^{ i \Phi \Theta(-\kappa)\Theta(t-m)}\ket{\tilde{\psi}^{\Phi}}
+\end{align}
+\maria{where $\Theta$ is the Heaviside theta function (?)}. 
+This gauge transformation places the branch cut at $m=t,\kappa < 0$, as can be seen in Fig.\ref{ssh_chern}(b)  (\carlos{need to double check signs}). 
+This is in fact the result one gets when using the gauge that allows one to compute the winding number from the Zak phase. \maria{???}
 
@@
 	\makebox[0pt]{
-		\subfloat[$\tpo-L/2$]{
+		\subfloat[$\tilde{\mathcal{P}}_{\rm open}-L/2$]{
 			\includegraphics[width=50mm]{windingSSH1.pdf}
 		}
-		\subfloat[$\tpo-L/2$]{
+		\subfloat[$\tilde{\mathcal{P}}_{\rm open,+}-L/2$]{
+			\includegraphics[width=45mm]{windingSSH2.pdf}
+		}
+	}\hspace{0mm}
+
+	\makebox[0pt]{
+		\subfloat[$\tilde{\mathcal{P}}_{\rm open}-L/2$]{
 			\includegraphics[width=45mm]{windingtp1.pdf}
 		}
+		\subfloat[$\tilde{\mathcal{P}}_{\rm open,+}-L/2$]{
+			\includegraphics[width=45mm]{windingtp2.pdf}
 		}
-	\caption{Polarization $\tpo$ computed for $t=1,\kappa'=0$ and system size $L=160$, for the SSH chain $t'=0$ in (a) and for $t'=-2$ in (b) }
+	}
+	\caption{Polarization $\tilde{\mathcal{P}}_{\rm open}$ computed for three different gauge choices, for $t=1,\kappa'=0$ and system size $L=160$. (a) and (b) with $t'=0$ and (c) and (d) for $t'=-2$. }
 	\label{ssh_chern}
@@
 	\makebox[0pt]{
-		\subfloat[$2\tilde{\mathcal{P}}_{\rm open}-L$($m$ should go from 0 to 3)]{
-			\includegraphics[width=45mm]{windingtpdis.pdf}
+		\subfloat[$\tilde{\mathcal{P}}_{\rm open}$]{
+			\includegraphics[width=50mm]{windingtpdis.pdf}
 		}
@@
 This is no longer true when $t' \neq 0$ allowing for a $\nu = 2$ phase.
-In Fig.\ref{ssh_chern}.(b) we see $\tpo$ for the dotted cut in the phase diagram with an added symmetry breaking $\kappa$ term.
-We can see that a loop around the two degenerate points will have a Chern number of $C=2$.
+In Fig.\ref{ssh_chern}.(c) we see $\tpo$ for the dotted cut in the phase diagram with an added symmetry breaking $\kappa$.
+We see clearly that $\tpo$ distinguishes between the $\nu=0$ and $\nu = 2$ phases.
+Similar to before, there is a gauge choice that makes $\tpo$ continuous in $\kappa$.
+However, which gauge transformation does this is a priori not known, so in practice we can only compute the limit $\lim_{\kappa \rightarrow 0} \tpo$ in the original gauge.
 
-A Chern number C calculated for a gapped loop crossing the trivial phase and some point at $(m,\kappa=0)$ requires the presence of at least C zero-modes.
-In this case we can see that the equivalence between the Chern number and the number of zero-modes, or the winding number, is exact.
-It is also interesting to point out that even far away from the symmetric point $\tpo$ still shows plateaus that point to a nearby topological transition.
-Note , however, that there are many ways of breaking chiral symmetry and one can find terms where a loop crossing two points in the $\nu=0$ and $\nu=2$ has Chern number $C=0$ such that these two phases would appear equivalent.
-\carlos{write term}
-
-Since $\tpo$ is naturally described in position space we can also use this method for a disordered system. Consider now that we add disorder in the $t$ and $m$ parameters
+Our expression for the Chern number \eqref{eq:Chern_number} is also valid without translational invariance. Consider now that we add disorder in the $t$ and $m$ parameters
 \begin{align}
@@
 In order to avoid it we break chiral symmetry locally in the edge of the open chain so it only affects the zero modes.
-If any of the other states cross $E=0$ they do it in pairs such that $\tpo$ is not affected by it.
-We show $\tpo$ in Fig.\ref{winding}.(a) using the local symmetry breaking term for a strong disorder, $W=6t$, where the gap is indeed filled with states.
-For the chain without disorder this local symmetry-breaking term would result in the same Chern number.
-The main difference being that since it only affects the zero-modes, $\tpo$ is quantized to multiples of $1/2$.
-
+If any of the other states cross $E=0$ they do it in pairs such that $\tilde{\mathcal{P}}_{\rm open}$ is not affected by it.
+We show $\tilde{\mathcal{P}}_{\rm open}$ in Fig.\ref{winding}.(a) using the local symmetry breaking term for a strong disorder, $W=6t$, where the gap is indeed filled with states.
 In Fig.\ref{winding}.(b) we show $C$ for the continuous line in the phase diagram and increasing disorder strength.
 The Chern number as defined in \eqref{eq:Chern_number} indeed provides a topological invariant which can be compared with the winding number obtained in reference \cite{Song2014} for the same system.
-Note that for large enough systems the spectrum of $C_A$ is not affected by the local chiral symmetry breaking term other than the occupation of the zero-energy modes such that $C$ is still quantized, however near the transitions it is smoothen by the disorder average.
-We show the error in the quantization of $C$ in figure \ref{winding}.(c) .
-\carlos{Fig.6.a seems important to discuss how with local symmetry braking term the relation between the chern number and the number of zero modes still remains.
-Maybe it would be enough to show a cut of 6.b }
+Note that for large enough systems the spectrum of $C_A$ is not affected by the local chiral symmetry breaking term other than the occupation of the zero-energy modes such that $C$ is still quantized, however near the transitions it is smoothen by the disorder average We show the error in the quantization of $C$ in figure \ref{winding}.(c) .
 
@@
 We can do the trick of splitting the edge-modes and arguing that the winding number should not change by doing that because its a  bulk property, but I'm not too sure about that, so it would be better if we found a way of computing the winding number for open chains.
-}
-
-% - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - %
-
-\section{Chern Number in 2D}
-
-We have shown how $\tpo$ allows us to compute Chern numbers of 1D chains in a simpler way. This can be generalized to 2D systems. Consider the following Chern insulator
-\begin{align}
-H =& \sum_{im\alpha,jn\beta} c_{im\alpha}^\dagger H_{im\alpha,jn\beta} c_{jn\beta}, \\
-H_{im\alpha,jn\beta}=& \frac{1}{2}(i\sigma_x-\sigma_z)\delta_{i,j+1}\delta_{mn}+\frac{1}{2}(-i\sigma_x-\sigma_z)\delta_{i,j-1}\delta_{mn} \nonumber\\
-+&\frac{1}{2}(i\sigma_y -\sigma_z)\delta_{ij}\delta_{m,n+1} +\frac{1}{2}(-i\sigma_y -\sigma_z)\delta_{ij}\delta_{m,n-1} \nonumber\\
-+&(2-m)\sigma_z\delta_{ij}\delta_{mn},
-\end{align}
-where $i,j$ describe the $x$ direction and $m,n$ the $y$ direction.
-The system with periodic boundary conditions is defined on a torus.
-In order to apply the method of computing $\tpo$ described in the last section we will open the torus into a cylinder periodic in $y$.
-We then fourier transform only the $y$ direction to obtain
-\begin{align}
-H =& \sum_{i\alpha,j\beta}\sum_{k_y} c_{i\alpha}^\dagger(k_y) H_{i\alpha,j\beta}(k_y) c_{j\beta}(k_y), \\
-H_{i\alpha,j\beta}(k_y)=& \frac{1}{2}(i\sigma_x-\sigma_z)\delta_{i,j+1}+\frac{1}{2}(-i\sigma_x-\sigma_z)\delta_{i,j-1} \nonumber\\
-+&\sin(k_y)\sigma_y+[2-m-\cos(k_y)]\sigma_z.
-\end{align}
-We can treat the Hamiltonian with elements $H_{i\alpha,j\beta}(k_y)$ as a 1D Hamiltonian with an extra parameter $k_y$ and compute $\tpo$, or equivalently $A_\Phi(\Phi,k_y)$, as in last section.
-But first lets analyze the entanglement occupancy spectrum of this system, shown in Fig.\ref{chern_insulator}(a) and (b) for $m=1,3$ respectively.
-For $0<m<2$ the EOS presents two edge modes that connect the occupied and empty bands that cross $\xi=1/2$ at $k_y=0$, in this region the system has a Chern number $C=+1$.
-For $2<m<4$ there are also edge modes but they cross $\xi=1/2$ at $k_y=\pi$ and the system has $C=-1$.
-For $m>4$ the system is a trivial insulator with $C=0$.
-When we open the chain one of the edge modes will be pushed to the bulk and will present a jump between $\xi=1,0$ whenever the edge-modes crossed zero energy.
-In this model, due to the symmetries, this is only possible at $k_y=0,\pi$.
-
-We can now take a look at $\tpo$ in the parameter space $(m,k_y)$, shown in figure \ref{chern_insulator}.(c).
-We see, again, that $\tpo$ remains continous except for the branch cuts that appear at the points where the physical edge-modes cross zero energy.
-The Chern number can be obtained as
-\begin{equation}
-C(m) = \int_{-\pi}^{\pi}\frac{dk_y}{2\pi} \partial_{k_y} \tpo(m,k_y),
-\label{eq:Zaletel_chern}
-\end{equation}
-However, since we know the position of the branch cuts we can integrate avoiding them and obtain
-\begin{align}
-C(m) =& [\tpo(m,k_y=0^-)-\tpo(m,k_y=-\pi^+)]\\
-&+[\tpo(m,k_y=\pi^-)-\tpo(m,k_y=0^+)].
-\end{align}
-
-\begin{figure}[t]
-	\centering
-	\makebox[0pt]{
-		\subfloat[Spectrum of $C_{A}$ for $m=1$]{
-			\includegraphics[width=45mm]{chernES0.pdf}
-		}
-		\subfloat[Spectrum of $C_{A}$ for $m=3$]{
-			\includegraphics[width=45mm]{chernESpi.pdf}
-		}
-	}\hspace{0mm}
-
-	\makebox[0pt]{
-		\subfloat[$\tpo-L/2$]{
-			\includegraphics[width=45mm]{cherntpo.pdf}
-		}
-		\subfloat[$\tilde{\mathcal{P}}$ mod 1]{
-			\includegraphics[width=45mm]{cherntp.pdf}
-		}
-	}\hspace{0mm}
-
-	\makebox[0pt]{
-		\subfloat[$\tpo-L/2$]{
-			\includegraphics[width=45mm]{cherntpo4.pdf}
-		}
-		\subfloat[$\tilde{\mathcal{P}}$ mod 1]{
-			\includegraphics[width=45mm]{cherntp4.pdf}
-		}
-	}
-
-	\caption{Polarization $\tpo$ computed for $t=1,\kappa'=0$ and system size $L=160$, for the SSH chain $t'=0$ in (a) and for $t'=-2$ in (b) CHANGE}
-	\label{chern_insulator}
-\end{figure}
-
-To ilustrate the advantage of using $\tpo$ instead of $\tilde{\mathcal{P}}$ we also show the latter in the same parameter space.
-In order to obtain the Chern number there is no way of avoiding having to compute the derivative and subsequent integration of
-\begin{align}
-C(m) =& [\tpo(m,k_y=0^-)-\tilde{\mathcal{P}}(m,k_y=-\pi^+)]\\
-&+[\tpo(m,k_y=\pi^-)-\tpo(m,k_y=0^+)],
-\end{align}
-as we no longer know in advance where the branch cuts will appear, due to it being defined modulo 1.
-This example seems a bit trivial as the Chern number is also easy to identify from the plot of $\tilde{\mathcal{P}}$ as well but in systems with higher Chern numbers $\tilde{\mathcal{P}}$ will wind around the circle several times which will make this more difficult.
-\carlos{Figures 7.e and f are computed for the same system where instead of nearest-neighbor hopping in the x direction we use a 4th neighbor hopping ($\delta_{i,j+1} \rightarrow \delta_{i,j+4}$), to illustrate the challenges of computing the Chern number when the integrand is only defined modulo 1.
-This is similar to the issues we had when doing the cut-average.
-One needs to identify whenever $\tilde{\mathcal{P}}$ (or $\chi$) crosses 1 or 0 to compute the derivative and integral properly.
-
-Note also that we are computing the Chern number in the torus defined by $(\Phi,k_y)$.
-It seems a bit weird to mix both flux and momentum but it works in the end.
-If we were to introduce another flux in the other direction the system would no longer be effectively 1D so I don't know if everything we do would still apply in the same way or if it would provide any advantage.
 }

### 2020-10-30 version2/draft.tex
Add files via upload

--- 2020-10-28 version2/draft.tex
@@
 The bulk polarization is a $\mathbb{Z}_2$ topological invariant for systems in one dimension with chiral or particle-hole symmetries.
-It is therefore related to the number of topological edge modes by the bulk-boundary correspondence.
 We show that the bulk polarization can always be decoded from the single-particle entanglement spectrum, even in the absence of symmetries that quantize it.
-From this relation the bulk-boundary correspondence can be derived in the presence of chiral and particle-hole symmetries.
-We use the bulk polarization to compute Chern numbers in 1D and 2D systems and show their relation to the entanglement spectrum.
+In the symmetric case the known relation between the bulk polarization and the number of topological edge modes is recovered.
+We use the bulk polarization to compute Chern numbers in 1D and 2D, their known relation to the entanglement spectrum is shown in a new way.
 Furthermore we propose an alternative bulk polarization that can carry more information about the surface spectrum than the conventional one and can simplify the calculation of Chern numbers.
@@
 
-Generically, any of the proposed tools provides some information about the phase naturally, while others seem inaccessible. 
+Generically, any of the proposed tools naturally provides only some information about the phase.
 It is not known, whether there exists a method to obtain the full topological content from the ground state.
@@
 For \emph{non-interacting} topological insulators/superconductors the ES for (gapped) periodic systems can be computed very efficiently, using methods developed by Peschel and others~\cite{Peschel2003}.
-The ES in these systems is equivalent to the band-flattened energy spectrum of the corresponding system with open boundaries~\cite{Fidkowski2010entanglement}.
+The ES in these systems is equivalent to the flat-band energy spectrum of the corresponding system with open boundaries~\cite{Fidkowski2010entanglement}.
 The same correspondence was also found for closely related gapless systems~\cite{Matern2018entanglement}.
 However, even for non-interacting systems, it is unclear which information (beyond the  `edge' spectrum) is encoded in the entanglement spectrum.
-In this manuscript, we show that the information about the Zak phase, or the bulk polarization, can be completely decoded from the entanglement spectrum and provide a simple formula to do so.
-We also show how to extend this method to compute the Chern number directly from the entanglement spectrum.
-In obtaining these topological invariants, the Zak phase and the Chern number, directly from the entanglement spectrum the \emph{bulk-boundary correspondence} is seen in an explicit way.
+In this manuscript, we show that the information about the Zak phase, or the bulk polarization, can be completely decoded from the entanglement spectrum and we provide a simple formula to do so.
+This is a general relation for non-interacting 1D gapped systems which, in the symmetric case where the bulk polarization is quantized, gives the known relation between the bulk polarization and the number of virtual topological edge states in the ES \cite{Ryu2006}.
+We also show how to extend this method to compute Chern numbers directly from the entanglement spectrum.
 
@@
 In Ref.~\cite{Zaletel2014}, the authors show that the Zak phase can be computed from the Schmidt decomposition of a translation invariant, infinite chain. 
-This demonstrates that there should be a direct relation between the single-particle entanglement energies and the Zak phase, although deriving that relation turned out to be non-trivial.
-A particular limit of our result for homogeneous (translationally invariant) systems~\eqref{eq:chi_integrable} was derived by Ryu and Hatsugai for a fully dimerized SSH chain~\cite{Ryu2006}.
-The similarity between the behavior of the Zak phase and the 'virtual edge modes' in the entanglement spectrum of certain Chern insulators, which was observed in Ref.~\cite{Huang2012,Huang2012-2}, can also be explained by our results.
-The bulk polarization is defined in $\mathbb{R}$ mod 1, which presents challenges when computing changes in the polarization, which are defined  in $\mathbb{R}$.
-We show how, using our method, one can define an alternative bulk polarization computed in the equivalent open chain that is defined in $\mathbb{R}$.
+This demonstrates that there should be a direct relation between the single-particle entanglement energies and the Zak phase for non-interacting systems.
+A particular limit of our result for homogeneous (translationally invariant) systems~\eqref{eq:chi_integrable} was derived
+%by Ryu and Hatsugai 
+for a fully dimerized SSH chain~\cite{Ryu2006}. 
+The similarity between the behavior of the Zak phase and the virtual topological edge states in the entanglement spectrum of certain Chern insulators, which was observed in Ref.~\cite{Huang2012,Huang2012-2}, can also be explained by our results.
+
+The bulk polarization is defined in $\mathbb{R}$ mod 1, which might present challenges when computing changes in the polarization, which are defined  in $\mathbb{R}$.
+We show how, using our method, one can define an alternative bulk polarization computed in the equivalent open chain that is continuous in $\mathbb{R}$ for gapped paths.
 The bulk polarization obtained this way can give more information about the edge spectrum than the usual one, and can be used to simplify the computation of Chern numbers.
@@
 
-\emph{Outline of the paper}: In section II we introduce the model used throughout the paper and we discuss its phase diagram.
-In section III we introduce the polarization, the different ways one can define it and its relation to the Zak phase.
-In section IV we introduce the entanglement spectrum and set the notation that will be used.
-In section V we introduce the method and show how it reproduces the Zak phase for both systems with and without translational invariance.
-In section VI we explain how the method applied to the opened chain allows us to compute Chern numbers for systems both in 1D and 2D.%and winding numbers
-and finally in section VII we present the conclusions and possible outlook.
+\emph{Outline of the paper}: In section II we introduce the 1D model used throughout the paper and we examine its phase diagram.
+In section III we discuss the bulk polarization, the different ways one can define it and its relation to the Zak phase.
+The ES is introduced in section IV, where we set the notation that will be used.
+In section V we present our method and show how we can reproduce the bulk polarization for both systems with and without translational invariance.
+In section VI we introduce an alternative bulk polarization constructed from the ES and use it to compute Chern numbers in 1D and 2D.
+Finally in section VII we conclude and discuss the possible outlook.
 
@@
 To showcase our results we consider as a starting point a model in the BDI class~\cite{Song2014}. 
-This is a simple model which nevertheless supports a rich phase diagram containing topological phases with winding numbers up to $\nu =2$. 
+This is a simple model which nevertheless supports a rich phase diagram containing topological phases with winding numbers $\nu =0,1$ and $2$. 
 In order to show that our results are not limited to any particular symmetry class~\cite{ryu2010topological}, we also include two symmetry breaking terms. 
@@
 \end{alignat}
-The corresponding phase diagram \cite{Song2014} is shown in Fig.\ref{fig:bdi_phase_diagram} for $t=1$. 
+The corresponding phase diagram \cite{Song2014} is shown in Fig.~\ref{fig:bdi_phase_diagram} for $t=1$.
+
 \begin{figure}[t]
@@
 \end{figure}
+
 The BDI class  is characterized by a $\mathbb{Z}$ invariant, the winding number,  in one spatial dimension.
@@
 The behavior of the  edge modes  is  shown in Fig.~\ref{fig:zero_E_modes}, using the dotted path ($t=1,t'=-2$) marked in the phase diagram. 
-Figure~\ref{fig:zero_E_modes}.a shows the edge spectrum in the BDI class, with 4 (resp. 2) symmetry-protected zero modes for winding number 2 (1). 
+Figure~\ref{fig:zero_E_modes}(a) shows the edge spectrum in the BDI class, with 4 (resp. 2) symmetry-protected zero modes for winding number 2 (1). 
 %We also a cut through the three phases (dotted line in the phase diagram) for different values of $\kappa,\kappa'$.
@@
 Consequently, adding such a term causes the zero modes in the $\nu=2$ phase to split  pairwise and the $\nu = 2$ and $\nu=0$ phases become equivalent. 
-This is shown in figure~\ref{fig:zero_E_modes}.b, where one clearly sees that the zero modes below $m=1$ are split.
+This is shown in Fig.~\ref{fig:zero_E_modes}(b), where one clearly sees that the zero modes below $m=1$ are split.
 
 For $\kappa \neq 0$ only time-reversal $T$ is preserved and the system is in the AI class. 
-In one dimensions, this class is trivial and, thus,  all zero-energy modes split. 
+In 1D this class is trivial and, thus,  all zero-energy modes split from zero energy. 
 With both $\kappa \neq 0$ and $\kappa' \neq0$ the system has no local symmetry, thus belonging to the $A$ class.  
 Also this symmetry class is topologically trivial in one dimension. 
-Figure~\ref{fig:zero_E_modes}.c and d show the edge spectrum for class AI and A, respectively. In both cases, the edge modes are split from zero for all values of $m$.
+Fig.~\ref{fig:zero_E_modes}(c) and (d) show the edge spectrum for class AI and A, respectively. In both cases, the edge modes are split from zero for all values of $m$.
 
@@
 }
-\caption{Edge spectrum along the dotted path (i.e. $t=1,t'=-2$) of Fig.~\ref{fig:bdi_phase_diagram} for different values of $\kappa$ and $\kappa'$.}
+\caption{Edge spectrum along the dotted path (i.e. $t=1,t'=-2$) of Fig.~\eqref{fig:bdi_phase_diagram} for different values of $\kappa$ and $\kappa'$.}
 \label{fig:zero_E_modes}
@@
 
-In Table \ref{fig:class_table} we summarize the different classes that are hosted by this model together with the correspondent symmetries and topological invariants in 1D.
+In Table \ref{fig:class_table} we summarize the different classes that this model can belong to for the different parameters, together with the correspondent symmetries and topological invariants in 1D.
 
@@
 \end{tabular}
-\caption{Different Cartan classes hosted by the model in Eq.\ref{bdi_model}, with the correspondent symmetries and topological invariants in 1D.} 
+\caption{Different Cartan classes for the model in Eq.~\eqref{bdi_model}, with the correspondent symmetries and topological invariants in 1D.} 
 \label{fig:class_table}
@@
 In this section we review some aspects of the modern theory of polarization, which gave a geometrical description of the polarization \cite{Resta1992,KingSmith1993,Vanderbilt1993,Resta1997}. 
-We also discuss the results of a recent article by Watanabe and Oshikawa \cite{Watanabe2018} that clarified the meanings of the different polarizations one can define this way, as well as the relations between them.
+We also discuss the results of a recent article by Watanabe and Oshikawa \cite{Watanabe2018} that clarified the meaning of different bulk polarizations that are found in the literature, as well as the relations between them.
 For clarity we will use the same notation they use.
@@
 
-The Zak phase is typically regarded as just a phase because it is only gauge invariant modulo $2\pi$.
+The Zak phase is only gauge invariant modulo $2\pi$.
 However, for two different states defined by a parameter $\lambda$ --- assuming $A_k(\lambda)$ is smooth in the path connecting them --- the change in the Zak phase can be computed as
@@
 \end{align}
- is the Berry curvature. The Berry curvature is fully gauge invariant and, therefore, the change in the Zak phase in Eq.\eqref{eq:change_zak_phase} is defined in $\mathbb{R}$. %?
+ is the Berry curvature. The Berry curvature is fully gauge invariant and, therefore, the change in the Zak phase in Eq.~\eqref{eq:change_zak_phase} is defined in $\mathbb{R}$. %?
 The Zak phase itself, as defined in Eq.~\eqref{eq:zak_phase}, can be shown to be proportional to the winding number for a certain gauge (see Appendix \ref{appendix:winding_ssh}), which means that it carries physical information beyond modulo $2\pi$.
@@
 One can also compute the geometric phase obtained by threading a $U(1)$ flux through the ring \cite{Watanabe2018}, which allows us to compute a polarization in the absence of translational invariance.
-If the flux is introduced homogeneously through a vector potential $A_x = \Phi/L$ an equivalent polarization can be defined as
+If the flux is introduced homogeneously through a vector potential $A_x = \Phi/L$ another polarization can be defined as
 \begin{equation}
@@
 
-There is one last polarization we must consider. 
+There is one last bulk polarization we must consider. 
 It is obtained when the flux is introduced in the bond between sites $j=1$ and $j=L$ via twisted boundary conditions. This is equivalent to performing the gauge transformation
@@
 \end{align}
-which makes $\ket{\tilde{\Psi}^\Phi}$ fully periodic in $\Phi$. We can now define the polarization
+which makes $\ket{\tilde{\Psi}^\Phi}$ fully periodic in $\Phi$. We can now define the bulk polarization
 \begin{equation}
@@
 \end{equation}
-which is related to the current flowing through this bond \cite{Watanabe2018}. This polarization is related to the homogeneous polarization by 
+which is related to the current flowing through this bond \cite{Watanabe2018}. This bulk polarization is related to the previous one by 
 \begin{equation}
@@
 Let us now proceed to discuss the entanglement spectrum. 
-An important quantity is the correlation matrix, which in position space is defined by
+An important quantity to consider is the correlation matrix, which in position space is defined by
 \begin{align}
@@
 \end{align}
-This form is particularly useful for all our subsequent calculations, since it does not require diagonalizing the Hamiltonian.
+This form, which we derive in Appendix~\ref{app:CM}, is particularly useful since it does not require diagonalizing the Hamiltonian and we will use it in all subsequent calculations.
 
@@
 The EOS is in one-to-one correspondence to the entanglement spectrum (ES). 
-Following Ref.~\cite{Peschel2003}, one can write the  reduced density matrix as 
+Following Ref.~\cite{Peschel2003,Peschel2008} , one can write the  reduced density matrix as 
 \begin{align}\label{eq:red_dens_mat}
@@
 where $\mathcal{K}$ is a normalization constant and $\mathcal{H}$ is a quadratic Hamiltonian, referred to as entanglement Hamiltonian. 
-The eigenvalues $\epsilon_j$ of $\mathcal{H}$, referred to as entanglement energies,  are related to those of the EOS by 
+The single-particle eigenvalues $\epsilon_j$ of $\mathcal{H}$, referred to as entanglement energies,  are related to those of the EOS by 
 \begin{align}\label{eq:xi_eps}
@@
 the corresponding eigenstates are the same. 
-Eq.~\eqref{eq:xi_eps} implies that the full information of the \emph{many-body} ES (MBES) is contained in the single-particle spectrum of the reduced correlation matrix. 
+Eq.~\eqref{eq:xi_eps} implies that the full information of the \emph{many-body} ES (MBES) is contained in the single-particle spectrum of the reduced correlation matrix for non-interacting systems. 
 The latter is much simpler to interpret. 
@@
 We also provide a simple formula in terms of the EOS eigenvalues that is identical to the Zak phase in the thermodyamic limit.  
-However, for a generic, finite size system, there will be finite-size discrepancies between our formula and the Zak phase, see figure \ref{huang}.d. 
+However, for a generic, finite size system, there will be finite-size discrepancies between our formula and the Zak phase, see figure \ref{huang}(d). 
 In order to reduce these, we are going to choose subsystem $A$ as half the system in the remainder of the manuscript.
@@
 % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % 
-\section{Zak phase in the EOS}
-
-Let us first review previous results on the relation between the entanglement occupancy spectrum and the Zak phase. 
-For symmetry-protected topological systems in one dimension, the quantized Zak phase is zero whenever there is an even number  of eigenvalues per vitual edge at $\xi = 1/2$, and $\pi$ when the number is odd~\cite{Peschel2008}.
+\section{Bulk Polarization in the EOS}
+
+Let us first review previous results on the relation between the entanglement occupancy spectrum and the bulk polarization (or Zak phase). 
+For systems in 1D protected by chiral or particle-hole symmetries the Zak phase is quantized.
+It is zero whenever there is an even number  of eigenvalues per vitual edge at $\xi = 1/2$, and $\pi$ when the number is odd~\cite{Asboth2016}.
 In the absence of symmetry-protection, much less is known. 
@@
 Although they do not specify to which of the two eigenvalues it corresponds.
-Their result is a particular limit of Eq.~\eqref{eq:chi_integrable}. 
+Their result is a particular limit of equation~\eqref{eq:chi_integrable}. 
 %
@@
 
-Huang and Arovas considered a two-dimensional Chern insulator and noted that there was a similarity between the Wannier centers (and therefore the Zak phase) and the `virtual edge mode' that connects the EOS values at 0 to those at 1~\cite{Huang2012,Huang2012-2}.
+In Ref.s~\cite{Huang2012,Huang2012-2} the authors considered a two-dimensional Chern insulator and noted that there was a similarity between the Zak phase and the virtual topological edge states that connect the EOS values at 0 to those at 1.
 Interpreting this as a one-dimensional system with a parameter, we can explain this behavior by Eq.~\eqref{eq:chi_integrable}, noting that there is a single pair of eigenvalues that dominates the sum.
-Note that the observation of Ref.s~\cite{Huang2012,Huang2012-2} is particular to systems with Chern number 0 or $\pm 1$, and it generically fails for systems with higher Chern numbers. 
-In the latter case, there are several terms in Eq.~\eqref{eq:chi_integrable} with comparably large contributions and, consequently, the Zak phase deviates considerably from the 'virtual edge modes'.
-% This last sentece is not entirely ok.
-They observe flow in the zak phase that correspond s to flow in the EOS.
-
-\subsubsection{Systems with equispaced entanglement energies}
+Note that the observation of Ref.s~\cite{Huang2012,Huang2012-2} is again particular to systems with few midgap states in the EOS, and fails whenever there are several terms in Eq.~\eqref{eq:chi_integrable} with comparably large contributions and, consequently, the Zak phase deviates considerably from the virtual topological edge states.
+
+\subsection{Systems with equispaced entanglement energies}
 
@@
 This is irrelevant in the case of symmetry-protected states where the Zak phase is quantized to $0$ or $\pi$.
-In the special case where there is only one eigenvalue per edge that is $\neq 0,1$ in  the EOS, like in the fully dimerized limit of the SSH chain, we recover the equality between one of the two midgap eigenvalues and the Zak phase found in Ref.~\cite{Ryu2006}.
-
-In Fig.~\ref{huang} we show  the EOS along the dashed cut in the phase diagram of Fig.~\ref{fig:bdi_phase_diagram}. 
+In the special case where there is only one eigenvalue per edge that is $\neq 0,1$ in  the EOS, like in the fully dimerized limit of the SSH chain, we recover the equality between one of the two midgap eigenvalues and the Zak phase found in reference ~\cite{Ryu2006}.
+
+In Fig.~\ref{huang} we show  the EOS along the dashed cut in the phase diagram of Fig.~\ref{fig:bdi_phase_diagram}, for $t'=0$. 
 Along this line, our model is equivalent to an SSH chain. 
@@
 Because of chiral symmetry the sum of the two eigenvalues related by the symmetry gives always $1$ so they do not contribute to $\chi$.
-The only contribution to $\chi$ is the one from the eigenstates with $\xi=1/2$ that do not have a chiral partner, see figure~\ref{huang}.a.
+The only contribution to $\chi$ is the one from the eigenstates with $\xi=1/2$ that do not have a chiral partner, see figure~\ref{huang}(a).
 Here we see explicitely the known relation between the Zak phase and the number of $\xi = 1/2$ modes.
@@
 When breaking chiral symmetry, the EOS will no longer be doubly degenerate. 
-For small $m$, there are two dominant midgap eigenvalues, and the Zak phase follows the lower one very closely, see figure~\ref{huang}.b.
-When this midgap eigenvalue approaches  0, the contribution from the other modes becomes more relevant and the Zak phase starts to deviate from the midgap eigenvalue, see figure \ref{huang}.c.
+For small $m$, there are two dominant midgap eigenvalues, and the bulk polarization follows the lower one very closely, see figure~\ref{huang}(b).
+When this midgap eigenvalue approaches  0, the contribution from the other modes becomes more relevant and the bulk polarization starts to deviate from the dominant eigenvalue, see figure~\ref{huang}(c).
+In Fig.~\ref{huang}(d) we show how the difference between $\chi$ and the bulk polarization decreases with system size, implies that they are equal in the thermodynamic limit.
 
@@
 \includegraphics[width=86mm]{fig3comp.pdf}
-\caption{In (a) and (b)  we show the EOS (black), $\chi$ (green) and $\tilde{\mathcal{P}}^{\rm Bloch}$(red) for a cut in the phase diagram through the $\nu = 1 \rightarrow \nu = 0$ transition (dashed line in~\ref{fig:bdi_phase_diagram}), $t = 1,t'=0$, for parameters  (a) $\kappa =\kappa'=0$ and (b) $\kappa = 0.3, \kappa'=0$.
+\caption{In (a) and (b)  we show the EOS (black), $\chi$ (green) and $\tilde{\mathcal{P}}^{\rm Bloch}$ (red) for a cut in the phase diagram through the $\nu = 1 \rightarrow \nu = 0$ transition (dashed line in Fig.~\ref{fig:bdi_phase_diagram}, for $t = 1,t'=0$), for parameters  (a) $\kappa =\kappa'=0$ and (b) $\kappa = 0.3, \kappa'=0$.
 Computed for $L=40$.
@@
 
-\subsubsection{General case}
-
-In general we find that it is only the eigenstates that localize on the left edge that should contribute to $\chi$.
+\subsection{General case}
+
+In general we find that it is only the eigenstates that localize on the left edge ($A_L$ in Fig.\ref{fig:chain}) that should contribute to $\chi$.
 The inclusion of bulk modes is irrelevant since they have $\xi = 0,1$.
@@
 the sum is performed over the subspace $L$ formed by the eigenstates with $\expval{\hat{x}}<L/4$, where $L/4$ is the middle point of region $A$.
-It does not matter how we choose the eigenvalues we sum over as long as we do not include the ones related to the right edge with $\xi$ away from $1$.
+It does not matter what threshold we use as we include all left-edge states and exclude all right-edge states.
 In practice only a few eigenvalues give a significant contribution to $\chi$.
 
-In the thermodynamic limit, it is $\tilde{\mathcal{P}}$ that we actually identify with $\chi$
+In the thermodynamic limit, it is $\tilde{\mathcal{P}}$ we identify with $\chi$
 \begin{equation}
@@
 This case is trivial, as we know that there is one left and one right eigenstate but if the degeneracy is four-fold or higher one needs to obtain the localized eigenstates.
-Localizing the eigenstates that are exponentially close to $\xi = 0,1$ is more difficult but not doing so only results in an exponentially small error.
+Localizing the bulk eigenstates that are exponentially close to $\xi = 0,1$ is more difficult but not doing so only results in an exponentially small error.
 
@@
 This issue can be fixed by determining the localization of a single eigenstate.
-MinMax{]
 
@@
 
-In Fig.\ref{2} we show two examples that highlight the importance of the localization structure of the EOS. 
-In figures \ref{2}.a and \ref{2}.b, we compute the EOS for varying $m$ at  $t=1$, $t'=2$, indicated by the solid line  in the phase-diagram in Fig.~\ref{fig:bdi_phase_diagram}, in the presence of two different symmetry breaking terms: (a) $\kappa =0.3$, $\kappa'=0$ and (b) $\kappa=0$, $\kappa'=0.3$.
+In Fig.~\ref{2} we show two examples that highlight the importance of the localization structure of the EOS. 
+In figures~\ref{2}(a) and (b), we compute the EOS for varying $m$ at  $t=1$, $t'=2$, indicated by the solid line  in the phase-diagram in Fig.~\ref{fig:bdi_phase_diagram}, in the presence of two different symmetry breaking terms: (a) $\kappa =0.3$, $\kappa'=0$ and (b) $\kappa=0$, $\kappa'=0.3$.
 For $\kappa=\kappa'=0$, there is a phase transition $\nu = 2 \rightarrow 0 $ at $m=2$, whereas both choices of symmetry breaking render the system trivial  for all $m$.
@@
 While $\kappa'$ splits the four-fold degenerate states into two two-fold degenerate pairs at opposite edges whose contribution cancels (modulo 1), $\kappa$ split them into two $left-left$ and $right-right$ pairs, with a finite contribution to the Zak phase.
-In figures \ref{2}.c and d, we consider the same symmetry breaking terms, but now for $t=1$ and $t'=-2$, where the BDI systems shows a phase transition  $\nu=2\rightarrow 1\rightarrow 0$. 
-Note that the $\kappa'$ term does not render the $\nu=1$ phase trivial, and the Zak phase will remain quantized in that region, see Fig.~\ref{2}.d. 
-Also here, Eq.~\ref{eq:chi_left} correctly reproduces the Zak phase for all values of $m$. 
-Note that in all examples shown in this subsection, the observations made in references \cite{Huang2012,Huang2012-2}, i.e.
-that the behavior of the Zak phase resembles that of the entanglement occupancy midgap states, do not apply anymore.
+In figures~\ref{2}(c) and (d), we consider the same symmetry breaking terms, but now for $t=1$ and $t'=-2$, where the BDI systems shows a phase transition  $\nu=2\rightarrow 1\rightarrow 0$. 
+Note that the $\kappa'$ term does not render the $\nu=1$ phase trivial, and the Zak phase will remain quantized in that region, see figure~\ref{2}(d). 
+Also here, Eq.~\eqref{eq:chi_left} correctly reproduces the Zak phase for all values of $m$. 
+Note that in the $\nu=2$ phase the observation made in Ref. \cite{Huang2012}, i.e. that the there is a similarity between the Zak phase and the entanglement occupancy, does not apply anymore.
 
 Since $\chi$ is naturally defined in position space our results are also valid for systems without translationally invariance.
-We show this in Fig.\ref{2}.e where we show the EOS, $\chi$, $\mathcal{P}$ and $\tilde{\mathcal{P}}$ along the dotted cut in the phase diagram with an added position dependent symmetry breaking term $\kappa_i = (i+L/4 \, {\rm mod}\, L)/L$.
+We show this in Fig.~\ref{2}(e) where we show the EOS, $\chi$, $\mathcal{P}$ and $\tilde{\mathcal{P}}$ along the dotted cut in the phase diagram with an added position dependent symmetry breaking term $\kappa_i = (i+L/4 \, {\rm mod}\, L)/L$.
 We use this profile for $\kappa_i$ instead of simple disorder because with the disorder the difference between $\mathcal{P}$ and $\tilde{\mathcal{P}}$ is very small.
-We obtain that $\chi$ is indeed identified with $\tilde{\mathcal{P}}$ as we hinted at before, and not $\mathcal{P}$, although these two polarization only differ by $(L+1)/2$ in the translationally invariant case.
+We obtain that $\chi$ is indeed identified with $\tilde{\mathcal{P}}$ as we hinted at before, and not $\mathcal{P}$, although these two polarization only differ by $(L+1)/2$ in the translationally invariant case for this model.
 
@@
 
-\section{Chern number and Winding number}
-
-As mentioned above the bulk polarization is only defined modulo 1.
+\section{Alternative Bulk Polarization and Chern Numbers}
+
+As mentioned above, the bulk polarization is only defined modulo 1.
 This is obvious when we express it in terms of a geometric phase, which is only defined modulo $2\pi$.
@@
 Another issue due to it being defined in $\mathbb{R}$ mod $1$ appears when we try to follow its evolution over a path $C_\lambda$.
-Between any two points $\lambda$ and $\lambda'$ in this path it is not possible to tell if the bulk polarization has increased or decreased by only looking at $\tilde{\mathcal{P}}(\lambda)$, this becomes relevant when computing Chern numbers.
+Between any two points $\lambda$ and $\lambda'$ in this path it is not possible to tell if the bulk polarization has increased or decreased by only looking at $\tilde{\mathcal{P}}(\lambda)$ at these specific points, this becomes relevant when computing Chern numbers.
 
@@
 
-The idea is that in the process of opening the chain, which can be done adiabatically, all eigenvalues that are related to the right virtual edge should get pushed to $\xi=0,1$.
-As a result, the sum of all eigenvalues of $C_{A,o}$ is equal, modulo 1 and in the thermodynamic limit, to $\chi$ (See Eq.~\eqref{eq:chi_left}).
-The advantage of using $\tpo$ is that the contribution from the eigenvalues far from the virtual edge is always constant, as opposed to the case of $\chi$ where this is only true when taken modulo 1.
+The idea is that in the process of opening the chain, which can be done adiabatically, all eigenvalues that are related to the right virtual edge ($A_R$ in Fig.~\ref{fig:chain}) should get pushed to $\xi=0,1$.
+As a result, the sum of all eigenvalues of $C_{A,o}$ is equal, modulo 1 and in the thermodynamic limit (to avoid the finite size effects of opening the chain), to $\chi$ (See Eq.~\eqref{eq:chi_left}).
+Therefore, in the thermodynamic limit, we have
+\begin{equation}
+\tpo = \tilde{\mathcal{P}} \, {\rm mod} \, 1.
+\end{equation}
+
+The advantage of using $\tpo$ is that the contribution from the eigenvalues far from the left virtual edge is constant as long as no states cross the Fermi energy, as opposed to the case of $\chi$ where this is only true when taken modulo 1.
 The consequence is that for paths in parameter space, $C_\lambda$, which are gapped, i.e.
@@
 Note, however, that $\tpo$ is still gauge-invariant only when taken modulo 1.
+This is similar to the case of the Zak phase, which is known to be equal to the winding number for certain gauges (see Appendix~\ref{appendix:winding_ssh}).
+It encodes more information than what is accesible when considered modulo $2\pi$.
 
@@
 
-\subsubsection{SSH chain}
+\subsection{SSH chain}
 
@@
 Unless $\tpo(0) = L/2$, which is the result expected for a trivial insulator, the result will be discontinuous at $\kappa = 0$.
-We can see this in Fig.\ref{ssh_chern}.a where we show $\tpo$ in the parameter space $(m,\kappa)$.
+We can see this in Fig.~\ref{ssh_chern}(a) where we show $\tpo$ in the parameter space $(m,\kappa)$.
 The presence of the zero-mode for $m<1$ appears in $\tpo$ as a branch cut.
-The relation between the zero-modes and $\Tr[C_{A,o}]$ was already observed in the context of Chern insulators \cite{Alexandrinata2011}.
+The appearance of a branch cut in the $(m,\kappa)$ paramater space is natural but its position depends on the choice of gauge (see Appendix~\ref{appendix:winding_ssh}).
+
+The relation between the zero-modes and discontinuities in $\Tr[C_{A,o}]$ was already observed in the context of Chern insulators \cite{Alexandrinata2011}.
 This branch cut is in fact not an artifice of the open chain trick, but the result of the degenerate point at $(m=1,\kappa=0)$ having an associated Chern number $C=1$ \cite{Asboth2016}.
@@
 \end{align}
-where $C(m)$ is the Chern number of any loop around the degenerate point crossing $\kappa=0$ at $m$.
-Using the symmetry of the polarization described above this results in 
+where $C(m)$ is the Chern number of any loop around the degenerate point crossing $\kappa=0$ at $m$ and another point in the trivial region ($m>1$).
+Using the symmetry of the polarization described above gives
 \begin{align}
@@
 \end{align}
-which results in $+1$ when $m<1$ and $0$ when $m>1$, as it is known for the SSH chain.
-The only contribution to the Chern number comes from the discontinuity of $\tpo$ at the branch cut.
-In general the Chern number over a loop is equal to the sum of the discontinuities of $\tpo$ across the branch cuts.
+which results in $+1$ when $m<1$ and $0$ when $m>1$, as it is known for the SSH chain. The only contribution to the Chern number comes from the discontinuity of $\tpo$ at the branch cut.
 
@@
 \includegraphics[width=86mm]{fig5comp.pdf}
-\caption{Polarization $\tpo-L/2$ computed for $t=1,\kappa'=0$.
-(a) and (b) are computed for $L=40$ sites and parameters $t'=0$ and $t'=-2$, respectively.
+\caption{(a) and (b) show the bulk polarization $\tpo-L/2$ computed for $t=1,\kappa'=0$ and $L=40$ sites, with (a) $t'=0$ and (b) $t'=-2$.
 $\tpo$ presents branch cuts that indicate the presence of zero-energy modes.
 $\tpo$ is able to differentiate between the phases $\nu=2$ and $\nu=0$.
-(c) is computed for $t'=2$, with an added disorder to $t$ and $m$ with strength $W=3$, and $\kappa$ is applied only at the edges.
-We compute it for $L=400$, as the disorder increases the correlation length.
-The disorder opens a region with $\nu=1$ as it has been observed in \cite{Song2014} }
+In (c) we show $\tilde{\mathcal{P}}$ mod $1$ for the same parameters of (b) to illustrate the advantage of using $\tpo$.
+}
 	\label{ssh_chern}
@@
 
-% \begin{figure}[t]
-%	\centering
-%	\makebox[0pt]{
-%		\subfloat[$2\tilde{\mathcal{P}}_{\rm open}-L$($m$ should go from 0 to 3)]{
-%			\includegraphics[width=45mm]{fig61.pdf}
-%		}
-%		\subfloat[$C $]{
-%			\includegraphics[width=45mm]{fig62.pdf}
-%		}		
-%	}\hspace{0mm}
-%	
-%	\makebox[0pt]{
-%		\subfloat[${\rm Log}_{10} (C \, {\rm mod} \, 1)$ ]{
-%			\includegraphics[width=45mm]{fig63.pdf}
-%		}
-%	}
-%	\caption{We compute (a)$\tilde{\mathcal{P}}_{\rm open}$, (b) $\nu$ and (c) error in the quantization of $\nu$ for the continuous line in the phase diagram, $t=1,t'=2$ and system size $L=400$.
-(a) is computed for $W=6$ showing that the branch cut is not affected by the disorder.
-(b) and (c) are computed for increasing disorder strength, showing a good quantization of the resulting winding number far from the transitions.
-}
-%	\label{winding}
-%\end{figure}
-
-\subsubsection{Second neighbor hopping, $t' \neq 0$}
+\subsection{$\nu = 2$ and disordered case. }
 
@@
 This is no longer true when $t' \neq 0$ allowing for a $\nu = 2$ phase.
-In Fig.\ref{ssh_chern}.b we see $\tpo$ for the dotted cut in the phase diagram with an added symmetry breaking $\kappa$ term.
+In Fig.~\ref{ssh_chern}(b) we see $\tpo$ for the dotted cut in the phase diagram with an added symmetry breaking $\kappa$ term.
+We can see that $\tpo$ is continuous for any path avoiding the branch cut even though the range of values for $\tpo$ is larger than 1.
 We can see that a loop around the two degenerate points will have a Chern number of $C=2$.
@@
 It is also interesting to point out that even far away from the symmetric point $\tpo$ still shows plateaus that point to a possible nearby topological transition.
-Note , however, that there are many ways of breaking chiral symmetry.
+Note , however, that there are many ways of breaking the chiral symmetry.
 Replacing $\kappa$ in Eq.~\eqref{bdi_model} by $\kappa_i=\kappa (-1)^i$, the zero-energy modes split in such a way that $\tpo$ is continuous at $\kappa=0$.
 It cannot differenciate between the phases $\nu=0$ and $\nu=2$, and a loop like the one discussed above results in a Chern number $C=0$.
+To illustrate the advatange of $\tpo$ we also show $\tilde{\mathcal{P}}$ mod $1$ in the $(m,\kappa)$ parameter space in figure~\ref{disorder_chern}(a).
+As mentioned above the only contribution to the Chern number comes from discontinuities in the bulk polarization, however when it is computed using $\tilde{\mathcal{P}}$ we no longer know in advance where this discontinuities are present, due to it being defined modulo 1.
+One needs to look at the full path to know how the bulk polarization evolves along the loop.
 
@@
 where $\omega_i$ and $\omega_i'$ are selected from a uniform distribution with range $[-1/2,1/2]$.
-For very strong disorder where the gap fills with states $\tpo$ will present many branch cuts as we approach the limit $\lim_{\kappa \rightarrow 0}$, which makes it unpractical.
+For very strong disorder where the gap fills with bulk states $\tpo$ will present many branch cuts as we approach the limit $\lim_{\kappa \rightarrow 0}$, which makes it unpractical.
 In order to avoid it we break chiral symmetry locally in the edge of the open chain so it only affects the edge modes.
 If any of the other states cross $E=0$ they do it in  chiral-symmetric pairs such that $\tpo$ is not affected by it.
-We show $\tpo$ in Fig.\ref{ssh_chern}.c using the local symmetry breaking term for a strong disorder, $W=3t$, where the gap is indeed filled with states.
+We show $\tpo$ in Fig.\ref{disorder_chern}(a) using the local symmetry breaking term for a strong disorder, $W=6t$, where the gap is indeed filled with states.
 We see that it works just as in the case without disorder, where now due to the disorder a region of $\nu=1$ opens up \cite{Song2014}.
 The main difference is that since we break chiral symmetry only for the edge modes, $\tpo$ is quantized to multiples of $1/2$.
-
-%In Fig.\ref{winding}.(b) we show $C$ for the continuous line in the phase diagram and increasing disorder strength.
-The Chern number as defined in \eqref{eq:Chern_number} indeed provides a topological invariant which can be compared with the winding number obtained in reference \cite{Song2014} for the same system.
-Note that for large enough systems the spectrum of $C_A$ is not affected by the local chiral symmetry breaking term other than the occupation of the zero-energy modes such that $C$ is still quantized for each disorder configuration, however near the transitions it is smoothen by the disorder average.
-We show the error in the quantization of $C$ in figure \ref{winding}.(c) .
-\carlos{Fig.6.a seems important to discuss how with local symmetry braking term the relation between the chern number and the number of zero modes still remains.
-Maybe it would be enough to show a cut of 6.b }
-
-% - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - %
-
-\section{Chern Number in 2D}
-
-We have shown how $\tpo$ simplifies the computation of Chern numbers in 1D chains. This can be generalized to 2D systems. Consider the following model for a Chern insulator
-\begin{align}
-H =& \sum_{im\alpha,jn\beta} c_{im\alpha}^\dagger H_{im\alpha,jn\beta} c_{jn\beta}, \\
-H_{im\alpha,jn\beta}=& \frac{1}{2}(i\sigma_x-\sigma_z)\delta_{i,j+1}\delta_{mn}+\frac{1}{2}(-i\sigma_x-\sigma_z)\delta_{i,j-1}\delta_{mn} \nonumber\\
-+&\frac{1}{2}(i\sigma_y -\sigma_z)\delta_{ij}\delta_{m,n+1} +\frac{1}{2}(-i\sigma_y -\sigma_z)\delta_{ij}\delta_{m,n-1} \nonumber\\
-+&(2-m)\sigma_z\delta_{ij}\delta_{mn},
-\label{eq:chern_model}
-\end{align}
-where $i,j$ describe the $x$ direction and $m,n$ the $y$ direction.
+Note that, since the states that fill the gap are localized at the bulk they will not appear in the EOS and one can still identify the appearance of virtual topological edge states (see Fig.~\ref{disorder_chern}(b)), as oposed to topological edge states in the energy spectrum.
+
+\begin{figure}[t]
+\includegraphics[width=86mm]{fig8comp.pdf}
+\caption{(a) Bulk polarization $\tpo$ computed for $t=1,t'=2,\kappa'=0$, with an added disorder to $t$ and $m$ with strength $W=3$, and $\kappa$ is applied only at the edges.
+We compute it for $L=400$, as the disorder increases the finite size  effect.
+The disorder opens a region with $\nu=1$ as it has been observed in \cite{Song2014}.
+(b) EOS for the same parameters as (a), with $\kappa=0$, where we have highlighted the 4 eigenvalues closest to $\xi = 1/2$ to show the phase transitions.
+As opposed to the energy spectrum, the gap in the EOS does not fill with states such that the virtual topological edge states are still easily identifiable.
+}
+	\label{disorder_chern}
+\end{figure}
+
+\subsection{Chern Number in 2D}
+%
+%We have shown how $\tpo$ simplifies the computation of Chern numbers in 1D chains. This can be generalized to 2D systems. Consider the following model for a Chern insulator with inversion symmetry,
+%\begin{align}
+%H =& \sum_{im\alpha,jn\beta} c_{im\alpha}^\dagger H_{im\alpha,jn\beta} c_{jn\beta}, \\
+%H_{im\alpha,jn\beta}=& \frac{1}{2}(i\sigma_x-\sigma_z)\delta_{i,j+1}\delta_{mn}+\frac{1}{2}(-i\sigma_x-\sigma_z)\delta_{i,j-1}\delta_{mn} \nonumber\\
+%+&\frac{1}{2}(i\sigma_y -\sigma_z)\delta_{ij}\delta_{m,n+1} +\frac{1}{2}(-i\sigma_y -\sigma_z)\delta_{ij}\delta_{m,n-1} \nonumber\\
+%+&(2-m)\sigma_z\delta_{ij}\delta_{mn},
+%\label{eq:chern_model}
+%\end{align}
+%where $i,j$ describe the $x$ direction and $m,n$ the $y$ direction.
 The system with periodic boundary conditions is defined on a torus.
@@
 We then fourier transform only the $y$ direction to obtain
+We consider now the computation of Chern numbers in 2D systems with $\tpo$. Consider the Chern insulator with inversion symmetry given by the Hamiltonian
 \begin{align}
@@
 We can treat the Hamiltonian with elements $H_{i\alpha,j\beta}(k_y)$ as a 1D Hamiltonian with an extra parameter $k_y$ and compute $\tpo$, or equivalently $A_\Phi(\Phi,k_y)$, as in last section.
-But first lets analyze the entanglement occupancy spectrum of the system with PBC, shown in Fig.\ref{chern_insulator}.a and b for $m=1,3$ respectively.
+But first lets analyze the entanglement occupancy spectrum of the system with PBC, shown in Fig.~\ref{chern_insulator}(a) and (b) for $m=1$ and $3$, respectively.
 For $0<m<2$ the EOS presents two edge modes that connect the occupied and empty bands that cross $\xi=1/2$ at $k_y=0$, in this region the occupied bands have a Chern number $C=+1$.
@@
 For $m>4$ the system is a trivial insulator with $C=0$.
-When we open the chain the eigenvalue of one of the virtual edge modes will be pushed to $\xi=1,0$ as seen in figure~\ref{chern_insulator}.a and b.
+When we open the chain the eigenvalue of the right virtual topological edge mode will be pushed to $\xi=1,0$ as seen in figure~\ref{chern_insulator}(a) and (b).
 It will have a jump whenever the correspondent edge-mode crossed zero energy, going between $\xi=1$ and $\xi=0$, i.e.
-the number of eigenvalues at $\xi = 1$ is not constant across $k$.
-In this model, due to the symmetries, this is only possible at $k_y=0,\pi$.
+the number of eigenvalues at $\xi = 1$ is not constant across $k_y$.
+In this model, due to inversion symmetry, this is only possible at $k_y=0,\pi$.
 
@@
 \centering
-	\makebox[0pt]{
-		\subfloat[EOS for $m=1$]{
-			\includegraphics[width=45mm]{fig71.pdf}
-		}
-		\subfloat[EOS for $m=3$]{
-			\includegraphics[width=45mm]{fig72.pdf}
-		}
-	}\hspace{0mm}
-
-	\makebox[0pt]{
-		\subfloat[$\tpo-L/2$]{
-			\includegraphics[width=45mm]{fig73.pdf}
-		}
-		\subfloat[$\tilde{\mathcal{P}}$ mod 1]{
-			\includegraphics[width=45mm]{fig74.pdf}
-		}
-	}
-%	\hspace{0mm}
-%	
-%	\makebox[0pt]{
-%		\subfloat[$\tpo-L/2$]{
-%			\includegraphics[width=45mm]{fig75.pdf}
-%		}
-%		\subfloat[$\tilde{\mathcal{P}}$ mod 1]{
-%			\includegraphics[width=45mm]{fig76.pdf}
-%		}
-%	}
-%	
-	\caption{(a) and (b) EOS.
-The evolution of the edge state towards $\xi=1,0$ as the chain is opened adiabatically is shown in two intermediate points.
-(c) $\tpo$ and (d) $\tilde{\mathcal{P}}$.
-Computed for the Chern insulator given by the model in Eq.~\eqref{eq:chern_model}.
+\includegraphics[width=86mm]{fig7comp.pdf}
+\caption{(a) and (b) EOS of the Chern insulator in the $C=1$ ($m=1$) and $C=-1$ ($m=3$) phases, respectively.
+In gray and light gray lines we plot two intermediate points of the adiabatic process of opening the chain between sites $1$ and $L$, showing how the midgap state of the left virtual cut evolves towards the bulk bands, introducing a discontinuity in momentum.
+(c) $\tpo-L/2$ and (d) $\tilde{\mathcal{P}}$ mod $1$.
+Computed for the Chern insulator given by the model in Eq.~\eqref{eq:chern_model} for $L=40$ sites.
 }
@@
 
-We can now take a look at $\tpo$ in the parameter space $(m,k_y)$, shown in figure~\ref{chern_insulator}.c .
+We can now take a look at $\tpo$ in the parameter space $(m,k_y)$, shown in figure~\ref{chern_insulator}(c) .
 We see, again, that $\tpo$ remains continous except for the branch cuts that appear at the points where the physical edge-modes cross zero energy.
@@
 In reference \cite{Alexandrinata2011} it is shown to be equal to the Chern number by relating it to the Hall current.
-By understanding $\Tr[C_{A,o}]$ as a geometric quantity the connection between the trace index and the Chern number is made more explicit.
-
-To ilustrate the advantage of using $\tpo$ over $\tilde{\mathcal{P}}$ we also show the latter in the same parameter space.
+Here, the connection between the trace index and the Chern number is seen by framing $\Tr[C_{A,o}]$ as a geometric quantity
+
+%
+%
+%To ilustrate the advantage of using $\tpo$ over $\tilde{\mathcal{P}}$ we also show the latter in the same parameter space.
 In order to obtain the Chern number via $\tilde{\mathcal{P}}$ for PBC there is no way of avoiding having to compute the derivative and subsequent integration of
-\begin{equation}
-C(m) = \int_{-\pi}^{\pi}\frac{dk_y}{2\pi} \partial_{k_y} \tilde{\mathcal{P}}(m,k_y),
-\end{equation}
-as we no longer know in advance where the branch cuts will appear, due to it being defined modulo 1.
-
-% This example might seem insufficient as the Chern number is also easy to identify from the plot of $\tilde{\mathcal{P}}$ as well but in systems with higher Chern numbers $\tilde{\mathcal{P}}$ will wind around the circle several times which will make this more difficult.
-We exemplify this in figures \ref{chern_insulator}.(e) and (f).
-To obtain these we change the nearest neighbor hoping in the $x$ direction into a third neighbor hopping, $\delta_{i,j+1} \rightarrow \delta_{i,j+3}$.
-This results in the same phase diagram with the Chern number, and the number of edge modes, multiplied by 3.
-As we see, while the plot of $\tpo$ remains simple, $\tilde{\mathcal{P}}$ becomes more complex, with the appearance of additional branch cuts.
-
+%\begin{equation}
+%C(m) = \int_{-\pi}^{\pi}\frac{dk_y}{2\pi} \partial_{k_y} \tilde{\mathcal{P}}(m,k_y),
+%\end{equation}
+%as we no longer know in advance where the branch cuts will appear, due to it being defined modulo 1.
 %
-%   \carlos{Figures 7.e and f are computed for the same system where instead of nearest-neighbor hopping in the x direction we use a 4th neighbor hopping ($\delta_{i,j+1} \rightarrow \delta_{i,j+4}$), to illustrate the challenges of computing the Chern number when the integrand is only defined modulo 1.
-This is similar to the issues we had when doing the cut-average.
-One needs to identify whenever $\tilde{\mathcal{P}}$ (or $\chi$) crosses 1 or 0 to compute the derivative and integral properly.
-%
-%Note also that we are computing the Chern number in the torus defined by $(\Phi,k_y)$.
-It seems a bit weird to mix both flux and momentum but it works in the end.
-If we were to introduce another flux in the other direction the system would no longer be effectively 1D so I don't know if everything we do would still apply in the same way or if it would provide any advantage.
-}
 
@@
 
-It was previously known that the Zak phase, or the geometric phase for $U(1)$ flux insertion, was encoded in the many-body entanglement spectrum.
-Here we have shown how it is encoded in the single-body entanglement spectrum.
-This relation is simpler than the one involving the MBES and it provides more insight into the ES and EOS, as well as it provides another simple method for computing the Zak phase for systems without translational invariance.
-Based on this method we show that one can obtain a polarization that is defined in the real axis.
-Although it is not gauge invariant it provides additional information and also simplifies the method for computing Chern numbers for systems in 1D without translational invariance or systems in 2D with translational invariance in only one direction.
+It was previously known that the bulk polarization (Zak phase, or geometric phase for $U(1)$ flux insertion) was encoded in the many-body entanglement spectrum.
+Here we have shown how it can be decoded from the single-particle entanglement spectrum.
+This relation is simpler than the one involving the MBES and it provides more insight into the ES and EOS.
+In particular in the topological case, where the relation between the quantized bulk polarization (or the Chern number in 2D) and the number of virtual topological edge states is shown in a new way.
+It also provides another simple method for computing the bulk polarization for systems without translational invariance.
+
+Based on this method we show that one can obtain an alternative bulk polarization that is continuous in $\mathbb{R}$ for gapped paths, although it is only gauge invariant when taken modulo 1, where it is equal to the conventional bulk polarization.
+It provides additional information about the edge spectrum, similar to how the Zak phase is equal to the winding number when computed in a particular gauge.
+This alternative bulk polarization simplifies the calculation of changes in the bulk polarization, with a particular application to the computation of Chern numbers.
+
+In the application of our results to topological systems we have focused on systems with translational invariance, both clean and with disorder.
+In general spatially inhomogeneous systems, not necesseraly with disorder, the bulk-boundary correspondence can appear in more interesting ways, such as in the case of quasicrystals in 1D \cite{Lang2012,Kraus2012}, which show properties characteristic of 2D topological insulators.
+In these systems our results might be able to give more insight into the relation between bulk and boundary.
+Another extension of our work can be in the field of higher-order topological insulators, where some systems can be characterized by a physical generalization of the bulk polarization to higher dimensions \cite{Benalcazar2017,Kang2019}.
+Our results for 1D might be able to be generalized to higher dimensions as well.
 
@@
 
-\section{Computing the Zak phase from the Schmidt decomposition}\label{app:pollmann}
-
-In this appendix, we review  how to compute the Zak phase from the Schmidt decomposition~\cite{Zaletel2014}and relate it to our own results.
+\section{Computing the Bulk polarization from the Schmidt decomposition}\label{app:pollmann}
+
+In this appendix we review  how to compute the bulk polarization from the Schmidt decomposition~\cite{Zaletel2014} and relate it to our own results.
 
@@
 \end{equation}
-The Schmidt indeces $p$ and $q$ label the fluctuations at the cuts in $i=1$ and $i=L/2$, respectively.
+The Schmidt indeces $p$ and $q$ label the fluctuations at the left ($i=1$) and right ($i=L/2$) cuts, respectively.
 The convention used for the states $\ket{p,q}_{R}$ is that the first (second) index labels the state near the left-most (right-most) region of $R$, where $R$ can be $A$ or $B$.
@@
 we will refer to these subspaces of the Hilbert space as $A_L$ ($A_R$).
-Note that the remaining subspace $A_{\rm bulk}$ is composed of eigenstates that have eigenvalues $\xi = 0,1$.
-The eigenvalues of $\rho_A$, $\lambda_{pq}=s_p^2 s_q^{\prime \, 2}$ can be obtained as
+Note that the remaining subspace $A_{\rm bulk}$ is composed of eigenstates that have eigenvalues exponentially close to $\xi = 0,1$.
+The eigenvalues of $\rho_A$, $\lambda_{pq}=s_p^2 s_q^{\prime \, 2}$ can be obtained as \cite{Alexandrinata2011}
 \begin{align}
@@
 where $N_{A_L}^p = \sum_{\mu \in A_L}n_\mu^p$.
+Note that there is an ambiguity in how the flux is introduced as one can always include bulk modes into $A_L$.
 As mentioned above, the label $p$ describes the set of occupation numbers $\{n_\mu^p\}$.
@@
 \tilde{\mathcal{P}} =& \sum_{\substack{\{n_{i \neq p \in A_L}\} = 0,1 \\ n_p=0}} \left(\sum_{j\neq p \in A_{L}} n_j \right) (1-\xi_p)\left(\prod_{k \neq p\in A_L} \lambda_k \right) \\
-&+ \sum_{\substack{\{n_{i \neq p \in A_L}\} = 0,1 \\ n_p=0}} \left(\sum_{j\neq p \in A_{L}} n_j+1 \right)\xi_p \left(\prod_{k \neq p\in A_L} \lambda_k \right) \\
+&+ \sum_{\substack{\{n_{i \neq p \in A_L}\} = 0,1 \\ n_p=1}} \left(\sum_{j\neq p \in A_{L}} n_j+1 \right)\xi_p \left(\prod_{k \neq p\in A_L} \lambda_k \right) \\
 =& \sum_{\substack{\{n_{i \neq p \in A_L}\} = 0,1 }} \left(\sum_{j\neq p \in A_{L}} n_j \right) \left(\prod_{k \neq p\in A_L} \lambda_k \right) \\
@@
 \end{align}
-which we can see is $1$ if we continue this procedure for all other occupation numbers. Therefore in terms of the spectrum of $C_A$ the first term of the polarization greatly simplifies to 
+which we can see is $1$ if we continue this procedure for all other occupation numbers. Therefore in terms of the spectrum of $C_A$ the bulk polarization greatly simplifies to 
 \begin{align}
@@
 \end{align}
-The subspace $A_L$ is not well-defined, as one can always include bulk modes.
-As a result the polarization obtained this way is defined only modulo 1.
-In practice we extend $A_L$ to include all eigenstates whose average position lies in the left half of $A$ and we finally get
+As mentioned above, the subspace $A_L$ is not well-defined, as one can always include bulk modes, however since the bulk polarization is defined modulo 1, this issue is irrelevant.
+In practice we extend $A_L$ to include all eigenstates whose average position lies in the left half of $A$ (which we denote by $L$) and we finally get
 \begin{align}
@@
 In this appendix, we show how to express the correlation matrix in terms of the Hamiltonian, used in Eq.~\eqref{eq:corr_mat2}. 
-We consider the generic, quadratic Hamiltonian in one dimension of Eq.~\ref{eq:quadr_Ham}, which is diagonalized by a unitary matrix $U$ with 
+We consider the generic, quadratic Hamiltonian in one dimension of Eq.~\eqref{eq:quadr_Ham}, which is diagonalized by a unitary matrix $U$ with 
 	\begin{align}\label{eq:D}
@@
 \end{align}
-The first term of the last line is simply a kronecker delta between both sets of indices.
-
-The second term can be rewritten, using Eq.~\ref{eq:D}, as
+The first term of the last line is simply a kronecker delta between both sets of indices. The second term can be rewritten, using Eq.~\eqref{eq:D}, as
 \begin{align}
@@
 \end{equation}
-and the resulting Zak phase can be seen in Fig.\ref{fig:ssh_zak}, which presents a branch cut for $m=0,\kappa<0$ for this particular gauge.
+and the resulting Zak phase can be seen in Fig.~\ref{fig:ssh_zak}, which presents a branch cut for $m=0,\kappa<0$ for this particular gauge.
 
@@
 	\centering
-	\includegraphics[width=50mm]{sshwindinganalytical.pdf}
+	\includegraphics[width=50mm]{figApp.pdf}
 	\caption{$\gamma/2\pi$ obtained in the smooth gauge that provides the winding number in the chiral symmetric limit.}

### 2020-10-30 version2/draft.tex
Final lic version

@@
 
-In the application of our results to topological systems we have focused on systems with translational invariance, both clean and with disorder.
-In general spatially inhomogeneous systems, not necesseraly with disorder, the bulk-boundary correspondence can appear in more interesting ways, such as in the case of quasicrystals in 1D \cite{Lang2012,Kraus2012}, which show properties characteristic of 2D topological insulators.
-In these systems our results might be able to give more insight into the relation between bulk and boundary.
-Another extension of our work can be in the field of higher-order topological insulators, where some systems can be characterized by a physical generalization of the bulk polarization to higher dimensions \cite{Benalcazar2017,Kang2019}.
-Our results for 1D might be able to be generalized to higher dimensions as well.
+\emph{Outlook:} In the application of our results to topological systems in 1D we have focused on systems with translational invariance even though our results do not require it.
+For general spatially inhomogeneous systems the same bulk can present different boundaries such that the relation between bulk and boundary is not as straighforward as in the translational invariant case.
+Our results might be able to give more insight into this relation.
+
+In the field of higher-order topological insulators some systems can be characterized by a physical generalization of the bulk polarization to higher dimensions \cite{Benalcazar2017,Kang2019}.
+There is an open question wether a formulation in terms of twisted boundary conditions can be obtained, in which case our results might be able to generalize to higher dimensions.
 

### 2020-10-31 version2/draft.tex
Add files via upload

@@
 \title{On the relation of the entanglement spectrum to the bulk polarization \\ DRAFT}
-\author{Carlos Ortega Taberner}
+\author{Carlos Ortega-Taberner}
 \affiliation{Department of Physics, Stockholm University, AlbaNova University Center, SE-106 91 Stockholm, Sweden}
@@
 %The previous discussion about the Zak phase being defined only modulo $2\pi$ although its changes can be defined as a real number applies in the same way to the polarization obtained in \eqref{eq:p_tilde}.
-Similar to the discussion about the Bloch polarization below Eq.~\eqref{eq:Bloch_polarization}, both $\mathcal{P}$ and $\tilde{\mathcal{P}}$ are defined modulo 1, while their changes are defiend in $\mathbb{R}$.
+Similar to the discussion about the Bloch polarization below Eq.~\eqref{eq:Bloch_polarization}, both $\mathcal{P}$ and $\tilde{\mathcal{P}}$ are defined modulo 1, while their changes are defined in $\mathbb{R}$.
 For translationally invariant systems one finds that $\tilde{\mathcal{P}} = \tilde{\mathcal{P}}^{\rm Bloch}$ \cite{Watanabe2018}.
@@
 In the following sections, we show how the Zak phase can be recovered from the EOS.
-We also provide a simple formula in terms of the EOS eigenvalues that is identical to the Zak phase in the thermodyamic limit.  
+We also provide a simple formula in terms of the EOS eigenvalues that is identical to the Zak phase in the thermodynamic limit.  
 However, for a generic, finite size system, there will be finite-size discrepancies between our formula and the Zak phase, see figure \ref{huang}(d). 
@@
 For systems in 1D protected by chiral or particle-hole symmetries the Zak phase is quantized.
-It is zero whenever there is an even number  of eigenvalues per vitual edge at $\xi = 1/2$, and $\pi$ when the number is odd~\cite{Asboth2016}.
+It is zero whenever there is an even number  of eigenvalues per virtual edge at $\xi = 1/2$, and $\pi$ when the number is odd~\cite{Asboth2016}.
 In the absence of symmetry-protection, much less is known. 
@@
 The only contribution to $\chi$ is the one from the eigenstates with $\xi=1/2$ that do not have a chiral partner, see figure~\ref{huang}(a).
-Here we see explicitely the known relation between the Zak phase and the number of $\xi = 1/2$ modes.
+Here we see explicitly the known relation between the Zak phase and the number of $\xi = 1/2$ modes.
 
@@
 First note that for translation invariant systems, the ES of the right and the left virtual edge are related by an overall '-' sign, i.e.
-for each eigenvalue $\epsilon$ on the right virutal edge, there is a corresponding value $-\varepsilon$ on the left one.
+for each eigenvalue $\epsilon$ on the right virtual edge, there is a corresponding value $-\varepsilon$ on the left one.
 Since the spectra is equispaced, this implies that the two virtual edge spectra are merely shifted with respect to each other.
@@
 This is similar to the case of the Zak phase, which is known to be equal to the winding number for certain gauges (see Appendix~\ref{appendix:winding_ssh}).
-It encodes more information than what is accesible when considered modulo $2\pi$.
+It encodes more information than what is accessible when considered modulo $2\pi$.
 
@@
 The presence of the zero-mode for $m<1$ appears in $\tpo$ as a branch cut.
-The appearance of a branch cut in the $(m,\kappa)$ paramater space is natural but its position depends on the choice of gauge (see Appendix~\ref{appendix:winding_ssh}).
+The appearance of a branch cut in the $(m,\kappa)$ parameter space is natural but its position depends on the choice of gauge (see Appendix~\ref{appendix:winding_ssh}).
 
@@
 Replacing $\kappa$ in Eq.~\eqref{bdi_model} by $\kappa_i=\kappa (-1)^i$, the zero-energy modes split in such a way that $\tpo$ is continuous at $\kappa=0$.
-It cannot differenciate between the phases $\nu=0$ and $\nu=2$, and a loop like the one discussed above results in a Chern number $C=0$.
-To illustrate the advatange of $\tpo$ we also show $\tilde{\mathcal{P}}$ mod $1$ in the $(m,\kappa)$ parameter space in figure~\ref{disorder_chern}(a).
+It cannot differentiate between the phases $\nu=0$ and $\nu=2$, and a loop like the one discussed above results in a Chern number $C=0$.
+To illustrate the advantage of $\tpo$ we also show $\tilde{\mathcal{P}}$ mod $1$ in the $(m,\kappa)$ parameter space in figure~\ref{disorder_chern}(a).
 As mentioned above the only contribution to the Chern number comes from discontinuities in the bulk polarization, however when it is computed using $\tilde{\mathcal{P}}$ we no longer know in advance where this discontinuities are present, due to it being defined modulo 1.
@@
 The main difference is that since we break chiral symmetry only for the edge modes, $\tpo$ is quantized to multiples of $1/2$.
-Note that, since the states that fill the gap are localized at the bulk they will not appear in the EOS and one can still identify the appearance of virtual topological edge states (see Fig.~\ref{disorder_chern}(b)), as oposed to topological edge states in the energy spectrum.
+Note that, since the states that fill the gap are localized at the bulk they will not appear in the EOS and one can still identify the appearance of virtual topological edge states (see Fig.~\ref{disorder_chern}(b)), as opposed to topological edge states in the energy spectrum.
 
@@
 We can now take a look at $\tpo$ in the parameter space $(m,k_y)$, shown in figure~\ref{chern_insulator}(c) .
-We see, again, that $\tpo$ remains continous except for the branch cuts that appear at the points where the physical edge-modes cross zero energy.
+We see, again, that $\tpo$ remains continuous except for the branch cuts that appear at the points where the physical edge-modes cross zero energy.
 The Chern number can be obtained as
@@
 \emph{Outlook:} In the application of our results to topological systems in 1D we have focused on systems with translational invariance even though our results do not require it.
-For general spatially inhomogeneous systems the same bulk can present different boundaries such that the relation between bulk and boundary is not as straighforward as in the translational invariant case.
+For general spatially inhomogeneous systems the same bulk can present different boundaries such that the relation between bulk and boundary is not as straightforward as in the translational invariant case.
 Our results might be able to give more insight into this relation.
@@
 In the field of higher-order topological insulators some systems can be characterized by a physical generalization of the bulk polarization to higher dimensions \cite{Benalcazar2017,Kang2019}.
-There is an open question wether a formulation in terms of twisted boundary conditions can be obtained, in which case our results might be able to generalize to higher dimensions.
+There is an open question whether a formulation in terms of twisted boundary conditions can be obtained, in which case one might be able to generalize the results of this paper to higher dimensions.
 
@@
 \end{equation}
-The Schmidt indeces $p$ and $q$ label the fluctuations at the left ($i=1$) and right ($i=L/2$) cuts, respectively.
+The Schmidt indices $p$ and $q$ label the fluctuations at the left ($i=1$) and right ($i=L/2$) cuts, respectively.
 The convention used for the states $\ket{p,q}_{R}$ is that the first (second) index labels the state near the left-most (right-most) region of $R$, where $R$ can be $A$ or $B$.
@@
 It gives the correct results in translationally invariant systems, however it is very unpractical.}).
-Note that a generalization of \eqref{eq:winding_k} to position space has also been developed which not necesseraly gives the same results as \eqref{eq:winding_flux} (\carlos{The only comparison is in the Prodan paper where they talk about practical numerical implementations but don't adress if both formulations are equivalent.}) Both formulations however reduce to \eqref{eq:winding_k} for translationally invariant systems \cite{Gong2018}.
+Note that a generalization of \eqref{eq:winding_k} to position space has also been developed which not necessarily gives the same results as \eqref{eq:winding_flux} (\carlos{The only comparison is in the Prodan paper where they talk about practical numerical implementations but don't address if both formulations are equivalent.}) Both formulations however reduce to \eqref{eq:winding_k} for translationally invariant systems \cite{Gong2018}.
 In the following we will omit the flux parameter for simplicity.
@@
 \end{align}
-We can not explicitely construct the correlation matrix with flux as
+We can not explicitly construct the correlation matrix with flux as
 \begin{align*}

### 2020-11-03 version2/draft.tex
Final lic version
Changed some notation

@@
 \begin{align}
-\sum_{j\beta}H_{ij,\alpha\beta} u_{p\mu}^{j\beta} = E_{p\mu} u^{i\alpha}_{p\mu},
-\end{align}
-where $[U]_{j\beta,p\mu} = u_{p\mu}^{j\beta}$ is the unitary matrix that diagonalizes $H$.
-In translational invariant systems $u_{k\mu}^{j\beta} = e^{ikj}\tilde{u}_{k \mu}^{\beta}$, where $\tilde{u}_{k \mu}^{\beta}$ are the components of the eigenstates of the Bloch Hamiltonian, $\ket{\tilde{u}_{k\mu}}$ and $k$ is the momentum.
-One can now compute the Zak phase \cite{Zak1989}, which for a two-band  model is defined as the geometric phase acquired by the occupied state $\ket{\tilde{u}_k}$ as it winds around the Brillouin zone,
+\sum_{j\beta}H_{ij,\alpha\beta} \psi_{p\mu}^{j\beta} = E_{p\mu} \psi^{i\alpha}_{p\mu},
+\end{align}
+where $[U]_{j\beta,p\mu} = \psi_{p\mu}^{j\beta}$ is the unitary matrix that diagonalizes $H$.
+In translational invariant systems $\psi_{k\mu}^{j\beta} = e^{ikj}u_{k \mu}^{\beta}$, where $u_{k \mu}^{\beta}$ are the components of the eigenstates of the Bloch Hamiltonian, $\ket{u_{k\mu}}$ and $k$ is the momentum.
+For simplicity we consider a unit cell with a single site.
+One can now compute the Zak phase \cite{Zak1989}, which for a two-band  model is defined as the geometric phase acquired by the occupied state $\ket{u_k}$ as it winds around the Brillouin zone,
 \begin{equation}
@@
 \end{equation}
-where $A_k = \bra{\tilde{u}_k}\partial_k \ket{\tilde{u}_k}$ is the Berry connection \cite{Berry1984}. 
+where $A_k = \bra{u_k}\partial_k \ket{u_k}$ is the Berry connection \cite{Berry1984}. 
 The generalization to higher-band models is straightforward.
@@
 \begin{equation}\label{eq:Bloch_polarization}
-\partial_\lambda \tilde{\mathcal{P}}^{\rm Bloch} = \int_{0}^{2\pi} \frac{dk}{2\pi} \, \Omega_{\lambda k},
-\end{equation}
-which allows us to make the identification $ \tilde{\mathcal{P}}^{\rm Bloch} = \gamma/2\pi$. 
+\partial_\lambda \mathcal{P}^{\rm Bloch} = \int_{0}^{2\pi} \frac{dk}{2\pi} \, \Omega_{\lambda k},
+\end{equation}
+which allows us to make the identification $ \mathcal{P}^{\rm Bloch} = \gamma/2\pi$. 
 As discussed above,  this polarization is defined only modulo 1.
@@
 \begin{equation}
-S_{p\mu,q\nu} = \sum_{j\alpha} u_{p\mu}^{j \alpha \, \ast} e^{i\frac{2\pi}{L}j}u_{q\nu}^{j \alpha},
+S_{p\mu,q\nu} = \sum_{j\alpha} \psi_{p\mu}^{j \alpha \, \ast} e^{i\frac{2\pi}{L}j}\psi_{q\nu}^{j \alpha},
 \end{equation}
@@
 Similar to the discussion about the Bloch polarization below Eq.~\eqref{eq:Bloch_polarization}, both $\mathcal{P}$ and $\tilde{\mathcal{P}}$ are defined modulo 1, while their changes are defined in $\mathbb{R}$.
-For translationally invariant systems one finds that $\tilde{\mathcal{P}} = \tilde{\mathcal{P}}^{\rm Bloch}$ \cite{Watanabe2018}.
+For translationally invariant systems one finds that $\mathcal{P} = \mathcal{P}^{\rm Bloch}+\frac{L-1}{2}\nu$, here $\nu$ is the number of occupied bands \cite{Watanabe2018}.
 
@@
 \begin{align}
-& \gamma_{i\alpha} = \sum_{j\beta}u_{j\beta}^{i\alpha \, \ast} c_{j\beta},
-%& c_{i\alpha} = \sum_{j\beta} u^{j\beta}_{i\alpha} \gamma_{j\beta}
+& \gamma_{i\alpha} = \sum_{j\beta}\psi_{j\beta}^{i\alpha \, \ast} c_{j\beta},
+%& c_{i\alpha} = \sum_{j\beta} \psi^{j\beta}_{i\alpha} \gamma_{j\beta}
 \end{align}
@@
 \begin{align}\label{eq:corr_mat}
-C_{ij}^{\alpha \beta} =& \sum_{pq,\mu\nu} u^{q\nu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} \expval{\gamma^\dagger_{p\mu} \gamma_{q\nu} }\nonumber \\
-=&  \sum_{p\mu} u^{p\mu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} \expval{\gamma^\dagger_{p\mu} \gamma_{p\mu} }\nonumber \\
-=&  \sum_{p\mu} u^{p\mu }_{j\beta} u^{p\mu \, \ast}_{i\alpha} [1-{\rm sign}(E_{p\mu})]/2.
+C_{ij}^{\alpha \beta} =& \sum_{pq,\mu\nu} \psi^{q\nu }_{j\beta} \psi^{p\mu \, \ast}_{i\alpha} \expval{\gamma^\dagger_{p\mu} \gamma_{q\nu} }\nonumber \\
+=&  \sum_{p\mu} \psi^{p\mu }_{j\beta} \psi^{p\mu \, \ast}_{i\alpha} \expval{\gamma^\dagger_{p\mu} \gamma_{p\mu} }\nonumber \\
+=&  \sum_{p\mu} \psi^{p\mu }_{j\beta} \psi^{p\mu \, \ast}_{i\alpha} [1-{\rm sign}(E_{p\mu})]/2.
 \end{align}

### 2021-01-12 version2/draft_mh.tex
partial changes

--- 2020-11-03 version2/draft.tex
@@
 \begin{abstract}
-The bulk polarization is a $\mathbb{Z}_2$ topological invariant for systems in one dimension with chiral or particle-hole symmetries.
+The bulk polarization is a $\mathbb{Z}_2$ topological invariant characterizing non-interacting systems in one dimension with chiral or particle-hole symmetries.
 We show that the bulk polarization can always be decoded from the single-particle entanglement spectrum, even in the absence of symmetries that quantize it.
-In the symmetric case the known relation between the bulk polarization and the number of topological edge modes is recovered.
-We use the bulk polarization to compute Chern numbers in 1D and 2D, their known relation to the entanglement spectrum is shown in a new way.
-Furthermore we propose an alternative bulk polarization that can carry more information about the surface spectrum than the conventional one and can simplify the calculation of Chern numbers.
+In the symmetric case, the known relation between the bulk polarization and the number of topological edge modes is recovered.
+We use the bulk polarization to compute Chern numbers in 1D and 2D, wcich illuminates their known relation to the entanglement spectrum  in a new way.
+Furthermore we discuss an alternative bulk polarization that can carry more information about the surface spectrum than the conventional one and, thus, can simplify the calculation of Chern numbers.
 \end{abstract}
@@
 There are few tools to identify topological phases. 
-A very efficient one is the entanglement entropy~\cite{Kitaev2006topological, Levin2006detecting}, which allows one to identify the total quantum dimension of the underlying topological quantum field theory.
+A very efficient one is the entanglement entropy~\cite{srednicki1993entropy}, which allows one to identify the total quantum dimension of the underlying topological quantum field theory~\cite{Kitaev2006topological, Levin2006detecting}.
 As such, the entanglement entropy can only be used for interacting systems. 
@@
 
-For \emph{non-interacting} topological insulators/superconductors the ES for (gapped) periodic systems can be computed very efficiently, using methods developed by Peschel and others~\cite{Peschel2003}.
+For \emph{non-interacting} topological insulators  and superconductors the ES for (gapped) periodic systems can be computed very efficiently, using methods developed by Peschel and others~\cite{Peschel2003}.
 The ES in these systems is equivalent to the flat-band energy spectrum of the corresponding system with open boundaries~\cite{Fidkowski2010entanglement}.
 The same correspondence was also found for closely related gapless systems~\cite{Matern2018entanglement}.
-However, even for non-interacting systems, it is unclear which information (beyond the  `edge' spectrum) is encoded in the entanglement spectrum.
-In this manuscript, we show that the information about the Zak phase, or the bulk polarization, can be completely decoded from the entanglement spectrum and we provide a simple formula to do so.
-This is a general relation for non-interacting 1D gapped systems which, in the symmetric case where the bulk polarization is quantized, gives the known relation between the bulk polarization and the number of virtual topological edge states in the ES \cite{Ryu2006}.
+However, even for non-interacting systems, it is unclear which information (beyond the protected `edge' spectrum) is encoded in the entanglement spectrum.
+In this manuscript, we show that the information about the Zak phase, or alternatively the bulk polarization, can be completely decoded from the entanglement spectrum and we provide a simple formula to do so.
+This is a general property of non-interacting 1D gapped systems which, in the symmetric case where the bulk polarization is quantized, gives the known relation between the bulk polarization and the number of virtual topological edge states in the ES \cite{Ryu2006}.
 We also show how to extend this method to compute Chern numbers directly from the entanglement spectrum.
@@
 The bulk polarization is defined in $\mathbb{R}$ mod 1, which might present challenges when computing changes in the polarization, which are defined  in $\mathbb{R}$.
-We show how, using our method, one can define an alternative bulk polarization computed in the equivalent open chain that is continuous in $\mathbb{R}$ for gapped paths.
+We show how, using our method, one can define an alternative bulk polarization \mariac{REF?} computed in the equivalent open chain that is continuous in $\mathbb{R}$ for gapped paths.
 The bulk polarization obtained this way can give more information about the edge spectrum than the usual one, and can be used to simplify the computation of Chern numbers.
@@
 	\includegraphics[width=66mm]{fig1.pdf}
-	\caption{Phase diagram for the system in the $BDI$ class showing the winding number of each region.
-Plotted for parameters $t = 1,\kappa =\kappa'=0$.
-Three cuts of the phase diagram are shown with continuous, dashed and dotted lines which will be used in the rest of the figures.}
+	\caption{Phase diagram for the system in the $BDI$ class showing the winding number of each region --- plotted for parameters $t = 1,\kappa =\kappa'=0$.
+Three cuts of the phase diagram, which will be used in later figures,  are shown with continuous, dashed and dotted lines.}
 \label{fig:bdi_phase_diagram}
@@
 
-The BDI class  is characterized by a $\mathbb{Z}$ invariant, the winding number,  in one spatial dimension.
-When open boundary conditions are imposed on the system, the number of symmetry-protected zero energy edge modes is equal to the winding number. 
+The BDI class  in one spatial dimension  is characterized by a $\mathbb{Z}$ invariant, the winding number.
+When open boundary conditions are imposed on the system, the number of symmetry-protected zero-energy edge modes is equal to the winding number. 
 The behavior of the  edge modes  is  shown in Fig.~\ref{fig:zero_E_modes}, using the dotted path ($t=1,t'=-2$) marked in the phase diagram. 
@@
 where $[U]_{j\beta,p\mu} = \psi_{p\mu}^{j\beta}$ is the unitary matrix that diagonalizes $H$.
-In translational invariant systems $\psi_{k\mu}^{j\beta} = e^{ikj}u_{k \mu}^{\beta}$, where $u_{k \mu}^{\beta}$ are the components of the eigenstates of the Bloch Hamiltonian, $\ket{u_{k\mu}}$ and $k$ is the momentum.
+In translational invariant systems $\psi_{k\mu}^{j\beta} = e^{ikj}u_{k \mu}^{\beta}$, where $u_{k \mu}^{\beta}$ are the components of the eigenstates, $\ket{u_{k\mu}}$, of the Bloch Hamiltonian and $k$ is the momentum.
 For simplicity we consider a unit cell with a single site.
-One can now compute the Zak phase \cite{Zak1989}, which for a two-band  model is defined as the geometric phase acquired by the occupied state $\ket{u_k}$ as it winds around the Brillouin zone,
+One can now compute the Zak phase \cite{Zak1989}.
+ For a two-band  model, it is defined as the geometric phase acquired by the occupied state $\ket{u_k}$ as it winds around the Brillouin zone,
 \begin{equation}
@@
 
-The derivative of the polarization with respect to $\lambda$ was obtained \cite{KingSmith1993} as
+The derivative of the polarization with respect to $\lambda$ was obtained in Ref.~\cite{KingSmith1993} as
 \begin{equation}\label{eq:Bloch_polarization}
@@
 which allows us to make the identification $ \mathcal{P}^{\rm Bloch} = \gamma/2\pi$. 
-As discussed above,  this polarization is defined only modulo 1.
-Its changes, however, are defined in $\mathbb{R}$. The polarization itself is not a physical observable. Only its changes, related to the current flowing through the system, are.
+As discussed above,  this polarization is defined only modulo 1, while its changes are  defined in $\mathbb{R}$.
+Note also that the polarization itself is not a physical observable.
+Only its changes, related to the current flowing through the system, are.
 
@@
 If the flux is introduced homogeneously through a vector potential $A_x = \Phi/L$ another polarization can be defined as
-\begin{equation}
+\begin{equation}\label{polarization_inhom}
 \mathcal{P} = \int_0^{2\pi}\frac{d\Phi}{2\pi} i\bra{\Psi^\Phi}\partial_\Phi \ket{\Psi^\Phi} + \frac{1}{2\pi}\Im\ln \bra{\Psi^0}e^{2\pi i \hat{P}} \ket{\Psi^{2\pi}}, 
 \end{equation}
-where $\ket{\Psi^{\Phi}}$ is the ground state in the presence of flux and $\hat{P}=\frac{1}{L}\sum_{j\alpha} j\hat{n}_{j\alpha}$ is the polarization operator, where $\hat{n}_{j\alpha}$ is the number operator.
+where $\ket{\Psi^{\Phi}}$ is the ground state in the presence of flux and $\hat{P}=\frac{1}{L}\sum_{j\alpha} j\hat{n}_{j\alpha}$ is the polarization operator, with $\hat{n}_{j\alpha}$ being the number operator.
 This is equivalent to the polarization obtained by Resta \cite{Resta1997} as
@@
 \end{align}
-This expression is particularly useful because it can be easily expressed in terms of single-particle eigenstates as
+
+Expression \eqref{polarization_inhom} is particularly useful because it can be easily expressed in terms of single-particle eigenstates as
 \begin{align}
@@
 There is one last bulk polarization we must consider. 
-It is obtained when the flux is introduced in the bond between sites $j=1$ and $j=L$ via twisted boundary conditions. This is equivalent to performing the gauge transformation
+It is obtained when the flux is not introduced homogeneously, but instead only in the bond between sites $j=1$ and $j=L$ via twisted boundary conditions.
+This is equivalent to performing the gauge transformation
 \begin{align}\label{eq:Psi_tilde}
@@
 Similar to the discussion about the Bloch polarization below Eq.~\eqref{eq:Bloch_polarization}, both $\mathcal{P}$ and $\tilde{\mathcal{P}}$ are defined modulo 1, while their changes are defined in $\mathbb{R}$.
-For translationally invariant systems one finds that $\mathcal{P} = \mathcal{P}^{\rm Bloch}+\frac{L-1}{2}\nu$, here $\nu$ is the number of occupied bands \cite{Watanabe2018}.
+For translationally invariant systems one finds that $\mathcal{P} = \mathcal{P}^{\rm Bloch}+\frac{L-1}{2}\nu$, where here $\nu$ denotes the number of occupied bands \cite{Watanabe2018}.
 
@@
 Let us now proceed to discuss the entanglement spectrum. 
-An important quantity to consider is the correlation matrix, which in position space is defined by
+An important quantity to consider is the correlation matrix, which in position space is defined by the groujd state expectation value
 \begin{align}
@@
 
-In this manuscript, we are mainly interested in the spectrum of the correlation matrix, when restricted to a spatial sub-region $A$ (its complement will be denoted by $B$ in the following). 
+In this manuscript, we are mainly interested in the spectrum of the correlation matrix, when restricted to a spatial sub-region $A$ (its complement will be denoted by $B$ in the following), see Fig.~\ref{fig:chain}.
 Following Ref.s~\cite{Huang2012,Huang2012-2}, we refer to this spectrum as the entanglement occupancy spectrum (EOS).
 We denote the eigenvalues of the EOS by $\xi_j$. 
-Topological phases are characterized by (symmetry-protected) zero-energy modes in the edge spectrum and $\xi=\frac 1 2$ modes in the EOS~\cite{Fidkowski2010entanglement}.
-
-The EOS is in one-to-one correspondence to the entanglement spectrum (ES). 
+Topological phases are characterized by (symmetry-protected) zero-energy modes in the edge spectrum and $\xi=\frac 1 2$ modes in the EOS. 
+Consequently, the EOS is in one-to-one correspondence to the entanglement spectrum (ES)~\cite{Fidkowski2010entanglement}.
+
 Following Ref.~\cite{Peschel2003,Peschel2008} , one can write the  reduced density matrix as 
@@
 We also provide a simple formula in terms of the EOS eigenvalues that is identical to the Zak phase in the thermodynamic limit.  
-However, for a generic, finite size system, there will be finite-size discrepancies between our formula and the Zak phase, see figure \ref{huang}(d). 
+However, for a generic, finite size system, there will be finite-size discrepancies between our formula and the Zak phase --- for a finite-size scaling analysis see figure \ref{huang}(d). 
 In order to reduce these, we are going to choose subsystem $A$ as half the system in the remainder of the manuscript.

### 2021-01-12 version2/draft_mh.tex
changes and comments

@@
 
-A few comments are needed regarding  bipartition. 
+A few comments are needed regarding our choice of  bipartition. 
 In the following sections, we show how the Zak phase can be recovered from the EOS.
 We also provide a simple formula in terms of the EOS eigenvalues that is identical to the Zak phase in the thermodynamic limit.  
-However, for a generic, finite size system, there will be finite-size discrepancies between our formula and the Zak phase --- for a finite-size scaling analysis see figure \ref{huang}(d). 
+However, for a generic, finite-size system, there will be finite-size discrepancies between our formula and the Zak phase --- for a finite-size scaling analysis see figure \ref{huang}(d). 
 In order to reduce these, we are going to choose subsystem $A$ as half the system in the remainder of the manuscript.
@@
 % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % 
-\section{Bulk Polarization in the EOS}
-
-Let us first review previous results on the relation between the entanglement occupancy spectrum and the bulk polarization (or Zak phase). 
+\section{Bulk Polarization in the EOS}\label{sec:bulkpolarization}
+
+Let us first review previous results on the relation between the EOS and the bulk polarization (or Zak phase). 
 For systems in 1D protected by chiral or particle-hole symmetries the Zak phase is quantized.
@@
 Ryu and Hatsugai considered the fully dimerized SSH chain with broken chiral symmetry~\cite{Ryu2006}.  
-This model is rather special in that there is only a single pair of eigenvalues in the EOS that is not strictly identical to $0$ or $1$, they are $\xi$ and $1-\xi$, related due to translational symmetry.
-The authors could show that the value of one of these `midgap states' is identical to the Zak phase divided by $2\pi$ or, alternatively, $\tilde{\mathcal{P}}^{\rm Bloch}$.
-Although they do not specify to which of the two eigenvalues it corresponds.
+This model is rather special in that there is only a single pair of eigenvalues in the EOS that is not strictly identical to $0$ or $1$: these are $\xi$ and $1-\xi$, related due to translational symmetry.
+The authors could show that the value of one of these `midgap states' is identical to the Zak phase divided by $2\pi$ --- or alternatively $\tilde{\mathcal{P}}^{\rm Bloch}$ --- although they do not specify to which of the two eigenvalues it corresponds.
 Their result is a particular limit of equation~\eqref{eq:chi_integrable}. 
@@
 Interpreting this as a one-dimensional system with a parameter, we can explain this behavior by Eq.~\eqref{eq:chi_integrable}, noting that there is a single pair of eigenvalues that dominates the sum.
-Note that the observation of Ref.s~\cite{Huang2012,Huang2012-2} is again particular to systems with few midgap states in the EOS, and fails whenever there are several terms in Eq.~\eqref{eq:chi_integrable} with comparably large contributions and, consequently, the Zak phase deviates considerably from the virtual topological edge states.
+Note that the observation of Ref.s~\cite{Huang2012,Huang2012-2} is  particular to systems with Chern number 0 or $\pm 1$, i.e. where there are only few midgap states in the EOS. 
+It fails for systems with higher Chern numbers,  for which there are several terms in Eq.~\eqref{eq:chi_integrable} with comparably large contributions and, consequently, the Zak phase deviates considerably from the virtual topological edge states.
 
@@
 That is, they are given in the thermodynamic limit by $\varepsilon_{n\alpha} = \varepsilon_{0\alpha}+n\delta_\alpha$, where $\alpha = L,R$.
-This is a feature of integrable systems \cite{Peschel1999} for e.g.
-nearest neighbor hopping models such as the SSH chain, and applies to the Hamiltonian in Eq.~\eqref{bdi_model} when $t'=0$.
+This is a feature of integrable systems \cite{Peschel1999},  e.g. nearest neighbor hopping models such as the SSH chain, and it applies to the Hamiltonian in Eq.~\eqref{bdi_model} when $t'=0$. 
 In this case, one can obtain the Zak phase in a very simple fashion: We reorder the eigenvalues of the EOS by magnitude, $\xi_1 < \xi_2 < ...< \xi_{L_AM}$, and compute
@@
 \end{equation}
-This generalizes the observation made by Ryu and Hatsugai \cite{Ryu2006} beyond the fully dimerized limit, $m=0$, but the ambiguity remains. This issue will be resolved in the next section. 
+This generalizes the observation made by Ryu and Hatsugai \cite{Ryu2006} beyond the fully dimerized limit, $m=0$.
+
+\mariac{Rewrite paragraph}
+The issue concerning the ambiguity will be discussed and resolved in the next section. 
 For translational invariant systems the EOS is symmetric such that a crossing of the eigenvalues $\xi$ results in $\chi$ changing sign.
@@
 Along this line, our model is equivalent to an SSH chain. 
-In the presence of chiral and translation symmetry, all eigenvalues of the EOS are doubly degenerate --- one eigenvalue from each virtual edge, such that $\chi$ from Eq.\eqref{eq:chi_integrable} will pick up one eigenvalue from each pair.
-Because of chiral symmetry the sum of the two eigenvalues related by the symmetry gives always $1$ so they do not contribute to $\chi$.
+In the presence of chiral and translation symmetry, all eigenvalues of the EOS are doubly degenerate: one eigenvalue from each virtual edge, such that $\chi$ from Eq.\eqref{eq:chi_integrable} will pick up one eigenvalue from each pair.
+Because of chiral symmetry, the sum of the two eigenvalues related by the symmetry gives always $1$ so they do not contribute to $\chi$.
 The only contribution to $\chi$ is the one from the eigenstates with $\xi=1/2$ that do not have a chiral partner, see figure~\ref{huang}(a).
 Here we see explicitly the known relation between the Zak phase and the number of $\xi = 1/2$ modes.
-
 Note that the full sum  in Eq.~\eqref{eq:chi_integrable}  evaluates to zero in the trivial regime %up to finite-size corrections% \carlos{It is actually exact due to symmetry}.
@@
 When this midgap eigenvalue approaches  0, the contribution from the other modes becomes more relevant and the bulk polarization starts to deviate from the dominant eigenvalue, see figure~\ref{huang}(c).
-In Fig.~\ref{huang}(d) we show how the difference between $\chi$ and the bulk polarization decreases with system size, implies that they are equal in the thermodynamic limit.
+In Fig.~\ref{huang}(d) we show how the difference between $\chi$ and the bulk polarization decreases with system size, showing that they are equal in the thermodynamic limit.
+Note that their discrepancy is very small already for small system sizes.
 
@@
 This is obvious when we express it in terms of a geometric phase, which is only defined modulo $2\pi$.
-In our formulation (See Eq.~\eqref{eq:ptilde_eq_chi}) this is reflected in an ambiguity in the number of bulk modes with $\xi = 1$ that are included in $\chi$.
+In our formulation inEq.~\eqref{eq:ptilde_eq_chi}, this is reflected in an ambiguity in the number of bulk modes with $\xi = 1$ that are included in $\chi$.
 Being defined only modulo $1$, the bulk polarization cannot differentiate the phases $\nu=0$ and $\nu=2$, even though the EOS of both phases is distinct.
 Another issue due to it being defined in $\mathbb{R}$ mod $1$ appears when we try to follow its evolution over a path $C_\lambda$.
-Between any two points $\lambda$ and $\lambda'$ in this path it is not possible to tell if the bulk polarization has increased or decreased by only looking at $\tilde{\mathcal{P}}(\lambda)$ at these specific points, this becomes relevant when computing Chern numbers.
-
-In an attempt to solve this issues we introduce an alternative bulk polarization given by
+Between any two points $\lambda$ and $\lambda'$ in this path it is not possible to tell if the bulk polarization has increased or decreased by only looking at $\tilde{\mathcal{P}}(\lambda)$ at these specific points.
+ This latter drawback becomes relevant when computing Chern numbers.
+
+In an attempt to solve these issues we introduce an alternative bulk polarization given by
 \begin{equation}\label{eq:tpo}
@@
 \end{equation}
-where $\ketpsio{{\Phi=0}}$ is the ground state of the periodic chain opened between sites $j=L/2$ and $j=L/2+1$ and 
+where $\ketpsio{{\Phi=0}}$ is the ground state when the (previously periodic) chain is opened between sites $j=L/2$ and $j=L/2+1$ and 
 \begin{equation}
@@
 When introducing the flux in the system with PBC (See Eq.~\eqref{eq:Ground_state_flux}) there is an ambiguity due to the bulk modes. 
-This ambiguity dissapears when we introduce the flux as above in the open chain.
+\mariac{Correct?} This ambiguity dissapears when we introduce the flux as above in the open chain --- we simply include all the modes. 
 Note that $\Phi$ is not a magnetic flux but should be treated as an additional parameter, and therefore $\tpo$ is not necessarily the physical polarization of the open chain.
@@
 The idea is that in the process of opening the chain, which can be done adiabatically, all eigenvalues that are related to the right virtual edge ($A_R$ in Fig.~\ref{fig:chain}) should get pushed to $\xi=0,1$.
-As a result, the sum of all eigenvalues of $C_{A,o}$ is equal, modulo 1 and in the thermodynamic limit (to avoid the finite size effects of opening the chain), to $\chi$ (See Eq.~\eqref{eq:chi_left}).
-Therefore, in the thermodynamic limit, we have
+As a result, the sum of all eigenvalues of $C_{A,o}$ (modulo 1) is equal to $\chi$ \eqref{eq:chi_left}.
+\footnote{ This is strictly speaking only true in the thermodynamic limit, where we avoid the finite size effects due to opening the chain.} Therefore, in the thermodynamic limit, we have
 \begin{equation}
@@
 
-The advantage of using $\tpo$ is that the contribution from the eigenvalues far from the left virtual edge is constant as long as no states cross the Fermi energy, as opposed to the case of $\chi$ where this is only true when taken modulo 1.
-The consequence is that for paths in parameter space, $C_\lambda$, which are gapped, i.e.
-no states cross the Fermi energy, $\tpo (\lambda)$ is defined in $\mathbb{R}$, such that the total change in $\tpo$ along this path can be obtained knowing $\tpo$ only in the initial and final states, $\Delta \tpo ({C_\lambda}) = \tpo(\lambda_f) - \tpo(\lambda_i)$.
+The advantage of using $\tpo$ is that the contribution from the 'bulk' modes is constant as long as no states cross the Fermi energy, i.e. the system remains gapped.  
+The quantity $\chi$ \eqref{eq:chi_left}, on the other hand, is only independent of the bulk mode contribution when the modulo 1 is included. 
+The consequence is that for paths $C_\lambda$ in parameter space, for which the energy spectrum is gapped, $\tpo (\lambda)$ is defined in $\mathbb{R}$, such that the total change in $\tpo$ along this path can be obtained knowing $\tpo$ only in the initial and final states, $\Delta \tpo ({C_\lambda}) = \tpo(\lambda_f) - \tpo(\lambda_i)$.
 Note, however, that $\tpo$ is still gauge-invariant only when taken modulo 1.
@@
 
-There are, however, two caveats to this method.
-When we open the chain, if the system is in a topological non-trivial state, zero-energy modes will appear.
+This method resolves most of the issues found in section \ref{sec:bulkpolarization}, but unfortunately it cannot be applied when the system is in the topological phase.
+When we open the chain for a system in a topological non-trivial state, zero-energy modes will appear.
 The ground state of the closed chain has several equivalent degenerate ground states for the open chain, so that $C_{A,\rm o}$ is not well-defined at these points.
-The second issue is that, even if we allow for symmetry-breaking terms so that we can compute $\tpo$, there might be integer jumps in $\tpo$ where the edge-modes cross zero-energy.
-However, as we will see below these jumps in $\tpo$ due to the zero modes are branch cuts that provide the correct description of the system.
+One might think that this problem can be circumvented by introducing symmetry-breaking term that render the system trivial and extrapolating $\tpo$ towards the topological phase. 
+However, this extrapolation is again not necessarily unique, as 
+%The second issue is that, even if we allow for symmetry-breaking terms so that we can compute $\tpo$, 
+there might be integer jumps in $\tpo$ when the edge-modes cross zero-energy. 
+As we will see below these jumps in $\tpo$are branch cuts that provide the correct description of the system.
 
@@
 
-Consider first the simple case of the SSH chain, with $t'=0$ in the model. When we include the chiral-symmetry braking parameter $\kappa$ the Hamiltonian now has the symmetry
+Consider first the simple case of the SSH chain. When we include the chiral-symmetry breaking parameter $\kappa$ the Hamiltonian now has the symmetry
 \begin{equation}
@@
 The integral in Eq.\eqref{eq:chern} is performed on the surface of the torus defined by $(\theta,\Phi)$.
-Because of the branch cut the Berry connection $A_\Phi(\theta)$ cannot be made smooth for the whole torus and therefore we cannot directly apply Stokes theorem.
+Because of the branch cut, the Berry connection $A_\Phi(\theta)$ cannot be made smooth for the whole torus and therefore we cannot directly apply Stokes theorem.
 Instead, we split the torus in two cylinders, $S_1$ with $\theta\in[\theta_0,2\pi-\theta_0]$ and  $S_2$ with $\theta\in[2\pi-\theta_0,2\pi+\theta_0]$ and we change the gauge in $\boldsymbol{A}(\theta)$ such that it is smooth in each cylinder.
@@
 where $C(m)$ is the Chern number of any loop around the degenerate point crossing $\kappa=0$ at $m$ and another point in the trivial region ($m>1$).
-Using the symmetry of the polarization described above gives
+Using the symmetry of the polarization described above \mariac{make 'above' more concrete?}  gives
 \begin{align}
@@
 \caption{(a) and (b) show the bulk polarization $\tpo-L/2$ computed for $t=1,\kappa'=0$ and $L=40$ sites, with (a) $t'=0$ and (b) $t'=-2$.
-$\tpo$ presents branch cuts that indicate the presence of zero-energy modes.
+$\tpo$ exhibits branch cuts that indicate the presence of zero-energy modes.
 $\tpo$ is able to differentiate between the phases $\nu=2$ and $\nu=0$.
@@
 For this model this equivalence is exact and the jump of $\tpo$ at $\kappa = 0$ gives the number of topological zero-modes, or the winding number.
-It is also interesting to point out that even far away from the symmetric point $\tpo$ still shows plateaus that point to a possible nearby topological transition.
-Note , however, that there are many ways of breaking the chiral symmetry.
-Replacing $\kappa$ in Eq.~\eqref{bdi_model} by $\kappa_i=\kappa (-1)^i$, the zero-energy modes split in such a way that $\tpo$ is continuous at $\kappa=0$.
+It is also interesting to point out that even far away from the symmetric point, $\tpo$ still shows plateaus that point to a possible nearby topological transition.
+Note, however, that there are many ways of breaking the chiral symmetry. 
+For instance, replacing $\kappa$ in Eq.~\eqref{bdi_model} by $\kappa_i=\kappa (-1)^i$, the zero-energy modes split in such a way that $\tpo$ is continuous at $\kappa=0$.
 It cannot differentiate between the phases $\nu=0$ and $\nu=2$, and a loop like the one discussed above results in a Chern number $C=0$.
-To illustrate the advantage of $\tpo$ we also show $\tilde{\mathcal{P}}$ mod $1$ in the $(m,\kappa)$ parameter space in figure~\ref{disorder_chern}(a).
-As mentioned above the only contribution to the Chern number comes from discontinuities in the bulk polarization, however when it is computed using $\tilde{\mathcal{P}}$ we no longer know in advance where this discontinuities are present, due to it being defined modulo 1.
+\mariac{We need some concluding remark here!}
+
+To illustrate the advantage of $\tpo$ we also show $\tilde{\mathcal{P}}$ mod $1$ in the $(m,\kappa)$ parameter space in figure~\ref{ssh_chern}(c).
+As mentioned above the only contribution to the Chern number comes from discontinuities in the bulk polarization.
+However, when it is computed using $\tilde{\mathcal{P}}$ we no longer know in advance where this discontinuities are present, due to it being defined modulo 1.
 One needs to look at the full path to know how the bulk polarization evolves along the loop.
@@
 where $\omega_i$ and $\omega_i'$ are selected from a uniform distribution with range $[-1/2,1/2]$.
-For very strong disorder where the gap fills with bulk states $\tpo$ will present many branch cuts as we approach the limit $\lim_{\kappa \rightarrow 0}$, which makes it unpractical.
-In order to avoid it we break chiral symmetry locally in the edge of the open chain so it only affects the edge modes.
+For very strong disorder where the gap fills with bulk states, $\tpo$ will present many branch cuts as we approach the limit $\lim_{\kappa \rightarrow 0}$, which makes it unpractical.
+In order to avoid this, we break chiral symmetry locally in the edge of the open chain so it only affects the edge modes.
 If any of the other states cross $E=0$ they do it in  chiral-symmetric pairs such that $\tpo$ is not affected by it.
@@
 The main difference is that since we break chiral symmetry only for the edge modes, $\tpo$ is quantized to multiples of $1/2$.
-Note that, since the states that fill the gap are localized at the bulk they will not appear in the EOS and one can still identify the appearance of virtual topological edge states (see Fig.~\ref{disorder_chern}(b)), as opposed to topological edge states in the energy spectrum.
+Note that, since the states that fill the gap are localized in the bulk, they will not appear in the EOS and one can still easily identify the appearance of virtual topological edge states (see Fig.~\ref{disorder_chern}(b)).
+In contrast, the topological edge states in the energy spectrum are completely masked by the (localized) bulk modes that fill the gap.
 
@@
 We then fourier transform only the $y$ direction to obtain
-We consider now the computation of Chern numbers in 2D systems with $\tpo$. Consider the Chern insulator with inversion symmetry given by the Hamiltonian
+We now discuss  the computation of Chern numbers in 2D systems with $\tpo$. Consider the Chern insulator with inversion symmetry given by the Hamiltonian
 \begin{align}
@@
 We can treat the Hamiltonian with elements $H_{i\alpha,j\beta}(k_y)$ as a 1D Hamiltonian with an extra parameter $k_y$ and compute $\tpo$, or equivalently $A_\Phi(\Phi,k_y)$, as in last section.
-But first lets analyze the entanglement occupancy spectrum of the system with PBC, shown in Fig.~\ref{chern_insulator}(a) and (b) for $m=1$ and $3$, respectively.
-For $0<m<2$ the EOS presents two edge modes that connect the occupied and empty bands that cross $\xi=1/2$ at $k_y=0$, in this region the occupied bands have a Chern number $C=+1$.
+But first lets analyze the EOS of the system with PBC, shown in Fig.~\ref{chern_insulator}(a) and (b) for $m=1$ and $3$, respectively.
+For $0<m<2$ the EOS presents two edge modes that connect the occupied and empty bands that cross $\xi=1/2$ at $k_y=0$.
+In this region the occupied bands have a Chern number $C=+1$.
 For $2<m<4$ there are also edge modes but they cross $\xi=1/2$ at $k_y=\pi$ and the system has $C=-1$.
-For $m>4$ the system is a trivial insulator with $C=0$.
+For $m<0$ or $m>4$ the system is a trivial insulator with $C=0$.
 When we open the chain the eigenvalue of the right virtual topological edge mode will be pushed to $\xi=1,0$ as seen in figure~\ref{chern_insulator}(a) and (b).
-It will have a jump whenever the correspondent edge-mode crossed zero energy, going between $\xi=1$ and $\xi=0$, i.e.
+It \mariac{what does 'it' refer to here?} will have a jump whenever the correspondent edge-mode crossed zero energy, going between $\xi=1$ and $\xi=0$, i.e.
 the number of eigenvalues at $\xi = 1$ is not constant across $k_y$.
@@
 
-We can now take a look at $\tpo$ in the parameter space $(m,k_y)$, shown in figure~\ref{chern_insulator}(c) .
+We can now take a look at $\tpo$ in the parameter space $(m,k_y)$, shown in figure~\ref{chern_insulator}(c).
 We see, again, that $\tpo$ remains continuous except for the branch cuts that appear at the points where the physical edge-modes cross zero energy.
@@
 \begin{equation}
-C(m) = \int_{-\pi}^{\pi}\frac{dk_y}{2\pi} \partial_{k_y} \tpo(m,k_y),
+C(m) = \int_{-\pi}^{\pi}\frac{dk_y}{2\pi} \partial_{k_y} \tpo(m,k_y).
 \label{eq:Zaletel_chern}
@@
 Here we have shown how it can be decoded from the single-particle entanglement spectrum.
-This relation is simpler than the one involving the MBES and it provides more insight into the ES and EOS.
+This relation is simpler than the one involving the many-body entanglement spectrum and it provides more insight into the ES and EOS.
 In particular in the topological case, where the relation between the quantized bulk polarization (or the Chern number in 2D) and the number of virtual topological edge states is shown in a new way.
@@
 There is an open question whether a formulation in terms of twisted boundary conditions can be obtained, in which case one might be able to generalize the results of this paper to higher dimensions.
+\mariac{I am not sure what you want to say in the last section}
 

### 2021-01-13 version2/draft_mh.tex
another small change
Add files via upload

@@
 \mathcal{P} = \frac{1}{2\pi} \Im \,\ln\bra{\Psi^{\Phi=0}}e^{2\pi i \hat{P}}\ket{\Psi^{\Phi=0}}.
-\end{align}
-
-Expression \eqref{polarization_inhom} is particularly useful because it can be easily expressed in terms of single-particle eigenstates as
+\label{eq:p_resta}
+\end{align}
+
+Expression \eqref{eq:p_resta} is particularly useful because it can be easily expressed in terms of single-particle eigenstates as
 \begin{align}
@@
 Note that their discrepancy is very small already for small system sizes.
+Even when the chiral symmetry is broken the polarization displays two different plateaus that point to a possible topological phase transition nearby in parameter space.
 
@@
 For this model this equivalence is exact and the jump of $\tpo$ at $\kappa = 0$ gives the number of topological zero-modes, or the winding number.
-It is also interesting to point out that even far away from the symmetric point, $\tpo$ still shows plateaus that point to a possible nearby topological transition.
-Note, however, that there are many ways of breaking the chiral symmetry. 
+Note, however, that there are many ways of breaking the chiral symmetry and for some of them, $\top$ may undercount the zero-energy modes. 
 For instance, replacing $\kappa$ in Eq.~\eqref{bdi_model} by $\kappa_i=\kappa (-1)^i$, the zero-energy modes split in such a way that $\tpo$ is continuous at $\kappa=0$.
 It cannot differentiate between the phases $\nu=0$ and $\nu=2$, and a loop like the one discussed above results in a Chern number $C=0$.
-\mariac{We need some concluding remark here!}
 
@@
 For $m<0$ or $m>4$ the system is a trivial insulator with $C=0$.
-When we open the chain the eigenvalue of the right virtual topological edge mode will be pushed to $\xi=1,0$ as seen in figure~\ref{chern_insulator}(a) and (b).
-It \mariac{what does 'it' refer to here?} will have a jump whenever the correspondent edge-mode crossed zero energy, going between $\xi=1$ and $\xi=0$, i.e.
+When we open the chain the eigenvalue of the right virtual topological edge mode will be pushed to $\xi=1,0$.
+This is seen in figure~\ref{chern_insulator}(a) and (b) for intermediate points in the process of opening the chain.
+This eigenvalue has a jump whenever the correspondent edge-mode crosses zero energy, going between $\xi=1$ and $\xi=0$, i.e.
 the number of eigenvalues at $\xi = 1$ is not constant across $k_y$.

### 2021-01-13 version2/draft_mh.tex
..

--- 2021-01-12 version2/draft_mh.tex
@@
 The bulk polarization is defined in $\mathbb{R}$ mod 1, which might present challenges when computing changes in the polarization, which are defined  in $\mathbb{R}$.
-We show how, using our method, one can define an alternative bulk polarization \mariac{REF?} computed in the equivalent open chain that is continuous in $\mathbb{R}$ for gapped paths.
+We show how, using our method, one can define an alternative bulk polarization computed in the equivalent open chain that is continuous in $\mathbb{R}$ for gapped paths.
 The bulk polarization obtained this way can give more information about the edge spectrum than the usual one, and can be used to simplify the computation of Chern numbers.
@@
 \end{equation}
-This generalizes the observation made by Ryu and Hatsugai \cite{Ryu2006} beyond the fully dimerized limit, $m=0$.
-
-\mariac{Rewrite paragraph}
-The issue concerning the ambiguity will be discussed and resolved in the next section. 
-For translational invariant systems the EOS is symmetric such that a crossing of the eigenvalues $\xi$ results in $\chi$ changing sign.
-This is irrelevant in the case of symmetry-protected states where the Zak phase is quantized to $0$ or $\pi$.
+%This generalizes the observation made by Ryu and Hatsugai \cite{Ryu2006} beyond the fully dimerized limit, $m=0$. 
 In the special case where there is only one eigenvalue per edge that is $\neq 0,1$ in  the EOS, like in the fully dimerized limit of the SSH chain, we recover the equality between one of the two midgap eigenvalues and the Zak phase found in reference ~\cite{Ryu2006}.
+
+The issue concerning the sign ambiguity will be discussed and resolved in the next section. 
+The origin of this ambiguity can be traced back to certain properties of the EOS. 
+In particular, for translational invariant systems the EOS is symmetric such that a crossing of the eigenvalues $\xi$ results in $\chi$ changing sign. 
+This sign is irrelevant in the case of symmetry-protected states where the Zak phase is quantized to $0$ or $\pi$.
 
@@
 (c) and (d) for the dotted line in the phase diagram, $t=1,t'=-2$  with the symmetry-breaking parameters (c) $\kappa=0.3,\kappa'=0$ and (d) $\kappa=0,\kappa'=0.3$.
-(d) is the same as (c) but with a position dependent $\kappa_i=(i+L/4 \, {\rm mod}\, L)/L$.
+(e) is the same as (c) but with a position dependent $\kappa_i=(i+L/4 \, {\rm mod}\, L)/L$.
 We also plot $\mathcal{P}+1/2$ (blue) to showcase that $\chi$ is indeed equal to $\tilde{\mathcal{P}}$ and not $\mathcal{P}$.
@@
 Note that in the $\nu=2$ phase the observation made in Ref. \cite{Huang2012}, i.e. that the there is a similarity between the Zak phase and the entanglement occupancy, does not apply anymore. 
+\mariac{It is also interesting to point out that even far away from the symmetric point, $\tpo$ still shows plateaus that point to a possible nearby topological transition.}
 
@@
 When introducing the flux in the system with PBC (See Eq.~\eqref{eq:Ground_state_flux}) there is an ambiguity due to the bulk modes. 
-\mariac{Correct?} This ambiguity dissapears when we introduce the flux as above in the open chain --- we simply include all the modes. 
+This ambiguity dissapears when we introduce the flux as above in the open chain --- we simply include all the bulk modes. 
 Note that $\Phi$ is not a magnetic flux but should be treated as an additional parameter, and therefore $\tpo$ is not necessarily the physical polarization of the open chain.
@@
 Consider first the simple case of the SSH chain. When we include the chiral-symmetry breaking parameter $\kappa$ the Hamiltonian now has the symmetry
-\begin{equation}
+\begin{equation}\label{eq:sym}
 SH(\kappa)S^\dagger = -H(-\kappa), \quad S C_A(\kappa) S^\dagger = \mathbb{I}-C_A(-\kappa),
@@
 
+For simplicity's sake, assume that there is a single branch cut along $\kappa=0$, such as depicted in Fig.s~\ref{ssh_chern} (a) or (b).
+The generalization to more complicated scenarios is straightforward.
 Consider a counterclockwise loop around the degenerate point at $(m=t,\kappa=0)$ starting from the branch cut parametrized by $\theta$,
@@
 where $C(m)$ is the Chern number of any loop around the degenerate point crossing $\kappa=0$ at $m$ and another point in the trivial region ($m>1$).
-Using the symmetry of the polarization described above \mariac{make 'above' more concrete?}  gives
+Using the symmetry of the polarization described above (see the discussion below Eq.~\eqref{eq:sym}) gives
 \begin{align}
@@
 
-A jump of $n$ in $\tpo$ implies the existence of \emph{at least} $n$ zero-energy modes at that point.
-For this model this equivalence is exact and the jump of $\tpo$ at $\kappa = 0$ gives the number of topological zero-modes, or the winding number.
-It is also interesting to point out that even far away from the symmetric point, $\tpo$ still shows plateaus that point to a possible nearby topological transition.
-Note, however, that there are many ways of breaking the chiral symmetry. 
+Jumps in $\tpo$ originate from zero modes changing their occupation. 
+Thus, a jump of $n$ in $\tpo$ implies the existence of \emph{at least} $n$ zero-energy modes at that point. 
+For this model this equivalence is exact and the jump of $\tpo$ at $\kappa = 0$ gives the number of topological zero-modes, i.e. the winding number. 
+%It is also interesting to point out that even far away from the symmetric point, $\tpo$ still shows plateaus that point to a possible nearby topological transition.
+Note, however, that there are many ways of breaking the chiral symmetry and for some of them, $\tpo$ may not be related to the winding number any longer. 
 For instance, replacing $\kappa$ in Eq.~\eqref{bdi_model} by $\kappa_i=\kappa (-1)^i$, the zero-energy modes split in such a way that $\tpo$ is continuous at $\kappa=0$.
 It cannot differentiate between the phases $\nu=0$ and $\nu=2$, and a loop like the one discussed above results in a Chern number $C=0$.
-\mariac{We need some concluding remark here!}
 

### 2021-01-13 version2/draft_mh.tex
..
Add files via upload

@@
 In the symmetric case, the known relation between the bulk polarization and the number of topological edge modes is recovered.
-We use the bulk polarization to compute Chern numbers in 1D and 2D, wcich illuminates their known relation to the entanglement spectrum  in a new way.
+We use the bulk polarization to compute Chern numbers in 1D and 2D, which illuminates their known relation to the entanglement spectrum  in a new way.
 Furthermore we discuss an alternative bulk polarization that can carry more information about the surface spectrum than the conventional one and, thus, can simplify the calculation of Chern numbers.
@@
 Let us now proceed to discuss the entanglement spectrum. 
-An important quantity to consider is the correlation matrix, which in position space is defined by the groujd state expectation value
+An important quantity to consider is the correlation matrix, which in position space is defined by the ground state expectation value
 \begin{align}
@@
 When breaking chiral symmetry, the EOS will no longer be doubly degenerate.
+Note that even far away from the symmetric point the polarization shows plateaus that point to a possible nearby topological transition.
 For small $m$, there are two dominant midgap eigenvalues, and the bulk polarization follows the lower one very closely, see figure~\ref{huang}(b).
@@
 Note that in the $\nu=2$ phase the observation made in Ref. \cite{Huang2012}, i.e. that the there is a similarity between the Zak phase and the entanglement occupancy, does not apply anymore.
-\mariac{It is also interesting to point out that even far away from the symmetric point, $\tpo$ still shows plateaus that point to a possible nearby topological transition.}
 
@@
 For $m<0$ or $m>4$ the system is a trivial insulator with $C=0$.
-When we open the chain the eigenvalue of the right virtual topological edge mode will be pushed to $\xi=1,0$.
+When we open the chain, the eigenvalue of the right virtual topological edge mode will be pushed to $\xi=0$ or $1$.
 This is seen in figure~\ref{chern_insulator}(a) and (b) for intermediate points in the process of opening the chain.
-This eigenvalue has a jump whenever the correspondent edge-mode crosses zero energy, going between $\xi=1$ and $\xi=0$, i.e.
+This eigenvalue presents a jump whenever the correspondent edge-mode crosses zero energy, going between $\xi=1$ and $\xi=0$, i.e.
 the number of eigenvalues at $\xi = 1$ is not constant across $k_y$.
@@
 
-In the field of higher-order topological insulators some systems can be characterized by a physical generalization of the bulk polarization to higher dimensions \cite{Benalcazar2017,Kang2019}.
-There is an open question whether a formulation in terms of twisted boundary conditions can be obtained, in which case one might be able to generalize the results of this paper to higher dimensions.
-\mariac{I am not sure what you want to say in the last section}
+In the field of higher-order topological insulators some systems can be characterized by a physical generalization of the bulk polarization to two dimensions, the quadrupole moment \cite{Benalcazar2017,Kang2019}.
+This quadrupole moment is defined with the position operator and there is an open question whether a formulation in terms of twisted boundary conditions can be obtained.
+In this case, one might be able to construct this quadrupole moment using the entanglement spectrum, as we do here for the bulk polarization.
+
+\acknowledgments
+{\em Acknowledgments.--} 
+FUNDING
+We thank Krishanu XX and Thors Hans Hansson for a critical reading of the manuscript.
 

### 2021-02-02 version2/draft_mh.tex
Add files via upload

@@
 The bulk polarization is a $\mathbb{Z}_2$ topological invariant characterizing non-interacting systems in one dimension with chiral or particle-hole symmetries.
-We show that the bulk polarization can always be decoded from the single-particle entanglement spectrum, even in the absence of symmetries that quantize it.
+We show that the bulk polarization can always be determined from the single-particle entanglement spectrum, even in the absence of symmetries that quantize it.
 In the symmetric case, the known relation between the bulk polarization and the number of topological edge modes is recovered.
-We use the bulk polarization to compute Chern numbers in 1D and 2D, which illuminates their known relation to the entanglement spectrum  in a new way.
-Furthermore we discuss an alternative bulk polarization that can carry more information about the surface spectrum than the conventional one and, thus, can simplify the calculation of Chern numbers.
+We use the bulk polarization to compute Chern numbers in 1D and 2D, which illuminates their known relation to the entanglement spectrum.
+Furthermore we discuss an alternative bulk polarization that can carry more information about the surface spectrum than the conventional one and can simplify the calculation of Chern numbers.
 \end{abstract}
@@
 Topological phases of matter have attracted a lot of attention during the last decades, not the least because a large variety of relevant systems have been realized experimentally. 
-The early focus was mainly on \emph{topologically ordered} systems \cite{wenbook}, where interaction effects are crucial for stabilizing the phase. 
-The most notable examples are the fractional quantum Hall effect~\cite{Tsui1982} and quantum spin liquids~\cite{Balents2010spin}. 
-However, since 2005~\cite{kane2005quantum, roy2009topological} the focus has shifted to non-interacting topological phases, which can be characterized in terms of topological invariants~\cite{ryu2010topological}.
-Symmetries are often necessary to protect these topological phases, and determine which distinct topological phases can be realized for a given dimensionality.
-
-There are few tools to identify topological phases. 
-A very efficient one is the entanglement entropy~\cite{srednicki1993entropy}, which allows one to identify the total quantum dimension of the underlying topological quantum field theory~\cite{Kitaev2006topological, Levin2006detecting}.
-As such, the entanglement entropy can only be used for interacting systems. 
-Its main drawback is the fact that it can only provide one of the pertinent quantum numbers that characterizes the system, and is, in general,  not able to uniquely determine the topologically ordered phase at hand.
-In addition, it requires a careful scaling analysis, which can be challenging for strongly correlated systems. 
-Another, closely related tool, is the entanglement spectrum (ES), originally introduced for fractional quantum Hall systems~\cite{Li2008entanglement}. 
-It was conjectured to provide information about the edge spectrum, which was later shown to be a general feature of ground states with an effective topological quantum field theory description~\cite{Qi2012general}.
-It also proved to be an effective tool for fractional Chern insulators~\cite{Regnault2011fractional} and certain quantum spin liquids~\cite{yao2010entanglement}. %\mariac{Others?}
-
-Generically, any of the proposed tools naturally provides only some information about the phase.
-It is not known, whether there exists a method to obtain the full topological content from the ground state.
-For the simplest quantum Hall states, the Laughlin states,  it was shown that the ES does, in fact, contain the full information about the topological phase~\cite{hermanns2011haldane}.
-However, it is far from clear if this holds for more complicated, in particular nonabelian, states as well.
+The early focus was mainly on \emph{topologically ordered} systems \cite{wenbook}, where interaction effects are crucial for stabilizing the phases. 
+The most notable examples are the fractional quantum Hall liquids~\cite{Tsui1982} and quantum spin liquids~\cite{Balents2010spin}. 
+However, since 2005~\cite{kane2005quantum, roy2009topological} the focus has shifted to  \emph{symmetry-protected topological phases} (SPT) where symmetries are necessary to protect the topological phases and determine which distinct topological phases can be realized for a given dimensionality.
+These can be implemented as free-theories and can be characterized in terms of topological invariants~\cite{ryu2010topological}.
+
+Entanglement has played an important role in the understanding and characterization of topological systems \cite{Ludwig2015,Wen2017}.
+One can distinguish two types of states depending on their entanglement.
+Short range entangled states can be continuously transformed into a direct product state, while long range entangled states cannot.
+The latter correspond to topologically ordered states.
+Certain short range entangled states cannot be continuously transformed between themselves unless certain symmetries are broken.
+These are the SPT phases that we focus on in this paper.
+
+There are different tools based on entanglement that have been used to characterize topological phases.
+A very efficient one is the entanglement entropy~\cite{srednicki1993entropy}, which allows one to determine the total quantum dimension of the underlying topological quantum field theory~\cite{Kitaev2006topological, Levin2006detecting}, but it can only be used for topologically ordered phases and it cannot uniquely characterize the topological phase at hand.
+Another, closely related, tool is the entanglement spectrum (ES), originally introduced for fractional quantum Hall systems~\cite{Li2008entanglement}.
+It provides information about the edge spectrum and has proven useful for other topologically ordered phases such as fractional Chern insulators~\cite{Regnault2011fractional} and certain quantum spin liquids~\cite{yao2010entanglement}.
 
@@
 The same correspondence was also found for closely related gapless systems~\cite{Matern2018entanglement}.
-However, even for non-interacting systems, it is unclear which information (beyond the protected `edge' spectrum) is encoded in the entanglement spectrum.
-In this manuscript, we show that the information about the Zak phase, or alternatively the bulk polarization, can be completely decoded from the entanglement spectrum and we provide a simple formula to do so.
-This is a general property of non-interacting 1D gapped systems which, in the symmetric case where the bulk polarization is quantized, gives the known relation between the bulk polarization and the number of virtual topological edge states in the ES \cite{Ryu2006}.
-We also show how to extend this method to compute Chern numbers directly from the entanglement spectrum.
-
-The relation between the entanglement spectrum and various topological invariants has already been studied in the literature. 
+However, even for non-interacting systems, it is unclear which information (beyond the protected `edge' spectrum) is encoded in the ES.
+The aim of this paper is to show that other physical properties are also encoded in the entanglement spectrum, in particular the bulk polarization.
+
+The bulk, or macroscopic, polarization is a fundamental concept in physics, primarily used to describe the response of matter to electric fields.
+The modern theory of polarization \cite{Resta1992,KingSmith1993,Vanderbilt1993,Resta1997} related the bulk polarization to a geometric phase (Zak phase for translationally invariant systems).
+Due to this, it has also found its way into topological physics and in topological charge pumping.
+In certain symmetry classes the bulk polarization is quantized and serves as a $\mathbb{Z}_2$ topological invariant, where it is known to be related to the number of protected topological edge state in the ES \cite{Ryu2006}.
+Because of its relation to a geometric phase, the bulk polarization is also related to other topological invariants in higher dimensions such as the Chern number.
+
+We show how the bulk polarization can be decoded from the ES.
+This is a general property of non-interacting 1D gapped systems, which in the topological case reduces to the known relation between the bulk polarization and the ES described above \cite{Ryu2006}.
+We also apply this method to compute Chern numbers directly from the entanglement spectrum.
+The differences in the bulk polarization, which can be any real number, are the relevant quantities to study.
+They can, however, be difficult to compute from the bulk polarization itself, as it is only defined modulo $1$.
+We show how, using our method, one can define an alternative bulk polarization computed by using open boundary conditions which is continuous in $\mathbb{R}$ for gapped paths, simplifying the calculation of polarization differences.
+The bulk polarization defined this way can give more information about the edge spectrum than the usual one, and can be used to simplify the computation of Chern numbers.
+We also discuss the relation between this alternative bulk polarization and another topological invariant known as the trace index \cite{Alexandrinata2011}.
+
+The relation between the entanglement spectrum and the bulk polarization has already been studied. 
 In Ref.~\cite{Zaletel2014}, the authors show that the Zak phase can be computed from the Schmidt decomposition of a translation invariant, infinite chain. 
-This demonstrates that there should be a direct relation between the single-particle entanglement energies and the Zak phase for non-interacting systems.
+This suggests that there should be a direct relation between the single-particle entanglement energies and the Zak phase for non-interacting systems.
 A particular limit of our result for homogeneous (translationally invariant) systems~\eqref{eq:chi_integrable} was derived
@@
 for a fully dimerized SSH chain~\cite{Ryu2006}. 
-The similarity between the behavior of the Zak phase and the virtual topological edge states in the entanglement spectrum of certain Chern insulators, which was observed in Ref.~\cite{Huang2012,Huang2012-2}, can also be explained by our results.
-
-The bulk polarization is defined in $\mathbb{R}$ mod 1, which might present challenges when computing changes in the polarization, which are defined  in $\mathbb{R}$.
-We show how, using our method, one can define an alternative bulk polarization computed in the equivalent open chain that is continuous in $\mathbb{R}$ for gapped paths.
-The bulk polarization obtained this way can give more information about the edge spectrum than the usual one, and can be used to simplify the computation of Chern numbers.
-We also discuss the relation between this alternative bulk polarization and another topological invariant known as the trace index \cite{Alexandrinata2011}.
+The similarity between the behavior of the Zak phase and the protected topological edge states in the entanglement spectrum of certain Chern insulators, which was observed in Ref.~\cite{Huang2012,Huang2012-2}, is also explained by our results.
 
 \emph{Outline of the paper}: In section II we introduce the 1D model used throughout the paper and we examine its phase diagram.
-In section III we discuss the bulk polarization, the different ways one can define it and its relation to the Zak phase.
-The ES is introduced in section IV, where we set the notation that will be used.
-In section V we present our method and show how we can reproduce the bulk polarization for both systems with and without translational invariance.
+In section III we discuss the bulk polarization; the different ways one can define it and its relation to the Zak phase.
+The ES is introduced in section IV, where we also set our notation.
+In section V we present our method and show how we can reproduce the bulk polarization for systems with and without translational invariance.
 In section VI we introduce an alternative bulk polarization constructed from the ES and use it to compute Chern numbers in 1D and 2D.
-Finally in section VII we conclude and discuss the possible outlook.
+Finally in section VII we conclude and discuss possible extensions of our work.
 
 % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % 
-\section{Model}
-To showcase our results we consider as a starting point a model in the BDI class~\cite{Song2014}. 
-This is a simple model which nevertheless supports a rich phase diagram containing topological phases with winding numbers $\nu =0,1$ and $2$. 
-In order to show that our results are not limited to any particular symmetry class~\cite{ryu2010topological}, we also include two symmetry breaking terms. 
-The Hamiltonian is then given by
+\section{The model}
+To illustrate our results we consider a model in the BDI class~\cite{Song2014}.
+Although simple, it supports a rich phase diagram containing topological phases with winding numbers $\nu =0,1$ and $2$.
+In order to show that our results are not limited to this particular symmetry class~\cite{ryu2010topological}, we also include two symmetry breaking terms. 
+The Hamiltonian is then
 \begin{align}
@@
 \end{align}
-where $H_{ij} $ is given by 
+where 
 \begin{align}
@@
 \end{align}
-and the corresponding Bloch Hamiltonian is given by
+and the corresponding Bloch Hamiltonian is 
 \begin{align}
@@
 
-For $\kappa = \kappa' = 0$ this Hamiltonian is in the BDI class, i.e. it has time-reversal ($T$), particle-hole ($C$), and chiral symmetry ($S$):
+For $\kappa = \kappa' = 0$, it is in the BDI class, i.e. it has time-reversal ($T$), particle-hole ($C$), and chiral symmetry ($S$):
 \begin{alignat}{2}
@@
 When open boundary conditions are imposed on the system, the number of symmetry-protected zero-energy edge modes is equal to the winding number. 
-The behavior of the  edge modes  is  shown in Fig.~\ref{fig:zero_E_modes}, using the dotted path ($t=1,t'=-2$) marked in the phase diagram. 
+The behavior of the  edge modes  is  shown in Fig.~\ref{fig:zero_E_modes}, using the dotted path ($t=1,t'=-2$) marked in the phase diagram in figure~\ref{fig:bdi_phase_diagram}. 
 Figure~\ref{fig:zero_E_modes}(a) shows the edge spectrum in the BDI class, with 4 (resp. 2) symmetry-protected zero modes for winding number 2 (1). 
@@
 In one spatial dimension, the latter is characterized by a $\mathbb{Z}_2$ invariant.  
-Consequently, adding such a term causes the zero modes in the $\nu=2$ phase to split  pairwise and the $\nu = 2$ and $\nu=0$ phases become equivalent. 
+Consequently, adding such a term causes the zero modes in the $\nu=2$ phase to split pairwise and the phase becomes topologically trivial, indistinguishable from the phase with $\nu=0$.
 This is shown in Fig.~\ref{fig:zero_E_modes}(b), where one clearly sees that the zero modes below $m=1$ are split.
@@
 
-In this section we review some aspects of the modern theory of polarization, which gave a geometrical description of the polarization \cite{Resta1992,KingSmith1993,Vanderbilt1993,Resta1997}. 
-We also discuss the results of a recent article by Watanabe and Oshikawa \cite{Watanabe2018} that clarified the meaning of different bulk polarizations that are found in the literature, as well as the relations between them.
-For clarity we will use the same notation they use.
+The bulk polarization is a property which characterizes topological insulators in 1D.
+The bulk polarization is proportional to the surface charges of the system with open boundary conditions.
+In a topological insulator these are quantized because of the appearance of zero-energy edge states.
+Because of this, the bulk polarization is itself quantized, and therefore serves as a topological invariant.
+
+In this section we review some relevant aspects of the bulk polarization and the modern theory of polarization, which gives a geometrical description of the polarization \cite{Resta1992,KingSmith1993,Vanderbilt1993,Resta1997}.
 
@@
 \end{equation}
-We can obtain its single-particle eigenstates as
+We obtain its single-particle eigenstates by
 \begin{align}
@@
 where $A_k = \bra{u_k}\partial_k \ket{u_k}$ is the Berry connection \cite{Berry1984}. 
-The generalization to higher-band models is straightforward.
+The generalization to multi-band models is straightforward.
 
@@
 which allows us to make the identification $ \mathcal{P}^{\rm Bloch} = \gamma/2\pi$.
-As discussed above,  this polarization is defined only modulo 1, while its changes are  defined in $\mathbb{R}$.
-Note also that the polarization itself is not a physical observable.
-Only its changes, related to the current flowing through the system, are.
-
-One can also compute the geometric phase obtained by threading a $U(1)$ flux through the ring \cite{Watanabe2018}, which allows us to compute a polarization in the absence of translational invariance.
-If the flux is introduced homogeneously through a vector potential $A_x = \Phi/L$ another polarization can be defined as
+The polarization itself is however not an observable.
+The relevant quantities are the derivatives or changes in the polarization with respect to external parameters.
+These result in currents and charge transport, which are the only measurable quantities.
+Similar to how only difference in the Zak phases between two states can be measured \cite{Atala2013}.
+
+Considering a system with periodic boundary conditions, one can also compute the geometric phase obtained by threading a $U(1)$ flux through the ring.
+This allows us to define a polarization in the absence of translational invariance.
+Depending on how the flux is introduced we obtain different polarizations with different physical meanings.
+These different polarizations, their relations and physical interpretations were studied in a recent article by Watanabe and Oshikawa \cite{Watanabe2018}.
+We focus on two polarizations, $\mathcal{P}$ and $\tilde{\mathcal{P}}$, obtained by introducing the flux in two different ways.
+
+whose derivatives give the average current along the chain and the current at the seam (where the PBC are implemented), respectively.
+
+If the flux is introduced homogeneously through a vector potential $A_x = \Phi/L$ we can define the polarization
 \begin{equation}\label{polarization_inhom}
@@
 where $\ket{\Psi^{\Phi}}$ is the ground state in the presence of flux and $\hat{P}=\frac{1}{L}\sum_{j\alpha} j\hat{n}_{j\alpha}$ is the polarization operator, with $\hat{n}_{j\alpha}$ being the number operator.
+The second term is needed to make the expression gauge invariant.
+The derivative of this polarization gives the average current along the chain.
+
 This is equivalent to the polarization obtained by Resta \cite{Resta1997} as
@@
 \end{align}
-
 Expression \eqref{eq:p_resta} is particularly useful because it can be easily expressed in terms of single-particle eigenstates as
@@
 
-There is one last bulk polarization we must consider. 
-It is obtained when the flux is not introduced homogeneously, but instead only in the bond between sites $j=1$ and $j=L$ via twisted boundary conditions.
-This is equivalent to performing the gauge transformation
+If the flux is introduced via twisted boundary conditions at the seam, in the bond between sites $j=1$ and $j=L$, this is equivalent to performing the gauge transformation
 \begin{align}\label{eq:Psi_tilde}
@@
 \end{equation}
-which is related to the current flowing through this bond \cite{Watanabe2018}. This bulk polarization is related to the previous one by 
-\begin{equation}
-\mathcal{P}= \tilde{\mathcal{P}}+\tilde{\mathcal{P}}_0,
+whose derivative gives the current flowing through the seam \cite{Watanabe2018}. This bulk polarization is related to the previous one by 
+\begin{equation}
+\mathcal{P}= \tilde{\mathcal{P}}+\bar{\mathcal{P}}_0,
 \end{equation} 
@@
 \begin{equation}
-\tilde {\mathcal{P}}_0= \int_0^{2\pi}\frac{d\Phi}{2\pi} \bra{\tilde{\Psi}^\Phi}\hat{P}\ket{\tilde{\Psi}^\Phi}.
+\bar {\mathcal{P}}_0= \int_0^{2\pi}\frac{d\Phi}{2\pi} \bra{\tilde{\Psi}^\Phi}\hat{P}\ket{\tilde{\Psi}^\Phi}.
 \end{equation} 
@@
 Similar to the discussion about the Bloch polarization below Eq.~\eqref{eq:Bloch_polarization}, both $\mathcal{P}$ and $\tilde{\mathcal{P}}$ are defined modulo 1, while their changes are defined in $\mathbb{R}$.
-For translationally invariant systems one finds that $\mathcal{P} = \mathcal{P}^{\rm Bloch}+\frac{L-1}{2}\nu$, where here $\nu$ denotes the number of occupied bands \cite{Watanabe2018}.
+For translationally invariant systems one finds that
+\begin{equation}
+\mathcal{P} = \mathcal{P}^{\rm Bloch}+\frac{L-1}{2}\nu \, {\rm mod} \, 1
+\end{equation}
+where here $\nu$ denotes the number of occupied bands \cite{Watanabe2018}.
 
@@
 \end{align}
-For a system where all single-particle eigenstates with energies $E<0$ are occupied, the correlation matrix can be written  in terms of the Hamiltonian as
+For a system where all single-particle eigenstates with energies $E<0$ are occupied, the correlation matrix can be written (see  Appendix~\ref{app:CM}) in terms of the Hamiltonian as
 \begin{align}\label{eq:corr_mat2}
@@
 \end{align}
-This form, which we derive in Appendix~\ref{app:CM}, is particularly useful since it does not require diagonalizing the Hamiltonian and we will use it in all subsequent calculations.
-
-In this manuscript, we are mainly interested in the spectrum of the correlation matrix, when restricted to a spatial sub-region $A$ (its complement will be denoted by $B$ in the following), see Fig.~\ref{fig:chain}.
+
+In this paper, we are mainly interested in the spectrum of the correlation matrix when restricted to a spatial sub-region $A$ (its complement will be denoted by $B$ in the following), see Fig.~\ref{fig:chain}.
 Following Ref.s~\cite{Huang2012,Huang2012-2}, we refer to this spectrum as the entanglement occupancy spectrum (EOS).
@@
 the corresponding eigenstates are the same. 
-Eq.~\eqref{eq:xi_eps} implies that the full information of the \emph{many-body} ES (MBES) is contained in the single-particle spectrum of the reduced correlation matrix for non-interacting systems. 
+Eq.~\eqref{eq:xi_eps} implies that, for non-interacting systems, the full information of the \emph{many-body} ES (MBES) is contained in the single-particle spectrum of the reduced correlation matrix. 
 The latter is much simpler to interpret. 
-Consequently, we will focus on the EOS in the remainder of the manuscript and only mention the ES when necessary.
-
-One property of the subsystem correlation matrix we will make use of later is that, for sufficiently large systems, its eigenstates are found to be localized on either virtual edge if $\xi$ is away from $0$ and $1$, or they are found to be bulk modes if the correspondent eigenvalues are exponentially close to $\xi = 0,1$ \cite{Peschel2008}.
+Consequently, we will focus on the EOS in the remainder of the paper and only mention the ES when necessary.
+
+One property of the subsystem correlation matrix we will make use of later is that, for sufficiently large systems, its eigenstates are found to be exponentially localized on either virtual edge (edges of sub-region A) if $\xi$ is away from $0$ and $1$, or they are found to be bulk modes if the correspondent eigenvalues are exponentially close to $\xi = 0,1$ \cite{Peschel2008}.
 
@@
 Let us first review previous results on the relation between the EOS and the bulk polarization (or Zak phase). 
-For systems in 1D protected by chiral or particle-hole symmetries the Zak phase is quantized.
+For systems in 1D protected by chiral or particle-hole symmetries the Zak phase is a topological invariant.
 It is zero whenever there is an even number  of eigenvalues per virtual edge at $\xi = 1/2$, and $\pi$ when the number is odd~\cite{Asboth2016}.
@@
 
-We first discuss systems for which the entanglement energies of a \emph{single} virtual edge are equispaced.
+We first discuss systems for which the entanglement energies of a \emph{either} virtual edge are equispaced.
 That is, they are given in the thermodynamic limit by $\varepsilon_{n\alpha} = \varepsilon_{0\alpha}+n\delta_\alpha$, where $\alpha = L,R$.
@@
 
-When breaking chiral symmetry, the EOS will no longer be doubly degenerate.
-Note that even far away from the symmetric point the polarization shows plateaus that point to a possible nearby topological transition.
-For small $m$, there are two dominant midgap eigenvalues, and the bulk polarization follows the lower one very closely, see figure~\ref{huang}(b).
+When breaking chiral symmetry, as shown in figure~\ref{huang}(b), the EOS will no longer be doubly degenerate.
+Note that, even with the broken symmetry, the bulk polarization shows plateaus that point to a possible nearby topological transition.
+For small $m$, there are two dominant midgap eigenvalues, and the bulk polarization follows the lower one very closely, see .
 When this midgap eigenvalue approaches  0, the contribution from the other modes becomes more relevant and the bulk polarization starts to deviate from the dominant eigenvalue, see figure~\ref{huang}(c).
@@
 Note that their discrepancy is very small already for small system sizes.
-Even when the chiral symmetry is broken the polarization displays two different plateaus that point to a possible topological phase transition nearby in parameter space.
 
@@
 
-In general we find that it is only the eigenstates that localize on the left edge ($A_L$ in Fig.\ref{fig:chain}) that should contribute to $\chi$.
+In general we find (see Appendix \ref{app:pollmann}) that it is only the eigenstates that localize on the left virtual edge ($A_L$ in Fig.\ref{fig:chain}) that should contribute to $\chi$.
 The inclusion of bulk modes is irrelevant since they have $\xi = 0,1$.
@@
 the sum is performed over the subspace $L$ formed by the eigenstates with $\expval{\hat{x}}<L/4$, where $L/4$ is the middle point of region $A$.
-It does not matter what threshold we use as we include all left-edge states and exclude all right-edge states.
+It does not matter what threshold we use as long as we include all left-edge states and exclude all right-edge states.
 In practice only a few eigenvalues give a significant contribution to $\chi$.
 
-In the thermodynamic limit, it is $\tilde{\mathcal{P}}$ we identify with $\chi$
-\begin{equation}
-\lim_{L \rightarrow \infty} \tilde{\mathcal{P}} = \lim_{L \rightarrow \infty} \chi.
+In the thermodynamic limit we obtain
+\begin{equation}
+\lim_{L \rightarrow \infty} \tilde{\mathcal{P}} = \lim_{L \rightarrow \infty} \chi \, {\rm mod } \, 1.
 \label{eq:ptilde_eq_chi}
@@
 It might be --- like in Fig.\ref{huang}(a) --- that the spectrum has degenerate pairs of left and right eigenstates and the numerical diagonalization mixes them.
-This case is trivial, as we know that there is one left and one right eigenstate but if the degeneracy is four-fold or higher one needs to obtain the localized eigenstates.
+This case is trivial, since we know that there is one left and one right eigenstate, but if the degeneracy is four-fold or higher we must obtain the localized eigenstates.
 Localizing the bulk eigenstates that are exponentially close to $\xi = 0,1$ is more difficult but not doing so only results in an exponentially small error.
 
-Before proceeding lets review again the result in the previous section.
+Before proceeding, lets review again the result so far.
 First note that for translation invariant systems, the ES of the right and the left virtual edge are related by an overall '-' sign, i.e.
 for each eigenvalue $\epsilon$ on the right virtual edge, there is a corresponding value $-\varepsilon$ on the left one.
-Since the spectra is equispaced, this implies that the two virtual edge spectra are merely shifted with respect to each other.
+Since the spectra are equispaced, this implies that the two virtual edge spectra are merely shifted with respect to each other.
 If the shift is zero, e.g.
 in presence of a symmetry, summing every other eigenvalue is equivalent, modulo 1, to summing all the left eigenvalues.
-With a finite shift summing all odd eigenvalues corresponds to either summing all left eigenvalues or all right eigenvalues.
-In general, it implies that $\chi$ as defined in Eq.~\eqref{eq:chi_integrable} only gives the Zak phase up to an overall sign.
-This issue can be fixed by determining the localization of a single eigenstate.
+For a finite shift, summing all odd eigenvalues corresponds to either summing all left eigenvalues or all right eigenvalues.
+In general, this implies that $\chi$, as defined in Eq.~\eqref{eq:chi_integrable}, only gives the Zak phase up to an overall sign.
+This issue can be resolved by determining the localization of a single eigenstate.
 
@@
 
-Since $\chi$ is naturally defined in position space our results are also valid for systems without translationally invariance.
+Since $\chi$ is defined in position space our results are also valid for systems without translationally invariance.
 We show this in Fig.~\ref{2}(e) where we show the EOS, $\chi$, $\mathcal{P}$ and $\tilde{\mathcal{P}}$ along the dotted cut in the phase diagram with an added position dependent symmetry breaking term $\kappa_i = (i+L/4 \, {\rm mod}\, L)/L$.
-We use this profile for $\kappa_i$ instead of simple disorder because with the disorder the difference between $\mathcal{P}$ and $\tilde{\mathcal{P}}$ is very small.
-We obtain that $\chi$ is indeed identified with $\tilde{\mathcal{P}}$ as we hinted at before, and not $\mathcal{P}$, although these two polarization only differ by $(L+1)/2$ in the translationally invariant case for this model.
+We show that Eq.~\ref{eq:ptilde_eq_chi} holds without translational invariance, where $\mathcal{P}$ now differs from $\chi$ and $\tilde{\mathcal{P}}$.
 
@@
 As mentioned above, the bulk polarization is only defined modulo 1.
-This is obvious when we express it in terms of a geometric phase, which is only defined modulo $2\pi$.
-In our formulation inEq.~\eqref{eq:ptilde_eq_chi}, this is reflected in an ambiguity in the number of bulk modes with $\xi = 1$ that are included in $\chi$.
+This is obvious when it is defined in terms of a geometric phase, which is only defined modulo $2\pi$.
+In our formulation in Eq.~\eqref{eq:ptilde_eq_chi}, this is reflected in an ambiguity in the number of bulk modes with $\xi = 1$ that are included in $\chi$.
 Being defined only modulo $1$, the bulk polarization cannot differentiate the phases $\nu=0$ and $\nu=2$, even though the EOS of both phases is distinct.
@@
 This ambiguity dissapears when we introduce the flux as above in the open chain --- we simply include all the bulk modes. 
-Note that $\Phi$ is not a magnetic flux but should be treated as an additional parameter, and therefore $\tpo$ is not necessarily the physical polarization of the open chain.
+Note that $\Phi$ is not a magnetic flux but should be treated as an additional parameter, and therefore $\tpo$ is not the polarization of the open chain.
 $\tpo$ can be obtained in an even simpler way than $\tilde{\mathcal{P}}$ as
@@
 
-The idea is that in the process of opening the chain, which can be done adiabatically, all eigenvalues that are related to the right virtual edge ($A_R$ in Fig.~\ref{fig:chain}) should get pushed to $\xi=0,1$.
+The idea is that in the process of opening the chain, which can be done adiabatically, all eigenvalues that are related to the right virtual edge ($A_R$ in Fig.~\ref{fig:chain}) are pushed to $\xi=0,1$.
 As a result, the sum of all eigenvalues of $C_{A,o}$ (modulo 1) is equal to $\chi$ \eqref{eq:chi_left}.
@@
 \begin{equation}
-\tpo = \tilde{\mathcal{P}} \, {\rm mod} \, 1.
+\lim_{L \rightarrow \infty }\tilde{\mathcal{P}} = \lim_{L \rightarrow \infty } \tpo \, {\rm mod} \, 1.
 \end{equation}
@@
 The consequence is that for paths $C_\lambda$ in parameter space, for which the energy spectrum is gapped, $\tpo (\lambda)$ is defined in $\mathbb{R}$, such that the total change in $\tpo$ along this path can be obtained knowing $\tpo$ only in the initial and final states, $\Delta \tpo ({C_\lambda}) = \tpo(\lambda_f) - \tpo(\lambda_i)$.
-Note, however, that $\tpo$ is still gauge-invariant only when taken modulo 1.
+Note, however, that $\tpo$ is gauge-invariant only when taken modulo 1.
 This is similar to the case of the Zak phase, which is known to be equal to the winding number for certain gauges (see Appendix~\ref{appendix:winding_ssh}).
@@
 
-This method resolves most of the issues found in section \ref{sec:bulkpolarization}, but unfortunately it cannot be applied when the system is in the topological phase.
+This method resolves most of the issues found in section \ref{sec:bulkpolarization}, but unfortunately it cannot be applied when the system is in a non-trivial topological phase.
 When we open the chain for a system in a topological non-trivial state, zero-energy modes will appear.
@@
 there might be integer jumps in $\tpo$ when the edge-modes cross zero-energy. 
-As we will see below these jumps in $\tpo$are branch cuts that provide the correct description of the system.
+As we will see below these jumps in $\tpo$ are branch cuts that provide the correct description of the system.
 
@@
 
-Consider first the simple case of the SSH chain. When we include the chiral-symmetry breaking parameter $\kappa$ the Hamiltonian now has the symmetry
+Consider first the simple case of the SSH chain. When we include the chiral-symmetry breaking parameter $\kappa$ the Hamiltonian has the symmetry
 \begin{equation}\label{eq:sym}
@@
 
-For simplicity's sake, assume that there is a single branch cut along $\kappa=0$, such as depicted in Fig.s~\ref{ssh_chern} (a) or (b).
-The generalization to more complicated scenarios is straightforward.
+For simplicity, assume that there is a single branch cut along $\kappa=0$, such as depicted in Fig.s~\ref{ssh_chern} (a) or (b). The generalization to more complicated scenarios is straightforward. 
 Consider a counterclockwise loop around the degenerate point at $(m=t,\kappa=0)$ starting from the branch cut parametrized by $\theta$,
@@
 As mentioned above the only contribution to the Chern number comes from discontinuities in the bulk polarization.
-However, when it is computed using $\tilde{\mathcal{P}}$ we no longer know in advance where this discontinuities are present, due to it being defined modulo 1.
+However, when it is computed using $\tilde{\mathcal{P}}$ we no longer know in advance where this discontinuities are located, due to it being defined modulo 1.
 One needs to look at the full path to know how the bulk polarization evolves along the loop.
 
-Since $\tpo$ is naturally described in position space we can also use this method for a disordered system. Consider now that we add disorder in the $t$ and $m$ parameters
+Since $\tpo$ is defined in position space we can also use this method for a disordered system. Assume now that we add disorder in the $t$ and $m$ parameters
 \begin{align}
@@
 where $\omega_i$ and $\omega_i'$ are selected from a uniform distribution with range $[-1/2,1/2]$.
-For very strong disorder where the gap fills with bulk states, $\tpo$ will present many branch cuts as we approach the limit $\lim_{\kappa \rightarrow 0}$, which makes it unpractical.
+For very strong disorder where the gap fills with bulk states, $\tpo$ will present many branch cuts as we approach the limit $\lim_{\kappa \rightarrow 0}$, which makes it impractical.
 In order to avoid this, we break chiral symmetry locally in the edge of the open chain so it only affects the edge modes.
@@
 \begin{align}
-H =& \sum_{i\alpha,j\beta}\sum_{k_y} c_{i\alpha}^\dagger(k_y) H_{i\alpha,j\beta}(k_y) c_{j\beta}(k_y), \\
+H =& \sum_{i\alpha,j\beta}\sum_{k_y} c_{i\alpha}^\dagger(k_y) H_{i\alpha,j\beta}(k_y) c_{j\beta}(k_y), \\ \nonumber
 H_{i\alpha,j\beta}(k_y)=& \frac{1}{2}(i\sigma_x-\sigma_z)\delta_{i,j+1}+\frac{1}{2}(-i\sigma_x-\sigma_z)\delta_{i,j-1} \nonumber\\
 +&(\sin(k_y)\sigma_y+[2-m-\cos(k_y)]\sigma_z) \delta_{ij}.
-\end{align}
+\label{eq:chern_model}
+\end{align}
+
 We can treat the Hamiltonian with elements $H_{i\alpha,j\beta}(k_y)$ as a 1D Hamiltonian with an extra parameter $k_y$ and compute $\tpo$, or equivalently $A_\Phi(\Phi,k_y)$, as in last section.
@@
 \begin{align}
-C(m) =& [\tpo(m,k_y=0^-)-\tpo(m,k_y=-\pi^+)]\\
+C(m) =& [\tpo(m,k_y=0^-)-\tpo(m,k_y=-\pi^+)] \nonumber \\ 
 &+[\tpo(m,k_y=\pi^-)-\tpo(m,k_y=0^+)]. 
@@
 It was previously known that the bulk polarization (Zak phase, or geometric phase for $U(1)$ flux insertion) was encoded in the many-body entanglement spectrum.
-Here we have shown how it can be decoded from the single-particle entanglement spectrum.
+Here we have shown how it can be obtained from the single-particle entanglement spectrum of a non-interacting gapped system.
 This relation is simpler than the one involving the many-body entanglement spectrum and it provides more insight into the ES and EOS.
-In particular in the topological case, where the relation between the quantized bulk polarization (or the Chern number in 2D) and the number of virtual topological edge states is shown in a new way.
-It also provides another simple method for computing the bulk polarization for systems without translational invariance.
-
-Based on this method we show that one can obtain an alternative bulk polarization that is continuous in $\mathbb{R}$ for gapped paths, although it is only gauge invariant when taken modulo 1, where it is equal to the conventional bulk polarization.
-It provides additional information about the edge spectrum, similar to how the Zak phase is equal to the winding number when computed in a particular gauge.
-This alternative bulk polarization simplifies the calculation of changes in the bulk polarization, with a particular application to the computation of Chern numbers.
+In particular in the topological case, the relation between the quantized bulk polarization (or the Chern number in 2D) and the number of virtual topological edge states is demonstrated in a particularly transparent way.
+It also provides a new simple method for computing the bulk polarization for systems without translational invariance.
+
+We also define a novel bulk polarization that is continuous in $\mathbb{R}$ for gapped paths.
+This provides additional information about the edge spectrum, similar to how the Zak phase is equal to the winding number when computed in a particular gauge.
+Our new bulk polarization simplifies the calculation of changes in the bulk polarization, and provides a new route to the computation of Chern numbers.
 
 \emph{Outlook:} In the application of our results to topological systems in 1D we have focused on systems with translational invariance even though our results do not require it.
-For general spatially inhomogeneous systems the same bulk can present different boundaries such that the relation between bulk and boundary is not as straightforward as in the translational invariant case.
-Our results might be able to give more insight into this relation.
+For general spatially inhomogeneous systems the same bulk can support different boundaries so the relation between bulk and boundary is not as straightforward as in the translational invariant case.
+Our results might give more insight into this relation.
 
 In the field of higher-order topological insulators some systems can be characterized by a physical generalization of the bulk polarization to two dimensions, the quadrupole moment \cite{Benalcazar2017,Kang2019}.
-This quadrupole moment is defined with the position operator and there is an open question whether a formulation in terms of twisted boundary conditions can be obtained.
-In this case, one might be able to construct this quadrupole moment using the entanglement spectrum, as we do here for the bulk polarization.
+This quadrupole moment is defined via the position operator and it is an open question whether there is a formulation in terms of twisted boundary conditions.
+In this case, one might be able to construct this quadrupole moment using the entanglement spectrum, as we did
+ here for the bulk polarization.
 
@@
 \begin{align}
-\lambda_{pq} =& \prod_{\mu \in A} (1-\xi_\mu)\left(\frac{\xi_\mu}{1-\xi_\mu}\right)^{n_\mu^{pq}} \\
-=&\left(\prod_{\mu\in A_L} (1-\xi_\mu)\left(\frac{\xi_\mu}{1-\xi_\mu}\right)^{n_\mu^{p}} \right) \\
+\lambda_{pq} =& \prod_{\mu \in A} (1-\xi_\mu)\left(\frac{\xi_\mu}{1-\xi_\mu}\right)^{n_\mu^{pq}}\nonumber \\
+=&\left(\prod_{\mu\in A_L} (1-\xi_\mu)\left(\frac{\xi_\mu}{1-\xi_\mu}\right)^{n_\mu^{p}} \right) \nonumber\\
 &\cross \left(\prod_{\nu \in A_R} (1-\xi_\nu)\left(\frac{\xi_\nu}{1-\xi_\nu}\right)^{n_\nu^{q}} \right) ,
@@
 from which we identify 
-\begin{align*}
-&s_p^2 = \prod_{\mu\in A_L} (1-\xi_\mu)\left(\frac{\xi_\mu}{1-\xi_\mu}\right)^{n_\mu^{p}} \\
+\begin{align}
+&s_p^2 = \prod_{\mu\in A_L} (1-\xi_\mu)\left(\frac{\xi_\mu}{1-\xi_\mu}\right)^{n_\mu^{p}} \nonumber \\
 &s_q^{\prime 2} = \prod_{\nu \in A_R} (1-\xi_\nu)\left(\frac{\xi_\nu}{1-\xi_\nu}\right)^{n_\nu^{q}}.
 \label{eq:spsq_A}
-\end{align*}
+\end{align}
 
@@
 \begin{align}
-\tilde{\mathcal{P}} =& \int_0^{2\pi} \frac{d\Phi}{2\pi}i\bra{\Psi^\Phi}\partial_\Phi\ket{\Psi^\Phi} \\
-=& \int_0^{2\pi} \frac{d\Phi}{2\pi}i\sum_{p,q} s_p^2 s_q^{\prime \, 2} e^{i\Phi N_{A_L}^p}\partial_{\Phi}e^{-i\Phi  N_{A_L}^p}\\
-=&\int_0^{2\pi} \frac{d\Phi}{2\pi}\sum_{p,q} s_p^2 s_q^{\prime \, 2}  N_{A_L}^p\\
+\tilde{\mathcal{P}} =& \int_0^{2\pi} \frac{d\Phi}{2\pi}i\bra{\Psi^\Phi}\partial_\Phi\ket{\Psi^\Phi} \nonumber\\
+=& \int_0^{2\pi} \frac{d\Phi}{2\pi}i\sum_{p,q} s_p^2 s_q^{\prime \, 2} e^{i\Phi N_{A_L}^p}\partial_{\Phi}e^{-i\Phi  N_{A_L}^p}\nonumber\\
+=&\int_0^{2\pi} \frac{d\Phi}{2\pi}\sum_{p,q} s_p^2 s_q^{\prime \, 2}  N_{A_L}^p \nonumber\\
 =&\sum_{p} s_p^2 \sum_{\mu \in A_L}n^p_\mu,
@@
 \begin{align}
-\tilde{\mathcal{P}} =& \sum_{\substack{\{n_{i \neq p \in A_L}\} = 0,1 \\ n_p=0}} \left(\sum_{j\neq p \in A_{L}} n_j \right) (1-\xi_p)\left(\prod_{k \neq p\in A_L} \lambda_k \right) \\
-&+ \sum_{\substack{\{n_{i \neq p \in A_L}\} = 0,1 \\ n_p=1}} \left(\sum_{j\neq p \in A_{L}} n_j+1 \right)\xi_p \left(\prod_{k \neq p\in A_L} \lambda_k \right) \\
-=& \sum_{\substack{\{n_{i \neq p \in A_L}\} = 0,1 }} \left(\sum_{j\neq p \in A_{L}} n_j \right) \left(\prod_{k \neq p\in A_L} \lambda_k \right) \\
+\tilde{\mathcal{P}} =& \sum_{\substack{\{n_{i \neq p \in A_L}\} = 0,1 \\ n_p=0}} \left(\sum_{j\neq p \in A_{L}} n_j \right) (1-\xi_p)\left(\prod_{k \neq p\in A_L} \lambda_k \right) \nonumber \\
+&+ \sum_{\substack{\{n_{i \neq p \in A_L}\} = 0,1 \\ n_p=1}} \left(\sum_{j\neq p \in A_{L}} n_j+1 \right)\xi_p \left(\prod_{k \neq p\in A_L} \lambda_k \right) \nonumber \\
+=& \sum_{\substack{\{n_{i \neq p \in A_L}\} = 0,1 }} \left(\sum_{j\neq p \in A_{L}} n_j \right) \left(\prod_{k \neq p\in A_L} \lambda_k \right) \nonumber \\
 &+ \xi_p \sum_{\substack{\{n_{i \neq p \in A_L}\} = 0,1 }} \left(\prod_{k \neq p\in A_L} \lambda_k \right) 
@@
 \begin{align}
-&(1-\xi_{p'})\sum_{\substack{\{n_{i \neq p,p' \in A_L}\} = 0,1 \\ n_{p'}=0}} \left(\prod_{k \neq p,p'\in A_L} \lambda_k\right) \\
+&(1-\xi_{p'})\sum_{\substack{\{n_{i \neq p,p' \in A_L}\} = 0,1 \\ n_{p'}=0}} \left(\prod_{k \neq p,p'\in A_L} \lambda_k\right) \nonumber \\
 &+ \xi_{p'}\sum_{\substack{\{n_{i \neq p,p' \in A_L}\} = 0,1 \\ n_{p'}=1}} \left(\prod_{k \neq p,p'\in A_L}\lambda_k\right) =1,
@@
 Therefore,
-\begin{align*}
+\begin{align}
 U (D^2)^{-1/2} U^\dagger =&(U D^2 U^\dagger)^{-1/2} 
-\end{align*}
+\end{align}
 and we conclude that
@@
 The Berry connections of the occupied state, $A_\lambda = \bra{-}\partial_\lambda \ket{-}$, can be obtained as
-\begin{align*}
+\begin{align}
 A_{\lambda}=& i\frac{\cos(\theta)-1}{2}\partial_q \phi \\
-=&i\frac{\cos(\theta)-1}{2} \cos(\phi)^2 \partial_q  \tan(\phi) \\
-=&\frac{i}{2}\left(\frac{\kappa}{\sqrt{\abs{f(k)}^2+\kappa^2}}-1\right) \left( \frac{\Re [f(k)]}{\abs{f(k)}}\right)^2 \partial_\lambda  \left( -\frac{\Im [f(k)]}{\Re [f(k)]} \right) \\
-=&\frac{-1}{2}\left(\frac{\kappa}{\sqrt{\abs{f(k)}^2+\kappa^2}}-1\right) \frac{f(k)^\dagger \partial_\lambda f(k)-f(k)\partial_\lambda f(k)^\dagger}{2\abs{f(k)}^2}
-\end{align*}
+=&i\frac{\cos(\theta)-1}{2} \cos(\phi)^2 \partial_q  \tan(\phi) \nonumber \\
+=&\frac{i}{2}\left(\frac{\kappa}{\sqrt{\abs{f(k)}^2+\kappa^2}}-1\right) \left( \frac{\Re [f(k)]}{\abs{f(k)}}\right)^2 \partial_\lambda  \left( -\frac{\Im [f(k)]}{\Re [f(k)]} \right) \nonumber \\
+=&\frac{-1}{2}\left(\frac{\kappa}{\sqrt{\abs{f(k)}^2+\kappa^2}}-1\right) \frac{f(k)^\dagger \partial_\lambda f(k)-f(k)\partial_\lambda f(k)^\dagger}{2\abs{f(k)}^2} \nonumber
+\end{align}
 The Berry connection with respect to momentum gives
@@
 Note that in the chiral symmetric case, for $\kappa = 0$, the Zak phase is
-\begin{align*}
-\gamma =& \int_0^{2\pi} dk\,\frac{1}{2}\frac{f(k)^\dagger \partial_k f(k)-f(k)\partial_k f(k)^\dagger}{2\abs{f(k)}^2} \\
+\begin{align}
+\gamma =& \int_0^{2\pi} dk\,\frac{1}{2}\frac{f(k)^\dagger \partial_k f(k)-f(k)\partial_k f(k)^\dagger}{2\abs{f(k)}^2} \nonumber \\
 =& \int_0^{2\pi} dk\,\frac{1}{2}q(k)^\dagger \partial_k q(k),
-\end{align*}
+\end{align}
 where $q(k) = f(k)/\abs{f(k)}$. This is nothing else than the winding number \cite{ryu2010topological}.

### 2021-02-03 version2/draft_mh.tex
added acknowledgement

@@
 {\em Acknowledgments.--} 
-FUNDING
-We thank Krishanu XX and Thors Hans Hansson for a critical reading of the manuscript.
+The research in this grant was supported by the Swedish Research Council under grant no. 2017-05162 and the Knut and Alice Wallenberg foundation under grant no. 2017.0157. 
+We thank Krishanu Roychowdhury and Thors Hans Hansson for helpful discussions and a critical reading of the manuscript.
 

### 2021-02-03 version2/draft_mh.tex
Add files via upload

@@
 We show that the bulk polarization can always be determined from the single-particle entanglement spectrum, even in the absence of symmetries that quantize it.
-In the symmetric case, the known relation between the bulk polarization and the number of topological edge modes is recovered.
+In the symmetric case, the known relation between the bulk polarization and the number of virtual topological edge modes is recovered.
 We use the bulk polarization to compute Chern numbers in 1D and 2D, which illuminates their known relation to the entanglement spectrum.
@@
 The bulk, or macroscopic, polarization is a fundamental concept in physics, primarily used to describe the response of matter to electric fields.
-The modern theory of polarization \cite{Resta1992,KingSmith1993,Vanderbilt1993,Resta1997} related the bulk polarization to a geometric phase (Zak phase for translationally invariant systems).
+The modern theory of polarization \cite{Resta1992,KingSmith1993,Vanderbilt1993,Resta1997} related the bulk polarization to a geometric phase, which is nothing but the Zak phase for translationally invariant systems.
 Due to this, it has also found its way into topological physics and in topological charge pumping.
-In certain symmetry classes the bulk polarization is quantized and serves as a $\mathbb{Z}_2$ topological invariant, where it is known to be related to the number of protected topological edge state in the ES \cite{Ryu2006}.
+In certain symmetry classes the bulk polarization is quantized and serves as a $\mathbb{Z}_2$ topological invariant, where it is known to be related to certain feature of the ES \cite{Ryu2006}.
 Because of its relation to a geometric phase, the bulk polarization is also related to other topological invariants in higher dimensions such as the Chern number.
 
-We show how the bulk polarization can be decoded from the ES.
-This is a general property of non-interacting 1D gapped systems, which in the topological case reduces to the known relation between the bulk polarization and the ES described above \cite{Ryu2006}.
+We show that there is much more information encoded in the \emph{single-particle} ES than was previously known.
+In particular, we show how the bulk polarization can be decoded from the single-particle ES, even when it is not quantized by the symmetries.
+This is a general property of non-interacting 1D gapped systems.
 We also apply this method to compute Chern numbers directly from the entanglement spectrum.
@@
 In Ref.~\cite{Zaletel2014}, the authors show that the Zak phase can be computed from the Schmidt decomposition of a translation invariant, infinite chain.
-This suggests that there should be a direct relation between the single-particle entanglement energies and the Zak phase for non-interacting systems.
+The latter is related to the many-particle entanglement spectrum.
+This suggests that, for non-interacting systems, there should be a direct relation between the Zak phase and the single-particle entanglement spectrum, which is substantially easier to compute than the many-body entanglement spectrum.
 A particular limit of our result for homogeneous (translationally invariant) systems~\eqref{eq:chi_integrable} was derived
@@
 for a fully dimerized SSH chain~\cite{Ryu2006}. 
-The similarity between the behavior of the Zak phase and the protected topological edge states in the entanglement spectrum of certain Chern insulators, which was observed in Ref.~\cite{Huang2012,Huang2012-2}, is also explained by our results.
+The similarity between the behavior of the Zak phase and the entanglement spectrum of certain Chern insulators, which was observed in Ref.~\cite{Huang2012,Huang2012-2}, is also explained by our results.
 
@@
 
-The Zak phase is only gauge invariant modulo $2\pi$.
+Since the Berry connection is not necessarily continuous, the Zak phase is only gauge invariant modulo $2\pi$.
 However, for two different states defined by a parameter $\lambda$ --- assuming $A_k(\lambda)$ is smooth in the path connecting them --- the change in the Zak phase can be computed as
@@
 
-whose derivatives give the average current along the chain and the current at the seam (where the PBC are implemented), respectively.
-
 If the flux is introduced homogeneously through a vector potential $A_x = \Phi/L$ we can define the polarization
@@
 
-If the flux is introduced via twisted boundary conditions at the seam, in the bond between sites $j=1$ and $j=L$, this is equivalent to performing the gauge transformation
+If the flux is introduced via twisted boundary conditions at the seam, i.e. in the bond between sites $j=1$ and $j=L$, this is equivalent to performing the gauge transformation
 \begin{align}\label{eq:Psi_tilde}
@@
 Following Ref.s~\cite{Huang2012,Huang2012-2}, we refer to this spectrum as the entanglement occupancy spectrum (EOS).
-We denote the eigenvalues of the EOS by $\xi_j$. 
-Topological phases are characterized by (symmetry-protected) zero-energy modes in the edge spectrum and $\xi=\frac 1 2$ modes in the EOS. 
+We denote the eigenvalues of the EOS by $\xi_j$. When restricted to sub-region A, the boundaries of A act as virtual boundaries to the system.
+Topological phases are characterized by (symmetry-protected) zero-energy modes in the edge spectrum and $\xi=\frac 1 2$ virtual edge modes in the EOS. 
 Consequently, the EOS is in one-to-one correspondence to the entanglement spectrum (ES)~\cite{Fidkowski2010entanglement}.
@@
 the corresponding eigenstates are the same. 
-Eq.~\eqref{eq:xi_eps} implies that, for non-interacting systems, the full information of the \emph{many-body} ES (MBES) is contained in the single-particle spectrum of the reduced correlation matrix. 
+Eq.~\eqref{eq:xi_eps} implies that, for non-interacting systems, the full information of the many-body ES is contained in the single-particle spectrum of the reduced correlation matrix. 
 The latter is much simpler to interpret. 
@@
 
-One property of the subsystem correlation matrix we will make use of later is that, for sufficiently large systems, its eigenstates are found to be exponentially localized on either virtual edge (edges of sub-region A) if $\xi$ is away from $0$ and $1$, or they are found to be bulk modes if the correspondent eigenvalues are exponentially close to $\xi = 0,1$ \cite{Peschel2008}.
+One property of the subsystem correlation matrix we will make use of later is that, for sufficiently large systems, its eigenstates are found to be exponentially localized on either virtual edge if $\xi$ is away from $0$ and $1$, or they are found to be bulk modes if the correspondent eigenvalues are exponentially close to $\xi = 0,1$ \cite{Peschel2008}.
 
@@
 The authors could show that the value of one of these `midgap states' is identical to the Zak phase divided by $2\pi$ --- or alternatively $\tilde{\mathcal{P}}^{\rm Bloch}$ --- although they do not specify to which of the two eigenvalues it corresponds.
-Their result is a particular limit of equation~\eqref{eq:chi_integrable}. 
+Their result is a particular limit of equation~\eqref{eq:chi_integrable}, which we discuss in the next section. 
 %
@@
 \end{equation}
-When introducing the flux in the system with PBC (See Eq.~\eqref{eq:Ground_state_flux}) there is an ambiguity due to the bulk modes. 
+When introducing the flux in the system with periodic boundary conditions (See Eq.~\eqref{eq:Ground_state_flux}) there is an ambiguity due to the bulk modes. 
 This ambiguity dissapears when we introduce the flux as above in the open chain --- we simply include all the bulk modes. 
@@
 We can treat the Hamiltonian with elements $H_{i\alpha,j\beta}(k_y)$ as a 1D Hamiltonian with an extra parameter $k_y$ and compute $\tpo$, or equivalently $A_\Phi(\Phi,k_y)$, as in last section.
-But first lets analyze the EOS of the system with PBC, shown in Fig.~\ref{chern_insulator}(a) and (b) for $m=1$ and $3$, respectively.
+But first lets analyze the EOS of the system with periodic boundary conditions, shown in Fig.~\ref{chern_insulator}(a) and (b) for $m=1$ and $3$, respectively.
 For $0<m<2$ the EOS presents two edge modes that connect the occupied and empty bands that cross $\xi=1/2$ at $k_y=0$.
@@
 
-It was previously known that the bulk polarization (Zak phase, or geometric phase for $U(1)$ flux insertion) was encoded in the many-body entanglement spectrum.
-Here we have shown how it can be obtained from the single-particle entanglement spectrum of a non-interacting gapped system.
-This relation is simpler than the one involving the many-body entanglement spectrum and it provides more insight into the ES and EOS.
+It was previously known that the bulk polarization (Zak phase, or geometric phase for $U(1)$ flux insertion) was encoded in the many-body ES.
+Here we have shown how it can be obtained from the single-particle entanglement spectrum of a non-interacting gapped system, showing that there is substantially more information in the single-particle ES than previously realized.
+This relation is simpler than the one involving the many-body ES, which is much more difficult to compute, and it provides more insight into the ES and EOS.
 In particular in the topological case, the relation between the quantized bulk polarization (or the Chern number in 2D) and the number of virtual topological edge states is demonstrated in a particularly transparent way.
@@
 Our results might give more insight into this relation.
+
+As it stands, the relation derived in this paper only works for non-interacting systems.
+With interactions, the correlation matrix does not have the full information of the ground state.
+However, there might be certain systems, perhaps for weakly interacting systems, where the correlation matrix encodes the topological information.
+This would be worth exploring since, even for interacting systems, the correlation matrix is much easier to compute than the density matrix.
 

### 2021-02-03 version2/draft_mh.tex
small changes in the outlook

@@
 
-\title{On the relation of the entanglement spectrum to the bulk polarization \\ DRAFT}
+\title{On the relation of the entanglement spectrum to the bulk polarization}
 \author{Carlos Ortega-Taberner}
@@
 Here we have shown how it can be obtained from the single-particle entanglement spectrum of a non-interacting gapped system, showing that there is substantially more information in the single-particle ES than previously realized.
-This relation is simpler than the one involving the many-body ES, which is much more difficult to compute, and it provides more insight into the ES and EOS.
+Our formulation is both simpler than the one involving the many-body ES, which is much more difficult to compute, and it provides additional insight into the ES and EOS.
 In particular in the topological case, the relation between the quantized bulk polarization (or the Chern number in 2D) and the number of virtual topological edge states is demonstrated in a particularly transparent way.

### 2021-04-14 polarization_ES_prb_submission/polarization_ES.tex
pro submission

--- 2021-02-03 version2/draft_mh.tex
@@
+\pdfoutput=1
 \documentclass[twocolumn,amsmath,longbibliography,amssymb,superscriptaddress]{revtex4-1}
@@
 
-\newcommand{\carlos}[1]{{\color{red} #1}}
-\newcommand{\maria}[1]{{\color{blue} #1}}
-\newcommand{\mariac}[1]{{\it\color{cyan}#1}}
-\newcommand{\changed}[1]{{\color{orange}#1}}
 \newcommand{\tpo}{\tilde{\mathcal{P}}_{\rm o}}
@@
 There are different tools based on entanglement that have been used to characterize topological phases.
-A very efficient one is the entanglement entropy~\cite{srednicki1993entropy}, which allows one to determine the total quantum dimension of the underlying topological quantum field theory~\cite{Kitaev2006topological, Levin2006detecting}, but it can only be used for topologically ordered phases and it cannot uniquely characterize the topological phase at hand.
+A very efficient one is the entanglement entropy~\cite{srednicki1993entropy}, which allows one to determine the total quantum dimension of the underlying topological quantum field theory~\cite{Kitaev2006topological, Levin2006detecting}.
+However, it can only be used for topologically ordered phases and it cannot uniquely characterize the topological phase at hand.
 Another, closely related, tool is the entanglement spectrum (ES), originally introduced for fractional quantum Hall systems~\cite{Li2008entanglement}.
@@
 However, even for non-interacting systems, it is unclear which information (beyond the protected `edge' spectrum) is encoded in the ES.
-The aim of this paper is to show that other physical properties are also encoded in the entanglement spectrum, in particular the bulk polarization.
+The aim of this paper is to show that other physical properties are also encoded in the ES, in particular the bulk polarization.
 
@@
 The modern theory of polarization \cite{Resta1992,KingSmith1993,Vanderbilt1993,Resta1997} related the bulk polarization to a geometric phase, which is nothing but the Zak phase for translationally invariant systems.
-Due to this, it has also found its way into topological physics and in topological charge pumping.
+Due to this, it has also found its way into topological physics \cite{Kudin2007}.
 In certain symmetry classes the bulk polarization is quantized and serves as a $\mathbb{Z}_2$ topological invariant, where it is known to be related to certain feature of the ES \cite{Ryu2006}.
@@
 This is a general property of non-interacting 1D gapped systems.
-We also apply this method to compute Chern numbers directly from the entanglement spectrum.
+We also apply this method to compute Chern numbers directly from the single-particle ES.
 The differences in the bulk polarization, which can be any real number, are the relevant quantities to study.
 They can, however, be difficult to compute from the bulk polarization itself, as it is only defined modulo $1$.
-We show how, using our method, one can define an alternative bulk polarization computed by using open boundary conditions which is continuous in $\mathbb{R}$ for gapped paths, simplifying the calculation of polarization differences.
-The bulk polarization defined this way can give more information about the edge spectrum than the usual one, and can be used to simplify the computation of Chern numbers.
+We show how, using our method, one can define an alternative bulk polarization, computed by using open boundary conditions, which is continuous in $\mathbb{R}$ for gapped paths in parameter space, simplifying the calculation of polarization differences.
+The bulk polarization defined this way can give more information about the edge spectrum than the conventional one, and can be used to simplify the computation of Chern numbers.
 We also discuss the relation between this alternative bulk polarization and another topological invariant known as the trace index \cite{Alexandrinata2011}.
 
-The relation between the entanglement spectrum and the bulk polarization has already been studied. 
-In Ref.~\cite{Zaletel2014}, the authors show that the Zak phase can be computed from the Schmidt decomposition of a translation invariant, infinite chain.
-The latter is related to the many-particle entanglement spectrum.
-This suggests that, for non-interacting systems, there should be a direct relation between the Zak phase and the single-particle entanglement spectrum, which is substantially easier to compute than the many-body entanglement spectrum.
-A particular limit of our result for homogeneous (translationally invariant) systems~\eqref{eq:chi_integrable} was derived
-%by Ryu and Hatsugai 
-for a fully dimerized SSH chain~\cite{Ryu2006}. 
-The similarity between the behavior of the Zak phase and the entanglement spectrum of certain Chern insulators, which was observed in Ref.~\cite{Huang2012,Huang2012-2}, is also explained by our results.
+The relation between the ES and the bulk polarization has already been studied. 
+In Ref.~\cite{Zaletel2014}, the authors show that the Zak phase can be computed from the Schmidt decomposition of a translationally invariant, infinite chain. The latter is related to the ES.
+This suggests that, for non-interacting systems, there should be a direct relation between the Zak phase and the single-particle ES, which is substantially easier to compute than the ES.
+A particular limit of our result was derived for a fully dimerized SSH chain~\cite{Ryu2006}.
+The similarity between the behavior of the Zak phase and the ES of certain Chern insulators, which was observed in Ref.~\cite{Huang2012,Huang2012-2}, is also explained by our results.
 
 \emph{Outline of the paper}: In section II we introduce the 1D model used throughout the paper and we examine its phase diagram.
-In section III we discuss the bulk polarization; the different ways one can define it and its relation to the Zak phase.
+In section III we discuss the bulk polarization; the different ways one can define it and its relation to the geometric phases.
 The ES is introduced in section IV, where we also set our notation.
@@
 Figure~\ref{fig:zero_E_modes}(a) shows the edge spectrum in the BDI class, with 4 (resp. 2) symmetry-protected zero modes for winding number 2 (1).
-%We also a cut through the three phases (dotted line in the phase diagram) for different values of $\kappa,\kappa'$.
 
@@
 With both $\kappa \neq 0$ and $\kappa' \neq0$ the system has no local symmetry, thus belonging to the $A$ class.  
-Also this symmetry class is topologically trivial in one dimension. 
+This symmetry class is also topologically trivial in one dimension. 
 Fig.~\ref{fig:zero_E_modes}(c) and (d) show the edge spectrum for class AI and A, respectively. In both cases, the edge modes are split from zero for all values of $m$.
-
-\iffalse
-\begin{figure}[h!]
-\centering
-\makebox[0pt]{
-\subfloat[$\kappa =\kappa'=0$]{
-  \includegraphics[width=43mm]{fig21.pdf}
-}
-\subfloat[$\kappa = 0, \kappa'=0.3$]{\includegraphics[width=43mm]{fig22.pdf}
-}
-}\hspace{0mm}
-
-\makebox[0pt]{
-\subfloat[$\kappa = 0.3,\kappa'=0$]{
-  \includegraphics[width=43mm]{fig23.pdf}
-}
-\subfloat[$\kappa = 0.3,\kappa'=0.3$]{\includegraphics[width=43mm]{fig24.pdf}
-}
-}
-\caption{Edge spectrum along the dotted path (i.e. $t=1,t'=-2$) of Fig.~\eqref{fig:bdi_phase_diagram} for different values of $\kappa$ and $\kappa'$.}
-\label{fig:zero_E_modes}
-\end{figure}
-\fi
 
@@
 \begin{table}[h]
-\centering % centering table
-\begin{tabular}{|c |rrr| c|c|} % creating eight columns
-\hline\hline %inserting double-line
+\centering 
+\begin{tabular}{|c |rrr| c|c|} 
+\hline\hline 
 Model &  $T$ & $C$ & $S$ & Class & inv.  \\  [1ex] 
@@
 $\kappa \neq 0,\kappa' \neq 0$ &  0 & 0 & 0 & A & 0  \\  [1ex] 
-\hline \hline%title of the table
+\hline \hline
 \end{tabular}
@@
 % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % 
-\section{Geometric phases and polarization}
+\section{Geometric phases and bulk polarization}
 
@@
 
-Since the Berry connection is not necessarily continuous, the Zak phase is only gauge invariant modulo $2\pi$.
+The Zak phase is only gauge invariant modulo $2\pi$.
 However, for two different states defined by a parameter $\lambda$ --- assuming $A_k(\lambda)$ is smooth in the path connecting them --- the change in the Zak phase can be computed as
@@
 \end{equation} 
-%The previous discussion about the Zak phase being defined only modulo $2\pi$ although its changes can be defined as a real number applies in the same way to the polarization obtained in \eqref{eq:p_tilde}.
 Similar to the discussion about the Bloch polarization below Eq.~\eqref{eq:Bloch_polarization}, both $\mathcal{P}$ and $\tilde{\mathcal{P}}$ are defined modulo 1, while their changes are defined in $\mathbb{R}$.
@@
 
-Let us now proceed to discuss the entanglement spectrum. 
+Let us now proceed to discuss the ES. 
 An important quantity to consider is the correlation matrix, which in position space is defined by the ground state expectation value
@@
 
-In this paper, we are mainly interested in the spectrum of the correlation matrix when restricted to a spatial sub-region $A$ (its complement will be denoted by $B$ in the following), see Fig.~\ref{fig:chain}.
+In this paper, we are mainly interested in the spectrum of the correlation matrix when restricted to a spatial sub-region $A$ (its complement will be denoted by $B$ in the following), see figure~\ref{fig:chain}.
+When restricted to sub-region A, the boundaries of A act as virtual boundaries to the system.
 Following Ref.s~\cite{Huang2012,Huang2012-2}, we refer to this spectrum as the entanglement occupancy spectrum (EOS).
-We denote the eigenvalues of the EOS by $\xi_j$. When restricted to sub-region A, the boundaries of A act as virtual boundaries to the system.
+We denote the eigenvalues of the EOS by $\xi_j$. 
 Topological phases are characterized by (symmetry-protected) zero-energy modes in the edge spectrum and $\xi=\frac 1 2$ virtual edge modes in the EOS.
-Consequently, the EOS is in one-to-one correspondence to the entanglement spectrum (ES)~\cite{Fidkowski2010entanglement}.
 
@@
 where $\mathcal{K}$ is a normalization constant and $\mathcal{H}$ is a quadratic Hamiltonian, referred to as entanglement Hamiltonian. 
-The single-particle eigenvalues $\epsilon_j$ of $\mathcal{H}$, referred to as entanglement energies,  are related to those of the EOS by 
+The spectrum of the reduced density matrix is the ES.
+The single-particle eigenvalues $\epsilon_j$ of $\mathcal{H}$, referred to as entanglement energies, form the single-particle ES.
+They are related to those of the EOS by
 \begin{align}\label{eq:xi_eps}
@@
 \end{align}
-the corresponding eigenstates are the same. 
-Eq.~\eqref{eq:xi_eps} implies that, for non-interacting systems, the full information of the many-body ES is contained in the single-particle spectrum of the reduced correlation matrix. 
+the corresponding eigenstates are the same. Consequently, the EOS is in one-to-one correspondence to the single-particle ES \cite{Fidkowski2010entanglement}.
+Eq.~\eqref{eq:xi_eps} implies that, for non-interacting systems, the full information of the ES is contained in the spectrum of the subsystem correlation matrix. 
 The latter is much simpler to interpret. 
@@
 One property of the subsystem correlation matrix we will make use of later is that, for sufficiently large systems, its eigenstates are found to be exponentially localized on either virtual edge if $\xi$ is away from $0$ and $1$, or they are found to be bulk modes if the correspondent eigenvalues are exponentially close to $\xi = 0,1$ \cite{Peschel2008}.
-
-%The single-particle entanglement spectrum (ES) is then obtained as the spectrum of the subsystem correlation function obtained after bisecting the system into two equal parts.
-
-%In the ES we can differentiate between two types of eigenvalues, which we denote by $\xi$.
-The ones at $\xi_b = 0,1$ correspond to bulk states while for the ones in between, $0<\xi_{l,r}<1$, we find their correspondent eigenstate localized in either the left, $\xi_l$, or right, $\xi_r$, virtual edge \cite{Peschel2008}.
-%\mariac{introduce entanglement energy, relation between C and energies}
 
@@
 Their result is a particular limit of equation~\eqref{eq:chi_integrable}, which we discuss in the next section.
-%
-%Beyond symmetry protected systems, there was observed that in the fully dimerized SSH chain, with only a single mode present in the ES, the Zak phase is equal to the midgap eigenvalue \cite{Ryu2006}.
 
 In Ref.s~\cite{Huang2012,Huang2012-2} the authors considered a two-dimensional Chern insulator and noted that there was a similarity between the Zak phase and the virtual topological edge states that connect the EOS values at 0 to those at 1.
-Interpreting this as a one-dimensional system with a parameter, we can explain this behavior by Eq.~\eqref{eq:chi_integrable}, noting that there is a single pair of eigenvalues that dominates the sum.
+Interpreting this as a one-dimensional system with a parameter, we can explain this behavior by Eq.~\eqref{eq:chi_left}, noting that there is a single pair of eigenvalues that dominates the sum. 
 Note that the observation of Ref.s~\cite{Huang2012,Huang2012-2} is  particular to systems with Chern number 0 or $\pm 1$, i.e. where there are only few midgap states in the EOS. 
-It fails for systems with higher Chern numbers,  for which there are several terms in Eq.~\eqref{eq:chi_integrable} with comparably large contributions and, consequently, the Zak phase deviates considerably from the virtual topological edge states.
+It fails for systems with higher Chern numbers,  for which there are several terms in Eq.~\eqref{eq:chi_left} with comparably large contributions and, consequently, the Zak phase deviates considerably from the virtual topological edge states.
 
@@
 
-We first discuss systems for which the entanglement energies of a \emph{either} virtual edge are equispaced.
-That is, they are given in the thermodynamic limit by $\varepsilon_{n\alpha} = \varepsilon_{0\alpha}+n\delta_\alpha$, where $\alpha = L,R$.
+We first discuss systems for which the entanglement energies of \emph{either} virtual edge are equispaced.
+That is, they are given in the thermodynamic limit by $\varepsilon_{n\alpha} = \varepsilon_{0\alpha}+n\delta_\alpha$, where $\alpha = L,R$ labels the two virtual edges.
 This is a feature of integrable systems \cite{Peschel1999},  e.g. nearest neighbor hopping models such as the SSH chain, and it applies to the Hamiltonian in Eq.~\eqref{bdi_model} when $t'=0$. 
@@
 \end{equation}
-%This generalizes the observation made by Ryu and Hatsugai \cite{Ryu2006} beyond the fully dimerized limit, $m=0$. 
+
 In the special case where there is only one eigenvalue per edge that is $\neq 0,1$ in  the EOS, like in the fully dimerized limit of the SSH chain, we recover the equality between one of the two midgap eigenvalues and the Zak phase found in reference ~\cite{Ryu2006}.
@@
 The issue concerning the sign ambiguity will be discussed and resolved in the next section. 
-The origin of this ambiguity can be traced back to certain properties of the EOS. 
-In particular, for translational invariant systems the EOS is symmetric such that a crossing of the eigenvalues $\xi$ results in $\chi$ changing sign. 
-This sign is irrelevant in the case of symmetry-protected states where the Zak phase is quantized to $0$ or $\pi$.
+The origin of this ambiguity can be traced back to certain properties of the EOS. This sign is irrelevant in the case of symmetry-protected states where the Zak phase is quantized to $0$ or $\pi$.
 
@@
 Here we see explicitly the known relation between the Zak phase and the number of $\xi = 1/2$ modes.
-Note that the full sum  in Eq.~\eqref{eq:chi_integrable}  evaluates to zero in the trivial regime %up to finite-size corrections% \carlos{It is actually exact due to symmetry}.
-despite the presence of modes with entanglement occupancy $0<\xi<1$.
+Note that the full sum  in Eq.~\eqref{eq:chi_integrable}  evaluates to zero in the trivial regime despite the presence of modes with entanglement occupancy $0<\xi<1$.
 
@@
 Note that, even with the broken symmetry, the bulk polarization shows plateaus that point to a possible nearby topological transition.
-For small $m$, there are two dominant midgap eigenvalues, and the bulk polarization follows the lower one very closely, see .
+For small $m$, there are two dominant midgap eigenvalues, and the bulk polarization follows the lower one very closely.
 When this midgap eigenvalue approaches  0, the contribution from the other modes becomes more relevant and the bulk polarization starts to deviate from the dominant eigenvalue, see figure~\ref{huang}(c).
@@
 Computed for $L=40$.
-In (c) we zoom in a region of the plot in (b) to show the splitting between $\tilde{\mathcal{P}}^{\rm Bloch}$ and the eigenvalue closest to $\xi=1/2$.
+In (c) we zoom in on a region of the plot in (b) to show the splitting between $\tilde{\mathcal{P}}^{\rm Bloch}$ and the eigenvalue closest to $\xi=1/2$.
 In (d) we show a scaling plot of the difference between $\chi$ and $\tilde{\mathcal{P}}^{\rm Bloch}$ for parameters $t=1,t'=0,m=1$, $\kappa = 0.3, \kappa'=0$ and increasing system size to show that the difference vanishes in the thermodynamic limit.
@@
 
-In general we find (see Appendix \ref{app:pollmann}) that it is only the eigenstates that localize on the left virtual edge ($A_L$ in Fig.\ref{fig:chain}) that should contribute to $\chi$.
+For general non-interacting gapped systems we find (see Appendix \ref{app:pollmann}) that it is only the eigenstates that localize on the left virtual edge ($A_L$ in Fig.\ref{fig:chain}) that should contribute to $\chi$.
 The inclusion of bulk modes is irrelevant since they have $\xi = 0,1$.
@@
 \includegraphics[width=86mm]{fig4comp.pdf}
-\caption{We show the Entanglement spectrum (black), $\chi$ (green) and $\tilde{\mathcal{P}}$ (red) computed for $L=40$ sites.
+\caption{We show the EOS (black), $\chi$ (green) and $\tilde{\mathcal{P}}$ (red) computed for $L=40$ sites.
 (a) and (b) for the continuous line in the phase diagram, $t=1,t'=2$  with the symmetry-breaking parameters (a) $\kappa=0.3,\kappa'=0$ and (b) $\kappa=0,\kappa'=0.3$.
@@
 In figures~\ref{2}(c) and (d), we consider the same symmetry breaking terms, but now for $t=1$ and $t'=-2$, where the BDI systems shows a phase transition  $\nu=2\rightarrow 1\rightarrow 0$. 
-Note that the $\kappa'$ term does not render the $\nu=1$ phase trivial, and the Zak phase will remain quantized in that region, see figure~\ref{2}(d). 
 Also here, Eq.~\eqref{eq:chi_left} correctly reproduces the Zak phase for all values of $m$. 
-Note that in the $\nu=2$ phase the observation made in Ref. \cite{Huang2012}, i.e. that the there is a similarity between the Zak phase and the entanglement occupancy, does not apply anymore.
+Note that in Fig.s~\ref{2}(a) and (b) the observation made in Ref. \cite{Huang2012}, i.e. that the there is a similarity between the Zak phase and the EOS, does not apply anymore.
 
@@
 We show this in Fig.~\ref{2}(e) where we show the EOS, $\chi$, $\mathcal{P}$ and $\tilde{\mathcal{P}}$ along the dotted cut in the phase diagram with an added position dependent symmetry breaking term $\kappa_i = (i+L/4 \, {\rm mod}\, L)/L$.
-We show that Eq.~\ref{eq:ptilde_eq_chi} holds without translational invariance, where $\mathcal{P}$ now differs from $\chi$ and $\tilde{\mathcal{P}}$.
+We show that Eq.~\eqref{eq:ptilde_eq_chi} holds without translational invariance, where $\mathcal{P}$ now differs from $\chi$ and $\tilde{\mathcal{P}}$.
 
-This relation between the Zak phase and the EOS is a consequence of an identity between the Zak phase and the many-body entanglement spectrum previously found for infinite chains \cite{Zaletel2014}.
+This relation between the Zak phase and the EOS is a consequence of an identity between the Zak phase and the many-body ES previously found for infinite chains \cite{Zaletel2014}.
 In Appendix A we modified this derivation to account for the periodic boundary conditions we use and show how their result simplifies greatly when expressing it in terms of the EOS, resulting in equation~\eqref{eq:ptilde_eq_chi}.
@@
 When introducing the flux in the system with periodic boundary conditions (See Eq.~\eqref{eq:Ground_state_flux}) there is an ambiguity due to the bulk modes. 
-This ambiguity dissapears when we introduce the flux as above in the open chain --- we simply include all the bulk modes. 
+This ambiguity disappears when we introduce the flux as above in the open chain --- we simply include all the bulk modes. 
 Note that $\Phi$ is not a magnetic flux but should be treated as an additional parameter, and therefore $\tpo$ is not the polarization of the open chain.
@@
 One might think that this problem can be circumvented by introducing symmetry-breaking term that render the system trivial and extrapolating $\tpo$ towards the topological phase. 
-However, this extrapolation is again not necessarily unique, as 
-%The second issue is that, even if we allow for symmetry-breaking terms so that we can compute $\tpo$, 
-there might be integer jumps in $\tpo$ when the edge-modes cross zero-energy. 
+However, this extrapolation is again not necessarily unique, as there might be integer jumps in $\tpo$ when the edge-modes cross zero-energy. 
 As we will see below these jumps in $\tpo$ are branch cuts that provide the correct description of the system.
@@
 The relation between the zero-modes and discontinuities in $\Tr[C_{A,o}]$ was already observed in the context of Chern insulators \cite{Alexandrinata2011}.
-This branch cut is in fact not an artifice of the open chain trick, but the result of the degenerate point at $(m=1,\kappa=0)$ having an associated Chern number $C=1$ \cite{Asboth2016}.
+The appearance of the branch cut is the result of the degenerate point at $(m=1,\kappa=0)$ having an associated Chern number $C=1$ \cite{Asboth2016}.
 
-For simplicity, assume that there is a single branch cut along $\kappa=0$, such as depicted in Fig.s~\ref{ssh_chern} (a) or (b). The generalization to more complicated scenarios is straightforward. 
+For simplicity, assume that the branch cuts appear along $\kappa=0$, such as depicted in Fig.s~\ref{ssh_chern} (a) or (b). The generalization to more complicated scenarios is straightforward. 
 Consider a counterclockwise loop around the degenerate point at $(m=t,\kappa=0)$ starting from the branch cut parametrized by $\theta$,
@@
 \begin{align}
-\brapsio{\Phi(\theta)}\partial_\theta \ketpsio{\Phi(\theta)} =& \sum_{lj} c^*_l(\theta) [\partial_\theta c_j(\theta)] \bra{l} e^{i\Phi (N_A^lN_A^j)}\ket{j}\nonumber\\
+\brapsio{\Phi(\theta)}\partial_\theta \ketpsio{\Phi(\theta)} =& \sum_{lj} c^*_l(\theta) [\partial_\theta c_j(\theta)] \bra{l} e^{i\Phi (N_A^l-N_A^j)}\ket{j}\nonumber\\
 =& \sum_j  c^*_j(\theta) \partial_\theta c_j(\theta) \nonumber\\
@@
 For this model this equivalence is exact and the jump of $\tpo$ at $\kappa = 0$ gives the number of topological zero-modes, i.e. the winding number. 
-%It is also interesting to point out that even far away from the symmetric point, $\tpo$ still shows plateaus that point to a possible nearby topological transition.
 Note, however, that there are many ways of breaking the chiral symmetry and for some of them, $\tpo$ may not be related to the winding number any longer. 
@@
 If any of the other states cross $E=0$ they do it in  chiral-symmetric pairs such that $\tpo$ is not affected by it.
-We show $\tpo$ in Fig.\ref{disorder_chern}(a) using the local symmetry breaking term for a strong disorder, $W=6t$, where the gap is indeed filled with states.
+We show $\tpo$ in Fig.\ref{disorder_chern}(a) using the local symmetry breaking term for a strong disorder, $W=3t$, where the gap is indeed filled with states.
 We see that it works just as in the case without disorder, where now due to the disorder a region of $\nu=1$ opens up \cite{Song2014}.
 The main difference is that since we break chiral symmetry only for the edge modes, $\tpo$ is quantized to multiples of $1/2$.
-Note that, since the states that fill the gap are localized in the bulk, they will not appear in the EOS and one can still easily identify the appearance of virtual topological edge states (see Fig.~\ref{disorder_chern}(b)).
+Note that, since the states that fill the gap are localized in the bulk, they will not appear in the EOS and one can still easily identify the appearance of virtual topological edge states (see figure~\ref{disorder_chern}(b)).
 In contrast, the topological edge states in the energy spectrum are completely masked by the (localized) bulk modes that fill the gap.
@@
 As opposed to the energy spectrum, the gap in the EOS does not fill with states such that the virtual topological edge states are still easily identifiable.
-}
+We show the two eigenvalues closest to $1/2$ in blue, and the next two in red.f }
 	\label{disorder_chern}
@@
 \subsection{Chern Number in 2D}
-%
-%We have shown how $\tpo$ simplifies the computation of Chern numbers in 1D chains. This can be generalized to 2D systems. Consider the following model for a Chern insulator with inversion symmetry,
-%\begin{align}
-%H =& \sum_{im\alpha,jn\beta} c_{im\alpha}^\dagger H_{im\alpha,jn\beta} c_{jn\beta}, \\
-%H_{im\alpha,jn\beta}=& \frac{1}{2}(i\sigma_x-\sigma_z)\delta_{i,j+1}\delta_{mn}+\frac{1}{2}(-i\sigma_x-\sigma_z)\delta_{i,j-1}\delta_{mn} \nonumber\\
-%+&\frac{1}{2}(i\sigma_y -\sigma_z)\delta_{ij}\delta_{m,n+1} +\frac{1}{2}(-i\sigma_y -\sigma_z)\delta_{ij}\delta_{m,n-1} \nonumber\\
-%+&(2-m)\sigma_z\delta_{ij}\delta_{mn},
-%\label{eq:chern_model}
-%\end{align}
-%where $i,j$ describe the $x$ direction and $m,n$ the $y$ direction.
-The system with periodic boundary conditions is defined on a torus.
-In order to apply the method of computing $\tpo$ described in the last section we will open the torus into a cylinder periodic in $y$.
-We then fourier transform only the $y$ direction to obtain
+
 We now discuss  the computation of Chern numbers in 2D systems with $\tpo$. Consider the Chern insulator with inversion symmetry given by the Hamiltonian
@@
 \caption{(a) and (b) EOS of the Chern insulator in the $C=1$ ($m=1$) and $C=-1$ ($m=3$) phases, respectively.
-In gray and light gray lines we plot two intermediate points of the adiabatic process of opening the chain between sites $1$ and $L$, showing how the midgap state of the left virtual cut evolves towards the bulk bands, introducing a discontinuity in momentum.
-(c) $\tpo-L/2$ and (d) $\tilde{\mathcal{P}}$ mod $1$.
+In gray and light gray lines we plot two intermediate points of the adiabatic process of opening the chain between sites $1$ and $L$, showing how the midgap state of the right virtual cut evolves towards the bulk bands, introducing a discontinuity in momentum.
+(c) $\tpo-L/2$ .
 Computed for the Chern insulator given by the model in Eq.~\eqref{eq:chern_model} for $L=40$ sites.
@@
 
-%
-%
-%To ilustrate the advantage of using $\tpo$ over $\tilde{\mathcal{P}}$ we also show the latter in the same parameter space.
-In order to obtain the Chern number via $\tilde{\mathcal{P}}$ for PBC there is no way of avoiding having to compute the derivative and subsequent integration of
-%\begin{equation}
-%C(m) = \int_{-\pi}^{\pi}\frac{dk_y}{2\pi} \partial_{k_y} \tilde{\mathcal{P}}(m,k_y),
-%\end{equation}
-%as we no longer know in advance where the branch cuts will appear, due to it being defined modulo 1.
-%
-
 % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - % - %
@@
 
-It was previously known that the bulk polarization (Zak phase, or geometric phase for $U(1)$ flux insertion) was encoded in the many-body ES.
-Here we have shown how it can be obtained from the single-particle entanglement spectrum of a non-interacting gapped system, showing that there is substantially more information in the single-particle ES than previously realized.
-Our formulation is both simpler than the one involving the many-body ES, which is much more difficult to compute, and it provides additional insight into the ES and EOS.
+It was previously known that the bulk polarization (Zak phase, or geometric phase for $U(1)$ flux insertion) was encoded in the ES.
+Here we have shown how it can be obtained from the single-particle ES of a non-interacting gapped system, even when it is not quantized, showing that there is substantially more information in the single-particle ES than previously realized.
+Our formulation is both simpler than the one involving the ES, which is much more difficult to compute, and it provides additional insight into the single-particle ES and EOS.
 In particular in the topological case, the relation between the quantized bulk polarization (or the Chern number in 2D) and the number of virtual topological edge states is demonstrated in a particularly transparent way.
@@
 This quadrupole moment is defined via the position operator and it is an open question whether there is a formulation in terms of twisted boundary conditions.
-In this case, one might be able to construct this quadrupole moment using the entanglement spectrum, as we did
+In this case, one might be able to construct this quadrupole moment using the ES, as we did
  here for the bulk polarization.
@@
 &(1-\xi_{p'})\sum_{\substack{\{n_{i \neq p,p' \in A_L}\} = 0,1 \\ n_{p'}=0}} \left(\prod_{k \neq p,p'\in A_L} \lambda_k\right) \nonumber \\
-&+ \xi_{p'}\sum_{\substack{\{n_{i \neq p,p' \in A_L}\} = 0,1 \\ n_{p'}=1}} \left(\prod_{k \neq p,p'\in A_L}\lambda_k\right) =1,
+&+ \xi_{p'}\sum_{\substack{\{n_{i \neq p,p' \in A_L}\} = 0,1 \\ n_{p'}=1}} \left(\prod_{k \neq p,p'\in A_L}\lambda_k\right) =1.
 \end{align}
-which we can see is $1$ if we continue this procedure for all other occupation numbers. Therefore in terms of the spectrum of $C_A$ the bulk polarization greatly simplifies to 
+If we continue this procedure for all other occupation numbers we arrive at
 \begin{align}
@@
 \end{align}
+In terms of the spectrum of $C_A$ the bulk polarization simplifies greatly.
 As mentioned above, the subspace $A_L$ is not well-defined, as one can always include bulk modes, however since the bulk polarization is defined modulo 1, this issue is irrelevant.
@@
 
-\iffalse
-\mariac{removed appendix D}
-\section{Twisted boundary condition Winding number}
-\carlos{Work in progress}
-For a system with chiral symmetry the Hamiltonian can be written in some basis as
-\begin{equation}
-H = \mqty(0 & h \\ h^\dagger & 0),
-\end{equation}
-where $h$ is in general a matrix. In translationally invariant systems the winding number is defined as 
-\begin{equation}
-\nu = -i\int_0^{2\pi} \frac{dk}{2\pi} {\rm Tr}[h(k)^{-1} \partial_k h(k)],
-\label{eq:winding_k}
-\end{equation}
-where $h(k)$ is the off-diagonal block of the Bloch Hamiltonian. In systems without translational invariance, one can similarly define the winding number
-\begin{equation}
-\nu = -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[h(\Phi)^{-1} \partial_\Phi h(\Phi)],
-\label{eq:winding_flux}
-\end{equation}
-for a $U(1)$ flux introduced (\carlos{I have only found a similar expression in the context of non-hermitian systems \cite{Gong2018}, but I assume this is the expression Prodan refers to.
-It gives the correct results in translationally invariant systems, however it is very unpractical.}).
-Note that a generalization of \eqref{eq:winding_k} to position space has also been developed which not necessarily gives the same results as \eqref{eq:winding_flux} (\carlos{The only comparison is in the Prodan paper where they talk about practical numerical implementations but don't address if both formulations are equivalent.}) Both formulations however reduce to \eqref{eq:winding_k} for translationally invariant systems \cite{Gong2018}.
-In the following we will omit the flux parameter for simplicity.
-
-Note that the winding number can also be expressed in terms of the flat-band Hamiltonian
-\begin{align*}
-Q =& H (H^2)^{-1/2} \\
-=& \mqty(0 & q \\ q^\dagger & 0).
-\end{align*}
-Note that since $Q$ only has eigenvalues $\pm 1 $ , $Q^2 = \mathbb{I}$ and therefore $q^\dagger = q^{-1}$. We can therefore write
-\begin{equation}
-\nu = -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[q^\dagger \partial_\Phi q].
-\end{equation}
-In this basis the chiral symmetry operator is given by $S = \sigma_z \otimes \mathbb{I}$ such that the winding number can be expressed in terms of the full flat-band Hamiltonian as
-\begin{align}
-\nu =& \frac{i}{2}\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S Q^\dagger \partial_\Phi Q] \\
-=& \frac{i}{2}\int_0^{2\pi} \frac{d\Phi}{2\pi} \left({\rm Tr}[q \partial_\Phi q^\dagger]-{\rm Tr}[q^\dagger \partial_\Phi q] \right)\\
-=& -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[q^\dagger \partial_\Phi q] 
-\end{align}
-where in the last step we use the fact that the winding number is a real number. We can now write the winding number in terms of the correlation matrix
-\begin{align}
-\nu =& \frac{i}{2}\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S (1-2C) \partial_\Phi (1-2C)] \\
-=& -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S (1-2C) \partial_\Phi C] .
-\end{align}
-Introducing the flux via twisted boundary conditions allows us to easily construct the ground state with flux in the case of the open chain \eqref{eq:gauge}, therefore we will employ this gauge to compute the correlation matrix.
-The correlation matrix can be written in position space as
-\begin{equation}
-C = \mqty(C_A & C_{AB} \\ C_{AB}^\dagger & C_B),
-\end{equation}
-where in this basis, assuming $S$ is a local symmetry it is also diagonal.
-Note that the flux will only appear in the off-diagonal blocks $C_{AB}$ and therefore the first term in the trace above will vanish, leaving the winding number as
-\begin{align}
-\nu =&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S C \partial_\Phi C] \\
-=&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} \left( {\rm Tr}[S C_{AB} \partial_\Phi C_{AB}^\dagger]+{\rm Tr}[S C_{AB}^\dagger \partial_\Phi C_{AB}] \right)
-\end{align}
-We can not explicitly construct the correlation matrix with flux as
-\begin{align*}
-C_{ij}^\Phi =& \bra{\tilde{\Psi}_{0,{\rm open}}^\Phi}c_i^\dagger c_j \ket{\tilde{\Psi}_{0,{\rm open}}^\Phi} \\
-=&\bra{\tilde{\Psi}_{0,{\rm open}}^0}e^{i\Phi \hat{N}_A} c_i^\dagger c_j e^{-i\Phi \hat{N}_A}\ket{\tilde{\Psi}_{0,{\rm open}}^0} .
-\end{align*}
-Depending on if $i,j$ belong to $A$ or not, commuting the exponentials will leave additional factors or not, resulting in
-\begin{align*}
-C^\Phi = \mqty(C_A^0 & e^{i\Phi} C_{AB}^0 \\e^{-i\Phi} C_{AB}^{0\,\dagger} &C_B^{0}),
-\end{align*}
-where we have made explicit the flux dependence. The winding number is then
-\begin{align}
-\nu =&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} {\rm Tr}[S C^\Phi \partial_\Phi C^\Phi] \\
-=&  -i\int_0^{2\pi} \frac{d\Phi}{2\pi} \left( {\rm Tr}[-iS C_{AB}^0 C_{AB}^{0\,\dagger}]+{\rm Tr}[iS C_{AB}^{0\,^\dagger} C_{AB}^0] \right) \\
-=&  {\rm Tr}[S C_B^0(1-C_B^0)]-{\rm Tr}[S C_A^0(1-C_A^0)] 
-\end{align}
-where in the last step we use one of the properties of the correlation function stemming from $C^2 = C$.
-(\carlos{So far this expression gives the correct result for translationally invariant systems however one needs to do the trick of breaking chiral symmetry slightly.
-However, unlike with $\tilde{\mathcal{P}}$ this expression is not discontinuous at $\kappa = 0$.
-})
-\fi
-
 \end{document}

### 2021-04-16 polarization_ES_prb_resubmission/polarization_ES.tex
Add files via upload

--- 2021-04-14 polarization_ES_prb_submission/polarization_ES.tex
@@
 One property of the subsystem correlation matrix we will make use of later is that, for sufficiently large systems, its eigenstates are found to be exponentially localized on either virtual edge if $\xi$ is away from $0$ and $1$, or they are found to be bulk modes if the correspondent eigenvalues are exponentially close to $\xi = 0,1$ \cite{Peschel2008}.
+The latter can be understood from the fact that the eigenstates of $C$ can be expressed as Wannier states.
+The Wannier states for the occupied ($\xi=1$) or empty ($\xi=0$) bands that are contained within region $A$ will also be eigenstates of $C_A$, with only a small deviation due to their exponential tails.
+However, in practice, all these eigenstates mix at $\xi=0,1$, forming the two bulk bands.
 
@@
 This case is trivial, since we know that there is one left and one right eigenstate, but if the degeneracy is four-fold or higher we must obtain the localized eigenstates.
-Localizing the bulk eigenstates that are exponentially close to $\xi = 0,1$ is more difficult but not doing so only results in an exponentially small error.
+One way of counting how many left eigenstates are in a degenerate subspace is discussed in Appendix \ref{app:loc}.
+Note that this is not necessary for the bulk eigenstates, exponentially close to $\chi=0,1$, as the error we make is also exponentially small once we take the result modulo 1.
 
@@
 \begin{equation}
-\ketpsio{\Phi} =e^{-i \Phi \hat{N}_A} \ketpsio{0}.
+\ketpsio{\Phi} =e^{-i \Phi \hat{N}_A} \ketpsio{0},
 \label{eq:gauge}
 \end{equation}
+where $\hat{N}_A$ is the number operator for particles in region $A$.
+
 When introducing the flux in the system with periodic boundary conditions (See Eq.~\eqref{eq:Ground_state_flux}) there is an ambiguity due to the bulk modes. 
 This ambiguity disappears when we introduce the flux as above in the open chain --- we simply include all the bulk modes. 
-Note that $\Phi$ is not a magnetic flux but should be treated as an additional parameter, and therefore $\tpo$ is not the polarization of the open chain.
 $\tpo$ can be obtained in an even simpler way than $\tilde{\mathcal{P}}$ as 
@@
 
-The idea is that in the process of opening the chain, which can be done adiabatically, all eigenvalues that are related to the right virtual edge ($A_R$ in Fig.~\ref{fig:chain}) are pushed to $\xi=0,1$.
+Note that the result is very easy to interpret as well.
+Changes in the polarization $\tilde{\mathcal{P}}$ on the periodic chain give the current across the seam.
+In the open chain, this is simply given by $J = \partial_t N_A(t)$.
+Therefore the equivalent polarization in the open chain is simply the number of particles in region $A$, $N_A$.
+It is, however, useful to define it in terms of the flux, as we are interested in studying the polarization as a geometric quantity.
+This description is more subtle when states cross the chemical potential, where $N_A$ will have a discontinuity without an associated current.
+Discontinuities in the trace of $C_A$ for open chains was already observed in the context of Chern insulators \cite{Alexandrinata2011}, where it was related to the appearance of gapless modes.
+Below we will employ similar arguments to study the topology of 1D chains.
+
+The key idea to understand the connection of the EOS for open and periodic chains is that in the process of opening the chain, which can be done adiabatically, all eigenvalues that are related to the right virtual edge ($A_R$ in Fig.~\ref{fig:chain}) are pushed to $\xi=0,1$.
 As a result, the sum of all eigenvalues of $C_{A,o}$ (modulo 1) is equal to $\chi$ \eqref{eq:chi_left}.
@@
 
-The relation between the zero-modes and discontinuities in $\Tr[C_{A,o}]$ was already observed in the context of Chern insulators \cite{Alexandrinata2011}.
 The appearance of the branch cut is the result of the degenerate point at $(m=1,\kappa=0)$ having an associated Chern number $C=1$ \cite{Asboth2016}.
@@
 
-We now discuss  the computation of Chern numbers in 2D systems with $\tpo$. Consider the Chern insulator with inversion symmetry given by the Hamiltonian
+We now discuss the computation of Chern numbers in 2D systems with $\tpo$ by using the same method we described for 1D systems.
+A similar (related?) discussion can be found in ref.\cite{Alexandrinata2011}.
+ Consider the Chern insulator with inversion symmetry given by the Hamiltonian
 \begin{align}
@@
 
+\section{Degenerate EOS}\label{app:loc}
+
+In order apply Eq.\eqref{eq:chi_left} one needs to identify to which virtual cut each eigenstate belongs to.
+In the case where the EOS is degenerate, eigenstates from the two different virtual cuts might mix, making this identification not straightforward .
+One solution to this problem is as follows.
+Consider the subspace spanned by the degenerate eigenstates $\ket{\psi_\mu}$.
+These are a mixture of states $\ket{\tilde{\psi}_\nu}$ that are localized either on the left or on the right half of region $A$.
+Consider the projector onto this subspace,
+\begin{align}
+P_{deg} =& \sum_\mu \ket{\psi_\mu}\bra{\psi_{\mu}} \nonumber \\
+=&  \sum_{\nu \in L} \ket{\tilde{\psi}_\mu}\bra{\tilde{\psi}_{\mu}}+ \sum_{\nu \in R} \ket{\tilde{\psi}_\mu}\bra{\tilde{\psi}_{\mu}} 
+\end{align}
+We want to count the number of these localized eigenstates in region $L$. This can be done simply by taking the trace over region $L$, such that
+\begin{align}
+\Tr_L [P_{deg} ]=&  \Tr_L[ \sum_{\nu \in L} \ket{\tilde{\psi}_\mu}\bra{\tilde{\psi}_{\mu}}]+ \Tr_L[ \sum_{\nu \in R} \ket{\tilde{\psi}_\mu}\bra{\tilde{\psi}_{\mu}} ] \\
+=&  \sum_{\nu \in L} \bra{\tilde{\psi}_{\mu}} \ket{\tilde{\psi}_\mu} .
+\end{align}
 \section{Winding number in the SSH chain}

### 2021-04-19 polarization_ES_prb_resubmission/polarization_ES.tex
Add files via upload

@@
 One way of counting how many left eigenstates are in a degenerate subspace is discussed in Appendix \ref{app:loc}.
-Note that this is not necessary for the bulk eigenstates, exponentially close to $\chi=0,1$, as the error we make is also exponentially small once we take the result modulo 1.
+Note that this is not necessary for the bulk eigenstates, exponentially close to $\xi=0,1$, as the error we make is also exponentially small once we take the result modulo 1.
 
@@
 Discontinuities in the trace of $C_A$ for open chains was already observed in the context of Chern insulators \cite{Alexandrinata2011}, where it was related to the appearance of gapless modes.
-Below we will employ similar arguments to study the topology of 1D chains.
+Below we will employ similar arguments to study topology and topological charge pumping on 1D chains.
 
@@
 In reference \cite{Alexandrinata2011} it is shown to be equal to the Chern number by relating it to the Hall current.
-Here, the connection between the trace index and the Chern number is seen by framing $\Tr[C_{A,o}]$ as a geometric quantity
+In this section we have shown numerically how the EOS from the closed and open chain are related for a Chern insulator.
+Applying the methods from last section we recover the connection between the trace index and the Chern number, seen here by framing $\Tr[C_{A,o}]$ as a geometric quantity
 
@@
 \end{align}
+
+As long as the states considered are exponentially localized on the edges, $\Tr_L [P_{deg} ]$ will be exponentially close to an integer.
+
 \section{Winding number in the SSH chain}

### 2021-04-20 polarization_ES_prb_resubmission/polarization_ES.tex
small changes

@@
 
-We now discuss the computation of Chern numbers in 2D systems with $\tpo$ by using the same method we described for 1D systems.
-A similar (related?) discussion can be found in ref.\cite{Alexandrinata2011}.
+We now discuss the computation of Chern numbers in 2D systems with $\tpo$ by using the same method we described for 1D systems. A similar  discussion can be found in ref.\cite{Alexandrinata2011}.
  Consider the Chern insulator with inversion symmetry given by the Hamiltonian
@@
 In this section we have shown numerically how the EOS from the closed and open chain are related for a Chern insulator.
-Applying the methods from last section we recover the connection between the trace index and the Chern number, seen here by framing $\Tr[C_{A,o}]$ as a geometric quantity
+Applying the methods from the last section we recover the connection between the trace index and the Chern number, seen here by framing $\Tr[C_{A,o}]$ as a geometric quantity
 
@@
 
-In order apply Eq.\eqref{eq:chi_left} one needs to identify to which virtual cut each eigenstate belongs to.
-In the case where the EOS is degenerate, eigenstates from the two different virtual cuts might mix, making this identification not straightforward .
-One solution to this problem is as follows.
-Consider the subspace spanned by the degenerate eigenstates $\ket{\psi_\mu}$.
+In order apply Eq.\eqref{eq:chi_left} one needs to identify  which virtual cut each eigenstate belongs to.
+In the case where the EOS is degenerate, eigenstates from the two different virtual cuts might mix, making this identification not straightforward.
+One solution to this problem is as follows: Consider the subspace spanned by the degenerate eigenstates $\ket{\psi_\mu}$.
 These are a mixture of states $\ket{\tilde{\psi}_\nu}$ that are localized either on the left or on the right half of region $A$.
@@
 \end{align}
-
 As long as the states considered are exponentially localized on the edges, $\Tr_L [P_{deg} ]$ will be exponentially close to an integer.
